### 2015-12-10 draft.tex
add tex code and PDF for version of paper submitted to JIE in May 2014

\documentclass[10pt]{article}

\addtolength{\textwidth}{1.3in}
\addtolength{\oddsidemargin}{-.65in} %left margin
\addtolength{\evensidemargin}{-.65in}
\setlength{\textheight}{9in}
\setlength{\topmargin}{-.5in}
\setlength{\headheight}{0.0in}
\setlength{\footskip}{.375in}
\renewcommand{\baselinestretch}{1.0}
%\linespread{1.5}
\usepackage{setspace}
\doublespacing

\usepackage{titling}

\usepackage[pdftex]{graphicx}

\usepackage[usenames,dvipsnames]{color}
\usepackage{cite}
\usepackage{times, verbatim,bm,pifont,pdfsync}
\usepackage{caption}
\usepackage{subcaption}

% disables chapter, section and subsection numbering
%\setcounter{secnumdepth}{-1}

\usepackage{amsbsy,amssymb, amsmath, amsthm, MnSymbol,bbding}
\usepackage[hang,flushmargin]{footmisc}

\usepackage[pdftex,
bookmarks=true,
bookmarksnumbered=false,
pdfview=fitH,
bookmarksopen=true]{hyperref}

\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem*{lemma*}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{assumption}{Assumption}
\newtheorem{fact}{Fact}
\newtheorem{result}{Result}

\newcommand{\ve}{\theta}
\newcommand{\ta}{\theta}
\newcommand{\ov}{\overline}
\newcommand{\un}{\underline}
\newcommand{\al}{\alpha}
\newcommand{\Ta}{\Theta}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\Bt}{B(\bm{\tau^a})}
\newcommand{\bta}{\bm{\tau^a}}
\newcommand{\bte}{\bm{\tau^E}}
\newcommand{\btn}{\bm{\tau^n}}
\newcommand{\ga}{\gamma}

\begin{document}
\title{\vskip-0.6in Trade Agreements in the Shadow of Lobbying\footnote{Formerly circulated under the title ``Trade Agreements, Lobbying and Separation of Powers''}}
\thanksmarkseries{alph}
\author{Kristy Buzard\thanks{Syracuse University, Economics Department, 110 Eggers Hall, Syracuse, NY 13244, USA. Ph: 315-443-4079. Fax: 315-443-3717. Email: kbuzard@syr.edu. http://faculty.maxwell.syr.edu/kbuzard.}} 
\date{\vskip-.1in \today}
\maketitle

\begin{center} {\bf Abstract} \end{center}

\begin{quote}
{This paper presents a model of international trade agreements in which the executive branches of each government negotiate agreements while the legislative branches, subject to political pressure from firms, can disrupt them. Lobbying is in the style of Grossman and Helpman (1994) with a new feature: all actors face uncertainty arising from the complexity of the legislative process. I demonstrate that the lower the executives set tariffs in a trade agreement, the more effort lobbies put forth to prevent its ratification. Thus trade agreements act as a domestic political commitment device: executives set relatively high tariffs to discourage lobbying and increase the chance that the agreement will be ratified. This reconciles the result from tests of Grossman and Helpman's model that protection levels are high relative to contributions given estimates of governments' social-welfare weights.

\textit{JEL classification:} D72, D80, F13, F53, F55 \\
\textit{Keywords:} trade policy, lobbying, political economy, protection, international trade agreements, uncertainty}
\end{quote}

\bigskip
\section{Introduction}

%http://www.whitehouse.gov/sites/default/files/fact_sheet_increasing_us_auto_exports_us_korea_free_trade_agreement.pdf

Empirical investigations of Grossman and Helpman's (1994) `Protection for Sale' model (henceforth `PFS') broadly support the predictions of the theory qualitatively, yet their quantitative estimates present several interesting puzzles.\footnote{cfr. Goldberg and Maggi (1999); Gawande and Bandyopadhyay (2000); Mitra, Thomakos and Ulubasoglu (2002); McCalman (2004).\label{fn:gh_lobby}} These studies have consistently found the weight governments place on social welfare to be many times that which they place on lobbying effort, while estimates indicate that the deadweight losses caused by trade distortions are several orders of magnitude larger than lobbying expenditures.\footnote{Feenstra (1992) assembles estimates from the mid-1980s that place a floor of $\$$8 billion per year on the deadweight losses from protection, whereas Bombardini and Trebbi (2012) find that total annual lobbying expenditure on trade issues in 1999-2001 when this data first became available was about $\$$200 million.} This raises the question: how could governments value social welfare so highly, yet grant these quantities of protection at such a low price?

Attempts to more fully model the lobbying process have successfully reduced the parameter estimates for the government's welfare considerations somewhat, but the smallest estimates still indicate that the U.S. government, for instance, values social welfare about twenty times as much as contributions.\footnote{cfr. Gawande, Krishna and Robbins (2006); Mitra, Thomakos and Ulubasoglu (2006); Bombardini (2008); and Gawande, Krishna and Olarreaga (2012) among others.} Taking a different approach, Gawande and Hoekman (2006) suggest that one can reconcile the empirical results with the model by acknowledging the complexity of the policy-making process and the uncertainty that arises from it.

Since so much of trade policy is determined in the context of international agreements that must be ratified by domestic legislative bodies, I incorporate PFS-style lobbying into a model of trade agreements in which the executive branches of the governments set tariff levels in anticipation of political pressure upon the legislatures. This political pressure may be interpreted either as lobbying against ratification of the trade agreement or lobbying to break the agreement once in force. I find that the separation-of-powers government structure can explain the empirical puzzle surrounding governments' welfare weights and that the addition of political uncertainty adds important realism so that the model presented here can speak to important trade policy-making events such as the four-year delay in the ratification of the U.S.-Korea Free Trade Agreement.

Consider first the case of political certainty, and assume that the executive is more pro-social than the legislature, as has been the case throughout the post-war period in the United States.\footnote{The standard assumption that the executive is less protectionist than the legislature is a special case of the finding that susceptibility to special interests generally declines with constituency size. One simple illustration from the realm of trade policy is the following: a legislator whose district has a large concentration of a particular industry does not take into account the impact of tariffs on the welfare of consumers in other districts, while the executive, whose constituency encompasses the whole country, will internalize these diffuse consumption effects. See Lohmann and O'Halloran (1994) for a detailed argument and Grossman and Helpman (2005) for a formal model that demonstrates the point for the legislature.\label{fn:ga_l_e}} For any trade-agreement tariff set by the executives, the lobby can calculate the effort level required to prevent the agreement from being ratified: it must shift the preferences of the median legislator just far enough so that he will choose the non-cooperative tariffs over the trade agreement tariffs. Because the lobby will pay this price so long as its benefit outweighs the cost, a trade war will occur unless the executive sets the trade-agreement tariff so that lobbying is not worthwhile.\footnote{As will be made clear below, non-ratification is equivalent to breaking a trade agreement once in place; likewise, the non-cooperative tariffs that result from failing to ratify the trade agreement are the same as the tariffs that would arise in trade war, so this model applies to both situations.}

Thus, any trade agreement under political certainty involves a tariff high enough to disengage the lobby completely; that is, even welfare-maximizing executives set high tariffs and induce zero contributions in equilibrium because they set policy in the shadow of a ratification decision influenced by lobbying. Careful modeling of the political process demonstrates that, while tariff-setting behavior is closely linked to deadweight-loss calculations, the empirical estimates of the responsiveness of protective measures to import penetration and the elasticity of import demand do not necessarily map directly onto the PFS parameter for welfare-mindedness.

In the non-unitary government model, these estimates and the earlier-cited stylized facts become consistent with each other. The ``case of the missing contributions,'' as Gawande and Hoekman (2006) refer to it, turns out to be well explained by an equilibrium phenomenon arising from the agenda-setting power of the executive branch. In particular, this model can explain the approximately 15$\%$ of sectors that receive protection in spite of apparently putting forth no lobbying effort.

The fact that \textit{zero} contributions are predicted in equilibrium is not particularly satisfying. The addition of political uncertainty provides the missing realism, smooths the results and delivers additional intuition. I assume all actors are uncertain about the weight the median legislator in a non-unitary legislature places on the profits of the lobbying industry and that this is attributable to the complexity of the legislative process; say, for instance, because a small number of lobbied legislators cannot deliver the votes of their non-lobbied colleagues with certainty. In this case, I show that the legislatures disrupt trade agreements with a higher probability and set higher trade war tariffs when lobbying effort increases. Because the lobbies respond to higher trade agreement tariffs by decreasing effort (and therefore the probability the agreement will fail to be ratified or will be broken), the executives must trade off the level of social welfare derived while a trade agreement is in force---since they prefer low tariffs---and the chances that the agreement will remain in force.

The executives therefore set trade-agreement tariffs with an eye to discouraging lobbying activity and the accompanying probability of abrogation. It can do this because the trade agreement increases lobbying costs: the legislature now has to be compensated for the loss of the reduced foreign tariff if it breaks the trade agreement. Moreover, in addition to the standard terms-of-trade internalization, the executive's agenda-setting power allows it to reduce tariffs until lobbying activity is at the optimal level relative to a non-cooperative setting, reducing the amount of political pressure faced by the legislatures and the resultant expected tariffs.

One contribution of this model, therefore, is the idea that trade agreements can act as a kind of domestic political commitment device. Rather than a time inconsistency such as that in Maggi and Rodr\'{i}guez-Clare (2007) where the welfare of the unitary policymaker is improved via commitment to a trade agreement by changing the investment decisions of firms, here we have an inconsistency between the preferences of the executive and legislative branches.\footnote{Indeed, depending on the modeling choices and parameters, the utility of the legislative branch taken as a whole may be lower under the trade agreement scenario.} Although the legislature has the final word both with and without the trade agreement, the executives are able to use commitment to the trade agreement to reduce lobbying effort, thus muting the legislature's protectionist bias and bringing the outcome more in line with their relatively liberal preferences.

To the best of my knowledge, this is the first attempt to model lobbying specifically aimed at derailing trade agreements. Although this kind of politically-driven failure is not rare, prior models have not allowed exploration of the endogenous dynamics behind them. Careful consideration of the political process seems to be important for understanding why governments so often fail to cooperate. The recent ratification of the U.S.-Korea Free Trade Agreement is a case in point. The agreement, as originally negotiated in 2007, was not able to be ratified in the United States until President Obama gained concessions for U.S. automakers, after which they removed their objections and the agreement was ratified by Congress in 2011 and entered into force in 2012. In a certain world, this ratification failure should not have occurred, and it was remedied by making the trade agreement more appealing to a key lobby and thus reducing its lobbying effort. \footnote{See \url{http://www.ustr.gov/trade-agreements/free-trade-agreements/korus-fta}. In this case, automakers are both importers and exporters, and the concessions were concentrated on market access for auto exports in South Korea. For simplicity, the model presented herein focuses on lobbying by an import-competing industry in a model of interindustry trade; the central insights do not depend on this structure and can be extended to the case of multiple lobbies and/or intraindustry trade. In particular, adding an export lobby in the simplest case of no uncertainty predicts that import lobbies will remain disengaged while the executive will be able to reduce tariffs by recruiting support from exporters. The Panama Trade Promotion Act seems to follow this pattern, where there was virtually no lobbying against, significant lobbying in favor, and numerous industries receiving import tariff protection.\label{fn:lobby}}

Moreover, I show that trade-agreement tariffs, lobbying effort and the probability of trade disruptions vary in nuanced ways with the amount of political uncertainty present: they are influenced strongly by lobbying incentives, but not in the straightforward ways predicted by models with unitary governments. Saiegh (2009) argues that political uncertainty is in part related to governmental and political structure. Although they do not discuss uncertainty specifically, Gawande, Krishna and Olarreaga (2009) provide cross-country empirical evidence that measures such as the number of checks and balances on the power of the legislature and the gap between the policy positions of the main political parties have predictive power for the PFS ``welfare-mindedness'' parameter. Thus, although empirical tests are outside the scope of this paper, there is suggestive evidence that political uncertainty is indeed an important factor empirically.

Ultimately, it is the addition of a rich structure of government, tailored carefully to the trade policy making process, that resolves the main empirical puzzle surrounding the PFS model: in this framework, even a welfare-maximizing executive and lobbying effort at zero can be consistent with high tariff levels.

\subsection{Related Literature}
The foundations of this work rest on Grossman and Helpman's (1994) `Protection for Sale.' Two related papers `Trade Wars and Trade Talks' (Grossman and Helpman (1995b)) and `The Politics of Free-Trade Agreements' (Grossman and Helpman (1995a)) first employed the PFS approach in the study of trade agreements. The first considers ``Trade Wars'' and ``Trade Talks'' separately, whereas my approach allows the ``Trade Talks'' to be influenced by the desire to avoid a ``Trade War'' and thus views the ``Trade War'' as a crucial subgame. Their model of preferential trade agreements (Grossman and Helpman (1995a)), although treating the government as a unitary actor, is closer to this approach.

The extensive literature that explores the empirical implications of the PFS model is well-summarized by Gawande and Magee (2012). After the first generation of work consistently found very high estimates for the government's welfare-mindedness (see footnote~\ref{fn:gh_lobby} above), a second generation has explored several modifications of the model and estimating techniques. Most of these have focused on an improved accounting of the lobbying process, such as including lobbying by up- and downstream firms (Gawande, Krishna and Olarreaga (2012)), better identifying which firms in an industry lobby (Bombardini (2008)), taking into account lobbying by foreign interests (Gawande, Krishna and Robbins (2006)) or changing the way in which organized sectors are determined (Mitra, Thomakos and Ulubasoglu (2006)). Indeed, Imai, Katayama and Krishna (2009) argue that many of the classification schemes present serious challenges for the validity of results.

Another related literature considers the impact of exogenously-determined political uncertainty on the potential for trade cooperation. These studies (e.g. Feenstra and Lewis (1991); Milner and Rosendorff (2001); Bagwell and Staiger (2005); Beshkar (2010)) derive various implications for the design of international trade agreements using a Baldwin (1987)-style government welfare function with exogenous shocks to the political-pressure parameter. As explained in the next section, the model presented here is intended to be a bridge between this literature and that discussed above that focuses on endogenously-determined political economy weights. To preview, this model is closest to that of Bagwell and Staiger (2005), with two main changes: in place of their unitary government signing a trade agreement and having different preferences ex-ante and ex-post, this model has two branches of government with differing preferences, where---in the case of the legislature---the political economy weights are determined endogenously.

Several papers have studied the impact of executive/legislative interactions on international agreements in the case of exogenously-determined preferences. Mansfield, Milner and Rosendorff (2000) construct a model in which democracies trade more with each other because the domestic legislature's influence requires the foreign government to make deeper tariff cuts to ensure its cooperation. Dai (2006) typifies a second approach, which, in line with the Schelling conjecture, argues that a country's legislative constraint assists its own executive in maintaining higher barriers. Ethier (2002) treats the effect of a different separation of powers on the international trade system: that between trade negotiators and the government officials who administer protection.

The above work is patterned on the model of Milner and Rosendorff (1997), which explores how uncertainty can affect trade policy and the probability of ratification failure when political preferences, and therefore uncertainty, are exogenous;  Iida (1996) presents a similar model. Related is Le Breton and Salanie (2003), which studies lobbying when the lobby is uncertain about the preferences of a unitary decision maker. Le Breton and Zaporozhets (2007) go a step further and replace the unitary decision maker with a legislature with multiple actors. Song (2008) presents a model in which policy-making power is shared between an executive and a unitary legislature where lobbies are able to endogenously influence the political preferences of the legislature in a context of unilateral policy making with no uncertainty, and Coates and Ludema (2001) study trade policy leadership in a model with endogenous lobbying in the presence of political uncertainty with imperfect monitoring.

I will begin by describing the model in detail. Section~\ref{sec:main} then presents the main results. In Section~ \ref{sec:example}, an extended example demonstrates these results as well as the impact of changing the institutional environment. Section~\ref{sec:dis} explores the connection between the model under consideration here and the `Protection for Sale' theoretical framework and includes in-depth discussions of the roles of uncertainty and the separation-of-powers structure. Several extensions of the model are explored in Section~\ref{sec:ext} and Section~\ref{sec:concl} concludes.

\section{The Model}
\label{sec:model}
Following the literature on the political economy of trade agreements as exemplified by Bagwell and Staiger (2005), I employ a partial equilibrium model with two countries: home (no asterisk) and foreign (asterisk). The countries trade two goods, $X$ and $Y$, where $P_i$ denotes the home price of good $i \in \left\{X,Y\right\}$ and $P_i^*$ denotes the foreign price of good $i$. In each country, the demand functions are taken to be identical for both goods, respectively $D(P_i)$ in home and $D(P_i^*)$ in foreign and are assumed strictly decreasing and twice continuously differentiable.

The supply functions for good $X$ are $Q_X(P_X)$ and $Q_X^*(P_X^*)$ and are assumed strictly increasing and twice continuously differentiable for all prices that elicit positive supply. I also assume $Q_X^*(P_X) > Q_X(P_X)$ for any such $P_X$ so that the home country is a net importer of good $X$. The production structure for good $Y$ is taken to be symmetric, with both demand and supply such that the economy is separable in goods $X$ and $Y$. It is assumed that the production of each good requires the possession of a sector-specific factor that is available in inelastic supply and is non-tradable so that the income of owners of the specific factors is tied to the price of the good in whose production their factor is used.

For simplicity, I assume each government's only trade policy instrument is a specific tariff on its import-competing good: the home country levies a tariff $\tau$ on good $X$ while the foreign country applies a tariff $\tau^*$ to good $Y$. Local prices are then $P_X = P_X^W + \tau$, $P_X^* = P_X^W$, $P_Y = P_Y^W$ and $P_Y^* = P_Y^W + \tau^*$ where a $W$ superscript indicates world prices and equilibrium prices are determined by the market clearing conditions
$$M_X(P_X)= D(P_X)-Q_X(P_X) = Q_X^*(P_X^*) - D(P_X^*) = E_X^*(P_X^*)$$
$$E_Y(P_Y)=Q_Y(P_Y)-D(P_Y) = D(P_Y^*)-Q_Y^*(P_Y) = M_Y^*(P_Y^*)$$
where $M_X$ are home-county imports and $E_X^*$ are foreign exports of good $X$ and $E_Y$ are home-county exports and $M_Y^*$ are foreign imports of good $Y$.

It follows that $P_X^W$ and $P_Y^W$ are decreasing in $\tau$ and $\tau^*$ respectively, while $P_X$ and $P_Y^*$ are increasing in the respective domestic tariff. This gives rise to a standard terms-of-trade externality. As profits and producer surplus (identical in this model) in a sector are increasing in the price of its good, profits in the import-competing sector are also increasing in the domestic tariff. This economic fact, combined with the assumptions on specific factor ownership, is what motivates political activity.

I next describe the politically-relevant actors. In order to focus attention on protectionist political forces, I assume that only the import-competing industry in each country is politically-organized and able to lobby and that it is represented by a single lobbying organization (see Footnote~\ref{fn:lobby} for discussion of additional lobbies). Each country's government is composed of two branches: an executive who can conclude trade agreements and a legislature that has final say on trade policy. In summary, the political process is modeled as involving three players in each country: the lobby, the executive, and the legislature.

The timing is as follows. First, the executives set trade policy cooperatively within the context of an international agreement. I assume that an external authority can ensure enforcement of the agreement in the one-shot game analyzed here.\footnote{Buzard (2014b) extends the analysis to a repeated-game context and can therefore remove this restriction.} After the trade agreement is concluded, the lobbies attempt to persuade the legislators in their respective countries to vote ``no'' on the ratification of the agreement; or alternatively, if there is no ratification hurdle, to break the trade agreement.\footnote{As will be seen shortly, these two interpretations are equivalent in this setting because tariffs in a ``trade war'' are assumed to be set non-cooperatively in the same manner as if there were no trade agreement. For simplicity, I refer to this as the ``break'' decision throughout.} Next, the uncertainty about the the median legislator's identity is resolved and the legislatures decide whether to abide by the agreement or to provoke a trade war. In the event that the trade agreement does not enter into or remain in force, there is a final stage of lobbying, resolution of the uncertainty surrounding this decision, and voting to set the trade-war tariffs.\footnote{Alternatively, one can model the lobbying for the ratification decision as determining the potential non-cooperative tariff; this changes behavior slightly depending on the relationship between the lobby and the legislature's preferences, but does not change the fundamental dynamics of the model. Yet another possibility is to replace the trade-war phase with the ex-ante status quo, which would most naturally be the non-cooperative equilibrium. This is precisely the outcome that obtains in the trade-war phase, so the results of the model are unaltered.} Once all political decisions are taken, producers and consumers make their decisions.

Although uncertainty is present, information about it is symmetric, so the appropriate solution concept is subgame perfect Nash equilibrium. As this game is solved by backward induction, it is intuitive to start by describing the incentives of the legislatures, whose decisions I model as being taken by a median legislator. As the economy is fully separable and the economic and political structures are symmetric, I focus here on the home country and the $X$-sector. The details are analogous for $Y$ and foreign.

In modeling the objective of the legislature, I introduce a variant of the general political support function introduced by Hillman (1982). While Hillman is credited with first proposing that policy makers choose tariffs in order to maximize their own political support, the PFS model contributed the explicit specification that lobbying contributions enter the political support function with constant marginal utility as $C + aW$.

Dixit, Grossman and Helpman (1997) point out that linearity in contributions prevents complete analysis of distributional questions; the PFS form also restricts the returns to lobbying activity to be constant.\footnote{See Section \ref{sec:gh} for further discussion of this point and an in-depth discussion of the relationship between the functional form presented herein and the PFS form. Note that the assumption of diminishing returns to lobbying effort has been present in the literature going back at least to Findlay and Wellisz (1982).} They show that the PFS equilibrium analysis can be extended to a more general setting where, among other features, returns to lobbying are not restricted to be constant.

For the case of trade policy, I specify the welfare function of the home legislature as
\begin{equation}
  W_{\mathit{ML}} = \mathit{CS}_X(\tau) + \ga(e,\ve) \cdot \mathit{PS}_X(\tau) + \mathit{CS}_Y(\tau^*) + \mathit{PS}_Y(\tau^*) + \mathit{TR}(\tau)
  \label{eq:ml}
\end{equation}
where $\mathit{CS}$ is consumer surplus, $\mathit{PS}$ is producer surplus, $\ga(e,\ta)$ is the weight placed on producer surplus (profits) in the import-competing industry, $e$ is lobbying effort, $\ta$ is a random variable, and $\mathit{TR}$ is tariff revenue.

Notice that, aside from the endogeneity of the weight the legislature places on the lobbying industry's profits, this is precisely the \textit{deus ex machina} government objective function popularized by Baldwin (1987) that is commonly employed in the literature on the political economy of trade agreements. Since trade policies are overwhelmingly determined within the context of trade agreements, it is useful to have a framework to bring together the endogenous political pressure of PFS-style modeling with the trade agreements approach; the formulation in Equation~\ref{eq:ml} is intended as a bridge between the two. Buzard (2014a) for an example of how this form can be used to integrate the analysis of endogenous political activity and exogenous shocks to the political welfare of governments.

In this formulation, the weight the median legislator places on the profits of the import-competing industry, $\ga(e,\ve)$, is affected by the level of lobbying effort and a random variable $\ta$. This allows for the analysis of endogenous lobbying to be integrated with the standard political economy shocks in the trade agreements literature. Alternatively, as motivated below, it can represent reduced-form modeling of political uncertainty arising from a complex decision-making process within a legislature. Modeling the objective function so closely on the standard in the trade agreements literature allows for direct comparisons to the large extant body of work that studies exogenous shocks only, revealing cleanly the effects of the addition of endogenous lobbying---and if one wants---this different kind of political uncertainty.

I allow for the possibility of this political uncertainty, which can be interpreted as uncertainty over the identity of the median legislator, in order to model the idea that the lobbying and voting ``game'' that goes on within legislatures is often complex enough that none of the actors know precisely what the outcome will be---that is, no one can predict exactly which legislator will be the median and therefore what weight will be placed on import-industry profits at the time of the vote. Thus one way we can think about $\ta$ is as an additional influence on the legislature's preferences about which all parties---executives, legislators and lobbies alike---are uncertain. It is this interpretation of $\ta$ I take throughout the paper.

One can conceptualize this uncertainty as being a result of the lobby's strategic choice to engage with only some key members of the legislature who wield significant influence but who cannot deliver policy decisions without the cooperation of others within the legislature, which is in turn uncertain. Different assumptions on $\ga(\cdot,\cdot)$ correspond to different institutional features and will affect optimal lobbying- and tariff-setting behavior. While a structural model of the legislative process might be more satisfying, the reduced form can deliver the key insights without distracting from the central policy choice and lobbying dynamics that take place at the level of the executives and lobbies, to which we turn next.

Given its expectations and the legislature's preferences, the home lobby chooses its lobbying effort ($e_b$ to influence the break/ratification decision, which for simplicity, I will refer to henceforth simply as ``break,'' and $e_n$ to influence the trade war tariff) to maximize the welfare function:
\begin{equation}
  U_L = \Pr\left[ \text{TradeWar}(\bta,e_b,e_n) \right] \left[ \pi(\tau^{\mathit{n}}) - e_n \right] + \Pr\left[ \text{TradeAgreement}(\bta,e_b,e_n) \right] \pi(\tau^a) - e_b
  \label{eq:lv}
\end{equation}
where $\pi(\cdot)$ is the current-period profit and $\tau^a \ (\tau^\mathit{n})$ is the home country's tariff on the import good under a trade agreement (war). I use the convention throughout of representing a vector of tariffs for both countries $(\tau,\tau^*)$ as a single bold $\bm{\tau}$.

I assume the lobby's contribution is not observable to the foreign legislature. The implication is that the lobby can directly influence only the home legislature, and so the influence of one country's lobby on the other country's legislature occurs only through the tariffs selected.\footnote{cfr. Grossman and Helpman (1995b), page 685.}

In the first stage, the executives choose the trade agreement tariffs $\bta=\left(\tau^a,\tau^{*a} \right)$ via a negotiating process that I assume to be efficient. In this symmetric environment, it therefore maximizes the expected joint payoffs:
\begin{multline}
  W_E(\bta) + W_E^*(\bta) = \Pr\left[ \text{TradeWar}(\bta) \right] \left[W_E(\bm{\tau}^{\mathit{n}}) + W_E^*(\bm{\tau}^{\mathit{n}}) \right] + \\ \Pr\left[ \text{TradeAgreement}(\bta) \right] \left[W_E(\bta) + W_E^*(\bta) \right]
  \label{eq:jv}
\end{multline}

I model the executives' choice via the Nash bargaining solution where the disagreement point is the executives' welfare resulting from the Nash equilibrium in the non-cooperative game (i.e. in the absence of a trade agreement) between the legislatures.

The home executive's welfare is specified as follows:
\[
  W_E = \mathit{CS}_X(\tau) + \ga_E \cdot \mathit{PS}_X(\tau) + \mathit{CS}_Y(\tau^*) + \mathit{PS}_Y(\tau^*) + \mathit{TR}(\tau)
\]
Note that this is identical to the welfare function for the legislature aside from the weight $\ga_E$ the executive places on the profits of the import industry, which is not a function of lobbying effort. This construction permits me to integrate the influence of the import-competing industry on negotiations prior to the formation of the trade agreement while also reflecting the idea that the mechanisms through which political considerations influence the executives' preferences appear to be quite different from those at the legislative level (the electoral college and primary/caucus systems for instance).

This does \textit{not} require that the executives are not lobbied; only that their preferences are not directly altered in a significant way by lobbying over trade---that they do not sell protection in order to finance their re-election campaigns. In the case of the post-war United States, where the Congress has consistently been significantly more protectionist than the President, this seems to reasonably reflect the political reality. For trade policy, where there are concentrated benefits but harm is diffuse, there are good reasons for this to be the case. Because the President has the largest constituency possible, delegating authority to the executive branch may simply be a mechanism for ``concentrating'' the benefits since consumers seem unable to overcome the free-riding problem. In fact, a strong argument can be made that power over trade policy has been delegated to the executive branch precisely \textit{because} it is less susceptible to the influence of special interests (Destler (2005)).

Therefore, in line with both the theoretical and empirical literature, I make the following assumption:

\begin{assumption}
  $\ga(e,\ve) \geq \ga_E \geq 1 \ \forall \ve$.
  \label{as:ga_l_e}
\end{assumption}

That is, even for the least favorable outcome of the lobbying process, the legislature will be at least slightly more protectionist than the executive. This assumption is not essential, but it matches well the empirical findings that politicians with larger constituencies are less sensitive to special interests, and it simplifies the analysis considerably, in particularly ensuring that trade agreement tariffs are less than trade war tariffs.

Although the political process here most closely matches that of the United States in the post-war era,\footnote{In particular, the binary decision by the legislature about whether to abide by or break the trade agreement is modeled on the ``Fast Track Authority'' that the U.S. Congress granted to the Executive branch almost continuously from 1974-1994 and then again as ``Trade Promotion Authority'' from 2002-2007.} the model is broadly applicable because the central dynamic that emerges---that trade agreements are formed in the \textit{shadow} of lobbying pressure instead of in direct response to it---appears to be present in a wide range of countries in which the formation and maintenance of trade policy requires ratification by a politically susceptible body either at home or in a trading partner.

%I believe the one major fudge between my model and Destler's evidence is acceptable: that the executive is actually the one who breaks trade agreements, but they do it largely to prevent the legislature from doing something far worse.

\section{Main Results}
\label{sec:main}
To understand how the executives optimally structure trade agreements, we must first examine the incentives of the lobbies and how the legislatures make decisions regarding breach of the trade agreement, including how trade-war tariffs are set. The symmetric structure of the model permits restriction of attention to the home country.

\subsection{Trade-War Tariffs}
In the event that the trade agreement is broken, the legislature sets its tariff $\tau$ unilaterally by maximizing Equation~\ref{eq:ml} given $\tau^*$. Because there are no interactions between the home and foreign tariffs, the home and foreign trade-war tariffs are independent and the home country's tariff in a trade war maximizes weighted home-country welfare in the $X$-sector only. The foreign legislature's decision problem is analogous, and unilateral optimization leads to what I refer to as the (expected) Nash or trade-war tariffs $\tau^n$ as the solution to the following first order condition:\footnote{Note that the random variable in the median legislator's weight on import profits is written as $\ta_n$ at this stage to distinguish it from $\ta_b$ at the break stage: as two separate votes are taken, there are two distinct realizations of uncertainty. It is not necessary to restrict them to come from the same distribution, although in applications it would seem unlikely that the distribution would change.}
\[
		\frac{\partial \mathit{CS}_X(\tau)}{\partial \tau^n} + \ga(e_n,\ve_n) \cdot \frac{\partial \mathit{PS}_X(\tau)}{\partial \tau^n} +  \frac{\partial \mathit{TR}(\tau)}{\partial \tau^n} = 0
\]
Conditions under which the second order condition is satisfied are contained in Lemma 0 in the Appendix.

In accordance with intuition, because profits are increasing in the tariff, trade war tariffs are increasing in the weight attached to the profits of the import-competing industry, as the following result shows:

\begin{lemma}
  If prices are linear in tariffs, $\tau^n(\ga(e_n,\ve_n))$ is increasing in $\ga(e_n,\ve_n)$.
  \label{res:tninc}
\end{lemma}
All proofs are in the Appendix. Weaker conditions for this lemma are possible, but most examples in the literature satisfy this condition. The linearity serves to guarantee that the second order condition is satisifed.

\noindent In the event of a trade war, the lobby chooses its effort $e_n$ given this tariff-setting behavior by maximizing its profits net of effort: $\pi\left(\tau^n\left(\ga\left(e_n,\ve_n\right)\right)\right) - e_n$. (Note this is Equation~\ref{eq:lv} simplified by the resolution of uncertainty over the legislature's decision on the trade agreement). This implies a first order condition of
\begin{equation}
	\frac{\partial \pi(\tau^n(\ga(e_n,\ve_n)))}{\partial e_n} = 1
  \label{eq:lobtw}
\end{equation}
That is, at this stage, the lobby chooses the level of effort that equates its expected marginal increase in profits with its marginal payment. Further details about the political weighting function $\ga$ and assumed structure of uncertainty are provided in the next section.

\bigskip
\subsection{To Break or Not to Break?}
With trade-war decisions specified, we can proceed to analyze the incentives regarding the legislature's decision to uphold or break the trade agreement. I assume that the form of the political weighting function is the same in both cases, but the model could be extended to allow the two to differ. What is in principle different are the lobbying efforts toward influencing the break decision, denoted $e_b$ at this stage to distinguish them from those that are made to influence the Nash tariff, $e_n$. The realization of uncertainty about the median legislator's identity will similarly be denoted $\ve_b$ so as to distinguish it from $\ve_n$ at the trade-war stage.

I also assume that each legislature will have the opportunity to take a vote to break the agreement with probability less than one-half, and that the probability that both legislatures have the opportunity is vanishingly small. This allows me to focus here on the key interaction between the domestic actors and abstract from the strategic interaction between the lobbies in the two countries. I discuss an extension to the case in which both legislatures vote to break the agreement at the same time in Section~\ref{sec:asym}.

The legislature will break the agreement and set the tariff at $\tau^n$ if the median legislator's utility from the Nash tariffs is higher than his utility from the trade agreement tariffs, i.e. if
\begin{equation}
  W_{ML}(\btn,\ga(e_b,\ve_b)) > W_{ML}\left(\bta,\ga(e_b,\ve_b)\right)
  \label{eq:lwcg}
\end{equation}
\noindent Recall that a bold $\bm{\tau}$ represents a vector of tariffs for both countries $(\tau,\tau^*)$.

The legislature's decisions are stochastic, so the outcome of the vote on whether or not to break the trade agreement, as well as the level at which the tariff will be set in the case of a disruption, is not known to \textit{any} player until the uncertainty over the identity of the median legislator is resolved at each stage---that is, until the moment the vote takes place.\footnote{We might think of the trade-war tariff as being set in the legislation before a vote takes place in the case of administered protection for instance; the results are robust to this variant.} I represent the probability that the home legislature breaks the trade agreement and sets the tariff at $\tau^n$ as:
\begin{multline}
  b(e_b,\bta,\btn) = \expect_{\ga|e_b} \bm{1} [ W_{ML}(\btn,\ga(e_b,\ve_b)) > W_{ML}\left(\bta,\ga(e_b,\ve_b)\right) ] \\ = \Pr [ W_{ML}(\btn,\ga(e_b,\ve_b)) > W_{ML}\left(\bta,\ga(e_b,\ve_b)\right) | e_b]
  \label{eq:b}
\end{multline}

Note that the legislature's best response functions (both for whether to break the agreement, and at what level to set the tariff if it does) will be affected by the shape of $\ga(e,\ve)$. I make the following assumptions on the weight on import-industry profits:

\begin{assumption}
  $\ga(e,\ta)$ is increasing and concave in $e$ for every $\ta \in \Theta$.
  \label{as:ga_c}
\end{assumption}

\begin{assumption}
  $\ga(e,\ta)$ is increasing in $\ta$.
  \label{as:ga_ta}
\end{assumption}

\begin{assumption}
  $\ga(e,\ve) \geq \ga_E \geq 1 \ \forall \ve$.
  \label{as:ga_l_e}
\end{assumption}

Assumption~\ref{as:ga_c} formalizes the intuition that the legislature favors the import-competing industry more the higher is its lobbying effort, but that there are diminishing returns to lobbying activity.\footnote{The diminishing returns here take the form of declining increments to the lobby's influence as effort increases; in Ethier (2012), the returns to lobbying decline as the level of protection increases.} It allows the political uncertainty inherent in $\ta$ to be correlated in interesting ways with the level of $e$, ruling out only (i) that higher effort levels make lower $\ga$-weights more likely and (ii) that effort and uncertainty are correlated in such a way that increasing lobbying effort changes the structure of uncertainty so that higher $\ga$-weights become more likely at an accelerating pace.

Assumption~\ref{as:ga_ta} formalizes the intuition that larger realization of the uncertainty variable $\ta$ increase the value of the political economy weight.

Assumption~\ref{as:ga_l_e} ensures that $\tau^a < \tau^n$, and more generally, that the legislature's incentives are more closely aligned with the lobby's than are those of the executive. This is not essential but simplifies the analysis and matches well the empirical findings that politicians with larger constituencies are less sensitive to special interests (See Destler (2005) and footnote~\ref{fn:ga_l_e} above).

We are now in a position to examine the legislature's decision more closely. Of central concern is how the probability that the legislature will break the trade agreement varies with lobbying effort:

\begin{result}
  The probability that the legislature breaks a trade agreement is increasing and concave in $e_b$.
  \label{res:bincC}
\end{result}

\noindent Lobbying affects only the weight the legislature places on the profits of the import-competing industry. These profits are higher in a trade war than under a trade agreement, so given Assumption~\ref{as:ga_c}, the expectation of $\ga$ is increasing in lobbying effort, implying that the legislature becomes more favorably inclined toward the high trade-war tariff and associated profits as lobbying increases and therefore more likely to break the trade agreement. However, there are decreasing returns for the lobby: increasing effort leads to less than proportional increases in the probability that the trade agreement will be broken.

The probability of a trade war, just like the trade-war tariffs, is increasing in $\ga$, so it's the natural assumption that $\ga$ is increasing in lobbying effort that implies that the lobby can increase both its trade war tariff and the probability of a trade war by raising its lobbying effort.

Turning to the effects of the trade-agreement tariffs on the probability that the agreement will be abrogated, it is straightforward that the legislature always prefers lower levels of the foreign tariff:
\begin{lemma}
  Holding lobbying effort constant, the probability the legislature breaks a trade agreement is weakly increasing in $\tau^{*a}$.
  \label{lem:leg_astar}
\end{lemma}

Because the home country is an exporter of good $Y$, the lower world price for $Y$ has a negative effect on producer surplus that is larger than the positive effect on consumer surplus. Thus the net effect of an increase in foreign trade-agreement tariffs on legislative (and home country) welfare is negative, leading to an increased probability that the trade agreement will be broken.

The relationship between the home country's trade-agreement tariff $\tau^a$ and the legislature's probability of breaking the agreement is slightly more complex because $\tau^a$ interacts directly with $\ga(e_b,\ta_b)$. The intuition is as follows. The median legislator will vote to break the agreement if it receives higher welfare from the trade war tariff than the trade agreement tariff. For any given level of effort $e_b$, when the trade agreement tariff is very low, only a small set of realizations of $\ta_b$ that lead to the lowest values of $\ga(e_b,\ta_b)$ will lead to approval of the trade agreement. Thus the break probability is high. As the trade agreement tariff rises, larger values of $\ga(e_b,\ta_b)$ are consistent with approving the trade agreement, so the set of $\ta_b$'s that lead to trade agreement approval is larger. That is, the break probability decreases as $\tau^a$ increases. Even if the trade agreement tariff is so high that it is disliked by most, or even all, potential median legislators, it is preferable to the --- even higher --- trade war tariff, so the probability of breaking either becomes zero or continues to decrease in $\tau^a$.

\begin{lemma}
  Holding lobbying effort constant, the probability the legislature breaks a trade agreement is weakly decreasing in $\tau^a$.
  \label{res:leg_a}
\end{lemma}

In the symmetric equilibria that are focal in the analysis that follows, any increase in $\tau^a$ will be accompanied by an equal increase in $\tau^{*a}$, so it is useful to state a result about the combined impact of the home and foreign tariff when the two are constrained to be equal:
\begin{lemma}
	Holding lobbying effort constant, the probability the legislature breaks a trade agreement is weakly decreasing in $\bta$ (i.e. $\frac{\partial b(e,\bta)}{\partial  \tau^a} + \frac{\partial b(e,\bta)}{\partial  \tau^{*a}} \leq 0$).
	\label{res:bcomb}
\end{lemma}

\noindent When any change in one of the trade agreement tariffs must be met with an equal change in the other (as in a symmetric agreement), the negative impact on the legislature's propensity to break the trade agreement of raising the home tariff always outweighs the positive impact of raising the foreign tariff because of the legislature's bias toward the import-competing industry.

\subsubsection{Lobby}
\label{sec:lob_un}
The lobby chooses its level of effort $e_b$ as a function of $\bta$, given the implications of that choice on the legislature's probability of breaking the agreement, $b(e_b,\bta,\btn)$. I will henceforth suppress the Nash tariffs in the expression of the break probability to simplify notation.\footnote{Since the decisions over the Nash tariffs are taken in the final period, they do not vary from the point of view of earlier stages. Similarly, because the lobbying effort made in that period does not interact with other decision variables in earlier periods, I will suppress the dependence of the Nash tariffs on $e_n$.} Recalling that if the lobby has the chance to act, there is no chance for the foreign legislature to break the agreement, the lobby's decision is
\[
  \max_{e_b} b(e_b,\bta) \left[\pi(\tau^n) - e_n \right] + [1 - b(e_b,\bta)] \pi(\tau^a) - e_b
\]
The lobby simply maximizes probability-weighted profits net of effort. From its first order condition, we can see that the lobby balances the cost of an extra dollar of lobbying expenditure with the higher profits from a trade war weighted by the increase in the probability of receiving the higher profit level that expenditure induces:
\begin{equation}
	\frac{\partial b(e_b,\bta)}{\partial e_b} \left[ \pi(\tau^n) -e_n - \pi(\tau^a) \right] = 1 
	\label{eq:lobbyfoc}
\end{equation}
The second order condition is satisfied given Assumption~\ref{as:ga_l_e} and Result~\ref{res:bincC}. However, the solution is not guaranteed to be interior, for which we need
  \begin{equation}
	  \frac{\partial b(0,\bta)}{\partial e_b} \left[ \pi(\tau^n) -e_n- \pi(\tau^a) \right] > 1.
		\label{ine:lobint}	
  \end{equation}
For instance, if the executives were to set $\tau^a = \tau^n$, there would be no incentive for the lobby to make a positive contribution. We will see that there are some cases in which it may be in the executives' joint interest to set trade agreement tariffs so as to completely disengage the lobby. The following results thus only hold when Condition~\ref{ine:lobint} is satisfied, that is, when the marginal impact of the first lobbying dollar on the break probability is sufficiently high to make engaging in the political process worthwhile for the lobby.\footnote{This is primarily a condition on the curvature of the political weighting function but also involves the production and demand structure.}

We can now proceed to the central result concerning the lobby's behavior. 
\begin{result}
  When the trade agreement remains in force with positive probability and the distribution of $\ga(e,\ta)$ is weakly increasing in $e$, lobbying effort is weakly decreasing in trade agreement tariffs.\footnote{See Section~\ref{sec:uncertainty} for a detailed account of behavior when the lobby is able to induce a trade war with probability 1.}
  \label{res:lobby}
\end{result}

\noindent Raising the trade agreement tariff decreases the benefit of inducing a break in the trade agreement by raising profits under the trade agreement. This is in many ways the key result of this paper: the executives can reduce lobbying effort by setting higher tariffs in their trade agreement. We will see in the next section how this shapes the executives' joint decision.

Note that the symmetric increase in the foreign tariff does not impact profits in the import-competing industry, but it can impact lobbying incentives through the break probability. However, there is an analogous effect on the break probability from raising $\tau^a$ and the net effect is negative given the assumption on the probability distribution function of $\ga(e,\ta)$, which ensures that the effect of lobbying effort on the break probability does not increase as $\bta$ increases. This assumption can be significantly weakened: this effect cd be positive as long as it is smaller in absolute value than the negative effect on trade agreement profits.

\subsection{The Trade Agreement}
\label{sec:ta}
The executives choose trade agreement tariffs $\bta=\left(\tau^a,\tau^{*a} \right)$ to maximize Equation \ref{eq:jv}. I call the maximized joint value $MV(\bta)$ and write the division of surplus according to the Nash bargaining solution, which, due to symmetry, can be simplified from its general multiplicative form as
\[
  V_E(\bta) + t = W_E(\btn) + \frac{1}{2} \left( MV(\bta) - W_E(\btn) - W_E^*(\btn) \right)
\]
\[
  V_E^*(\bta) - t = W_E^*(\btn) + \frac{1}{2} \left( MV(\bta) - W_E(\btn) - W_E^*(\btn) \right)
\]
where expectation operators have been dropped for convenience and $t$ is an intergovernmental transfer.\footnote{While direct monetary transfers have to date rarely been used in practice, it seems appropriate to interpret linked concessions on non-trade issues as indirect transfers (Klimenko, Ramey and Watson (2008); Maggi and Staiger (2011)). Here, transfers do not occur as long as the full set of symmetry assumptions are maintained but are an important consideration in the environment of Section~\ref{sec:asym}.} The solution to this system of equations determines the tariffs $\bta$ that will be set in the trade agreement.

I simplify the problem by assuming that political constraints prevent the executives from choosing asymmetric tariffs. This restricts attention to symmetric solutions in the fully symmetric environment under consideration, which means that the trade agreement tariffs for which we're looking are simply those that maximize the joint welfare of the executives.\footnote{I have assumed that $W_E(\tau) = W_E^*(\tau) \ \forall \tau$; further in a symmetric equilibrium, $\tau^n =\tau^{*n}$ and $\tau^a =\tau^{*a}$, so 
\[
  V_E(\bta) + t = W_E(\btn) + \frac{1}{2} \left( MV(\bta) - W_E(\btn) - W_E^*(\btn) \right) = W_E(\btn) + \frac{1}{2} \left( MV(\bta) - 2W_E(\btn) \right) = \frac{1}{2}\expect MV(\bta)
\]}
In order to solve for them, I represent the probability that the trade agreement will be broken as $B(\bta)=b(e(\bta),\bta)$ where $e(\bta)$ is the best response function implicit in Equation \ref{eq:lobbyfoc}, the lobby's first order condition.

The problem then becomes to maximize the following modified version of Equation \ref{eq:jv}:
\begin{equation}
    \left\{ o \cdot B(\bta) + o^* \cdot B^*(\bta) \right\} \bm{W_E}(\btn) + \left\{ 1- o \cdot B(\bta) - o^* \cdot B^*(\bta) \right\} \bm{W_E}(\bta)  
  \label{eq:jv2}
\end{equation}
where I have written the sum of the home and foreign executives' utilities as $\bm{W_E}(\cdot)$. Here $o$ and $o^*$ are the probabilities that the home and foreign legislatures, respectively, will have the opportunity to break the trade agreement; these are assumed to be mutually exclusive events that are realized after the conclusion of the trade agreement. The case in which both legislatures can potentially break the agreement is treated in Section~\ref{sec:two}. One can think of $o$ and $o^*$ as being determined by events beyond the executives' or legislatures' control that determine whether or not the legislatures will be willing to consider the lobby's request; for example, they would be affected by the occurrence of an economic crisis that diverts the legislature's attention from less-pressing matters.

Of central concern is how the break probabilities behave as a function of the trade agreement tariffs given the lobby's optimal response. Recalling the definition of the total break probability, $B(\bta)=b(e_b(\bta),\bta)$, we must combine th direct effect of tariffs on the legislature's break decision with the indirect effect through the impact on the lobby's choice of $e_b$.

The direct effect is negative by Lemma~\ref{res:bcomb}. The fact that the indirect effect of raising tariffs leads to a reduction in lobbying (Result~\ref{res:lobby}) and therefore a further reduction in the propensity of the legislature to break the agreement (Result~\ref{res:bincC}) leads to the following key result: the overall probability that the trade agreement will be broken is decreasing in the trade agreement tariffs.

\begin{result}
	The total probability that the trade agreement will be broken is decreasing in $\bta$ (i.e. $\frac{\partial B(\bta)}{\partial  \tau^a} + \frac{\partial B(\bta)}{\partial  \tau^{*a}} \leq 0$).
	\label{res:bcomB}
\end{result}

\noindent That is, when the executives raise the trade agreement tariffs the legislature becomes less likely to abrogate the agreement, for three reasons. First, the legislature prefers a higher domestic tariff; second, the higher tariff discourages lobbying; finally, the lower lobbying effort reduces the legislature's preferred tariff further. Thus, beyond promising a lower tariff from the trading partner, we can think of the trade agreement as a sort of political commitment device that can be used to reduce political pressure and therefore get the legislature to maintain a lower tariff than it otherwise would.

We are now prepared to examine the executives' optimal choice of trade agreement tariffs. Because joint executive welfare is decreasing in trade agreement tariffs for $\bta$ above the executives' preferred tariffs, we have the following fundamental feature of the executives' problem:
\begin{lemma}
  The executives face the following trade off when choosing $\bta$: higher tariffs decrease the probability that the trade agreement will be broken but also decrease welfare when the agreement is in force.
  \label{res:to}
\end{lemma}
I define the tariff the executives prefer in the absence of legislative constraints as $\bm{\tau^E}$ and show in the \hyperlink{int_soln}{Appendix} that it is never optimal in a symmetric equilibrium for the executives to choose $\bm{\tau^a} < \bm{\tau^E}$ if there is a positive probability that the legislatures will have an opportunity to break the trade agreement. This is the result of an envelope-like result: there are only second-order losses from raising tariffs slightly from the most-preferred level yet there are first-order gains in reducing the probability that the agreement will be broken.

The executives will also always choose $\bta < \btn$ unless the legislature breaks even agreements with tariffs very close to the Nash level with certainty, in which case the problem is not interesting so I will ignore this case. Thus there is an interior solution in all cases of interest, but this solution may be of two different forms. First, it could be at the point that maximizes the concave portion of the executives' welfare function (Equation~\ref{eq:jv2}). Here, the optimal tariffs are chosen to balance concerns for trade-agreement welfare against those for discouraging lobbying activity aimed at breaking the agreement.

But there is an alternative. For some specifications, high enough tariffs can disengage the lobby sufficiently so that the probability that the trade agreement will be broken is reduced to zero. Interestingly, if this occurs at a low enough tariff level, executive welfare can be maximized at this point. 
\begin{result}
  The executives maximize their welfare by either (a) raising tariffs sufficiently high to ensure that the trade agreement will remain in force or (b) trading off reductions in the probability that the agreement will be broken with reductions in welfare under the agreement.
  \label{res:execsoln}
\end{result}
%Can I say when one would happen versus the other?
This seems to accord well with observations of trade policy politics, in that some lobbies appear to exert significant effort toward disrupting trade agreements (often in the form of some kind of individual administrative remedy such as anti-dumping duties) whereas others apparently do not contest the levels of protection granted them. The current model points to differences across industries in production and demand structure, as well as political weighting ($\ga$), to help explain these variations.

\section{An Example}
\label{sec:example}
It is instructive to examine a simple parameterization of the model economy. The fundamentals here are chosen to match those of Bagwell and Staiger (2005). Home country demand, supply and profits are given by $D(P_i) = 1 - P_i$, $Q_X(P_X) = \frac{P_X}{2}$, $Q_Y(P_Y) = P_Y$, $\Pi_X(P_X) = \frac{(P_X)^2}{4}$, and $\Pi_Y(P_Y) = \frac{(P_Y)^2}{2}$ where $P_i$ is the price of good $i$ in the home country market. %Note that production function consistent with these supply functions are $F_X = \sqrt{l}$ and $F_Y = \sqrt{2l}$ where we assume $w=1$. 
Foreign is taken to be symmetric.

This implies Home-country imports of $X$ and exports of $Y$ of $M_X(P_X)= 1 - \frac{3}{2}P_X$ and $E_Y(P_Y)= 2P_Y -1$, with foreign imports of $Y$ and exports of $X$ given by $M_Y^*(P_Y^*)= 1 - \frac{3}{2}P_Y^*$ and $E_X(P_X^*)= 2P_X^* -1$. With the only trade policy instruments being tariffs on import competing goods, world prices are $P_X = P_X^W + \tau$, $P_X^* = P_X^W$, $P_Y^* = P_Y^W + \tau^*$, and $P_Y = P_Y^W$. Market clearing implies that world and home prices of $X$ are $P_X^W = \frac{4-3\tau}{7}$ and $P_X = \frac{4+4\tau}{7}$, symmetric for $Y$. \\

\subsection{Trade War Tariffs}
The median legislator's welfare can be written as 
\[
  W_{\mathit{ML}}^X(\tau,\ga(e,\ve)) + W_{\mathit{ML}}^Y(\tau^*) = \left\{\frac{9}{98} - \frac{5}{49}\tau - \frac{34}{49}\tau^2 +\frac{1}{98}\ga(e,\ve)\left[ 8 + 16\tau + 8\tau^2 \right] \right\}+ \frac{25}{98} - \frac{3}{49}\tau^* + \frac{9}{49}(\tau^*)^2.
\]
%where $W_{\mathit{ML}}^X(\tau,\ga(e,\ve))$ is the utility derived from consumer surplus, producer surplus and tariff revenues in the import-competing industry and $W_{\mathit{ML}}^Y(\tau^*)$ is the utility derived from consumer and producer surplus in the exporting industry.

When setting the trade-war tariff, the legislature maximizes $W_{\mathit{ML}}(\tau, \tau^*) = W_{\mathit{ML}}^X(\tau) + W_{\mathit{ML}}^Y(\tau^*)$ by choice of $\tau$ given $\tau^*$. As there are no interactions between $\tau$ and $\tau^*$, the legislature maximizes $W_{\mathit{ML}}^X(\tau)$ only (in curly braces above) and sets the trade war tariff
\[
  \tau^n = \frac{8\ga(e,\ve)-5}{68-8\ga(e,\ve)}
\]
I refer to this as the Nash tariff because it is the result of Nash equilibrium in the non-cooperative game between the legislatures. $\tau^n$ is increasing in $e$ and the second order condition is satisfied for all realizations of $\ga < 17/2$. Because $\ga = 7/4$ is enough to achieve the prohibitive tariff of $1/6$, I will assume this condition is satisfied.

In the event of a trade war and facing this tariff-setting behavior by the legislature, the lobby maximizes $\pi\left(\tau^n\left(\ga\left(e_n,\ve_n\right)\right)\right) - e_n$. In order to predict the Nash tariff, the political weighting function and form of political uncertainty must be specified. Take for instance $\ga(e,\ve) = 1.25 + e^{.2} + \ta$ with $\ve$ distributed uniformly on $[-0.25,0.25]$. Facing this specification of the political process, the lobby maximizes its objective function at $e_n = 0.00166$, which produces an expected Nash tariff of $0.129$.

\subsection{Break Decision}

\begin{figure}
\begin{center}
\includegraphics[height=3in, width=4in]{brprob2.jpg}
\end{center}
\caption{Probability Trade Agreement will be Broken\label{fig:br}}
\end{figure}

\begin{comment}
\begin{figure}
\begin{center}
\includegraphics[height=2.50in, width=3in]{contributions.jpg}
\end{center}
\caption{Lobbying Effort\label{fig:cont}}
\end{figure}
%graph is created in lobby_fix.m
\end{comment}

\begin{comment}
\begin{figure}
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.4\linewidth]{brprob2.jpg}
  \caption{Probability Trade Agreement will be Broken}
  \label{fig:br}
  %\label{fig:sub1}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.4\linewidth]{contributions.jpg}
  \caption{Lobbying Effort}
  \label{fig:cont}
  %\label{fig:sub2}
\end{subfigure}
%\caption{A figure with two subfigures}
%\label{fig:test}
\end{figure}
\end{comment}

Next we move to the legislature's decision on whether or not to break the trade agreement. Figure~\ref{fig:br} depicts the probability that the legislature will vote to break the trade agreement as a function of the tariff levels set in the trade agreement (with the restriction that $\tau^a = \tau^{*a}$) and the lobby's effort. It is strongly increasing in lobbying effort and decreasing in the level of tariffs set in the trade agreements.

Given the impact of lobbying on the legislature's break decision, the lobby's optimal contribution level is strongly decreasing in the trade agreement tariffs.%, as shown in Figure~\ref{fig:cont}.

\subsection{Trade Agreement}
Because the probability of legislative break never reaches zero for this parameterization, the executives' welfare function is concave everywhere. Assuming that each legislature has the opportunity to vote to break the agreement with probability $\frac{1}{2}$ and that the executives are social-welfare maximizers (i.e. $\ga_E = \ga_E^* = 1$), they will set trade agreements tariffs of $\tau^a = \tau^{*a} = 0.078$, with lobbying expenditures of $0.0007$ and a total break probability of $0.505$. The expected tariff is then $0.103$.

We can compare this against several different benchmarks. The most stark is the trade-war outcome itself, which can also be interpreted as the outcome that would prevail in the absence of executive involvement in trade policy \textit{and} the absence of any effort to make a trade agreement. The tariff in the executive-formed trade agreement is about 60$\%$ of the Nash tariff of $0.129$, and the expected level given the probability that the agreement will be broken is 80$\%$ of the Nash level. Lobbying expenditures in the agreement-maintenance phase are about 40$\%$ of the those in the Nash game, while expected expenditures in the executive-led trade-agreement scenario are 90$\%$ of what they would be if the legislatures set tariffs unilaterally.

Although the welfare-maximizing governments here are not able to set and maintain tariffs at zero as they would like, they are able to achieve significant reductions in tariff levels through the use of the trade agreement.

Another interesting benchmark is the scenario in which no executives are involved in trade policy, but in which the politically-susceptible legislatures can make use of a trade agreement. As in Bagwell and Staiger (2005), I find when maximizing their joint welfare in a legislature-led trade agreement ($W_{\mathit{ML}}(\tau, \tau^*) = W_{\mathit{ML}}^X(\tau) + W_{\mathit{ML}}^Y(\tau^*)$), the cooperative tariff levels will be set at
\[
  \tau^L = \frac{4\ga(e,\ve)-4}{25-4\ga(e,\ve)}, \ \tau^{*L} = \frac{4\ga(e^*,\ve)-4}{25-4\ga(e^*,\ve)}
\]
The legislatures are able to use the agreement to internalize the terms of trade externality, making political influence more expensive for the lobbies. Indeed, I find that lobbying expenditures rise to $0.0027$---60$\%$ higher than in the Nash case---while the agreement tariffs are set at $0.118$, only slightly lower than the Nash tariffs of $0.129$ and still higher than even the expected tariffs from the trade agreement scenario in which the executives choose the tariffs.

If one interprets the addition of a second policy-making actor as consistent with a move from autocracy to democracy, this result helps to explain the extensive empirical evidence that democracies trade more with each other than autocracies.\footnote{cfr. Laver and Shepsle (1991); Henisz and Mansfield (2006); Aidt and Gassebner (2010).}

\section{Discussion}
\label{sec:dis}

\subsection{Separation of Powers}
\label{sec:sep_powers}
Consider the model of Section~\ref{sec:main} for the case of $\ve_b=0$---that is, no uncertainty at the break stage (as will be discussed in Section~\ref{sec:uncertainty}, in the case of mean-zero uncertainty, strategic behavior at the trade-war stage is not altered by uncertainty). This allows the isolation of the results that derive from the assumption that power over trade policy is shared between the executive and legislative branches. Again, we focus on the home lobby.

Now $\ga(e_b)$ is deterministic, so the lobby knows the precise contribution it must make for any given trade agreement tariffs $\bta$ to induce the legislature to break the agreement (see Inequality~\ref{eq:lwcg}). Here, the lobby's contribution \textit{increases} in $\bta$ since the higher are the trade agreement tariffs, the larger is the political weight that is required to induce the legislature to find them unsatisfactorily low. As long as trade war profits net of the required contribution are greater than trade agreement profits, the lobby will induce a trade war; otherwise, it is not in the lobby's interest to make any contribution. Facing this behavior, the executives set the lowest trade agreement tariffs that make it prohibitively expensive for the lobby to have the agreement broken.

I return to the example of Section~\ref{sec:example} to illustrate the effects of changing the environment to one without political uncertainty. Recall that in the main example with $\ve$ distributed uniformly on $[-0.25,0.25]$, the expected Nash tariff is $0.129$ and the executive-led trade agreement tariff is $0.078$, with lobbying expenditures of $0.0007$ and a total break probability of $0.505$. The expected tariff is then $0.103$.

In this example, if the executives set the trade agreement facing no uncertainty, the tariff level is $0.106$, the lobby exerts zero effort, and the agreement remain in force with probability 1. If instead the executives were to set trade-agreement tariffs of $0.105$, the lobby would contribute $0.0025$ and the legislature would break the agreement with probability 1. This demonstrates both the agenda-setting power of the executives and the stark discontinuities induced when political uncertainty is not an issue.

Moreover, this case highlights the manipulation of the tariff level to discourage lobbying and therefore the disconnect that can arise between the tariffs that are chosen and the preferences of the governmental actors. In the example, the executive is a social welfare maximizer and so prefers zero tariffs, while the legislators with $\ga(e) = 1.25 + e^{.2}$ with $e_b = 0$ prefer $\tau^a = 0.05$. However, the trade agreement tariffs are set at $0.106$ precisely to ensure that $e_b =0$ so that we have a relatively free-trading legislature that upholds the agreement.

\subsection{Relation to Grossman and Helpman's `Protection for Sale'}
\label{sec:gh}
The legislative welfare function employed in this paper permits flexibility to take account of a wide range of political processes within a non-unitary legislature; alternatively, it can be a model of a unitary policy-maker when special interests face non-constant returns to lobbying activity. As noted in Section 2, the non-linear relationship between lobbying effort and tariffs follows Dixit, Grossman and Helpman (1997) and is a departure from the `Protection for Sale' welfare function, which assumes that a unitary government maximizes the sum of contributions and some fraction, $a$, of social welfare.

To see the relationship between the two forms, write the PFS welfare function (replacing $C$ with $e$) as
\[
  e + aW = e + a \left[ \mathit{CS}_X(\tau) + \mathit{PS}_X(\tau) + \mathit{CS}_Y(\tau^*) + \mathit{PS}_Y(\tau^*) + \mathit{TR}(\tau) \right]
\]
The relationship between $e$ and the weight the legislature places on profits is endogenous; because the PFS government welfare function has such an elegant and completely specified form, it is possible to solve for equilibrium behavior to retrieve this relationship directly.

It is easy to describe the equilibrium behavior and thus how lobbying effort will be related to the weight on profits in equilibrium in the PFS model when there is one lobby who makes a take-it-or-leave-it offer. Take the case of no uncertainty and assume that the legislature chooses $\tau=0$ in the absence of lobbying and that $\ga(0)=1$. For any $\tau$ that the lobby desires, it must pay according to the  indifference condition $e(\tau) = a \left[ W(0) - W(\tau) \right]$. That is, the lobby must pay for the $a$-weighted welfare loss caused by the tariff it requests.

In the model of this paper, this indifference condition, rearranged to solve for $\ga(e)$, is
		    \[
		      \ga(e) = \frac{\mathit{CS}_X(0) + \pi_X(0) + \mathit{CS}_Y(0) + \pi_Y(0) + \mathit{TR}(0) - \mathit{CS}_X(\tau) - \mathit{CS}_Y(\tau^*) - \pi_Y(\tau^*) - \mathit{TR}(\tau)}{\pi_X(\tau)}
		    \] 
In order to match the form of the PFS framework, we must therefore have
		    \begin{equation}
		      \ga(e) = 1 + \frac{e}{a\pi_X(\tau)}
		      \label{eq:gh}
		    \end{equation}

Choosing the political weighting function identified in Equation~\ref{eq:gh} aligns the political objective function used in the current model with the PFS framework. Although using the $\ga$-weighted government welfare function is a reduced-form approach in the sense that legislative dynamics are not fully modeled, a drawback of the PFS welfare function is that it embodies the restriction that returns to lobbying effort must be constant.

As noted in Dixit, Grossman and Helpman (1997), the linear form also implies transferable utility between the lobby and unitary government; this leads to equilibrium indeterminancies and an inability to speak to distributional questions, prompting them to generalize the PFS form. From their results, we have that the same equilibrium analysis holds in the more general environment.

For the trade policy literature, generalizing the linear form in a sense would obviate many of the questions that have been raised by empirical investigations of the PFS model about the relationship of the magnitude of welfare losses relative to the weight placed on social welfare by governments as the $a$ parameter for the constant relative marginal utility provided by social welfare would no longer be present in the model.

The linear PFS form provides for an almost unmatched simplicity and elegance that is able to capture the key cross-industry predictions on import tariffs. A perhaps unintended byproduct in the subsequent literature seems to have arisen in the form of a focus on matching fine details of governments' welfare-mindedness using a model that has a very simple unitary government and is thus better suited for other purposes.

It is clear that trade policy is often shaped in a complicated process involving multiple actors (see, for instance, Destler (2005) for the U.S.). The model presented herein is a first attempt to more richly model the policy-making process and thus shed light on open questions surrounding how government preferences and the details of political institutions impact trade policy outcomes.

\subsection{The Role of Political Uncertainty}
\label{sec:uncertainty}
In this model, if there were no political uncertainty, there would be no lobbying in equilibrium. Thus, if nothing else, adding the realism of political uncertainty allows for positive lobbying on the equilibrium path. If we examine more closely the assumption that there is significant uncertainty surrounding the legislative lobbying process, additional insights come to light. It is most clear in the case of mean-zero uncertainty, so I will assume this throughout this section unless otherwise noted.

In the event of a trade war, the lobbies' behavior, and therefore the expected trade-war tariffs, are not altered by mean-zero uncertainty (see Equation~\ref{eq:lobtw}).

At the earlier stage of lobbying, uncertainty alters optimizing behavior by both the lobby and the executives. In contrast to the previous section with no uncertainty, consider adding a very small amount of uncertainty to the previous example: take $\ve$ distributed uniformly on $[-0.01,0.01]$. Here, the executives find it optimal to set the tariffs to completely disengage the lobby and ensure the agreement remains in force. However, the tariff level will be different because instead of making the decision to lobby according to the certainty condition
\[\pi(\tau^n) - e_n - \pi(\tau^a) > e_b,
\]
the lobby makes this decision according to Condition~\ref{ine:lobint}, which can be re-written as
\[
\pi(\tau^n) - e_n - \pi(\tau^a) > \left[\frac{\partial b(0,\bta,\btn(e_n,e_n^*))}{\partial e_b}\right]^{-1}.
\]

Although the relationship between the tariff under certainty and a small amount of uncertainty will depend on the structure of the economy and the legislative process, in this case, the tariff can be reduced slightly to $0.105$. As we have seen above, with significant uncertainty, the tariff in this example is reduced to $0.078$. Interestingly, the optimal tariff does not decrease monotonically between these two values as uncertainty increases.

\begin{figure}
\begin{center}
\includegraphics[height=3in, width=4in]{lobby_br.jpg}
\end{center}
\caption{Lobby's Optimal Expenditure Function (Low and Intermediate Uncertainty)\label{fig:lobby_br}}
\end{figure}

It is particularly instructive to examine the lobby's reaction function at this very low level of uncertainty as well as the intermediate level when $\ve$ is distributed uniformly on $[-0.14,0.14]$ because it is at the latter that the executives set the highest tariff level of $0.107$. For the former, when facing very little uncertainty, the range of tariff levels to which the lobby responds with an intermediate contribution that leaves open the possibility of a break is very small: only those between 0.103 and 0.104. The lobby disengages completely at $\bta=0.105$ (see Figure~\ref{fig:lobby_br}).

In contrast, when uncertainty is increased to the interval $[-0.14,0.14]$, the lobby begins to leave open the possibility of a break at $\bta = 0.048$. It disengages at $0.107$ and this is the tariff level that the executives find optimal. Interestingly, at intermediate uncertainly levels, both portions of the best response curve are steeper and the lobby disengages at slightly lower tariff levels---as low as $0.099$ when $\ve \sim U[-0.06,0.06]$.

The preceding discussion illustrates Part (a) of Result~\ref{res:execsoln}: here the executives maximize their welfare by raising tariffs to the point where the import-competing industry ceases to lobby and thus the agreement remains in force for sure. Note, however, that although lobbying expenditures are zero, the potential for lobbying behavior is essential in shaping the trade agreement.

%\begin{figure}
%\begin{center}
%\includegraphics[height=3.8in, width=7in]{exec_welfare.jpg}
%\end{center}
%\caption{Differences in Executive Welfare when Uncertainty Increases\label{fig:exec_wel}}
%\end{figure}

When uncertainty rises above this threshold of $\ve \sim U[-0.14,0.14]$, the executives' choices are made according to Part (b) of Result~\ref{res:execsoln}: that is, they trade off reductions in welfare under the agreement with reductions in the probability that the agreement will be abrogated by the legislature. %In terms of the lobby's best response function, the executives now choose the optimal point on the downward-sloping portion of the curve instead of the point where it reaches zero. In terms of the welfare function for the executives, shown in Figure~\ref{fig:exec_wel}, they are now choosing the point that maximizes the concave portion of the curve instead of the ``spike'' that is created when the lobbies disengage and the chance that the agreement will be abrogated is removed altogether.\footnote{Note that the initial, flat portion of the curve is where the tariff is low enough that the lobby buys a trade war with certainty; we then enter the concave portion where the main results of Section~\ref{sec:main} are applicable; finally, we see the upward spike where the probability of a trade war is reduced to zero and the executives' welfare declines after that because there is no further reduction in break probability as the tariff increases.}

As uncertainty increases further, the general pattern is that tariffs, and therefore lobbying effort, are reduced, while the break probability increases. However, this pattern is not smooth or absolute, in contrast to the welfare level achieved by the executives, which increases monotonically in the amount of uncertainty present. %It is not clear how general a result this is, but it is particularly interesting in contrast to the non-monotone pattern over the lower-range of uncertainty where executive welfare is minimized at the edges of the range that lead to disengagement of the lobby and maximized in the middle of that range.

Several salient points emerge from this example. First, different amounts of uncertainty lead to very different outcomes---both in terms of the tariffs that are set in the trade agreement and the probability of disruption. Second, these outcomes vary in nuanced ways with the underlying behavior of the lobbies and the executives. The provided tariff levels are strongly influenced by lobbying incentives, but not in the straightforward ways predicted by unitary models.

\section{Extensions}
\label{sec:ext}
\begin{comment}
This analysis has been undertaken with the assumption that the import-competing industry is distributed evenly across legislative districts. By varying the way in which those profits enter the welfare function of the median legislator (before lobbying takes place), I can provide results about how district composition affects trade policy.

There are also several directions to explore in comparing different government structures. I start by describing the value of a trade agreement in an institutional setting of divided government.
\begin{result}
  Lobbying effort is reduced by the presence of executive branches who form a trade agreement (where the counterfactual is legislatures setting unilateral trade policy).
\end{result}
\textcolor{blue}{Don't yet know what assumptions I need to make this work} \\

Comparing this model to the unitary-government model is a delicate task. The most straightforward approach from the current structure is to disregard the executives and allow the legislative branches to form a trade agreement cooperatively after lobbying takes place. Note that we can retain the stochastic nature of the tariff-setting, but there is no vehicle for the agreement to be broken. In this case, the trade agreement tariff in home, $\tau_U^a$ will maximize
\[
  W_X(\tau,\ga_E) + W_X^*(\tau) = \frac{9}{98} - \frac{5}{49}\tau - \frac{34}{49}\tau^2 +\frac{1}{98}\ga(e,\ve)\left[ 8 + 16\tau + 8\tau^2 \right] + \frac{25}{98} - \frac{3}{49}\tau + \frac{9}{49}(\tau)^2 
\]
and will thus be set at
\[
  \tau_U^a = \frac{4\ga(e,\ve)-4}{25-4\ga(e,\ve)}.
\]
$\tau_U^a < \tau^n$ for all $\ga(e,\ve) < 7/4$, while $\ga(e,\ve)=7/4$ results in the prohibitive tariff of $1/6$. This is the standard result that a trade agreement internalized the terms of trade externality and results in a lower tariff than unilateral policy making.

\textcolor{blue}{will an executive inserted into this situation set higher or lower tariffs, what will happen to $e$?}

What if we add executive branches with essentially the same preferences as the legislature? That is, set $\ga_E = \expect[\ga(e,\ve)]$. \textcolor{blue}{It's hard to say how $e$ should be chosen here...still need to figure that out. But if we can get around the weird time inconsistency problem...} What agreement tariffs will the executives choose? It's easy to see that they will not choose $\tau^a < \tau_U^a$, as $\tau_U^a$ is their ideal point if there were no possibility of the agreement being broken, and lower tariffs only incentivize more lobbying and a higher probability that the agreement will be broken.

Thus, if an interior solution exists \textcolor{blue}{[should have conditions to insert here]}, the executive will exploit the same trade-off explored earlier, raising $\tau^a$ a bit to reduce the probability that the agreement will be abrogated. So interestingly, we can see that executives with the same political preferences as the legislature will choose higher trade agreement tariffs than the legislature would by itself due to the possibility that the agreement can be broken and the impact of the tariff choice on lobbying. We will also therefore have a reduced level of contributions, and if a trade war were possible, lower trade war tariffs because of it.

Along these lines, allowing for asymmetries in the preferences and political processes between the countries will provide important insights as well.

I also plan to use the richness of this model to parse out how much of the changes in tariffs comes from the fact that the executives place lower weight on producer profits and how much comes from the fact that they make an agreement---that is, compare the effects of structure versus preferences.
\end{comment}

\subsection{Asymmetric Trade Agreements}
\label{sec:asym}
The assumption that the legislatures do not have the opportunity to break a trade agreement simultaneously serves to simplify the analysis of the interaction between the lobbies and the consequences of relaxing it will be discussed in the following section. Maintaining it for now, however, facilitates examination of the assumption that trade agreements must be symmetric. This may be quite realistic in some settings where additional political constraints would make ``unfair'' arrangements unpalatable. However, it is important to point out that, for some parameter choices, the joint welfare of the executives is in fact maximized by an asymmetric agreement.

The parameterization from the example in Section~\ref{sec:example} is one such instance. The optimal unrestricted trade agreement is to set tariffs in one country (say home) at $0.062$ and those in the other (foreign) at 0. In this case, the equilibrium break probability in home, if the home legislature is afforded the opportunity, is zero, while the foreign legislature will break the agreement with probability 1 if given the chance. With the assumption from the example that each legislature gets the opportunity to repudiate the agreement half the time, this results in the trade agreement being broken with probability $0.50$, just less than the $0.505$ of the optimal symmetric agreement with both tariffs set at $0.078$. Because both the trade agreement tariffs and the expected chance that the agreement will remain in force are lower, expected welfare for the executives is higher.

What is behind the rather surprising result that a lower break probability can be achieved when both tariffs are strictly lower than in the symmetric agreement? Here, the executives relinquish any chance of passage of the trade agreement in the foreign country (in essence hoping that its legislature will not have a chance to act). Once they decide to pursue this course, it is optimal to reduce $\tau^{*a}$ to zero. This drastically reduces the home legislature's incentive to break the trade agreement. Because of the large direct impact of $\tau^{*a}$ in reducing the break probability, the home tariff is not required to make such a large contribution and can be lowered, which in turn reduces the lobby's incentive to exert effort (see Result~\ref{res:lobby}). This effect is strong enough that the executives can set a tariff level lower than that of the symmetric agreement that is sufficient to disengage the home lobby (see Inequality~\ref{ine:lobint}).

It is useful to keep in mind that this result occurs in a completely symmetric environment. Recall that, in general, the executives maximize joint welfare in Equation~\ref{eq:jv2}. If the chance that the legislatures will have the opportunity to break the agreement ($o$ and $o^*$) are not equal, or if another significant aspect of the economic or political situation is asymmetric, the importance of this result grows in a potentially substantial way. In particular, if only one country faces a legislative constraint, an asymmetric agreement seems highly likely as in the large body of literature investigating the Schelling conjecture.

\subsection{Break Opportunities in Both Countries}
\label{sec:two}
Although it changes their expected payoffs, extending the model to allow both legislatures the opportunity to break the trade agreement simultaneously does not impact the legislatures' incentives as each decision to break or not break the agreement is independent of the other.

However, the lobbies' incentives are altered because the total probability of a break with which they weight profits changes. Now, in addition to their own effort positively influencing their own legislature to break the agreement, the probability of a break in the trade agreement also increases in the effort of the other country's lobby. As might be expected, the lobbying effort in each country turns out to be a negative function of that in the trading partner; that is, to some extent, the lobbies have an incentive to free ride.

Without making stronger assumptions, it is not possible to say how much the effort of each lobby will be reduced therefore whether total lobbying effort rises or falls. What is clear is that the optimization of the executives becomes significantly more complex as they are now faced with the problem of how to best exploit the free-riding dynamics that occur between the lobbies.

\section{Conclusion}
\label{sec:concl}
I have shown that the legislature both breaks trade agreements with a higher probability and sets higher trade war tariffs when lobbying activity increases, while the probability with which it breaks agreements (or fails to ratify them) decreases in the domestic trade agreement tariff. Because the lobby decreases its effort in response to higher trade agreement tariffs, the executives face a trade-off between the welfare derived while a trade agreement is in force and the probability with which the agreement actually enters into force.

I have also shown that in a policy-making structure in which power is shared, a less politically-motivated executive can utilize an international trade agreement to reduce the political pressure on the ratifying body and therefore increase the probability that the agreement will remain in force. Thus, in a model with a richer description of government structure, a political-commitment role for trade agreements can arise.

The executives' incentive to raise tariffs in order to reduce lobbying effort helps to explain the empirical finding in the Protection for Sale literature that levels of protection and associated deadweight losses are too high relative to lobbying expenditure given the high estimates for governments' weighting of social welfare. I have shown that this serves to mediate the relationship between preferences for contributions relative to social welfare and the tariffs that are provided. The intuition is clear: the observed lobbying expenditure levels may in fact be low \textit{because} tariffs have been raised sufficiently high to prevent political pressure and the increased risk of a costly trade disruption it engenders.

That lobbying and tariff levels are related in systematic ways to the amount of political uncertainty present suggests interesting avenues for future empirical work. Several directions for future theoretical work also seem potentially fruitful, including removing the assumption of perfect enforceability and supporting cooperation through repeated interaction and generalizing the model to the case of multiple lobbies.

\section{Appendix}
\label{sec:appendix}
\begin{lemma*}[\textbf{0}]
  When prices are linear in tariffs, the second-order condition for the legislature's problem when setting trade-war tariffs holds for all non-prohibitive tariffs given that the tariff elasticity of supply is less than unity.
  \label{res:l1}
\end{lemma*}
Proof: The first order condition can be rewritten as
   \begin{equation}
        - D(P_{X}) \frac{\partial P_X}{\partial \tau} + \ga(e_n,\ve_n) Q_X(P_{X}) \frac{\partial P_X}{\partial \tau}
	    + \left[ D(P_{X})- Q_X(P_{X})\right] + \tau \left(\frac{\partial D(P_X)}{\partial \tau} - \frac{\partial Q_X(P_X)}{\partial \tau} \right) = 0
			\label{eq:foc}
	 \end{equation}

	 \vskip.1in
	 \noindent The required second order condition is that the following is negative:
	 \begin{multline*}
	    	- D(P_{X}) \frac{\partial^2 P_X}{\partial \tau^2} - \frac{\partial D(P_{X})}{\partial \tau}  \frac{\partial P_X}{\partial \tau} 
	    	+ \ga(e_n,\ve_n) Q_X(P_{X}) \frac{\partial^2 P_X}{\partial \tau^2} + \ga(e_n,\ve_n) \frac{\partial Q_X(P_{X})}{\partial \tau} \frac{\partial P_X}{\partial \tau} \\
	    + 2 \left[ \frac{\partial D(P_{X})}{\partial \tau} - \frac{\partial Q_X(P_{X})}{\partial \tau}\right] 
	    + \tau \left(\frac{\partial^2 D(P_X)}{\partial \tau^2} - \frac{\partial^2 Q_X(P_X)}{\partial \tau^2} \right)
	  \end{multline*}

	\vskip.1in  
	\noindent $\ga(e_n,\ve_n)$ is endogenous, so we must find a way to bound it. Because the lobby receives no gains if the tariff is raised beyond its prohibitive level (that is when $D(P_{X}) = Q_X(P_{X})$), it has no incentive to exert effort above the level that induces the $\ga(e_n,\ve_n)$ that leads to the prohibitive tariff. Setting $D(P_{X}) = Q_X(P_{X})$ in the first order condition, this is
    \[
	     \ga(e_n,\ve_n) = 1 - \frac{\tau \left(\frac{\partial D(P_X)}{\partial \tau} - \frac{\partial Q_X(P_X)}{\partial \tau} \right)}{Q_X(P_{X}) \frac{\partial P_X}{\partial \tau}}
	  \]

	\noindent If prices are linear in tariffs, the second order terms are zero, and substituting the above expression for $\ga(e_n,\ve_n)$, we arrive at
    \[
	    	- \frac{\partial D(P_{X})}{\partial \tau}  \frac{\partial P_X}{\partial \tau} + \frac{\partial Q_X(P_{X})}{\partial \tau} \frac{\partial P_X}{\partial \tau} - \tau \left(\frac{\partial D(P_X)}{\partial \tau} - \frac{\partial Q_X(P_X)}{\partial \tau} \right) \frac{ \frac{\partial Q_X(P_{X})}{\partial \tau}}{Q_X(P_{X})} + 2 \left[ \frac{\partial D(P_{X})}{\partial \tau} - \frac{\partial Q_X(P_{X})}{\partial \tau}\right]
	  \]
	  Because $P_X = P_X^W + \tau$, $\frac{\partial P_X}{\partial \tau} = \frac{\partial P_X^W}{\partial \tau} + 1$; together with $\frac{\partial P_X^W}{\partial \tau} < 0,$ this implies $\frac{\partial P_X}{\partial \tau} < 1$. Therefore the previous expression is bounded above by
	  \[
	    - \tau \left(\frac{\partial D(P_X)}{\partial \tau} - \frac{\partial Q_X(P_X)}{\partial \tau} \right) \frac{ \frac{\partial Q_X(P_{X})}{\partial \tau}}{Q_X(P_{X})} + \left[ \frac{\partial D(P_{X})}{\partial \tau} - \frac{\partial Q_X(P_{X})}{\partial \tau}\right]
	  \]
	 Since $\frac{\partial D(P_{X})}{\partial \tau} - \frac{\partial Q_X(P_{X})}{\partial \tau}<0$, the second order condition is negative as long as $\frac{\tau}{Q_X(P_{X})}\frac{\partial Q_X(P_{X})}{\partial \tau} < 1.$
	 If prices are not linear in the tariff, additional conditions are required. $\hfill\blacksquare$

\vskip.4in	 
\noindent \textbf{\hypertarget{tninc}{Proof of Lemma~\ref{res:tninc}}}:\\
By the implicit function theorem, $\frac{\partial \tau^n}{\partial \ga} = - \frac{\frac{\partial FOC}{\partial \ga}}{\frac{\partial FOC}{\partial \tau^n}}$ where \textit{FOC} is the first order condition for the legislature's problem, expanded in Equation~\ref{eq:foc}. We have $\frac{\partial FOC}{\partial \ga} = Q_X(P_{X}) \frac{\partial P_X}{\partial \tau}$ from the proof of Lemma 0. This is always positive. Lemma 0 also shows that $\frac{\partial FOC}{\partial \tau^n}$ is negative everywhere. Thus $\frac{\partial \tau^n}{\partial \ga}\geq 0$. $\hfill\blacksquare$

\vskip.4in
\noindent \textbf{\hypertarget{Pr_bincC}{Proof of Result~\ref{res:bincC}}}:\\
Substituting from Equation~\ref{eq:ml}, Equation~\ref{eq:b} can be re-written as
\begin{multline}  
  b(e_b,\bta,\btn) = \Pr [ \mathit{CS}_X(\tau^n) + \ga(e_b,\ve_b) \cdot \mathit{PS}_X(\tau^n) + \mathit{CS}_Y(\tau^{*n}) + \mathit{PS}_Y(\tau^{*n}) + \mathit{TR}(\tau^n) >  \\ \mathit{CS}_X(\tau^a) + \ga(e_b,\ve_b) \cdot \mathit{PS}_X(\tau^a) + \mathit{CS}_Y(\tau^{*a}) + \mathit{PS}_Y(\tau^{*a}) + \mathit{TR}(\tau^a) ]
\end{multline}
Rearranging, we have $b(e_b,\bta,\btn) = $
\begin{multline}  
  \textstyle \Pr \Big[ \frac{\mathit{CS}_X(\tau^n) + \mathit{PS}_X(\tau^n) + \mathit{CS}_Y(\tau^{*n}) + \mathit{PS}_Y(\tau^{*n}) + \mathit{TR}(\tau^n)  -\mathit{CS}_X(\tau^a) - \mathit{PS}_X(\tau^a) - \mathit{CS}_Y(\tau^{*a}) - \mathit{PS}_Y(\tau^{*a}) - \mathit{TR}(\tau^a)}{\mathit{PS}_X(\tau^a) - \mathit{PS}_X(\tau^n)} + 1 \\ \textstyle < \ga(e_b,\ve_b) \Big]
  \label{eq:b_ex}
\end{multline}
%\noindent The left side of the inequality in Expression~\ref{eq:b_ex} does not depend on $e_b$. By Assumption~\ref{as:ga_c}, the right side of the inequality is increasing and concave in $e_b$. Thus $b(e_b,\bta,\btn)$ is increasing and concave in $e_b$.    $\hfill\blacksquare$
\noindent The left side of the inequality in Expression~\ref{eq:b_ex} does not depend on $e_b$. Call it $Z$. Thus we have $b(e_b,\bta,\btn) = \Pr [Z < \ga(e_b,\ve_b) ] = 1 - F_{\ga}(\ga=Z)$ where $F_{\ga}(\ga=Z) = F_{\ta}(\ta=h^{-1}(\ga,e))$ by the Change of Variables Theorem and Assumption~\ref{as:ga_ta} with $\ga = h(e,\ta)$ giving the change of variable.

Then $\frac{\partial b}{\partial e} = -\frac{\partial F_{\ta}(\ta=h^{-1}(\ga,e))}{\partial \ta}\frac{\partial h^{-1}(\ga,e)}{\partial e} = -f_\ta(\ta) \frac{\partial h^{-1}(\ga,e)}{\partial e}$ and $\frac{\partial^2 b}{\partial e^2} = -f_\ta(\ta) \frac{\partial^2 h^{-1}(\ga,e)}{\partial e^2}$. Because $\ga = h(e,\ta)$ is increasing and concave in $e \ \forall \ta$ by Assumption~\ref{as:ga_c}, its inverse is decreasing and convex $\left(\frac{\partial h^{-1}(\ga,e)}{\partial e}\leq 0; \ \frac{\partial^2 h^{-1}(\ga,e)}{\partial e^2} \geq 0 \right)$ The pdf of $\ta$ is non-negative, so $\frac{\partial b}{\partial e} \geq 0$ and $\frac{\partial^2 b}{\partial e^2} \leq 0$. $\hfill\blacksquare$

\vskip.4in
\noindent \textbf{\hypertarget{Pr_leg_astar}{Proof of Lemma~\ref{lem:leg_astar}}}:\\
It must be shown that the left hand side of the inequality in Expression~\ref{eq:b_ex} is decreasing in $\tau^{*a}$. The derivative of this quantity with respect to $\tau^{*a}$ is 
\begin{equation}
  \frac{-\left(\mathit{PS}_X(\tau^a) - \mathit{PS}_X(\tau^n)\right)\left(\frac{\partial \mathit{CS}_Y(\tau^{*a})}{\partial \tau^{*a}} + \frac{\partial \mathit{PS}_Y(\tau^{*a})}{\partial \tau^{*a}}\right)}{\left(\mathit{PS}_X(\tau^a) - \mathit{PS}_X(\tau^n)\right)^2} = \frac{\frac{\partial \mathit{CS}_Y(\tau^{*a})}{\partial \tau^{*a}} + \frac{\partial \mathit{PS}_Y(\tau^{*a})}{\partial \tau^{*a}}}{\mathit{PS}_X(\tau^n) - \mathit{PS}_X(\tau^a)}.
  \label{appex:astar}
\end{equation}

\noindent The price of good $Y$ is decreasing in $\tau^{*a}$, so consumer surplus is increasing and producer surplus is decreasing in $\tau^{*a}$. Because $Y$ is being exported, the decrease in producer surplus is larger than the increase in consumer surplus, making the numerator negative. Since producer surplus in the denominator is increasing in $\tau$, the expression in Equation~\ref{appex:astar} is negative for all $\tau^{*a}$.  $\hfill\blacksquare$

\vskip.4in
\noindent \textbf{\hypertarget{Pr_leg_a}{Proof of Lemma~\ref{res:leg_a}}}:\\
Using the logic of the proof of Lemma~\ref{lem:leg_astar}, the effect on the break probability is determined by the sign of the derivative of the left hand side of the inequality in Expression~\ref{eq:b_ex} with respect to $\tau^a$; to show that the break probability is decreasing in $\tau^a$, I must demonstrate that this derivative is positive. Labeling the numerator of that expression $\left[ W(\btn)-W(\bta)\right]$ (for the change in social welfare), this derivative can be written
\begin{equation}
  \frac{\left(\mathit{PS}_X(\tau^n) - \mathit{PS}_X(\tau^a)\right)\left(\frac{\partial \mathit{CS}_X(\tau^{a})}{\partial \tau^{a}} + \frac{\partial \mathit{PS}_X(\tau^a)}{\partial \tau^{a}} + \frac{\partial \mathit{TR}(\tau^{a})}{\partial \tau^{a}}\right) - \left[ W(\btn)-W(\bta)\right] \frac{\partial \mathit{PS}_X(\tau^a)}{\partial \tau^a}}{\left(\mathit{PS}_X(\tau^a) - \mathit{PS}_X(\tau^n)\right)^2}.
  \label{appex:a}
\end{equation}

\noindent $\left(\mathit{PS}_X(\tau^n) - \mathit{PS}_X(\tau^a)\right)$ is always positive by Assumption~\ref{as:ga_l_e}. Because the optimal unilateral tariff for large welfare-maximizing governments is positive (call it $\tau^O$), $\left(\frac{\partial \mathit{CS}_X(\tau^{a})}{\partial \tau^{a}} + \frac{\partial \mathit{PS}_X(\tau^a)}{\partial \tau^{a}} + \frac{\partial \mathit{TR}(\tau^{a})}{\partial \tau^{a}}\right)$ is increasing up to $\tau^O$ and decreasing above it. Thus the first summand is increasing up until $\tau^O$ and decreasing thereafter.

Because total social welfare is maximized at $\tau^a = \tau^{*a} = 0$, $W(\btn)-W(\bta)$ is always negative, whereas producer surplus is increasing in $\tau^a$, so the second summand is positive everywhere. With a positive denominator, we thus have that the derivative is positive on $[0,\tau^O]$.

It is also positive over the remaining $(\tau^O,\tau^n)$. To see this, add $\left(\tilde{\Gamma} - 1 \right) \frac{\partial \mathit{PS}_X(\tau^a)}{\partial \tau^a} \left(\mathit{PS}_X(\tau^n) - \mathit{PS}_X(\tau^a)\right)$ to the first summand and subtract it from the second. For any particular value of $\tilde{\tau}^a$, one can choose the $\tilde{\Gamma}$ weight that would make $\tilde{\tau}^a$ the preferred unilateral tariff; this makes the derivative in the first summand zero. Having subtracted the same quantity from the second summand modifies the welfare difference in the second summand to be maximized at $\tilde{\tau}^a$ so that this term is always negative, thus ensuring the result.  $\hfill\blacksquare$

\vskip.4in
\noindent \textbf{\hypertarget{Pr_bcomb}{Proof of Lemma~\ref{res:bcomb}}}:\\
Again, I want to show how the inequality in Expression~\ref{eq:b_ex} changes, now with respect to both $\tau^a$ and $\tau^{*a}$, so I add the derivatives in Expressions~\ref{appex:astar} and \ref{appex:a} to get
\[
  \frac{\left(\mathit{PS}_X(\tau^n) - \mathit{PS}_X(\tau^a)\right)\left(\frac{\partial \mathit{W}_X(\bta)}{\partial \tau^{a}} + \frac{\partial \mathit{W}_X(\bta)}{\partial \tau^{*a}}\right) - \left[ W(\btn)-W(\bta)\right] \frac{\partial \mathit{PS}_X(\tau^a)}{\partial \tau^a}}{\left(\mathit{PS}_X(\tau^n) - \mathit{PS}_X(\tau^a)\right)^2}.
\]
where $\frac{\partial \mathit{W}_X(\bta)}{\partial \tau^{a}} + \frac{\partial \mathit{W}_X(\bta)}{\partial \tau^{*a}} = \frac{\partial \mathit{CS}_X(\tau^{a})}{\partial \tau^{a}} + \frac{\partial \mathit{PS}_X(\tau^a)}{\partial \tau^{a}} + \frac{\partial \mathit{TR}(\tau^{a})}{\partial \tau^{a}} + \frac{\partial \mathit{CS}_Y(\tau^{*a})}{\partial \tau^{*a}} + \frac{\partial \mathit{PS}_Y(\tau^{*a})}{\partial \tau^{*a}}$ is the total derivative of social welfare. Since social welfare is maximized at $\bta = (0,0)$,\footnote{Note, this is identical to the result that joint social welfare is maximized at zero tariffs because of symmetry.}  this is negative $\forall \bm{\tau} \in (\bm{0},\btn]$; note that it is 0 at $\bm{0}$ and vanishingly small for very small tariffs.

Thus the first summand in the numerator is zero at $\bta = \bm{0}$ and increasingly negative as $\bta$ increases. The second summand is positive everywhere because social welfare, $W$, is lowest at $\btn$ and producer surplus is increasing everywhere. Thus the numerator is positive at $\bm{0}$ and at least for very small $\bta$

It is also positive for all other values of $\bta$ strictly below $\btn$. Just as in the proof of Lemma~\ref{res:leg_a}, one one can add
$\left(\tilde{\Gamma} - 1 \right) \frac{\partial \mathit{PS}_X(\tau^a)}{\partial \tau^a} \left(\mathit{PS}_X(\tau^n) - \mathit{PS}_X(\tau^a)\right)$ to the first summand and subtract it from the second. For any particular value of $\tilde{\bta}$, one can choose the $\tilde{\Gamma}$ weight that would make $\tilde{\bta}$ the politically optimal tariff; this makes the derivative in the first summand zero. Having subtracted the same quantity from the second summand modifies the welfare difference in the second summand to be maximized at $\tilde{\bta}$ so that this term is always negative, thus ensuring the result.

Because the denominator is positive, the entire expression is positive for all $\bta < \btn$.   $\hfill\blacksquare$

\vskip.4in
\noindent \textbf{\hypertarget{Pr_lobby}{Proof of Result~\ref{res:lobby}}}:\\
Proof is via the Implicit Function Theorem using the lobby's first order condition, Equation~\ref{eq:lobbyfoc}, referred to here as $FOC_L$.
  \[
		\frac{\partial e_b}{\partial \tau^a} + \frac{\partial e_b}{\partial \tau^{*a}} = - \left[ \frac{\frac{\partial FOC_L}{\partial \tau^a}}{\frac{\partial FOC_L}{\partial e_b}} + \frac{\frac{\partial FOC_L}{\partial \tau^{*a}}}{\frac{\partial FOC_L}{\partial e_b}} \right] = \frac{ \frac{\partial b}{\partial e_b} \frac{\partial \pi(\bta)}{\partial \tau^a} + \frac{\partial b}{\partial e_b} \frac{\partial \pi(\bta)}{\partial \tau^{*a}} - \left\{\frac{\partial^2 b}{\partial e_b \partial \tau^a} + \frac{\partial^2 b}{\partial e_b \partial \tau^{*a}}   \right\}\left[ \pi(\tau^n) - e_n - \pi(\tau^a) \right]}{\frac{\partial^2 b}{\partial e_b{}^2} \left[ \pi(\tau^n) - e_n - \pi(\tau^a) \right]}
	\]
Beginning with the denominator: because $\pi(\tau)$ is increasing everywhere, $\left[\pi(\tau^n) - e_n - \pi(\tau^a) \right]$ is positive for all but very large values of $\tau^a$, that is for all $\tau^a$ such that $\pi(\tau^n) - e_n > \pi(\tau^a)$. When $\tau^a$ rises above this level, it is no longer in the lobby's interest to ask to have the agreement broken so $e_b=0$ and $\frac{\partial e_b}{\partial \tau^a} = 0$. $\frac{\partial^2 b}{\partial e_b{}^2}$ negative by Result~\ref{res:bincC}, so the denominator is negative for all values of $\tau^a$ at which $e_b$ is not constant.

$\frac{\partial b}{\partial e_b}$ is positive, $\frac{\partial \pi(\bta)}{\partial \tau^a}$ is positive by construction, and $\frac{\partial \pi(\bta)}{\partial \tau^{*a}}$ is zero: separability between the sectors implies that profits in the import-competing sector do not depend on $\tau^{*a}$.

We can rewrite $\frac{\partial b(e,\bta)}{\partial  \tau^a} + \frac{\partial b(e,\bta)}{\partial  \tau^{*a}} = -\frac{\partial F_\ga (Z(\bta))}{\partial \tau^a} -\frac{\partial F_\ga (Z(\bta))}{\partial \tau^{*a}} = -\frac{\partial F_\ga (Z(\bta))}{\partial Z(\bta)}\left[\frac{\partial Z(\bta)}{\partial \tau^a} + \frac{\partial Z(\bta)}{\partial \tau^{*a}} \right] = - f_\ga \left[\frac{\partial Z(\bta)}{\partial \tau^a} + \frac{\partial Z(\bta)}{\partial \tau^{*a}} \right]$, where $F_\ga (Z(\bta))$ is the CDF of $\ga$ and $f_\ga (Z(\bta))$ is the pdf of $\ga$ and $Z(\bta)$ represents the left hand side of the inequality in Expression~\ref{eq:b_ex}.

Then $\frac{\partial^2 b}{\partial e_b \partial \tau^a} + \frac{\partial^2 b}{\partial e_b \partial \tau^{*a}} = 	- \frac{\partial}{\partial e} \left( \frac{\partial F_\ga (Z(\bta))}{\partial \tau^a} + \frac{\partial F_\ga (Z(\bta))}{\partial \tau^{*a}} \right) = 
	   - f_\ga \left[\frac{\partial^2 Z(\bta)}{\partial \tau^a \partial e} + \frac{\partial^2 Z(\bta)}{\partial \tau^{*a} \partial e} \right] - \left[\frac{\partial Z(\bta)}{\partial \tau^a} + \frac{\partial Z(\bta)}{\partial \tau^{*a}} \right] \frac{\partial f_\ga}{\partial e} = - \left[\frac{\partial Z(\bta)}{\partial \tau^a} + \frac{\partial Z(\bta)}{\partial \tau^{*a}} \right] \frac{\partial f_\ga}{\partial e} 
$
where the above equality holds because $Z(\bta)$ does not depend on $e$. Lemma~\ref{res:bcomb} shows that $\frac{\partial Z(\bta)}{\partial \tau^a} + \frac{\partial Z(\bta)}{\partial \tau^{*a}} \geq 0$, so $\frac{\partial f_\ga}{\partial e} \geq 0$ ensures that $\left\{\frac{\partial^2 b}{\partial e_b \partial \tau^a} + \frac{\partial^2 b}{\partial e_b \partial \tau^{*a}} \right\} \geq 0$. Thus $\frac{\partial e_b}{\partial \tau^a} + \frac{\partial e_b}{\partial \tau^{*a}} \leq 0$.  $\hfill\blacksquare$

\vskip.4in	  	
\noindent \textbf{\hypertarget{Pr_bcomB}{Proof of Result~\ref{res:bcomB}}}: \\
$	\frac{\partial B(\bta)}{\partial \tau^a} + \frac{\partial B(\bta)}{\partial  \tau^{*a}} = \frac{\partial b}{\partial  e_b}\frac{\partial e_b}{\partial \tau^a} + \frac{\partial b}{\partial  e_b}\frac{\partial  e_b}{\partial  \tau^{*a}} + \frac{\partial b}{\partial \tau^a} + \frac{\partial b}{\partial  \tau^{*a}} = \frac{\partial b}{\partial  e_b} \left\{ \frac{\partial e_b}{\partial \tau^a} + \frac{\partial  e_b}{\partial  \tau^{*a}} \right\} + \frac{\partial b}{\partial \tau^a} + \frac{\partial b}{\partial  \tau^{*a}}.$ $\frac{\partial b}{\partial  e_b}$ is positive by Result~\ref{res:bincC}. $\left\{ \frac{\partial e_b}{\partial \tau^a} + \frac{\partial  e_b}{\partial  \tau^{*a}} \right\}$ is negative by Result~\ref{res:lobby}. Taken together, the final two summands are negative by Lemma~\ref{res:bcomb}. Thus the entire expression is negative.   $\hfill\blacksquare$

\vskip.4in
\noindent \textbf{\hypertarget{int_soln}{Conditions for Interior Solution to Executives' Problem}} \\
The first order condition for maximizing joint surplus with respect to $\tau^a$ and $\tau^{*a}$ when the two are constrained to be equal is
\begin{multline*}
  \textstyle \left\{ 1 - o \cdot B(\bta) - o^* \cdot B^*(\bta) \right\} \left\{\frac{\partial \bm{W_E}(\bta)}{\partial \tau^a} + \frac{\partial \bm{W_E}(\bta)}{\partial \tau^{*a}} \right\} \\
     \textstyle + \left\{ o \cdot \left[\frac{\partial B(\bta) }{\partial \tau^a} + \frac{\partial B(\bta) }{\partial \tau^{*a}} \right] + o^* \cdot \left[\frac{\partial B^*(\bta)}{\partial \tau^a} + \frac{\partial B^*(\bta)}{\partial \tau^{*a}} \right] \right\} \left[ \bm{W_E}(\btn)- \bm{W_E}(\bta) \right] = 0
\end{multline*}
Because there is no benefit to setting $\bta$ below the executives' preferred level, which I will denote $\bm{\tau^E}$, I will take the choice space to be $[\bte,\btn]$. Note that for $\ga_E = 1$, $\bte=0$.

To demonstrate that the executives do not choose $\bta=\bte$, I must show that the left side of the above equation is positive at $\bta=\bte$. Assumption~\ref{as:ga_l_e} and Result~\ref{res:bcomB} combined with symmetry ensure that the first term of the second summand is negative. That executive welfare is maximized at $\bte$ ensures that the term multiplying it is also negative as well as that $\frac{\partial \bm{W_E}(\bte)}{\partial \tau^a} + \frac{\partial \bm{W_E}(\bte)}{\partial \tau^{*a}}$ is zero. Therefore the derivative of joint executive welfare is positive at $\bte$.  $\hfill\blacksquare$

%\section{Acknowledgements}
%The author thanks the following people for their insightful comments: Joel Watson, Jim Rauch, Marc Muendler, Lawrence Broz, Peter Cowhey, Kyle Bagwell, Andrei Levchenko, Bob Staiger, two anonymous referees, and seminar participants at UC San Diego, Syracuse, Georgetown, Florida International, Brandeis, BGSU, U Mass Amherst, Swarthmore, SUNY Oswego, Kenyon College, and PEIO 2013. All remaining errors are my own.

\section{References}
\singlespacing

\begin{list}{}{\setlength{\leftmargin}{0.0in}\setlength{\rightmargin}{0.0in}\setlength{\itemindent}{0.0in}\setlength{\itemsep}{0.1in}}

\item Aidt, T., Gassebner, M., 2010. Do Autocratic States Trade Less? The World Bank Economic Review 24, 38-76.

\item Ansolabehere, S., de Figueiredo, J., Snyder Jr., J., 2003. Why Is There so Little Money in U.S. Politics? Journal of Economic Perspectives 17, 105-130.

\item Bagwell, K., Staiger, R., 2005. Enforcement, Private Political Pressure, and the General Agreement on Tariffs and Trade/World Trade Organization Escape Clause. Journal of Legal Studies 34, 471-513.

\item Baldwin, R.E., 1987. Politically realistic objective functions and trade policy: PROFs and tariffs. Economic Letters 24, 287-90.

\item Beshkar, M., 2010. Trade skirmishes safeguards: A theory of the WTO dispute settlement process. Journal of International Economics 82, 35-48.

\item Bombardini, M., 2008. Firm Heterogeneity and Lobby Participation. Journal of International Economics 75, 329-348.

\item Bombardini, M., Trebbi, F., 2012. Competition and Political Organization: Together or Alone in Lobbying for Trade Policy? Journal of International Economics 87, 18-26.

\item Buzard, K., 2014a. Endogenous Politics and the Design of Trade Agreements. Available at: \url{https://kbuzard.expressions.syr.edu/wp-content/uploads/Endogenous-Politics.pdf}.

\item Buzard, K., 2014b. Self-enforcing Trade Agreements, Dispute Settlement and Separation of Powers. Available at: \url{https://kbuzard.expressions.syr.edu/wp-content/uploads/2014/04/Self-enforcing-Trade_Agreements.pdf}.

\item Coates, D., Ludema, R., 2001. A Theory of Trade Policy Leadership. Journal of Development Economics 65, 1-29.

\item Dai, X., 2006. Dyadic Myth and Monadic Advantage: Conceptualizing the Effect of Democratic Constraints on Trade. Journal of Theoretical Politics 18, 267-297.

\item Destler, I.M., 2005. American Trade Politics. Institute for International Economics, Washington, DC.

\item Dixit, A., G. Grossman, Helpman, E., 1997. Common Agency and Coordination: General Theory and Application to Government Policy Making. Journal of Political Economy 105, 752-769.

\item Ethier, W., 2002. Unilateralism in a Multilateral World. The Economic Journal 112, 266-292.

\item Ethier, W., 2012. The Political-Support Approach to Protection. Global Journal of Economics 1, 1-14.

\item Feenstra, R., 1992. How Costly is Protectionism? Journal of Economic Perspectives 6, 159-178.

\item Feenstra, R., Lewis, T., 1991. Negotiated Trade Restrictions with Private Political Pressure. Quarterly Journal of Economics 106, 1287-1307.

\item Findlay, R., Wellisz, S., 1982. Endogenous Tariffs and the Political Economy of Trade Restrictions and Welfare. In Jagdish Bhagwati (ed.) Import Competition and Response, Chicago, IL: University of Chicago, 1982

\item Gawande, K., Bandyopadhyay, U., 2000. Is Protection for Sale? Evidence on the Grossman-Helpman Theory of Endogenous Protection. The Review of Economics and Statistics 82, 139-152.

\item Gawande, K., Hoekman, B., 2006. Lobbying and Agricultural Trade Policy in the United States. International Organization 60, 527-561.

\item Gawande, K., Magee, C., 2012. Free Riding on Protection for Sale. International Studies Quarterly 56, 735-747.

\item Gawande, K., Krishna, P., Robbins, M., 2006. Foreign Lobbies and U.S. Trade Policy. Review of Economics and Statistics 88, 563-571.

\item Gawande, K., Krishna, P., Olarreaga, M., 2009. What Governments Maximize and Why: The View from Trade. International Organization 63, 491-532.

\item Gawande, K., Krishna, P., Olarreaga, M., 2012. Lobbying Competition Over Trade Policy. International Economic Review 53, 115-132.

\item Goldberg, P., Maggi, G., 1999. Protection for Sale: An Empirical Investigation. American Economic Review 89, 1135-1155.

\item Grossman, G., Helpman, E., 1994. Protection for Sale. The American Economic Review 84, 833-850.

\item Grossman, G., Helpman, E., 1995a. The Politics of Free-Trade Agreements. The American Economic Review 85, 667-690.

\item Grossman, G., Helpman, E., 1995b. Trade Wars and Trade Talks. The Journal of Political Economy 103, 675-708.

\item Grossman, G., Helpman, E., 2005. A Protectionist Bias in Majoritarian Politics. The Quarterly Journal of Economics 120, 1239-1282.

\item Henisz, W., Mansfield, E., 2006. Votes and Vetoes: The Political Determinants of Commercial Openness. International Studies Quarterly 50, 189-211.

\item Hillman, A., 1982. Declining Industries and Political-Support Protectionist Motives. The American Economic Review 5, 1180-1187.

\item Iida, K., 1996. Involuntary Defection in Two-Level Games. Public Choice 89, 283-303.

\item Imai, S., Katayama, H., Krishna, K., 2009. Is protection really for sale? A survey and directions for future research. International Review of Economics and Finance 18, 181-191.

%\item Kibris, A., 2012. Uncertainty and Ratification Failure. Public Choice 150, 439-467.

\item Klimenko, M., Ramey, G., Watson, J., 2008. Recurrent Trade Agreements and the Value of External Enforcement. Journal of International Economics 74, 475-499.

%\item Kono, D., 2006. Optimal Obfuscation: Democracy and Trade Policy Transparency. American Political Science Review 100, 369-384.

\item Laver, M., Shepsle, K., 1991. Divided Government: America is Not `Exceptional'. Governance: An International Journal of Policy and Administration 4, 250-269.

\item Le Breton, M., Salanie, F., 2003. Lobbying under political uncertainty. Journal of Public Economics 87, 2589-2610.

\item Le Breton, M., Zaporozhets, V., 2007. Legislative Lobbying under Political Uncertainty. Available at SSRN: \url{http://ssrn.com/abstract=1024686}.

\item Lohmann, S., O'Halloran, S., 1994. Divided Government and US Trade Policy. International Organization 48, 595-632.

\item Maggi, G., Rodr\'{i}guez-Clare, A., 2007. A Political-Economy Theory of Trade Agreements. The American Economic Review 97, 1374-1406.

\item Maggi, G., Staiger, R., 2011. The Role of Dispute Settlement Procedures in International Trade Agreements. Quarterly Journal of Economics 126, 475-515.

\item Mansfield, E., Milner, H., Rosendorff, B.P., 2000. Free to Trade: Democracies, Autocracies, and International Trade. The American Political Science Review 94, 305-321.

%\item Milner, H., Kubota, K., 2005. Why the Move to Free Trade? Democracy and Trade Policy in the Developing Countries. International Organization 59, 107-143.

\item Milner, H., Rosendorff, B.P., 1997. Democratic Politics and International trade Negotiations: Elections and Divided Government as Constraints on Trade Liberalization. Journal of Conflict Resolution 41, 117-147.

\item Milner, H., Rosendorff, B.P., 2001. The Optimal Design of International Trade Institutions: Uncertainty and Escape. International Organization 55, 829-857.

\item Mitra, D., Thomakos, D., Ulubasoglu, M., 2002. `Protection for Sale' in a Developing Country: Democracy vs. Dictatorship. Review of Economics and Statistics 84, 497-508.

\item Mitra, D., Thomakos, D., Ulubasoglu, M., 2006. Can We Obtain Realistic Parameter Estimates for the Protection for Sale Model? Canadian Journal of Economics 39, 187-210.

%\item Morrow, J., Siverson, R., Tabares, T., 1999. Correction to: ``The Political Determinants of International Trade,' American Political Science Review 94, 931-933.

\item McCalman, P., 2004. Protection for Sale and Trade Liberalization: an Empirical Investigation. Review of International Economics 12, 81-94.

%\item Paltseva, E., 2011. Protection for Sale to Oligopolists. Available at: \url{http://web.econ.ku.dk/okoep/Paltseva_P4S2O_June2011_1.pdf}.

\item Saiegh, S., 2009. Political  Prowess or Lady Luck? Evaluating Chief Executives' Legislative Success Rates. The Journal of Politics 71, 1342-1356.

\item Song, Y., 2008. Protection for Sale: Agenda-Setting and Ratification in the Presence of Lobbying. Korea Institute for International Trade Policy Working Paper Series. Vol. 2008-22, August 2008.

%\item Tarar, A., 2001. International Bargaining with Two-Sided Domestic Constraints. Journal of Conflict Resolution 45, 320-340.

%\item Ward, M., Hoff, P., 2007. Persistent Patterns of International Commerce. Journal of Peace Research 44, 157-175.

\end{list}

\end{document}

### 2015-12-10 to-do.tex
add blank to-do list

\documentclass[12pt]{article}

\addtolength{\textwidth}{1.4in}
\addtolength{\oddsidemargin}{-.7in} %left margin
\addtolength{\evensidemargin}{-.7in}
\setlength{\textheight}{8.5in}
\setlength{\topmargin}{0.0in}
\setlength{\headsep}{0.0in}
\setlength{\headheight}{0.0in}
\setlength{\footskip}{.5in}
\renewcommand{\baselinestretch}{1.0}
\setlength{\parindent}{0pt}
\linespread{1.1}

\usepackage{amssymb, amsmath, amsthm, bm}
\usepackage{graphicx,csquotes,verbatim}
\usepackage[backend=biber,block=space,style=authoryear]{biblatex}
\setlength{\bibitemsep}{\baselineskip}
\usepackage[american]{babel}
%dell laptop
\addbibresource{C:/Users/Kristy/Dropbox/Research/xBibs/tradeagreements.bib}
%\addbibresource{C:/Users/Kristy/Documents/Dropbox/Research/xBibs/tradeagreements.bib}
\renewcommand{\newunitpunct}{,}
\renewbibmacro{in:}{}

\DeclareMathOperator*{\argmax}{arg\,max}
\usepackage{xcolor}
\hbadness=10000

\newcommand{\ve}{\varepsilon}
\newcommand{\ov}{\overline}
\newcommand{\un}{\underline}
\newcommand{\ta}{\theta}
\newcommand{\al}{\alpha}
\newcommand{\Ta}{\Theta}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\Bt}{B(\bm{\tau^a})}
\newcommand{\bta}{\bm{\tau^a}}
\newcommand{\btn}{\bm{\tau^n}}
\newcommand{\ga}{\gamma}
\newcommand{\Ga}{\Gamma}
\newcommand{\de}{\delta}

\begin{document}
\begin{center}
RIE R$\&$R of SOP
\end{center}

\vskip.3in

\end{document}

### 2015-12-10 to-do.tex
fill out to-do list with comments from editor and referee

@@
 
-\vskip.3in
+\vskip.1in
+Editor
+\begin{enumerate}
+	\item The introduction needs improvement and the rest of the paper needs substantial overhaul. 
+		\begin{itemize}
+			\item An efficient introduction formula starts with saying what the paper does and what it finds (two paragraphs) and then says how it achieves this and how it contributes to the extant literature.
+Do not overreach.
+In particular, I agree with Referee 1 that you do not solve the PFS empirical puzzle of high weights to social welfare.
+			\item (Referee) In the absence of any empirical estimates, I would refrain from such strong statements.
+Suffice to say that the paper is providing additional insights into an empirical puzzle by tailoring the political process more carefully.
+				\begin{itemize}
+					\item 8. Again, (5) makes it clear that your setting is definitely not PFS, for contributions are not specified (and need be to make the comparison).
+				\end{itemize}
+		\end{itemize}
+	\item Structure sections 2 and following. Right now, it is a seemingly random bag of assumptions, results, and discussions of both.
+	\item Define the game much more formally and more clearly. 
+		\begin{enumerate}
+			\item players, their strategy sets, the payoffs, and the timing. Right now, these ingredients are scattered all over the place.
+			\item A figure with the extensive form of the game would also come in handy.
+		\end{enumerate}
+	\item You need quasi-linear preferences in X and Y as well as in an outside good, Z.
+Preferences should be linear in Z, increasing and concave in Y and Y.
+Z should be produced with CRS and be freely traded.
+Otherwise, Foreign’s decisions have an impact on Home’s via the balance of trade condition.
+This problem is currently hidden by the fact that you impose symmetric countries (so that trade in X balances trade in Y in the symmetric equilibrium) but it remains true that X and Y are related (this is most visible off the symmetric equilibrium path).
+	\item The LHS term and the term in the last square bracket of the RHS of Equation 3 need different notations (you use the same notation for both).
+	\item The fact that $\gamma_E$ is a constant does not imply that protection is not for sale to the executive, as you write on page 10. It actually is, by the second equality of Assumption 1.
+		\begin{itemize}
+			\item Assumption 1 confusingly gets relabeled Ass. 4 later.
+			\item You actually impose these inequalities (in Assumption 1, I think) for all $\theta$ and for all e (not just $\theta$).
+		\end{itemize}
+	\item You need to set out the sufficient SOCs in the text.
+But I really don’t understand all the fuss you are making about these around Lemma 1 and (4).
+Indeed, the second order condition requires that the sum of the second derivatives of CS, PS, and TR wrt $\tau$ is negative.
+Thus, totally differentiating the FOC wrt to $\tau$ and $\gamma$ yields $\frac{d\tau}{d\gamma} = - \frac{\frac{\partial PS}{\partial\tau}}{SOC}$, which is positive (PS is increasing in prices and thus in $\tau$, “SOC”<0 by the SOC).
+In turn, the FOC to the lobbying effort decision can be written as $\frac{\partial PS}{\partial\tau} \cdot \frac{d\tau}{d\gamma} \frac{ d\gamma}{de} = 1$, which implies $\frac{d\tau}{d e} = - \frac{\frac{\partial PS}{ \partial\tau}}{ SOC} * \gamma_e$, which is larger than zero by assumption, $\gamma_e>0$, and by the SOCs.
+	\addtocounter{enumi}{1}
+	\item Since the game is symmetric, assuming that the other legislature is not being lobbied and/or does not risk making a ‘break’ decision is not the most natural one.
+I can leave with it, but then I would simply assume this explicitly and openly when writing the extensive form of the game, without the current, twisted discussion of it.
+	\item I am not sure that section 6 adds much to the paper.
+\end{enumerate}
 
+\vskip.4in
+Referee 1
+\begin{enumerate}
+	\item The paper is derailing at the start.
+It initially reads like political uncertainty will be incorporated into Grossman and Helpman (1994) framework through modelling separation of powers in that setting.
+But later it comes up that the paper is actually about ratification of free trade agreements, in particular, the lobbying process that aims to prevent the agreement from being ratified.
+A better start that is consistent with the main contribution of the paper is in order.
+The author can relate the main contribution to the ‘unrealistic’ PFS estimates  following the introduction of the main contribution.
+	\item Also, because the main contribution of the paper is the introduction of political uncertainty into the lobbying process, some more intuition in the Introduction about political uncertainty would be helpful.
+	\item While researchers have found high estimates in the case of the US, they also found high estimates for Turkey (Mitra, Thomakos and Ulubasoglu 2002) and Australia (McCalman 2004).
+The issue is that both Turkey and Australia are parliamentary democracies.
+Additionally, in Turkey the legislative organ is uni-cameral, but in Australia it is bi-cameral.
+The point here is that these parliamentary democracies do not exhibit strict separation of powers as does the US, but still the PFS model in these countries delivers high empirical estimates.
+I expect that the author provide some insights into this issue.
+\end{enumerate}
 \end{document}

### 2016-01-20 unbalanced_trade.m
no longer need long matlab version of unbalanced-trade code

@@ -1,80 +1,0 @@
-%Maintaining symmetry, I want to try a bigger intercept for the gamma
-%function PLUS probabilistic one-sided breaking (same gamma as lobby4).
-
-%gamma: 1.25 + cn^.2
-%cn = .00166
-% tn = (8*(1 + .00156^.1)-5)/(68-8*(1 + .00156^.1)) = 0.12592
-
-tn = (8*(1.25 + .00166^.2)-5)/(68-8*(1.25 + .00166^.2));
-
-a=.2;
-tns=tn;
-ta = [0:.001:.1259];
-tad = size(ta,2);
-c=[0:.00005:.005];
-tac = size(c,2);
-[ta,c]=meshgrid(ta,c);
-
-lhs = (8*(1.25 + (c).^a)-5).*(tn-ta) + (4*(1.25 + (c).^a)-34).*(tn^2-ta.^2)-3.*(tns-ta)+9.*(tns^2-ta.^2);
-
-%add e to gamma
-e1=.25; e2=.25;
-up = (8*(1.25 + (c).^a +e1)-5).*(tn-ta) + (4*(1.25 + (c).^a +e1)-34).*(tn^2-ta.^2)-3.*(tns-ta)+9.*(tns^2-ta.^2);
-low = (8*(1.25 + (c).^a -e2)-5).*(tn-ta) + (4*(1.25 + (c).^a -e2)-34).*(tn^2-ta.^2)-3.*(tns-ta)+9.*(tns^2-ta.^2);
-spread = up - low;
-b = up./spread;
-for j=1:tad
-    if lhs(1,j)< 0
-        b(1,j) = 0;
-        else
-        b(1,j) = 1;
-     end; 
-     for i=2:101
-        if up(i,j) < 0
-        b(i,j) = 0;
-        elseif low(i,j) > 0
-        b(i,j) = 1;
-      end
-    end
-end
-
-l = b;
-l2 = 4/49*((1+tn).^2);
-l3 = l.*l2;
-l4 = 4/49*((1+ta).^2);
-l5 = l3 + (1-l).*l4 - c;
-[M,I] = max(l5,[],1);
-opt_C = (I -1)/20000;
-%plot(TA,opt_C);
-
-
-TA = [0:.001:.125];
-
-for j=1:tad
-  B(j) = b(I(j),j);
-end
-
-E=B;
-E2 = 21/49*(2-2*tn^2);
[…]

### 2016-01-20 submitted_example.r
no idea why I need a separate commit after changing extension on submitted_example to r

@@ -5,7 +5,7 @@
 %cn = .00166
 % tn = (8*(1 + .00156^.1)-5)/(68-8*(1 + .00156^.1)) = 0.12592
 
-tn = (8*(1.25 + .00166^.2)-5)/(68-8*(1.25 + .00166^.2));
+tn = (8*(1.25 + .00166^.2)-5)/(68-8*(1.25 + .00166^.2))
 
 a=.2;
 tns=tn;

### 2016-01-21 submitted_example.r
continue changing over 'submitted_example' from matlab to R code

@@ -1,28 +1,37 @@
-%Maintaining symmetry, I want to try a bigger intercept for the gamma
-%function PLUS probabilistic one-sided breaking (same gamma as lobby4).
+#Maintaining symmetry, I want to try a bigger intercept for the gamma
+#function PLUS probabilistic one-sided breaking (same gamma as lobby4).
 
-%gamma: 1.25 + cn^.2
-%cn = .00166
-% tn = (8*(1 + .00156^.1)-5)/(68-8*(1 + .00156^.1)) = 0.12592
+#gamma: 1.25 + cn^.2
+#cn = .00166
+#tn = (8*(1 + .00156^.1)-5)/(68-8*(1 + .00156^.1)) = 0.12592
 
 tn = (8*(1.25 + .00166^.2)-5)/(68-8*(1.25 + .00166^.2))
 
-a=.2;
-tns=tn;
-ta = [0:.001:.1259];
-tad = size(ta,2);
-c=[0:.00005:.005];
-tac = size(c,2);
-[ta,c]=meshgrid(ta,c);
+a=.2
+tns=tn
+ta_vector = seq(0,.1259,.001)
+#tad = NROW(ta)
+c_vector=seq(0,.005,.00005)
+#tac = NROW(c)
+#[ta,c]=meshgrid(ta,c);
 
-lhs = (8*(1.25 + (c).^a)-5).*(tn-ta) + (4*(1.25 + (c).^a)-34).*(tn^2-ta.^2)-3.*(tns-ta)+9.*(tns^2-ta.^2);
+#add e to gamma
+e1=.25
+e2=.25
 
-%add e to gamma
-e1=.25; e2=.25;
-up = (8*(1.25 + (c).^a +e1)-5).*(tn-ta) + (4*(1.25 + (c).^a +e1)-34).*(tn^2-ta.^2)-3.*(tns-ta)+9.*(tns^2-ta.^2);
-low = (8*(1.25 + (c).^a -e2)-5).*(tn-ta) + (4*(1.25 + (c).^a -e2)-34).*(tn^2-ta.^2)-3.*(tns-ta)+9.*(tns^2-ta.^2);
-spread = up - low;
-b = up./spread;
+param <- expand.grid("ta" = ta_vector, "c" = c_vector)
+
+l <- function(ta,c) {
[…]

### 2016-01-22 draft.tex
compile draft for first time, fix footnote on title

@@
 \begin{document}
-\title{\vskip-0.6in Trade Agreements in the Shadow of Lobbying\footnote{Formerly circulated under the title ``Trade Agreements, Lobbying and Separation of Powers''}}
+\title{\vskip-0.6in Trade Agreements in the Shadow of Lobbying\protect\footnote{Formerly circulated under the title ``Trade Agreements, Lobbying and Separation of Powers''}}
 \thanksmarkseries{alph}

### 2016-01-25 submitted_example.r
add comments to submitted_example.r

@@ -52,7 +52,7 @@ res <- Map(l, ta = param$ta, c = param$c)
 soln <- lapply(seq_along(res), function(x) res[[x]]$value)
 soln <- do.call("rbind", soln)
 
-
+#lobby's profits
 l = b;
 l2 = 4/49*((1+tn).^2);
 l3 = l.*l2;
@@ -69,6 +69,7 @@ for j=1:tad
   B(j) = b(I(j),j);
 end
 
+#exec utility with ga=1 and tn=tns
 E=B;
 E2 = 21/49*(2-2*tn^2);
 E3 = E*E2;

### 2016-01-25 to-do.tex
fix typo in to-do list

@@
 Indeed, the second order condition requires that the sum of the second derivatives of CS, PS, and TR wrt $\tau$ is negative.
-Thus, totally differentiating the FOC wrt to $\tau$ and $\gamma$ yields $\frac{d\tau}{d\gamma} = - \frac{\frac{\partial PS}{\partial\tau}}{SOC}$, which is positive (PS is increasing in prices and thus in $\tau$, “SOC”<0 by the SOC).
+Thus, totally differentiating the FOC wrt to $\tau$ and $\gamma$ yields $\frac{d\tau}{d\gamma} = - \frac{\frac{\partial PS}{\partial\tau}}{SOC}$, which is positive (PS is increasing in prices and thus in $\tau$, “SOC”$<0$ by the SOC).
 In turn, the FOC to the lobbying effort decision can be written as $\frac{\partial PS}{\partial\tau} \cdot \frac{d\tau}{d\gamma} \frac{ d\gamma}{de} = 1$, which implies $\frac{d\tau}{d e} = - \frac{\frac{\partial PS}{ \partial\tau}}{ SOC} * \gamma_e$, which is larger than zero by assumption, $\gamma_e>0$, and by the SOCs.

### 2016-01-25 draft.tex
remove lemma 0 from draft

@@
 \label{sec:appendix}
-\begin{lemma*}[\textbf{0}]
-  When prices are linear in tariffs, the second-order condition for the legislature's problem when setting trade-war tariffs holds for all non-prohibitive tariffs given that the tariff elasticity of supply is less than unity.
-  \label{res:l1}
-\end{lemma*}
-Proof: The first order condition can be rewritten as
+\noindent \textbf{\hypertarget{tninc}{Proof of Lemma~\ref{res:tninc}}}:\\
+Expand the first order condition for the legislature's problem as follows:
    \begin{equation}
@@
 	 \end{equation}
-
-	 \vskip.1in
-	 \noindent The required second order condition is that the following is negative:
-	 \begin{multline*}
-	    	- D(P_{X}) \frac{\partial^2 P_X}{\partial \tau^2} - \frac{\partial D(P_{X})}{\partial \tau}  \frac{\partial P_X}{\partial \tau} 
-	    	+ \ga(e_n,\ve_n) Q_X(P_{X}) \frac{\partial^2 P_X}{\partial \tau^2} + \ga(e_n,\ve_n) \frac{\partial Q_X(P_{X})}{\partial \tau} \frac{\partial P_X}{\partial \tau} \\
-	    + 2 \left[ \frac{\partial D(P_{X})}{\partial \tau} - \frac{\partial Q_X(P_{X})}{\partial \tau}\right] 
-	    + \tau \left(\frac{\partial^2 D(P_X)}{\partial \tau^2} - \frac{\partial^2 Q_X(P_X)}{\partial \tau^2} \right)
-	  \end{multline*}
-
-	\vskip.1in  
-	\noindent $\ga(e_n,\ve_n)$ is endogenous, so we must find a way to bound it.
-Because the lobby receives no gains if the tariff is raised beyond its prohibitive level (that is when $D(P_{X}) = Q_X(P_{X})$), it has no incentive to exert effort above the level that induces the $\ga(e_n,\ve_n)$ that leads to the prohibitive tariff.
-Setting $D(P_{X}) = Q_X(P_{X})$ in the first order condition, this is
-    \[
-	     \ga(e_n,\ve_n) = 1 - \frac{\tau \left(\frac{\partial D(P_X)}{\partial \tau} - \frac{\partial Q_X(P_X)}{\partial \tau} \right)}{Q_X(P_{X}) \frac{\partial P_X}{\partial \tau}}
-	  \]
-
-	\noindent If prices are linear in tariffs, the second order terms are zero, and substituting the above expression for $\ga(e_n,\ve_n)$, we arrive at
-    \[
-	    	- \frac{\partial D(P_{X})}{\partial \tau}  \frac{\partial P_X}{\partial \tau} + \frac{\partial Q_X(P_{X})}{\partial \tau} \frac{\partial P_X}{\partial \tau} - \tau \left(\frac{\partial D(P_X)}{\partial \tau} - \frac{\partial Q_X(P_X)}{\partial \tau} \right) \frac{ \frac{\partial Q_X(P_{X})}{\partial \tau}}{Q_X(P_{X})} + 2 \left[ \frac{\partial D(P_{X})}{\partial \tau} - \frac{\partial Q_X(P_{X})}{\partial \tau}\right]
-	  \]
-	  Because $P_X = P_X^W + \tau$, $\frac{\partial P_X}{\partial \tau} = \frac{\partial P_X^W}{\partial \tau} + 1$; together with $\frac{\partial P_X^W}{\partial \tau} < 0,$ this implies $\frac{\partial P_X}{\partial \tau} < 1$.
-Therefore the previous expression is bounded above by
-	  \[
-	    - \tau \left(\frac{\partial D(P_X)}{\partial \tau} - \frac{\partial Q_X(P_X)}{\partial \tau} \right) \frac{ \frac{\partial Q_X(P_{X})}{\partial \tau}}{Q_X(P_{X})} + \left[ \frac{\partial D(P_{X})}{\partial \tau} - \frac{\partial Q_X(P_{X})}{\partial \tau}\right]
-	  \]
-	 Since $\frac{\partial D(P_{X})}{\partial \tau} - \frac{\partial Q_X(P_{X})}{\partial \tau}<0$, the second order condition is negative as long as $\frac{\tau}{Q_X(P_{X})}\frac{\partial Q_X(P_{X})}{\partial \tau} < 1.$
-	 If prices are not linear in the tariff, additional conditions are required. $\hfill\blacksquare$
-
-\vskip.4in	 
-\noindent \textbf{\hypertarget{tninc}{Proof of Lemma~\ref{res:tninc}}}:\\
-By the implicit function theorem, $\frac{\partial \tau^n}{\partial \ga} = - \frac{\frac{\partial FOC}{\partial \ga}}{\frac{\partial FOC}{\partial \tau^n}}$ where \textit{FOC} is the first order condition for the legislature's problem, expanded in Equation~\ref{eq:foc}.
+I will refer to this equation as \textit{FOC}.
+
+By the implicit function theorem, $\frac{\partial \tau^n}{\partial \ga} = - \frac{\frac{\partial FOC}{\partial \ga}}{\frac{\partial FOC}{\partial \tau^n}}$.
 We have $\frac{\partial FOC}{\partial \ga} = Q_X(P_{X}) \frac{\partial P_X}{\partial \tau}$ from the proof of Lemma 0.

### 2016-01-25 draft.tex
remove duplicate assumption ga_l_e and merged textual description

@@
 That is, even for the least favorable outcome of the lobbying process, the legislature will be at least slightly more protectionist than the executive.
-This assumption is not essential, but it matches well the empirical findings that politicians with larger constituencies are less sensitive to special interests, and it simplifies the analysis considerably, in particularly ensuring that trade agreement tariffs are less than trade war tariffs.
+This assumption is not essential, but it matches well the empirical findings that politicians with larger constituencies are less sensitive to special interests, and it simplifies the analysis considerably, in particular ensuring that trade agreement tariffs are less than trade war tariffs (See Destler (2005) and footnote~\ref{fn:ga_l_e} above).
 
@@
 Note that the legislature's best response functions (both for whether to break the agreement, and at what level to set the tariff if it does) will be affected by the shape of $\ga(e,\ve)$.
-I make the following assumptions on the weight on import-industry profits:
+I make the following further assumptions on the weight on import-industry profits:
 
@@
 
-\begin{assumption}
-  $\ga(e,\ve) \geq \ga_E \geq 1 \ \forall \ve$.
-  \label{as:ga_l_e}
-\end{assumption}
-
 Assumption~\ref{as:ga_c} formalizes the intuition that the legislature favors the import-competing industry more the higher is its lobbying effort, but that there are diminishing returns to lobbying activity.\footnote{The diminishing returns here take the form of declining increments to the lobby's influence as effort increases; in Ethier (2012), the returns to lobbying decline as the level of protection increases.} It allows the political uncertainty inherent in $\ta$ to be correlated in interesting ways with the level of $e$, ruling out only (i) that higher effort levels make lower $\ga$-weights more likely and (ii) that effort and uncertainty are correlated in such a way that increasing lobbying effort changes the structure of uncertainty so that higher $\ga$-weights become more likely at an accelerating pace.
@@
 
-Assumption~\ref{as:ga_l_e} ensures that $\tau^a < \tau^n$, and more generally, that the legislature's incentives are more closely aligned with the lobby's than are those of the executive.
+%Assumption~\ref{as:ga_l_e} ensures that $\tau^a < \tau^n$, and more generally, that the legislature's incentives are more closely aligned with the lobby's than are those of the executive.
 This is not essential but simplifies the analysis and matches well the empirical findings that politicians with larger constituencies are less sensitive to special interests (See Destler (2005) and footnote~\ref{fn:ga_l_e} above).

### 2016-01-25 draft.tex
rewrite sec:twt to integrate SOC into text and simplify/remove formal proof of lemma 1

@@
 \subsection{Trade-War Tariffs}
+\label{sec:twt}
 In the event that the trade agreement is broken, the legislature sets its tariff $\tau$ unilaterally by maximizing Equation~\ref{eq:ml} given $\tau^*$.
@@
 It is not necessary to restrict them to come from the same distribution, although in applications it would seem unlikely that the distribution would change.}
-\[
-		\frac{\partial \mathit{CS}_X(\tau)}{\partial \tau^n} + \ga(e_n,\ve_n) \cdot \frac{\partial \mathit{PS}_X(\tau)}{\partial \tau^n} +  \frac{\partial \mathit{TR}(\tau)}{\partial \tau^n} = 0
-\]
-Conditions under which the second order condition is satisfied are contained in Lemma 0 in the Appendix.
-
-In accordance with intuition, because profits are increasing in the tariff, trade war tariffs are increasing in the weight attached to the profits of the import-competing industry, as the following result shows:
+\begin{equation}
+		\frac{\partial \mathit{CS}_X(\tau)}{\partial \tau^n} + \ga(e_n,\ve_n) \cdot \frac{\partial \mathit{PS}_X(\tau)}{\partial \tau^n} +  \frac{\partial \mathit{TR}(\tau)}{\partial \tau^n} = 0 .
+		\label{eq:legfoc}
+\end{equation}
+
+In the event of a trade war, the lobby chooses its effort $e_n$ given this tariff-setting behavior by maximizing its profits net of effort: $\pi\left(\tau^n\left(\ga\left(e_n,\ve_n\right)\right)\right) - e_n$.
+(Note this is Equation~\ref{eq:lv} simplified by the resolution of uncertainty over the legislature's decision on the trade agreement).
+This implies a first order condition for the lobby of
+\begin{equation}
+	\frac{\partial \pi(\tau^n)}{\partial \tau^n}\frac{\partial \tau^n}{\partial \ga} \frac{\partial \ga}{\partial e_n} = 1
+  \label{eq:lobtw}
+\end{equation}
+That is, at this stage, the lobby chooses the level of effort that equates its expected marginal increase in profits with its marginal payment.
+Further details about the political weighting function $\ga$ and assumed structure of uncertainty are provided in the next section.
+
+In accordance with intuition, because profits are increasing in the tariff, trade war tariffs are increasing in the weight attached to the profits of the import-competing industry.
+This can be seen by rearranging Equation~\ref{eq:lobtw} as
+\begin{equation}
+  \frac{\partial \tau^n}{\partial \ga} = \frac{1}{\frac{\partial \pi(\tau^n)}{\partial \tau^n} \frac{\partial \ga}{\partial e_n}}
+	\label{eq:lem1}
+\end{equation}
+since $\frac{\partial \ga}{\partial e_n}$ is positive by Assumption~\ref{as:ga_c} and profits are increasing in prices which are in turn increasing in the tariff.
+This expression can also demonstrate that the second order condition for the legislature's problem is always satisfied.
+By the implicit function theorem, $\frac{\partial \tau^n}{\partial \ga} = -\frac{\frac{\partial \text{Equation}~\ref{eq:legfoc}}{\partial \ga}}{\frac{\partial \text{Equation}~\ref{eq:legfoc}}{\partial \tau^n}} = -\frac{\frac{\partial \pi(\tau^n)}{\partial \tau^n}}{SOC}$ where ``SOC'' stands for `Second Order Condition', the second derivative of the legislature's objective function which should be everywhere negative.
+Setting this expression equal to Equation~\ref{eq:lem1}, the second order condition must be satisfied since $\frac{\partial \ga}{\partial e_n}$ is positive by Assumption~\ref{as:ga_c}.
 
 \begin{lemma}
-  If prices are linear in tariffs, $\tau^n(\ga(e_n,\ve_n))$ is increasing in $\ga(e_n,\ve_n)$.
+  $\tau^n(\ga(e_n,\ve_n))$ is increasing in $\ga(e_n,\ve_n)$.
   \label{res:tninc}
@@
 All proofs are in the Appendix.
-Weaker conditions for this lemma are possible, but most examples in the literature satisfy this condition.
-The linearity serves to guarantee that the second order condition is satisifed.
-
-\noindent In the event of a trade war, the lobby chooses its effort $e_n$ given this tariff-setting behavior by maximizing its profits net of effort: $\pi\left(\tau^n\left(\ga\left(e_n,\ve_n\right)\right)\right) - e_n$.
-(Note this is Equation~\ref{eq:lv} simplified by the resolution of uncertainty over the legislature's decision on the trade agreement).
-This implies a first order condition of
-\begin{equation}
-	\frac{\partial \pi(\tau^n(\ga(e_n,\ve_n)))}{\partial e_n} = 1
-  \label{eq:lobtw}
-\end{equation}
-That is, at this stage, the lobby chooses the level of effort that equates its expected marginal increase in profits with its marginal payment.
-Further details about the political weighting function $\ga$ and assumed structure of uncertainty are provided in the next section.
 

### 2016-01-25 draft.tex
remove formal lemma 1 and its proof; move statement about proofs in appendix to after result 1

@@
 
-\begin{lemma}
-  $\tau^n(\ga(e_n,\ve_n))$ is increasing in $\ga(e_n,\ve_n)$.
-  \label{res:tninc}
-\end{lemma}
-All proofs are in the Appendix.
-
 \bigskip
@@
 
-\noindent Lobbying affects only the weight the legislature places on the profits of the import-competing industry.
+\noindent All proofs are in the Appendix.
+
+Lobbying affects only the weight the legislature places on the profits of the import-competing industry.
 These profits are higher in a trade war than under a trade agreement, so given Assumption~\ref{as:ga_c}, the expectation of $\ga$ is increasing in lobbying effort, implying that the legislature becomes more favorably inclined toward the high trade-war tariff and associated profits as lobbying increases and therefore more likely to break the trade agreement.
@@
 \label{sec:appendix}
-\noindent \textbf{\hypertarget{tninc}{Proof of Lemma~\ref{res:tninc}}}:\\
-Expand the first order condition for the legislature's problem as follows:
-   \begin{equation}
-        - D(P_{X}) \frac{\partial P_X}{\partial \tau} + \ga(e_n,\ve_n) Q_X(P_{X}) \frac{\partial P_X}{\partial \tau}
-	    + \left[ D(P_{X})- Q_X(P_{X})\right] + \tau \left(\frac{\partial D(P_X)}{\partial \tau} - \frac{\partial Q_X(P_X)}{\partial \tau} \right) = 0
-			\label{eq:foc}
-	 \end{equation}
-I will refer to this equation as \textit{FOC}.
-
-By the implicit function theorem, $\frac{\partial \tau^n}{\partial \ga} = - \frac{\frac{\partial FOC}{\partial \ga}}{\frac{\partial FOC}{\partial \tau^n}}$.
-We have $\frac{\partial FOC}{\partial \ga} = Q_X(P_{X}) \frac{\partial P_X}{\partial \tau}$ from the proof of Lemma 0.
-This is always positive.
-Lemma 0 also shows that $\frac{\partial FOC}{\partial \tau^n}$ is negative everywhere.
-Thus $\frac{\partial \tau^n}{\partial \ga}\geq 0$.
-$\hfill\blacksquare$
-
-\vskip.4in
 \noindent \textbf{\hypertarget{Pr_bincC}{Proof of Result~\ref{res:bincC}}}:\\

### 2016-01-25 draft.tex
begin letter to editor/referee and fix one tiny typo in draft

@@
 
-It is also positive for all other values of $\bta$ strictly below $\btn$. Just as in the proof of Lemma~\ref{res:leg_a}, one one can add
+It is also positive for all other values of $\bta$ strictly below $\btn$. Just as in the proof of Lemma~\ref{res:leg_a}, one can add
 $\left(\tilde{\Gamma} - 1 \right) \frac{\partial \mathit{PS}_X(\tau^a)}{\partial \tau^a} \left(\mathit{PS}_X(\tau^n) - \mathit{PS}_X(\tau^a)\right)$ to the first summand and subtract it from the second.

### 2016-01-25 letter.tex
begin letter to editor/referee and fix one tiny typo in draft

\documentclass[12pt]{article}

\addtolength{\textwidth}{1in}
\addtolength{\oddsidemargin}{-.5in} %left margin
\addtolength{\evensidemargin}{-.5in}
\setlength{\textheight}{8.5in}
\setlength{\topmargin}{-.5in}
\setlength{\headheight}{18.0pt}
\setlength{\footskip}{.375in}
\renewcommand{\baselinestretch}{1.0}
\linespread{1.2}
\usepackage{fancyhdr}
\pagestyle{fancy}
\renewcommand{\headrulewidth}{1pt}
\lhead{Trade Agreements in the Shadow of Lobbying}
%\chead{}
\rhead{Kristy Buzard}
%\fancyfoot[C]{}
%\fancyhead[R]{}

\usepackage[pdftex,
bookmarks=true,
bookmarksnumbered=false,
pdfview=fitH,
bookmarksopen=true]{hyperref}
\usepackage[pdftex]{graphicx}
\setcounter{secnumdepth}{-1}

\usepackage[usenames,dvipsnames]{color}
\usepackage{cite}
\usepackage{times, verbatim,bm,pifont}
\usepackage{amsbsy,amssymb, amsmath, amsthm, MnSymbol,bbding}

\begin{document}

\begin{center}
\large Response to Editor and Referee \\
\normalsize February 2, 2016
\end{center}

Editor's point $\#$7: Thank you for pointing out that the second order condition holds everywhere and that the proof of what was formerly Lemma 1 can be drastically simplified. I moved the discussion of the second order condition into the text and integrated it with the simple result on the increasingness of trade war tariff in the political economy weight. As this did not need to be referenced elsewhere and no longer requires a proof in the appendix, it is no longer labeled Lemma 1. Likewise, Lemma 0 is no longer necessary and has been removed. Note that this advances the numbering of the lemmas in the revised draft relative to the previous draft.

\end{document}

### 2016-01-26 draft.tex
revise first paragraph of model section to integrate numeraire

@@
 \label{sec:model}
-Following the literature on the political economy of trade agreements as exemplified by Bagwell and Staiger (2005), I employ a partial equilibrium model with two countries: home (no asterisk) and foreign (asterisk).
-The countries trade two goods, $X$ and $Y$, where $P_i$ denotes the home price of good $i \in \left\{X,Y\right\}$ and $P_i^*$ denotes the foreign price of good $i$.
-In each country, the demand functions are taken to be identical for both goods, respectively $D(P_i)$ in home and $D(P_i^*)$ in foreign and are assumed strictly decreasing and twice continuously differentiable.
+I begin by describing the basic economic setting within which trade occurs.
+It is a three-good model with two countries: home (no asterisk) and foreign (asterisk).
+In each country, preferences are linear in good $N$, which is denoted the numeraire, while the demand functions for $X$ and $Y$ are assumed strictly decreasing and twice continuously differentiable.
+Further, the demand functions for $X$ and $Y$ are taken to be identical and denoted $D(P_i)$ in home and $D(P_i^*)$ in foreign.
+$P_i$ denotes the home price of good $i \in \left\{N,X,Y\right\}$ and $P_i^*$ denotes the foreign price of good $i$.
 

### 2016-01-26 draft.tex
revise paragraphs 2-4 of sec:model to include numeraire; add footnote about analyzing only x, home and y,foreign cases

@@
 
+Good $N$ is produced with labor alone with constant returns to scale technology so that $Q_N = l_N$.
+I assume the aggregate labor supply is large enough to ensure that the output of good $N$ is large enough to guarantee balanced trade.
 The supply functions for good $X$ are $Q_X(P_X)$ and $Q_X^*(P_X^*)$ and are assumed strictly increasing and twice continuously differentiable for all prices that elicit positive supply.
@@
 The production structure for good $Y$ is taken to be symmetric, with both demand and supply such that the economy is separable in goods $X$ and $Y$.
-It is assumed that the production of each good requires the possession of a sector-specific factor that is available in inelastic supply and is non-tradable so that the income of owners of the specific factors is tied to the price of the good in whose production their factor is used.
+It is assumed that the production of goods $X$ and $Y$ requires, in addition to labor, the possession of a sector-specific factor that is available in inelastic supply and is non-tradable so that the income of owners of the specific factors is tied to the price of the good in whose production their factor is used.
 
 For simplicity, I assume each government's only trade policy instrument is a specific tariff on its import-competing good: the home country levies a tariff $\tau$ on good $X$ while the foreign country applies a tariff $\tau^*$ to good $Y$.
-Local prices are then $P_X = P_X^W + \tau$, $P_X^* = P_X^W$, $P_Y = P_Y^W$ and $P_Y^* = P_Y^W + \tau^*$ where a $W$ superscript indicates world prices and equilibrium prices are determined by the market clearing conditions
+Local prices are then $P_X = P_X^W + \tau$, $P_X^* = P_X^W$, $P_Y = P_Y^W$ and $P_Y^* = P_Y^W + \tau^*$ where a $W$ superscript indicates world prices.
+Equilibrium prices are determined by the market clearing conditions
 $$M_X(P_X)= D(P_X)-Q_X(P_X) = Q_X^*(P_X^*) - D(P_X^*) = E_X^*(P_X^*)$$
@@
 where $M_X$ are home-county imports and $E_X^*$ are foreign exports of good $X$ and $E_Y$ are home-county exports and $M_Y^*$ are foreign imports of good $Y$.
-
-It follows that $P_X^W$ and $P_Y^W$ are decreasing in $\tau$ and $\tau^*$ respectively, while $P_X$ and $P_Y^*$ are increasing in the respective domestic tariff.
+Given the production technology and no trade taxes, the price of the numeraire is equal to one in both countries and on the world market.
+
+$P_X^W$ and $P_Y^W$ are decreasing in $\tau$ and $\tau^*$ respectively, while $P_X$ and $P_Y^*$ are increasing in the respective domestic tariff.
 This gives rise to a standard terms-of-trade externality.
@@
 As the economy is fully separable and the economic and political structures are symmetric, I focus here on the home country and the $X$-sector.
-The details are analogous for $Y$ and foreign.
+The details are analogous for $Y$ and foreign.\footnote{These are the relevant sectors for analysis because they are the sectors in which there are lobbies.}
 

### 2016-01-26 draft.tex
add footnote on why labor income not included in welfare

@@
 Good $N$ is produced with labor alone with constant returns to scale technology so that $Q_N = l_N$.
-I assume the aggregate labor supply is large enough to ensure that the output of good $N$ is large enough to guarantee balanced trade.
+I assume the aggregate labor supply $l$ is large enough to ensure that the output of good $N$ is large enough to guarantee balanced trade.
 The supply functions for good $X$ are $Q_X(P_X)$ and $Q_X^*(P_X^*)$ and are assumed strictly increasing and twice continuously differentiable for all prices that elicit positive supply.
@@
 \end{equation}
-where $\mathit{CS}$ is consumer surplus, $\mathit{PS}$ is producer surplus, $\ga(e,\ta)$ is the weight placed on producer surplus (profits) in the import-competing industry, $e$ is lobbying effort, $\ta$ is a random variable, and $\mathit{TR}$ is tariff revenue.
+where $\mathit{CS}$ is consumer surplus, $\mathit{PS}$ is producer surplus, $\ga(e,\ta)$ is the weight placed on producer surplus (profits) in the import-competing industry, $e$ is lobbying effort, $\ta$ is a random variable, and $\mathit{TR}$ is tariff revenue.\footnote{Labor income $l$ could also be included in both median legislator and executive welfare.
+This alters none of the results and so would only serve to complicate the exposition.}
 

### 2016-01-26 letter.tex
add response to letter re:numeraire (E4)

@@
 
+Editor's point $\#$4: I have added a traded numeraire good to address the concern about balanced trade.
+As all the results hinge on the incentives of the lobby, the executive and the median legislator---none of which are impacted by the addition of the numeraire good---this does not change any of the results.
+In carefully examining this issue, I realized that it would be more complete to include labor income in the welfare functions of the executive and lobby (again, this would change no results, as they all involve either differencing two welfare terms or taking derivatives with respect to a tariff).
+However, since much of the literature I'm following abstracts from a balanced trade condition and a numeraire sector with labor input, this would complicate the exposition and comparison to the extant literature.
+I have therefore chosen to insert Footnote [{\color{blue}ADD FOOTNOTE REFERENCE}] to this effect.
+It would be a simple matter to put labor income into the welfare functions if that is judged preferable.
+
 Editor's point $\#$7: Thank you for pointing out that the second order condition holds everywhere and that the proof of what was formerly Lemma 1 can be drastically simplified.

### 2016-01-26 draft.tex
extensions section removed, fn on asymmetry added

@@
 Section~\ref{sec:dis} explores the connection between the model under consideration here and the `Protection for Sale' theoretical framework and includes in-depth discussions of the roles of uncertainty and the separation-of-powers structure.
-Several extensions of the model are explored in Section~\ref{sec:ext} and Section~\ref{sec:concl} concludes.
+Section~\ref{sec:concl} concludes.
 
@@
 I also assume that each legislature will have the opportunity to take a vote to break the agreement with probability less than one-half, and that the probability that both legislatures have the opportunity is vanishingly small.
-This allows me to focus here on the key interaction between the domestic actors and abstract from the strategic interaction between the lobbies in the two countries.
-I discuss an extension to the case in which both legislatures vote to break the agreement at the same time in Section~\ref{sec:asym}.
+This allows me to focus here on the key interaction between the domestic actors and abstract from the strategic interaction between the lobbies in the two countries.\footnote{An incentive for each lobby to free ride on the effort of the other makes the optimization problem of the executives more complex as they are faced with the problem of how best to exploit the free-riding dynamic.}
 
@@
 where expectation operators have been dropped for convenience and $t$ is an intergovernmental transfer.\footnote{While direct monetary transfers have to date rarely been used in practice, it seems appropriate to interpret linked concessions on non-trade issues as indirect transfers (Klimenko, Ramey and Watson (2008); Maggi and Staiger (2011)).
-Here, transfers do not occur as long as the full set of symmetry assumptions are maintained but are an important consideration in the environment of Section~\ref{sec:asym}.} The solution to this system of equations determines the tariffs $\bta$ that will be set in the trade agreement.
+Here, transfers do not occur as long as the full set of symmetry assumptions are maintained but are an important consideration in asymmetric environments.} The solution to this system of equations determines the tariffs $\bta$ that will be set in the trade agreement.
 
@@
 Here $o$ and $o^*$ are the probabilities that the home and foreign legislatures, respectively, will have the opportunity to break the trade agreement; these are assumed to be mutually exclusive events that are realized after the conclusion of the trade agreement.
-The case in which both legislatures can potentially break the agreement is treated in Section~\ref{sec:two}.
 One can think of $o$ and $o^*$ as being determined by events beyond the executives' or legislatures' control that determine whether or not the legislatures will be willing to consider the lobby's request; for example, they would be affected by the occurrence of an economic crisis that diverts the legislature's attention from less-pressing matters.
@@
 
+\begin{comment}
 \section{Extensions}
 \label{sec:ext}
-\begin{comment}
+
 This analysis has been undertaken with the assumption that the import-competing industry is distributed evenly across legislative districts.
@@
 I also plan to use the richness of this model to parse out how much of the changes in tariffs comes from the fact that the executives place lower weight on producer profits and how much comes from the fact that they make an agreement---that is, compare the effects of structure versus preferences.
-\end{comment}
 
@@
 What is clear is that the optimization of the executives becomes significantly more complex as they are now faced with the problem of how to best exploit the free-riding dynamics that occur between the lobbies.
+\end{comment}
 

### 2016-01-26 letter.tex
add note on removing section 6 to letter
add explanation of E5 to letter

@@
 
+Editor's point $\#$5: I have added expectations operators to Equations (3) and (4) [{\color{blue}CHECK EQ:JV AND EQ:LV}] to address the redundant notation.
+Thank you for pointing out this needed clarification.
+
 Editor's point $\#$7: Thank you for pointing out that the second order condition holds everywhere and that the proof of what was formerly Lemma 1 can be drastically simplified.
@@
 
+Editor's point $\#10$: I have removed the extensions (Section 6) per your suggestion. {\color{blue}ADD discussion about fixing exposition around one break at a time}
+
 \end{document}

### 2016-01-27 draft.tex
add expectations operators to eq:lv and eq:jv; closes #5 / E5
bold changes that have been made to draft so far

@@
 \label{sec:model}
-I begin by describing the basic economic setting within which trade occurs.
+\textbf{I begin by describing the basic economic setting within which trade occurs.
 It is a three-good model with two countries: home (no asterisk) and foreign (asterisk).
@@
 Further, the demand functions for $X$ and $Y$ are taken to be identical and denoted $D(P_i)$ in home and $D(P_i^*)$ in foreign.
-$P_i$ denotes the home price of good $i \in \left\{N,X,Y\right\}$ and $P_i^*$ denotes the foreign price of good $i$.
-
-Good $N$ is produced with labor alone with constant returns to scale technology so that $Q_N = l_N$.
-I assume the aggregate labor supply $l$ is large enough to ensure that the output of good $N$ is large enough to guarantee balanced trade.
-The supply functions for good $X$ are $Q_X(P_X)$ and $Q_X^*(P_X^*)$ and are assumed strictly increasing and twice continuously differentiable for all prices that elicit positive supply.
+$P_i$ denotes the home price of good $i \in \left\{N,X,Y\right\}$ and $P_i^*$ denotes the foreign price of good $i$.}
+
+\textbf{Good $N$ is produced with labor alone with constant returns to scale technology so that $Q_N = l_N$.
+I assume the aggregate labor supply $l$ is large enough to ensure that the output of good $N$ is large enough to guarantee balanced trade.} The supply functions for good $X$ are $Q_X(P_X)$ and $Q_X^*(P_X^*)$ and are assumed strictly increasing and twice continuously differentiable for all prices that elicit positive supply.
 I also assume $Q_X^*(P_X) > Q_X(P_X)$ for any such $P_X$ so that the home country is a net importer of good $X$.
 The production structure for good $Y$ is taken to be symmetric, with both demand and supply such that the economy is separable in goods $X$ and $Y$.
-It is assumed that the production of goods $X$ and $Y$ requires, in addition to labor, the possession of a sector-specific factor that is available in inelastic supply and is non-tradable so that the income of owners of the specific factors is tied to the price of the good in whose production their factor is used.
+It is assumed that the production of goods $X$ and $Y$ requires, \textbf{in addition to labor,} the possession of a sector-specific factor that is available in inelastic supply and is non-tradable so that the income of owners of the specific factors is tied to the price of the good in whose production their factor is used.
 
@@
 where $M_X$ are home-county imports and $E_X^*$ are foreign exports of good $X$ and $E_Y$ are home-county exports and $M_Y^*$ are foreign imports of good $Y$.
-Given the production technology and no trade taxes, the price of the numeraire is equal to one in both countries and on the world market.
+\textbf{Given the production technology and no trade taxes, the price of the numeraire is equal to one in both countries and on the world market.}
 
@@
 As the economy is fully separable and the economic and political structures are symmetric, I focus here on the home country and the $X$-sector.
-The details are analogous for $Y$ and foreign.\footnote{These are the relevant sectors for analysis because they are the sectors in which there are lobbies.}
+The details are analogous for $Y$ and foreign.\footnote{\textbf{These are the relevant sectors for analysis because they are the sectors in which there are lobbies.}}
 
@@
 \end{equation}
-where $\mathit{CS}$ is consumer surplus, $\mathit{PS}$ is producer surplus, $\ga(e,\ta)$ is the weight placed on producer surplus (profits) in the import-competing industry, $e$ is lobbying effort, $\ta$ is a random variable, and $\mathit{TR}$ is tariff revenue.\footnote{Labor income $l$ could also be included in both median legislator and executive welfare.
-This alters none of the results and so would only serve to complicate the exposition.}
+where $\mathit{CS}$ is consumer surplus, $\mathit{PS}$ is producer surplus, $\ga(e,\ta)$ is the weight placed on producer surplus (profits) in the import-competing industry, $e$ is lobbying effort, $\ta$ is a random variable, and $\mathit{TR}$ is tariff revenue.\footnote{\textbf{Labor income $l$ could also be included in both median legislator and executive welfare.
+This alters none of the results and so would only serve to complicate the exposition.}}
 
@@
 \begin{equation}
-  U_L = \Pr\left[ \text{TradeWar}(\bta,e_b,e_n) \right] \left[ \pi(\tau^{\mathit{n}}) - e_n \right] + \Pr\left[ \text{TradeAgreement}(\bta,e_b,e_n) \right] \pi(\tau^a) - e_b
+  {\color{blue} \expect \left[U_L \right]} = \Pr\left[ \text{TradeWar}(\bta,e_b,e_n) \right] \left[ \pi(\tau^{\mathit{n}}) - e_n \right] + \Pr\left[ \text{TradeAgreement}(\bta,e_b,e_n) \right] \pi(\tau^a) - e_b
   \label{eq:lv}
@@
 \begin{multline}
-  W_E(\bta) + W_E^*(\bta) = \Pr\left[ \text{TradeWar}(\bta) \right] \left[W_E(\bm{\tau}^{\mathit{n}}) + W_E^*(\bm{\tau}^{\mathit{n}}) \right] + \\ \Pr\left[ \text{TradeAgreement}(\bta) \right] \left[W_E(\bta) + W_E^*(\bta) \right]
+  {\color{blue}\expect \left[W_E(\bta) \right] + \expect \left[ W_E^{*}(\bta) \right]} = \Pr\left[ \text{TradeWar}(\bta) \right] \left[W_E(\bm{\tau}^{\mathit{n}}) + W_E^*(\bm{\tau}^{\mathit{n}}) \right] + \\ \Pr\left[ \text{TradeAgreement}(\bta) \right] \left[W_E(\bta) + W_E^*(\bta) \right]
   \label{eq:jv}
@@
 That is, even for the least favorable outcome of the lobbying process, the legislature will be at least slightly more protectionist than the executive.
-This assumption is not essential, but it matches well the empirical findings that politicians with larger constituencies are less sensitive to special interests, and it simplifies the analysis considerably, in particular ensuring that trade agreement tariffs are less than trade war tariffs (See Destler (2005) and footnote~\ref{fn:ga_l_e} above).
+This assumption is not essential, but it matches well the empirical findings that politicians with larger constituencies are less sensitive to special interests, and it simplifies the analysis considerably, in particular ensuring that trade agreement tariffs are less than trade war tariffs \textbf{(See Destler (2005) and footnote~\ref{fn:ga_l_e} above)}.
 
@@
 (Note this is Equation~\ref{eq:lv} simplified by the resolution of uncertainty over the legislature's decision on the trade agreement).
-This implies a first order condition for the lobby of
+This implies a first order condition \textbf{for the lobby} of
 \begin{equation}
-	\frac{\partial \pi(\tau^n)}{\partial \tau^n}\frac{\partial \tau^n}{\partial \ga} \frac{\partial \ga}{\partial e_n} = 1
+	\bm{\frac{\partial \pi(\tau^n)}{\partial \tau^n}\frac{\partial \tau^n}{\partial \ga} \frac{\partial \ga}{\partial e_n} = 1}
   \label{eq:lobtw}
@@
 
-In accordance with intuition, because profits are increasing in the tariff, trade war tariffs are increasing in the weight attached to the profits of the import-competing industry.
+\textbf{In accordance with intuition, because profits are increasing in the tariff, trade war tariffs are increasing in the weight attached to the profits of the import-competing industry.
 This can be seen by rearranging Equation~\ref{eq:lobtw} as
@@
 By the implicit function theorem, $\frac{\partial \tau^n}{\partial \ga} = -\frac{\frac{\partial \text{Equation}~\ref{eq:legfoc}}{\partial \ga}}{\frac{\partial \text{Equation}~\ref{eq:legfoc}}{\partial \tau^n}} = -\frac{\frac{\partial \pi(\tau^n)}{\partial \tau^n}}{SOC}$ where ``SOC'' stands for `Second Order Condition', the second derivative of the legislature's objective function which should be everywhere negative.
-Setting this expression equal to Equation~\ref{eq:lem1}, the second order condition must be satisfied since $\frac{\partial \ga}{\partial e_n}$ is positive by Assumption~\ref{as:ga_c}.
+Setting this expression equal to Equation~\ref{eq:lem1}, the second order condition must be satisfied since $\frac{\partial \ga}{\partial e_n}$ is positive by Assumption~\ref{as:ga_c}.}
 
@@
 I also assume that each legislature will have the opportunity to take a vote to break the agreement with probability less than one-half, and that the probability that both legislatures have the opportunity is vanishingly small.
-This allows me to focus here on the key interaction between the domestic actors and abstract from the strategic interaction between the lobbies in the two countries.\footnote{An incentive for each lobby to free ride on the effort of the other makes the optimization problem of the executives more complex as they are faced with the problem of how best to exploit the free-riding dynamic.}
+This allows me to focus here on the key interaction between the domestic actors and abstract from the strategic interaction between the lobbies in the two countries.\footnote{\textbf{An incentive for each lobby to free ride on the effort of the other makes the optimization problem of the executives more complex as they are faced with the problem of how best to exploit the free-riding dynamic.}}
 
@@
 
-\noindent All proofs are in the Appendix.
+\noindent \textbf{All proofs are in the Appendix.}
 

### 2016-01-28 draft.tex
add generic tikzpicture to start extensive form

@@
 \usepackage{subcaption}
+
+\usepackage{tikz}
+\tikzset{
+  treenode/.style = {shape=rectangle, rounded corners,
+                     draw, align=center,
+                     top color=white, bottom color=blue!20},
+  root/.style     = {treenode, font=\Large, bottom color=red!30},
+  env/.style      = {treenode, font=\ttfamily\normalsize},
+  dummy/.style    = {circle,draw}
+}
 
@@
 This economic fact, combined with the assumptions on specific factor ownership, is what motivates political activity.
+
+\begin{figure}
+\begin{center}
+\begin{tikzpicture}
+  [
+    grow                    = right,
+    sibling distance        = 6em,
+    level distance          = 10em,
+    edge from parent/.style = {draw, -latex},
+    every node/.style       = {font=\footnotesize},
+    sloped
+  ]
+  \node [root] {Formula}
+    child { node [env] {equation}
+      edge from parent node [below] {single-line?} }
+    child { node [dummy] {}
+      child { node [dummy] {}
+        child { node [env] {align\\flalign}
+          edge from parent node [below] {at relation sign?} }
+        child { node [env] {alignat}
+          edge from parent node [above] {at several}
+                           node [below] {places?} }
+        child { node [env] {gather}
+                edge from parent node [above] {centered?} }
+        edge from parent node [below] {aligned?} }
+      child { node [env] {multline}
+              edge from parent node [above, align=center]
+                {first left,\\centered,}
+              node [below] {last right}}
+              edge from parent node [above] {multi-line?} };
+\end{tikzpicture}
+
+\end{center}
+\caption{Extensive Form\label{fig:ext}}
+\end{figure}
 

### 2016-01-29 draft.tex
add game-tree.tex and start of code for extensive form into draft.tex

--- 2016-01-27 draft.tex
@@
 \usepackage{subcaption}
+
+\usepackage{tikz}
+\tikzset{
+% Two node styles for game trees: solid and hollow
+solid node/.style={circle,draw,inner sep=1.5,fill=black},
+hollow node/.style={circle,draw,inner sep=1.5,fill=white}
+%non node/.style={circle,draw,inner sep=.1,fill=black}
+}
 
@@
 This economic fact, combined with the assumptions on specific factor ownership, is what motivates political activity.
+
+\begin{figure}
+\begin{center}
+\begin{tikzpicture}[scale=1.5,font=\footnotesize, grow=right]
+% Specify spacing for each level of the tree
+\tikzstyle{level 1}=[level distance=10mm,sibling distance=10mm]
+\tikzstyle{level 2}=[level distance=10mm,sibling distance=10mm]
+% The Tree
+\node(0)[solid node,label=above:{$E,E^*$}]{} 
+child{node(1)[solid node,label=below:{$0$}]{}
+}
+child{[white] node(2)[solid node,xshift=12]{}  %note that you need to adjust the yshift if you change the sibling distance
+child{[black] node(4)[solid node,label=below:{$0$}]{}
+}
+child{[white] node(7)[hollow node,xshift=12]{}  %note that you need to adjust the yshift if you change the sibling distance
+child{[black] node(9)[solid node,label=below:{$0$}]{} }
+child{[black] node(11)[solid node,label=above:{$\infty$}]{} }
+ edge from parent node[black, xshift=23]{$e_b$} %note that you need to adjust the yshift if you change the level distance
+}
+child{[black] node(6)[solid node,label=above:{$\infty$}]{}}
+ edge from parent node[black, xshift=23]{$\bta$} %note that you need to adjust the yshift if you change the level distance
+}
+child{node(3)[solid node,label=above:{$\infty$}]{}
+}
+;
+% information set
+    \draw[solid,bend right](1)to(3);
+		\draw[solid,bend right](4)to(6);
+		\draw[solid,bend right](9)to(11);
+
+\end{tikzpicture}
+
+\end{center}
+\caption{Extensive Form\label{fig:ext}}
+\end{figure}
 

### 2016-01-29 game-tree.tex
add game-tree.tex and start of code for extensive form into draft.tex

--- 2016-01-27 draft.tex
@@
 \usepackage{subcaption}
+
+\usepackage{tikz}
 
@@
 \begin{document}
-\title{\vskip-0.6in Trade Agreements in the Shadow of Lobbying\protect\footnote{Formerly circulated under the title ``Trade Agreements, Lobbying and Separation of Powers''}}
-\thanksmarkseries{alph}
-\author{Kristy Buzard\thanks{Syracuse University, Economics Department, 110 Eggers Hall, Syracuse, NY 13244, USA.
-Ph: 315-443-4079.
-Fax: 315-443-3717.
-Email: kbuzard@syr.edu.
-http://faculty.maxwell.syr.edu/kbuzard.}}
-\date{\vskip-.1in \today}
-\maketitle
 
-\begin{center} {\bf Abstract} \end{center}
-
-\begin{quote}
-{This paper presents a model of international trade agreements in which the executive branches of each government negotiate agreements while the legislative branches, subject to political pressure from firms, can disrupt them.
-Lobbying is in the style of Grossman and Helpman (1994) with a new feature: all actors face uncertainty arising from the complexity of the legislative process.
-I demonstrate that the lower the executives set tariffs in a trade agreement, the more effort lobbies put forth to prevent its ratification.
-Thus trade agreements act as a domestic political commitment device: executives set relatively high tariffs to discourage lobbying and increase the chance that the agreement will be ratified.
-This reconciles the result from tests of Grossman and Helpman's model that protection levels are high relative to contributions given estimates of governments' social-welfare weights.
-
-\textit{JEL classification:} D72, D80, F13, F53, F55 \\
-\textit{Keywords:} trade policy, lobbying, political economy, protection, international trade agreements, uncertainty}
-\end{quote}
-
-\bigskip
-\section{Introduction}
-
-%http://www.whitehouse.gov/sites/default/files/fact_sheet_increasing_us_auto_exports_us_korea_free_trade_agreement.pdf
-
-Empirical investigations of Grossman and Helpman's (1994) `Protection for Sale' model (henceforth `PFS') broadly support the predictions of the theory qualitatively, yet their quantitative estimates present several interesting puzzles.\footnote{cfr.
-Goldberg and Maggi (1999); Gawande and Bandyopadhyay (2000); Mitra, Thomakos and Ulubasoglu (2002); McCalman (2004).\label{fn:gh_lobby}} These studies have consistently found the weight governments place on social welfare to be many times that which they place on lobbying effort, while estimates indicate that the deadweight losses caused by trade distortions are several orders of magnitude larger than lobbying expenditures.\footnote{Feenstra (1992) assembles estimates from the mid-1980s that place a floor of $\$$8 billion per year on the deadweight losses from protection, whereas Bombardini and Trebbi (2012) find that total annual lobbying expenditure on trade issues in 1999-2001 when this data first became available was about $\$$200 million.} This raises the question: how could governments value social welfare so highly, yet grant these quantities of protection at such a low price?
-
-Attempts to more fully model the lobbying process have successfully reduced the parameter estimates for the government's welfare considerations somewhat, but the smallest estimates still indicate that the U.S.
-government, for instance, values social welfare about twenty times as much as contributions.\footnote{cfr.
-Gawande, Krishna and Robbins (2006); Mitra, Thomakos and Ulubasoglu (2006); Bombardini (2008); and Gawande, Krishna and Olarreaga (2012) among others.} Taking a different approach, Gawande and Hoekman (2006) suggest that one can reconcile the empirical results with the model by acknowledging the complexity of the policy-making process and the uncertainty that arises from it.
-
-Since so much of trade policy is determined in the context of international agreements that must be ratified by domestic legislative bodies, I incorporate PFS-style lobbying into a model of trade agreements in which the executive branches of the governments set tariff levels in anticipation of political pressure upon the legislatures.
-This political pressure may be interpreted either as lobbying against ratification of the trade agreement or lobbying to break the agreement once in force.
-I find that the separation-of-powers government structure can explain the empirical puzzle surrounding governments' welfare weights and that the addition of political uncertainty adds important realism so that the model presented here can speak to important trade policy-making events such as the four-year delay in the ratification of the U.S.-Korea Free Trade Agreement.
-
-Consider first the case of political certainty, and assume that the executive is more pro-social than the legislature, as has been the case throughout the post-war period in the United States.\footnote{The standard assumption that the executive is less protectionist than the legislature is a special case of the finding that susceptibility to special interests generally declines with constituency size.
-One simple illustration from the realm of trade policy is the following: a legislator whose district has a large concentration of a particular industry does not take into account the impact of tariffs on the welfare of consumers in other districts, while the executive, whose constituency encompasses the whole country, will internalize these diffuse consumption effects.
-See Lohmann and O'Halloran (1994) for a detailed argument and Grossman and Helpman (2005) for a formal model that demonstrates the point for the legislature.\label{fn:ga_l_e}} For any trade-agreement tariff set by the executives, the lobby can calculate the effort level required to prevent the agreement from being ratified: it must shift the preferences of the median legislator just far enough so that he will choose the non-cooperative tariffs over the trade agreement tariffs.
-Because the lobby will pay this price so long as its benefit outweighs the cost, a trade war will occur unless the executive sets the trade-agreement tariff so that lobbying is not worthwhile.\footnote{As will be made clear below, non-ratification is equivalent to breaking a trade agreement once in place; likewise, the non-cooperative tariffs that result from failing to ratify the trade agreement are the same as the tariffs that would arise in trade war, so this model applies to both situations.}
-
-Thus, any trade agreement under political certainty involves a tariff high enough to disengage the lobby completely; that is, even welfare-maximizing executives set high tariffs and induce zero contributions in equilibrium because they set policy in the shadow of a ratification decision influenced by lobbying.
-Careful modeling of the political process demonstrates that, while tariff-setting behavior is closely linked to deadweight-loss calculations, the empirical estimates of the responsiveness of protective measures to import penetration and the elasticity of import demand do not necessarily map directly onto the PFS parameter for welfare-mindedness.
-
-In the non-unitary government model, these estimates and the earlier-cited stylized facts become consistent with each other.
-The ``case of the missing contributions,'' as Gawande and Hoekman (2006) refer to it, turns out to be well explained by an equilibrium phenomenon arising from the agenda-setting power of the executive branch.
-In particular, this model can explain the approximately 15$\%$ of sectors that receive protection in spite of apparently putting forth no lobbying effort.
-
-The fact that \textit{zero} contributions are predicted in equilibrium is not particularly satisfying.
-The addition of political uncertainty provides the missing realism, smooths the results and delivers additional intuition.
-I assume all actors are uncertain about the weight the median legislator in a non-unitary legislature places on the profits of the lobbying industry and that this is attributable to the complexity of the legislative process; say, for instance, because a small number of lobbied legislators cannot deliver the votes of their non-lobbied colleagues with certainty.
-In this case, I show that the legislatures disrupt trade agreements with a higher probability and set higher trade war tariffs when lobbying effort increases.
-Because the lobbies respond to higher trade agreement tariffs by decreasing effort (and therefore the probability the agreement will fail to be ratified or will be broken), the executives must trade off the level of social welfare derived while a trade agreement is in force---since they prefer low tariffs---and the chances that the agreement will remain in force.
-
-The executives therefore set trade-agreement tariffs with an eye to discouraging lobbying activity and the accompanying probability of abrogation.
-It can do this because the trade agreement increases lobbying costs: the legislature now has to be compensated for the loss of the reduced foreign tariff if it breaks the trade agreement.
-Moreover, in addition to the standard terms-of-trade internalization, the executive's agenda-setting power allows it to reduce tariffs until lobbying activity is at the optimal level relative to a non-cooperative setting, reducing the amount of political pressure faced by the legislatures and the resultant expected tariffs.
-
-One contribution of this model, therefore, is the idea that trade agreements can act as a kind of domestic political commitment device.
-Rather than a time inconsistency such as that in Maggi and Rodr\'{i}guez-Clare (2007) where the welfare of the unitary policymaker is improved via commitment to a trade agreement by changing the investment decisions of firms, here we have an inconsistency between the preferences of the executive and legislative branches.\footnote{Indeed, depending on the modeling choices and parameters, the utility of the legislative branch taken as a whole may be lower under the trade agreement scenario.} Although the legislature has the final word both with and without the trade agreement, the executives are able to use commitment to the trade agreement to reduce lobbying effort, thus muting the legislature's protectionist bias and bringing the outcome more in line with their relatively liberal preferences.
-
-To the best of my knowledge, this is the first attempt to model lobbying specifically aimed at derailing trade agreements.
-Although this kind of politically-driven failure is not rare, prior models have not allowed exploration of the endogenous dynamics behind them.
-Careful consideration of the political process seems to be important for understanding why governments so often fail to cooperate.
-The recent ratification of the U.S.-Korea Free Trade Agreement is a case in point.
-The agreement, as originally negotiated in 2007, was not able to be ratified in the United States until President Obama gained concessions for U.S.
-automakers, after which they removed their objections and the agreement was ratified by Congress in 2011 and entered into force in 2012.
-In a certain world, this ratification failure should not have occurred, and it was remedied by making the trade agreement more appealing to a key lobby and thus reducing its lobbying effort.
-\footnote{See \url{http://www.ustr.gov/trade-agreements/free-trade-agreements/korus-fta}.
-In this case, automakers are both importers and exporters, and the concessions were concentrated on market access for auto exports in South Korea.
-For simplicity, the model presented herein focuses on lobbying by an import-competing industry in a model of interindustry trade; the central insights do not depend on this structure and can be extended to the case of multiple lobbies and/or intraindustry trade.
-In particular, adding an export lobby in the simplest case of no uncertainty predicts that import lobbies will remain disengaged while the executive will be able to reduce tariffs by recruiting support from exporters.
-The Panama Trade Promotion Act seems to follow this pattern, where there was virtually no lobbying against, significant lobbying in favor, and numerous industries receiving import tariff protection.\label{fn:lobby}}
-
-Moreover, I show that trade-agreement tariffs, lobbying effort and the probability of trade disruptions vary in nuanced ways with the amount of political uncertainty present: they are influenced strongly by lobbying incentives, but not in the straightforward ways predicted by models with unitary governments.
-Saiegh (2009) argues that political uncertainty is in part related to governmental and political structure.
-Although they do not discuss uncertainty specifically, Gawande, Krishna and Olarreaga (2009) provide cross-country empirical evidence that measures such as the number of checks and balances on the power of the legislature and the gap between the policy positions of the main political parties have predictive power for the PFS ``welfare-mindedness'' parameter.
-Thus, although empirical tests are outside the scope of this paper, there is suggestive evidence that political uncertainty is indeed an important factor empirically.
-
-Ultimately, it is the addition of a rich structure of government, tailored carefully to the trade policy making process, that resolves the main empirical puzzle surrounding the PFS model: in this framework, even a welfare-maximizing executive and lobbying effort at zero can be consistent with high tariff levels.
-
-\subsection{Related Literature}
-The foundations of this work rest on Grossman and Helpman's (1994) `Protection for Sale.' Two related papers `Trade Wars and Trade Talks' (Grossman and Helpman (1995b)) and `The Politics of Free-Trade Agreements' (Grossman and Helpman (1995a)) first employed the PFS approach in the study of trade agreements.
-The first considers ``Trade Wars'' and ``Trade Talks'' separately, whereas my approach allows the ``Trade Talks'' to be influenced by the desire to avoid a ``Trade War'' and thus views the ``Trade War'' as a crucial subgame.
-Their model of preferential trade agreements (Grossman and Helpman (1995a)), although treating the government as a unitary actor, is closer to this approach.
-
-The extensive literature that explores the empirical implications of the PFS model is well-summarized by Gawande and Magee (2012).
-After the first generation of work consistently found very high estimates for the government's welfare-mindedness (see footnote~\ref{fn:gh_lobby} above), a second generation has explored several modifications of the model and estimating techniques.
-Most of these have focused on an improved accounting of the lobbying process, such as including lobbying by up- and downstream firms (Gawande, Krishna and Olarreaga (2012)), better identifying which firms in an industry lobby (Bombardini (2008)), taking into account lobbying by foreign interests (Gawande, Krishna and Robbins (2006)) or changing the way in which organized sectors are determined (Mitra, Thomakos and Ulubasoglu (2006)).
-Indeed, Imai, Katayama and Krishna (2009) argue that many of the classification schemes present serious challenges for the validity of results.
-
-Another related literature considers the impact of exogenously-determined political uncertainty on the potential for trade cooperation.
-These studies (e.g.
-Feenstra and Lewis (1991); Milner and Rosendorff (2001); Bagwell and Staiger (2005); Beshkar (2010)) derive various implications for the design of international trade agreements using a Baldwin (1987)-style government welfare function with exogenous shocks to the political-pressure parameter.
-As explained in the next section, the model presented here is intended to be a bridge between this literature and that discussed above that focuses on endogenously-determined political economy weights.
-To preview, this model is closest to that of Bagwell and Staiger (2005), with two main changes: in place of their unitary government signing a trade agreement and having different preferences ex-ante and ex-post, this model has two branches of government with differing preferences, where---in the case of the legislature---the political economy weights are determined endogenously.
-
-Several papers have studied the impact of executive/legislative interactions on international agreements in the case of exogenously-determined preferences.
-Mansfield, Milner and Rosendorff (2000) construct a model in which democracies trade more with each other because the domestic legislature's influence requires the foreign government to make deeper tariff cuts to ensure its cooperation.
-Dai (2006) typifies a second approach, which, in line with the Schelling conjecture, argues that a country's legislative constraint assists its own executive in maintaining higher barriers.
-Ethier (2002) treats the effect of a different separation of powers on the international trade system: that between trade negotiators and the government officials who administer protection.
-
-The above work is patterned on the model of Milner and Rosendorff (1997), which explores how uncertainty can affect trade policy and the probability of ratification failure when political preferences, and therefore uncertainty, are exogenous;  Iida (1996) presents a similar model.
-Related is Le Breton and Salanie (2003), which studies lobbying when the lobby is uncertain about the preferences of a unitary decision maker.
-Le Breton and Zaporozhets (2007) go a step further and replace the unitary decision maker with a legislature with multiple actors.
-Song (2008) presents a model in which policy-making power is shared between an executive and a unitary legislature where lobbies are able to endogenously influence the political preferences of the legislature in a context of unilateral policy making with no uncertainty, and Coates and Ludema (2001) study trade policy leadership in a model with endogenous lobbying in the presence of political uncertainty with imperfect monitoring.
-
-I will begin by describing the model in detail.
-Section~\ref{sec:main} then presents the main results.
-In Section~ \ref{sec:example}, an extended example demonstrates these results as well as the impact of changing the institutional environment.
-Section~\ref{sec:dis} explores the connection between the model under consideration here and the `Protection for Sale' theoretical framework and includes in-depth discussions of the roles of uncertainty and the separation-of-powers structure.
-Section~\ref{sec:concl} concludes.
-
-\section{The Model}
-\label{sec:model}
-\textbf{I begin by describing the basic economic setting within which trade occurs.
-It is a three-good model with two countries: home (no asterisk) and foreign (asterisk).
-In each country, preferences are linear in good $N$, which is denoted the numeraire, while the demand functions for $X$ and $Y$ are assumed strictly decreasing and twice continuously differentiable.
-Further, the demand functions for $X$ and $Y$ are taken to be identical and denoted $D(P_i)$ in home and $D(P_i^*)$ in foreign.
-$P_i$ denotes the home price of good $i \in \left\{N,X,Y\right\}$ and $P_i^*$ denotes the foreign price of good $i$.}
-
-\textbf{Good $N$ is produced with labor alone with constant returns to scale technology so that $Q_N = l_N$.
-I assume the aggregate labor supply $l$ is large enough to ensure that the output of good $N$ is large enough to guarantee balanced trade.} The supply functions for good $X$ are $Q_X(P_X)$ and $Q_X^*(P_X^*)$ and are assumed strictly increasing and twice continuously differentiable for all prices that elicit positive supply.
-I also assume $Q_X^*(P_X) > Q_X(P_X)$ for any such $P_X$ so that the home country is a net importer of good $X$.
-The production structure for good $Y$ is taken to be symmetric, with both demand and supply such that the economy is separable in goods $X$ and $Y$.
-It is assumed that the production of goods $X$ and $Y$ requires, \textbf{in addition to labor,} the possession of a sector-specific factor that is available in inelastic supply and is non-tradable so that the income of owners of the specific factors is tied to the price of the good in whose production their factor is used.
-
-For simplicity, I assume each government's only trade policy instrument is a specific tariff on its import-competing good: the home country levies a tariff $\tau$ on good $X$ while the foreign country applies a tariff $\tau^*$ to good $Y$.
-Local prices are then $P_X = P_X^W + \tau$, $P_X^* = P_X^W$, $P_Y = P_Y^W$ and $P_Y^* = P_Y^W + \tau^*$ where a $W$ superscript indicates world prices.
-Equilibrium prices are determined by the market clearing conditions
-$$M_X(P_X)= D(P_X)-Q_X(P_X) = Q_X^*(P_X^*) - D(P_X^*) = E_X^*(P_X^*)$$
-$$E_Y(P_Y)=Q_Y(P_Y)-D(P_Y) = D(P_Y^*)-Q_Y^*(P_Y) = M_Y^*(P_Y^*)$$
-where $M_X$ are home-county imports and $E_X^*$ are foreign exports of good $X$ and $E_Y$ are home-county exports and $M_Y^*$ are foreign imports of good $Y$.
-\textbf{Given the production technology and no trade taxes, the price of the numeraire is equal to one in both countries and on the world market.}
-
-$P_X^W$ and $P_Y^W$ are decreasing in $\tau$ and $\tau^*$ respectively, while $P_X$ and $P_Y^*$ are increasing in the respective domestic tariff.
-This gives rise to a standard terms-of-trade externality.
-As profits and producer surplus (identical in this model) in a sector are increasing in the price of its good, profits in the import-competing sector are also increasing in the domestic tariff.
-This economic fact, combined with the assumptions on specific factor ownership, is what motivates political activity.
-
-I next describe the politically-relevant actors.
-In order to focus attention on protectionist political forces, I assume that only the import-competing industry in each country is politically-organized and able to lobby and that it is represented by a single lobbying organization (see Footnote~\ref{fn:lobby} for discussion of additional lobbies).
-Each country's government is composed of two branches: an executive who can conclude trade agreements and a legislature that has final say on trade policy.
-In summary, the political process is modeled as involving three players in each country: the lobby, the executive, and the legislature.
-
-The timing is as follows.
-First, the executives set trade policy cooperatively within the context of an international agreement.
-I assume that an external authority can ensure enforcement of the agreement in the one-shot game analyzed here.\footnote{Buzard (2014b) extends the analysis to a repeated-game context and can therefore remove this restriction.} After the trade agreement is concluded, the lobbies attempt to persuade the legislators in their respective countries to vote ``no'' on the ratification of the agreement; or alternatively, if there is no ratification hurdle, to break the trade agreement.\footnote{As will be seen shortly, these two interpretations are equivalent in this setting because tariffs in a ``trade war'' are assumed to be set non-cooperatively in the same manner as if there were no trade agreement.
-For simplicity, I refer to this as the ``break'' decision throughout.} Next, the uncertainty about the the median legislator's identity is resolved and the legislatures decide whether to abide by the agreement or to provoke a trade war.
-In the event that the trade agreement does not enter into or remain in force, there is a final stage of lobbying, resolution of the uncertainty surrounding this decision, and voting to set the trade-war tariffs.\footnote{Alternatively, one can model the lobbying for the ratification decision as determining the potential non-cooperative tariff; this changes behavior slightly depending on the relationship between the lobby and the legislature's preferences, but does not change the fundamental dynamics of the model.
-Yet another possibility is to replace the trade-war phase with the ex-ante status quo, which would most naturally be the non-cooperative equilibrium.
-This is precisely the outcome that obtains in the trade-war phase, so the results of the model are unaltered.} Once all political decisions are taken, producers and consumers make their decisions.
-
-Although uncertainty is present, information about it is symmetric, so the appropriate solution concept is subgame perfect Nash equilibrium.
-As this game is solved by backward induction, it is intuitive to start by describing the incentives of the legislatures, whose decisions I model as being taken by a median legislator.
-As the economy is fully separable and the economic and political structures are symmetric, I focus here on the home country and the $X$-sector.
-The details are analogous for $Y$ and foreign.\footnote{\textbf{These are the relevant sectors for analysis because they are the sectors in which there are lobbies.}}
-
-In modeling the objective of the legislature, I introduce a variant of the general political support function introduced by Hillman (1982).
-While Hillman is credited with first proposing that policy makers choose tariffs in order to maximize their own political support, the PFS model contributed the explicit specification that lobbying contributions enter the political support function with constant marginal utility as $C + aW$.
-
-Dixit, Grossman and Helpman (1997) point out that linearity in contributions prevents complete analysis of distributional questions; the PFS form also restricts the returns to lobbying activity to be constant.\footnote{See Section \ref{sec:gh} for further discussion of this point and an in-depth discussion of the relationship between the functional form presented herein and the PFS form.
-Note that the assumption of diminishing returns to lobbying effort has been present in the literature going back at least to Findlay and Wellisz (1982).} They show that the PFS equilibrium analysis can be extended to a more general setting where, among other features, returns to lobbying are not restricted to be constant.
-
-For the case of trade policy, I specify the welfare function of the home legislature as
-\begin{equation}
-  W_{\mathit{ML}} = \mathit{CS}_X(\tau) + \ga(e,\ve) \cdot \mathit{PS}_X(\tau) + \mathit{CS}_Y(\tau^*) + \mathit{PS}_Y(\tau^*) + \mathit{TR}(\tau)
-  \label{eq:ml}
-\end{equation}
-where $\mathit{CS}$ is consumer surplus, $\mathit{PS}$ is producer surplus, $\ga(e,\ta)$ is the weight placed on producer surplus (profits) in the import-competing industry, $e$ is lobbying effort, $\ta$ is a random variable, and $\mathit{TR}$ is tariff revenue.\footnote{\textbf{Labor income $l$ could also be included in both median legislator and executive welfare.
-This alters none of the results and so would only serve to complicate the exposition.}}
-
-Notice that, aside from the endogeneity of the weight the legislature places on the lobbying industry's profits, this is precisely the \textit{deus ex machina} government objective function popularized by Baldwin (1987) that is commonly employed in the literature on the political economy of trade agreements.
-Since trade policies are overwhelmingly determined within the context of trade agreements, it is useful to have a framework to bring together the endogenous political pressure of PFS-style modeling with the trade agreements approach; the formulation in Equation~\ref{eq:ml} is intended as a bridge between the two.
-Buzard (2014a) for an example of how this form can be used to integrate the analysis of endogenous political activity and exogenous shocks to the political welfare of governments.
-
-In this formulation, the weight the median legislator places on the profits of the import-competing industry, $\ga(e,\ve)$, is affected by the level of lobbying effort and a random variable $\ta$.
-This allows for the analysis of endogenous lobbying to be integrated with the standard political economy shocks in the trade agreements literature.
-Alternatively, as motivated below, it can represent reduced-form modeling of political uncertainty arising from a complex decision-making process within a legislature.
-Modeling the objective function so closely on the standard in the trade agreements literature allows for direct comparisons to the large extant body of work that studies exogenous shocks only, revealing cleanly the effects of the addition of endogenous lobbying---and if one wants---this different kind of political uncertainty.
-
-I allow for the possibility of this political uncertainty, which can be interpreted as uncertainty over the identity of the median legislator, in order to model the idea that the lobbying and voting ``game'' that goes on within legislatures is often complex enough that none of the actors know precisely what the outcome will be---that is, no one can predict exactly which legislator will be the median and therefore what weight will be placed on import-industry profits at the time of the vote.
-Thus one way we can think about $\ta$ is as an additional influence on the legislature's preferences about which all parties---executives, legislators and lobbies alike---are uncertain.
-It is this interpretation of $\ta$ I take throughout the paper.
-
-One can conceptualize this uncertainty as being a result of the lobby's strategic choice to engage with only some key members of the legislature who wield significant influence but who cannot deliver policy decisions without the cooperation of others within the legislature, which is in turn uncertain.
-Different assumptions on $\ga(\cdot,\cdot)$ correspond to different institutional features and will affect optimal lobbying- and tariff-setting behavior.
-While a structural model of the legislative process might be more satisfying, the reduced form can deliver the key insights without distracting from the central policy choice and lobbying dynamics that take place at the level of the executives and lobbies, to which we turn next.
-
-Given its expectations and the legislature's preferences, the home lobby chooses its lobbying effort ($e_b$ to influence the break/ratification decision, which for simplicity, I will refer to henceforth simply as ``break,'' and $e_n$ to influence the trade war tariff) to maximize the welfare function:
-\begin{equation}
-  {\color{blue} \expect \left[U_L \right]} = \Pr\left[ \text{TradeWar}(\bta,e_b,e_n) \right] \left[ \pi(\tau^{\mathit{n}}) - e_n \right] + \Pr\left[ \text{TradeAgreement}(\bta,e_b,e_n) \right] \pi(\tau^a) - e_b
-  \label{eq:lv}
-\end{equation}
-where $\pi(\cdot)$ is the current-period profit and $\tau^a \ (\tau^\mathit{n})$ is the home country's tariff on the import good under a trade agreement (war).
-I use the convention throughout of representing a vector of tariffs for both countries $(\tau,\tau^*)$ as a single bold $\bm{\tau}$.
-
-I assume the lobby's contribution is not observable to the foreign legislature.
-The implication is that the lobby can directly influence only the home legislature, and so the influence of one country's lobby on the other country's legislature occurs only through the tariffs selected.\footnote{cfr.
-Grossman and Helpman (1995b), page 685.}
-
-In the first stage, the executives choose the trade agreement tariffs $\bta=\left(\tau^a,\tau^{*a} \right)$ via a negotiating process that I assume to be efficient.
-In this symmetric environment, it therefore maximizes the expected joint payoffs:
-\begin{multline}
-  {\color{blue}\expect \left[W_E(\bta) \right] + \expect \left[ W_E^{*}(\bta) \right]} = \Pr\left[ \text{TradeWar}(\bta) \right] \left[W_E(\bm{\tau}^{\mathit{n}}) + W_E^*(\bm{\tau}^{\mathit{n}}) \right] + \\ \Pr\left[ \text{TradeAgreement}(\bta) \right] \left[W_E(\bta) + W_E^*(\bta) \right]
-  \label{eq:jv}
-\end{multline}
-
-I model the executives' choice via the Nash bargaining solution where the disagreement point is the executives' welfare resulting from the Nash equilibrium in the non-cooperative game (i.e.
-in the absence of a trade agreement) between the legislatures.
-
-The home executive's welfare is specified as follows:
-\[
-  W_E = \mathit{CS}_X(\tau) + \ga_E \cdot \mathit{PS}_X(\tau) + \mathit{CS}_Y(\tau^*) + \mathit{PS}_Y(\tau^*) + \mathit{TR}(\tau)
-\]
-Note that this is identical to the welfare function for the legislature aside from the weight $\ga_E$ the executive places on the profits of the import industry, which is not a function of lobbying effort.
-This construction permits me to integrate the influence of the import-competing industry on negotiations prior to the formation of the trade agreement while also reflecting the idea that the mechanisms through which political considerations influence the executives' preferences appear to be quite different from those at the legislative level (the electoral college and primary/caucus systems for instance).
-
-This does \textit{not} require that the executives are not lobbied; only that their preferences are not directly altered in a significant way by lobbying over trade---that they do not sell protection in order to finance their re-election campaigns.
-In the case of the post-war United States, where the Congress has consistently been significantly more protectionist than the President, this seems to reasonably reflect the political reality.
-For trade policy, where there are concentrated benefits but harm is diffuse, there are good reasons for this to be the case.
-Because the President has the largest constituency possible, delegating authority to the executive branch may simply be a mechanism for ``concentrating'' the benefits since consumers seem unable to overcome the free-riding problem.
-In fact, a strong argument can be made that power over trade policy has been delegated to the executive branch precisely \textit{because} it is less susceptible to the influence of special interests (Destler (2005)).
-
-Therefore, in line with both the theoretical and empirical literature, I make the following assumption:
-
-\begin{assumption}
-  $\ga(e,\ve) \geq \ga_E \geq 1 \ \forall \ve$.
-  \label{as:ga_l_e}
-\end{assumption}
-
-That is, even for the least favorable outcome of the lobbying process, the legislature will be at least slightly more protectionist than the executive.
-This assumption is not essential, but it matches well the empirical findings that politicians with larger constituencies are less sensitive to special interests, and it simplifies the analysis considerably, in particular ensuring that trade agreement tariffs are less than trade war tariffs \textbf{(See Destler (2005) and footnote~\ref{fn:ga_l_e} above)}.
-
-Although the political process here most closely matches that of the United States in the post-war era,\footnote{In particular, the binary decision by the legislature about whether to abide by or break the trade agreement is modeled on the ``Fast Track Authority'' that the U.S.
-Congress granted to the Executive branch almost continuously from 1974-1994 and then again as ``Trade Promotion Authority'' from 2002-2007.} the model is broadly applicable because the central dynamic that emerges---that trade agreements are formed in the \textit{shadow} of lobbying pressure instead of in direct response to it---appears to be present in a wide range of countries in which the formation and maintenance of trade policy requires ratification by a politically susceptible body either at home or in a trading partner.
-
-%I believe the one major fudge between my model and Destler's evidence is acceptable: that the executive is actually the one who breaks trade agreements, but they do it largely to prevent the legislature from doing something far worse.
-
-\section{Main Results}
-\label{sec:main}
-To understand how the executives optimally structure trade agreements, we must first examine the incentives of the lobbies and how the legislatures make decisions regarding breach of the trade agreement, including how trade-war tariffs are set.
-The symmetric structure of the model permits restriction of attention to the home country.
-
-\subsection{Trade-War Tariffs}
-\label{sec:twt}
-In the event that the trade agreement is broken, the legislature sets its tariff $\tau$ unilaterally by maximizing Equation~\ref{eq:ml} given $\tau^*$.
-Because there are no interactions between the home and foreign tariffs, the home and foreign trade-war tariffs are independent and the home country's tariff in a trade war maximizes weighted home-country welfare in the $X$-sector only.
-The foreign legislature's decision problem is analogous, and unilateral optimization leads to what I refer to as the (expected) Nash or trade-war tariffs $\tau^n$ as the solution to the following first order condition:\footnote{Note that the random variable in the median legislator's weight on import profits is written as $\ta_n$ at this stage to distinguish it from $\ta_b$ at the break stage: as two separate votes are taken, there are two distinct realizations of uncertainty.
-It is not necessary to restrict them to come from the same distribution, although in applications it would seem unlikely that the distribution would change.}
-\begin{equation}
-		\frac{\partial \mathit{CS}_X(\tau)}{\partial \tau^n} + \ga(e_n,\ve_n) \cdot \frac{\partial \mathit{PS}_X(\tau)}{\partial \tau^n} +  \frac{\partial \mathit{TR}(\tau)}{\partial \tau^n} = 0 .
-		\label{eq:legfoc}
-\end{equation}
-
-In the event of a trade war, the lobby chooses its effort $e_n$ given this tariff-setting behavior by maximizing its profits net of effort: $\pi\left(\tau^n\left(\ga\left(e_n,\ve_n\right)\right)\right) - e_n$.
-(Note this is Equation~\ref{eq:lv} simplified by the resolution of uncertainty over the legislature's decision on the trade agreement).
-This implies a first order condition \textbf{for the lobby} of
-\begin{equation}
-	\bm{\frac{\partial \pi(\tau^n)}{\partial \tau^n}\frac{\partial \tau^n}{\partial \ga} \frac{\partial \ga}{\partial e_n} = 1}
-  \label{eq:lobtw}
-\end{equation}
-That is, at this stage, the lobby chooses the level of effort that equates its expected marginal increase in profits with its marginal payment.
-Further details about the political weighting function $\ga$ and assumed structure of uncertainty are provided in the next section.
-
-\textbf{In accordance with intuition, because profits are increasing in the tariff, trade war tariffs are increasing in the weight attached to the profits of the import-competing industry.
-This can be seen by rearranging Equation~\ref{eq:lobtw} as
-\begin{equation}
-  \frac{\partial \tau^n}{\partial \ga} = \frac{1}{\frac{\partial \pi(\tau^n)}{\partial \tau^n} \frac{\partial \ga}{\partial e_n}}
-	\label{eq:lem1}
-\end{equation}
-since $\frac{\partial \ga}{\partial e_n}$ is positive by Assumption~\ref{as:ga_c} and profits are increasing in prices which are in turn increasing in the tariff.
-This expression can also demonstrate that the second order condition for the legislature's problem is always satisfied.
-By the implicit function theorem, $\frac{\partial \tau^n}{\partial \ga} = -\frac{\frac{\partial \text{Equation}~\ref{eq:legfoc}}{\partial \ga}}{\frac{\partial \text{Equation}~\ref{eq:legfoc}}{\partial \tau^n}} = -\frac{\frac{\partial \pi(\tau^n)}{\partial \tau^n}}{SOC}$ where ``SOC'' stands for `Second Order Condition', the second derivative of the legislature's objective function which should be everywhere negative.
-Setting this expression equal to Equation~\ref{eq:lem1}, the second order condition must be satisfied since $\frac{\partial \ga}{\partial e_n}$ is positive by Assumption~\ref{as:ga_c}.}
-
-\bigskip
-\subsection{To Break or Not to Break?}
-With trade-war decisions specified, we can proceed to analyze the incentives regarding the legislature's decision to uphold or break the trade agreement.
-I assume that the form of the political weighting function is the same in both cases, but the model could be extended to allow the two to differ.
-What is in principle different are the lobbying efforts toward influencing the break decision, denoted $e_b$ at this stage to distinguish them from those that are made to influence the Nash tariff, $e_n$.
-The realization of uncertainty about the median legislator's identity will similarly be denoted $\ve_b$ so as to distinguish it from $\ve_n$ at the trade-war stage.
-
-I also assume that each legislature will have the opportunity to take a vote to break the agreement with probability less than one-half, and that the probability that both legislatures have the opportunity is vanishingly small.
-This allows me to focus here on the key interaction between the domestic actors and abstract from the strategic interaction between the lobbies in the two countries.\footnote{\textbf{An incentive for each lobby to free ride on the effort of the other makes the optimization problem of the executives more complex as they are faced with the problem of how best to exploit the free-riding dynamic.}}
-
-The legislature will break the agreement and set the tariff at $\tau^n$ if the median legislator's utility from the Nash tariffs is higher than his utility from the trade agreement tariffs, i.e. if
-\begin{equation}
-  W_{ML}(\btn,\ga(e_b,\ve_b)) > W_{ML}\left(\bta,\ga(e_b,\ve_b)\right)
-  \label{eq:lwcg}
-\end{equation}
-\noindent Recall that a bold $\bm{\tau}$ represents a vector of tariffs for both countries $(\tau,\tau^*)$.
-
-The legislature's decisions are stochastic, so the outcome of the vote on whether or not to break the trade agreement, as well as the level at which the tariff will be set in the case of a disruption, is not known to \textit{any} player until the uncertainty over the identity of the median legislator is resolved at each stage---that is, until the moment the vote takes place.\footnote{We might think of the trade-war tariff as being set in the legislation before a vote takes place in the case of administered protection for instance; the results are robust to this variant.} I represent the probability that the home legislature breaks the trade agreement and sets the tariff at $\tau^n$ as:
-\begin{multline}
-  b(e_b,\bta,\btn) = \expect_{\ga|e_b} \bm{1} [ W_{ML}(\btn,\ga(e_b,\ve_b)) > W_{ML}\left(\bta,\ga(e_b,\ve_b)\right) ] \\ = \Pr [ W_{ML}(\btn,\ga(e_b,\ve_b)) > W_{ML}\left(\bta,\ga(e_b,\ve_b)\right) | e_b]
-  \label{eq:b}
-\end{multline}
-
-Note that the legislature's best response functions (both for whether to break the agreement, and at what level to set the tariff if it does) will be affected by the shape of $\ga(e,\ve)$.
-I make the following further assumptions on the weight on import-industry profits:
-
-\begin{assumption}
-  $\ga(e,\ta)$ is increasing and concave in $e$ for every $\ta \in \Theta$.
-  \label{as:ga_c}
-\end{assumption}
-
-\begin{assumption}
-  $\ga(e,\ta)$ is increasing in $\ta$.
-  \label{as:ga_ta}
-\end{assumption}
-
-Assumption~\ref{as:ga_c} formalizes the intuition that the legislature favors the import-competing industry more the higher is its lobbying effort, but that there are diminishing returns to lobbying activity.\footnote{The diminishing returns here take the form of declining increments to the lobby's influence as effort increases; in Ethier (2012), the returns to lobbying decline as the level of protection increases.} It allows the political uncertainty inherent in $\ta$ to be correlated in interesting ways with the level of $e$, ruling out only (i) that higher effort levels make lower $\ga$-weights more likely and (ii) that effort and uncertainty are correlated in such a way that increasing lobbying effort changes the structure of uncertainty so that higher $\ga$-weights become more likely at an accelerating pace.
-
-Assumption~\ref{as:ga_ta} formalizes the intuition that larger realization of the uncertainty variable $\ta$ increase the value of the political economy weight.
-
-%Assumption~\ref{as:ga_l_e} ensures that $\tau^a < \tau^n$, and more generally, that the legislature's incentives are more closely aligned with the lobby's than are those of the executive.
-This is not essential but simplifies the analysis and matches well the empirical findings that politicians with larger constituencies are less sensitive to special interests (See Destler (2005) and footnote~\ref{fn:ga_l_e} above).
-
-We are now in a position to examine the legislature's decision more closely. Of central concern is how the probability that the legislature will break the trade agreement varies with lobbying effort:
-
-\begin{result}
-  The probability that the legislature breaks a trade agreement is increasing and concave in $e_b$.
-  \label{res:bincC}
-\end{result}
-
-\noindent \textbf{All proofs are in the Appendix.}
-
-Lobbying affects only the weight the legislature places on the profits of the import-competing industry.
-These profits are higher in a trade war than under a trade agreement, so given Assumption~\ref{as:ga_c}, the expectation of $\ga$ is increasing in lobbying effort, implying that the legislature becomes more favorably inclined toward the high trade-war tariff and associated profits as lobbying increases and therefore more likely to break the trade agreement.
-However, there are decreasing returns for the lobby: increasing effort leads to less than proportional increases in the probability that the trade agreement will be broken.
-
-The probability of a trade war, just like the trade-war tariffs, is increasing in $\ga$, so it's the natural assumption that $\ga$ is increasing in lobbying effort that implies that the lobby can increase both its trade war tariff and the probability of a trade war by raising its lobbying effort.
-
-Turning to the effects of the trade-agreement tariffs on the probability that the agreement will be abrogated, it is straightforward that the legislature always prefers lower levels of the foreign tariff:
-\begin{lemma}
-  Holding lobbying effort constant, the probability the legislature breaks a trade agreement is weakly increasing in $\tau^{*a}$.
-  \label{lem:leg_astar}
-\end{lemma}
-
-Because the home country is an exporter of good $Y$, the lower world price for $Y$ has a negative effect on producer surplus that is larger than the positive effect on consumer surplus.
-Thus the net effect of an increase in foreign trade-agreement tariffs on legislative (and home country) welfare is negative, leading to an increased probability that the trade agreement will be broken.
-
-The relationship between the home country's trade-agreement tariff $\tau^a$ and the legislature's probability of breaking the agreement is slightly more complex because $\tau^a$ interacts directly with $\ga(e_b,\ta_b)$.
-The intuition is as follows.
-The median legislator will vote to break the agreement if it receives higher welfare from the trade war tariff than the trade agreement tariff.
-For any given level of effort $e_b$, when the trade agreement tariff is very low, only a small set of realizations of $\ta_b$ that lead to the lowest values of $\ga(e_b,\ta_b)$ will lead to approval of the trade agreement.
-Thus the break probability is high.
-As the trade agreement tariff rises, larger values of $\ga(e_b,\ta_b)$ are consistent with approving the trade agreement, so the set of $\ta_b$'s that lead to trade agreement approval is larger.
-That is, the break probability decreases as $\tau^a$ increases.
-Even if the trade agreement tariff is so high that it is disliked by most, or even all, potential median legislators, it is preferable to the --- even higher --- trade war tariff, so the probability of breaking either becomes zero or continues to decrease in $\tau^a$.
-
-\begin{lemma}
-  Holding lobbying effort constant, the probability the legislature breaks a trade agreement is weakly decreasing in $\tau^a$.
-  \label{res:leg_a}
-\end{lemma}
-
-In the symmetric equilibria that are focal in the analysis that follows, any increase in $\tau^a$ will be accompanied by an equal increase in $\tau^{*a}$, so it is useful to state a result about the combined impact of the home and foreign tariff when the two are constrained to be equal:
-\begin{lemma}
-	Holding lobbying effort constant, the probability the legislature breaks a trade agreement is weakly decreasing in $\bta$ (i.e.
-$\frac{\partial b(e,\bta)}{\partial  \tau^a} + \frac{\partial b(e,\bta)}{\partial  \tau^{*a}} \leq 0$).
-	\label{res:bcomb}
-\end{lemma}
-
-\noindent When any change in one of the trade agreement tariffs must be met with an equal change in the other (as in a symmetric agreement), the negative impact on the legislature's propensity to break the trade agreement of raising the home tariff always outweighs the positive impact of raising the foreign tariff because of the legislature's bias toward the import-competing industry.
-
-\subsubsection{Lobby}
-\label{sec:lob_un}
-The lobby chooses its level of effort $e_b$ as a function of $\bta$, given the implications of that choice on the legislature's probability of breaking the agreement, $b(e_b,\bta,\btn)$.
-I will henceforth suppress the Nash tariffs in the expression of the break probability to simplify notation.\footnote{Since the decisions over the Nash tariffs are taken in the final period, they do not vary from the point of view of earlier stages.
-Similarly, because the lobbying effort made in that period does not interact with other decision variables in earlier periods, I will suppress the dependence of the Nash tariffs on $e_n$.} Recalling that if the lobby has the chance to act, there is no chance for the foreign legislature to break the agreement, the lobby's decision is
-\[
-  \max_{e_b} b(e_b,\bta) \left[\pi(\tau^n) - e_n \right] + [1 - b(e_b,\bta)] \pi(\tau^a) - e_b
-\]
-The lobby simply maximizes probability-weighted profits net of effort.
-From its first order condition, we can see that the lobby balances the cost of an extra dollar of lobbying expenditure with the higher profits from a trade war weighted by the increase in the probability of receiving the higher profit level that expenditure induces:
-\begin{equation}
-	\frac{\partial b(e_b,\bta)}{\partial e_b} \left[ \pi(\tau^n) -e_n - \pi(\tau^a) \right] = 1 
-	\label{eq:lobbyfoc}
-\end{equation}
-The second order condition is satisfied given Assumption~\ref{as:ga_l_e} and Result~\ref{res:bincC}. However, the solution is not guaranteed to be interior, for which we need
-  \begin{equation}
-	  \frac{\partial b(0,\bta)}{\partial e_b} \left[ \pi(\tau^n) -e_n- \pi(\tau^a) \right] > 1.
-		\label{ine:lobint}	
-  \end{equation}
-For instance, if the executives were to set $\tau^a = \tau^n$, there would be no incentive for the lobby to make a positive contribution.
-We will see that there are some cases in which it may be in the executives' joint interest to set trade agreement tariffs so as to completely disengage the lobby.
-The following results thus only hold when Condition~\ref{ine:lobint} is satisfied, that is, when the marginal impact of the first lobbying dollar on the break probability is sufficiently high to make engaging in the political process worthwhile for the lobby.\footnote{This is primarily a condition on the curvature of the political weighting function but also involves the production and demand structure.}
-
-We can now proceed to the central result concerning the lobby's behavior. 
-\begin{result}
-  When the trade agreement remains in force with positive probability and the distribution of $\ga(e,\ta)$ is weakly increasing in $e$, lobbying effort is weakly decreasing in trade agreement tariffs.\footnote{See Section~\ref{sec:uncertainty} for a detailed account of behavior when the lobby is able to induce a trade war with probability 1.}
-  \label{res:lobby}
-\end{result}
-
-\noindent Raising the trade agreement tariff decreases the benefit of inducing a break in the trade agreement by raising profits under the trade agreement.
-This is in many ways the key result of this paper: the executives can reduce lobbying effort by setting higher tariffs in their trade agreement.
-We will see in the next section how this shapes the executives' joint decision.
-
-Note that the symmetric increase in the foreign tariff does not impact profits in the import-competing industry, but it can impact lobbying incentives through the break probability.
-However, there is an analogous effect on the break probability from raising $\tau^a$ and the net effect is negative given the assumption on the probability distribution function of $\ga(e,\ta)$, which ensures that the effect of lobbying effort on the break probability does not increase as $\bta$ increases.
-This assumption can be significantly weakened: this effect cd be positive as long as it is smaller in absolute value than the negative effect on trade agreement profits.
-
-\subsection{The Trade Agreement}
-\label{sec:ta}
-The executives choose trade agreement tariffs $\bta=\left(\tau^a,\tau^{*a} \right)$ to maximize Equation \ref{eq:jv}.
-I call the maximized joint value $MV(\bta)$ and write the division of surplus according to the Nash bargaining solution, which, due to symmetry, can be simplified from its general multiplicative form as
-\[
-  V_E(\bta) + t = W_E(\btn) + \frac{1}{2} \left( MV(\bta) - W_E(\btn) - W_E^*(\btn) \right)
-\]
-\[
-  V_E^*(\bta) - t = W_E^*(\btn) + \frac{1}{2} \left( MV(\bta) - W_E(\btn) - W_E^*(\btn) \right)
-\]
-where expectation operators have been dropped for convenience and $t$ is an intergovernmental transfer.\footnote{While direct monetary transfers have to date rarely been used in practice, it seems appropriate to interpret linked concessions on non-trade issues as indirect transfers (Klimenko, Ramey and Watson (2008); Maggi and Staiger (2011)).
-Here, transfers do not occur as long as the full set of symmetry assumptions are maintained but are an important consideration in asymmetric environments.} The solution to this system of equations determines the tariffs $\bta$ that will be set in the trade agreement.
-
-I simplify the problem by assuming that political constraints prevent the executives from choosing asymmetric tariffs.
-This restricts attention to symmetric solutions in the fully symmetric environment under consideration, which means that the trade agreement tariffs for which we're looking are simply those that maximize the joint welfare of the executives.\footnote{I have assumed that $W_E(\tau) = W_E^*(\tau) \ \forall \tau$; further in a symmetric equilibrium, $\tau^n =\tau^{*n}$ and $\tau^a =\tau^{*a}$, so
-\[
-  V_E(\bta) + t = W_E(\btn) + \frac{1}{2} \left( MV(\bta) - W_E(\btn) - W_E^*(\btn) \right) = W_E(\btn) + \frac{1}{2} \left( MV(\bta) - 2W_E(\btn) \right) = \frac{1}{2}\expect MV(\bta)
-\]}
-In order to solve for them, I represent the probability that the trade agreement will be broken as $B(\bta)=b(e(\bta),\bta)$ where $e(\bta)$ is the best response function implicit in Equation \ref{eq:lobbyfoc}, the lobby's first order condition.
-
-The problem then becomes to maximize the following modified version of Equation \ref{eq:jv}:
-\begin{equation}
-    \left\{ o \cdot B(\bta) + o^* \cdot B^*(\bta) \right\} \bm{W_E}(\btn) + \left\{ 1- o \cdot B(\bta) - o^* \cdot B^*(\bta) \right\} \bm{W_E}(\bta)  
-  \label{eq:jv2}
-\end{equation}
-where I have written the sum of the home and foreign executives' utilities as $\bm{W_E}(\cdot)$.
-Here $o$ and $o^*$ are the probabilities that the home and foreign legislatures, respectively, will have the opportunity to break the trade agreement; these are assumed to be mutually exclusive events that are realized after the conclusion of the trade agreement.
-One can think of $o$ and $o^*$ as being determined by events beyond the executives' or legislatures' control that determine whether or not the legislatures will be willing to consider the lobby's request; for example, they would be affected by the occurrence of an economic crisis that diverts the legislature's attention from less-pressing matters.
-
-Of central concern is how the break probabilities behave as a function of the trade agreement tariffs given the lobby's optimal response.
-Recalling the definition of the total break probability, $B(\bta)=b(e_b(\bta),\bta)$, we must combine th direct effect of tariffs on the legislature's break decision with the indirect effect through the impact on the lobby's choice of $e_b$.
-
-The direct effect is negative by Lemma~\ref{res:bcomb}.
-The fact that the indirect effect of raising tariffs leads to a reduction in lobbying (Result~\ref{res:lobby}) and therefore a further reduction in the propensity of the legislature to break the agreement (Result~\ref{res:bincC}) leads to the following key result: the overall probability that the trade agreement will be broken is decreasing in the trade agreement tariffs.
-
-\begin{result}
-	The total probability that the trade agreement will be broken is decreasing in $\bta$ (i.e. $\frac{\partial B(\bta)}{\partial  \tau^a} + \frac{\partial B(\bta)}{\partial  \tau^{*a}} \leq 0$).
-	\label{res:bcomB}
-\end{result}
-
-\noindent That is, when the executives raise the trade agreement tariffs the legislature becomes less likely to abrogate the agreement, for three reasons.
-First, the legislature prefers a higher domestic tariff; second, the higher tariff discourages lobbying; finally, the lower lobbying effort reduces the legislature's preferred tariff further.
-Thus, beyond promising a lower tariff from the trading partner, we can think of the trade agreement as a sort of political commitment device that can be used to reduce political pressure and therefore get the legislature to maintain a lower tariff than it otherwise would.
-
-We are now prepared to examine the executives' optimal choice of trade agreement tariffs.
-Because joint executive welfare is decreasing in trade agreement tariffs for $\bta$ above the executives' preferred tariffs, we have the following fundamental feature of the executives' problem:
-\begin{lemma}
-  The executives face the following trade off when choosing $\bta$: higher tariffs decrease the probability that the trade agreement will be broken but also decrease welfare when the agreement is in force.
-  \label{res:to}
-\end{lemma}
-I define the tariff the executives prefer in the absence of legislative constraints as $\bm{\tau^E}$ and show in the \hyperlink{int_soln}{Appendix} that it is never optimal in a symmetric equilibrium for the executives to choose $\bm{\tau^a} < \bm{\tau^E}$ if there is a positive probability that the legislatures will have an opportunity to break the trade agreement.
-This is the result of an envelope-like result: there are only second-order losses from raising tariffs slightly from the most-preferred level yet there are first-order gains in reducing the probability that the agreement will be broken.
-
-The executives will also always choose $\bta < \btn$ unless the legislature breaks even agreements with tariffs very close to the Nash level with certainty, in which case the problem is not interesting so I will ignore this case.
-Thus there is an interior solution in all cases of interest, but this solution may be of two different forms.
-First, it could be at the point that maximizes the concave portion of the executives' welfare function (Equation~\ref{eq:jv2}).
-Here, the optimal tariffs are chosen to balance concerns for trade-agreement welfare against those for discouraging lobbying activity aimed at breaking the agreement.
-
-But there is an alternative.
-For some specifications, high enough tariffs can disengage the lobby sufficiently so that the probability that the trade agreement will be broken is reduced to zero.
-Interestingly, if this occurs at a low enough tariff level, executive welfare can be maximized at this point.
-\begin{result}
-  The executives maximize their welfare by either (a) raising tariffs sufficiently high to ensure that the trade agreement will remain in force or (b) trading off reductions in the probability that the agreement will be broken with reductions in welfare under the agreement.
-  \label{res:execsoln}
-\end{result}
-%Can I say when one would happen versus the other?
-This seems to accord well with observations of trade policy politics, in that some lobbies appear to exert significant effort toward disrupting trade agreements (often in the form of some kind of individual administrative remedy such as anti-dumping duties) whereas others apparently do not contest the levels of protection granted them.
-The current model points to differences across industries in production and demand structure, as well as political weighting ($\ga$), to help explain these variations.
-
-\section{An Example}
-\label{sec:example}
-It is instructive to examine a simple parameterization of the model economy.
-The fundamentals here are chosen to match those of Bagwell and Staiger (2005).
-Home country demand, supply and profits are given by $D(P_i) = 1 - P_i$, $Q_X(P_X) = \frac{P_X}{2}$, $Q_Y(P_Y) = P_Y$, $\Pi_X(P_X) = \frac{(P_X)^2}{4}$, and $\Pi_Y(P_Y) = \frac{(P_Y)^2}{2}$ where $P_i$ is the price of good $i$ in the home country market.
-%Note that production function consistent with these supply functions are $F_X = \sqrt{l}$ and $F_Y = \sqrt{2l}$ where we assume $w=1$.
-Foreign is taken to be symmetric.
-
-This implies Home-country imports of $X$ and exports of $Y$ of $M_X(P_X)= 1 - \frac{3}{2}P_X$ and $E_Y(P_Y)= 2P_Y -1$, with foreign imports of $Y$ and exports of $X$ given by $M_Y^*(P_Y^*)= 1 - \frac{3}{2}P_Y^*$ and $E_X(P_X^*)= 2P_X^* -1$.
-With the only trade policy instruments being tariffs on import competing goods, world prices are $P_X = P_X^W + \tau$, $P_X^* = P_X^W$, $P_Y^* = P_Y^W + \tau^*$, and $P_Y = P_Y^W$.
-Market clearing implies that world and home prices of $X$ are $P_X^W = \frac{4-3\tau}{7}$ and $P_X = \frac{4+4\tau}{7}$, symmetric for $Y$.
-\\
-
-\subsection{Trade War Tariffs}
-The median legislator's welfare can be written as 
-\[
-  W_{\mathit{ML}}^X(\tau,\ga(e,\ve)) + W_{\mathit{ML}}^Y(\tau^*) = \left\{\frac{9}{98} - \frac{5}{49}\tau - \frac{34}{49}\tau^2 +\frac{1}{98}\ga(e,\ve)\left[ 8 + 16\tau + 8\tau^2 \right] \right\}+ \frac{25}{98} - \frac{3}{49}\tau^* + \frac{9}{49}(\tau^*)^2.
-\]
-%where $W_{\mathit{ML}}^X(\tau,\ga(e,\ve))$ is the utility derived from consumer surplus, producer surplus and tariff revenues in the import-competing industry and $W_{\mathit{ML}}^Y(\tau^*)$ is the utility derived from consumer and producer surplus in the exporting industry.
-
-When setting the trade-war tariff, the legislature maximizes $W_{\mathit{ML}}(\tau, \tau^*) = W_{\mathit{ML}}^X(\tau) + W_{\mathit{ML}}^Y(\tau^*)$ by choice of $\tau$ given $\tau^*$.
-As there are no interactions between $\tau$ and $\tau^*$, the legislature maximizes $W_{\mathit{ML}}^X(\tau)$ only (in curly braces above) and sets the trade war tariff
-\[
-  \tau^n = \frac{8\ga(e,\ve)-5}{68-8\ga(e,\ve)}
-\]
-I refer to this as the Nash tariff because it is the result of Nash equilibrium in the non-cooperative game between the legislatures.
-$\tau^n$ is increasing in $e$ and the second order condition is satisfied for all realizations of $\ga < 17/2$.
-Because $\ga = 7/4$ is enough to achieve the prohibitive tariff of $1/6$, I will assume this condition is satisfied.
-
-In the event of a trade war and facing this tariff-setting behavior by the legislature, the lobby maximizes $\pi\left(\tau^n\left(\ga\left(e_n,\ve_n\right)\right)\right) - e_n$.
-In order to predict the Nash tariff, the political weighting function and form of political uncertainty must be specified.
-Take for instance $\ga(e,\ve) = 1.25 + e^{.2} + \ta$ with $\ve$ distributed uniformly on $[-0.25,0.25]$.
-Facing this specification of the political process, the lobby maximizes its objective function at $e_n = 0.00166$, which produces an expected Nash tariff of $0.129$.
-
-\subsection{Break Decision}
-
-\begin{figure}
-\begin{center}
-\includegraphics[height=3in, width=4in]{brprob2.jpg}
-\end{center}
-\caption{Probability Trade Agreement will be Broken\label{fig:br}}
-\end{figure}
-
-\begin{comment}
-\begin{figure}
-\begin{center}
-\includegraphics[height=2.50in, width=3in]{contributions.jpg}
-\end{center}
-\caption{Lobbying Effort\label{fig:cont}}
-\end{figure}
-%graph is created in lobby_fix.m
-\end{comment}
-
-\begin{comment}
-\begin{figure}
-\centering
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=.4\linewidth]{brprob2.jpg}
-  \caption{Probability Trade Agreement will be Broken}
-  \label{fig:br}
-  %\label{fig:sub1}
-\end{subfigure}%
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=.4\linewidth]{contributions.jpg}
-  \caption{Lobbying Effort}
-  \label{fig:cont}
-  %\label{fig:sub2}
-\end{subfigure}
-%\caption{A figure with two subfigures}
-%\label{fig:test}
-\end{figure}
-\end{comment}
-
-Next we move to the legislature's decision on whether or not to break the trade agreement.
-Figure~\ref{fig:br} depicts the probability that the legislature will vote to break the trade agreement as a function of the tariff levels set in the trade agreement (with the restriction that $\tau^a = \tau^{*a}$) and the lobby's effort.
-It is strongly increasing in lobbying effort and decreasing in the level of tariffs set in the trade agreements.
-
-Given the impact of lobbying on the legislature's break decision, the lobby's optimal contribution level is strongly decreasing in the trade agreement tariffs.%, as shown in Figure~\ref{fig:cont}.
-
-\subsection{Trade Agreement}
-Because the probability of legislative break never reaches zero for this parameterization, the executives' welfare function is concave everywhere.
-Assuming that each legislature has the opportunity to vote to break the agreement with probability $\frac{1}{2}$ and that the executives are social-welfare maximizers (i.e.
-$\ga_E = \ga_E^* = 1$), they will set trade agreements tariffs of $\tau^a = \tau^{*a} = 0.078$, with lobbying expenditures of $0.0007$ and a total break probability of $0.505$.
-The expected tariff is then $0.103$.
-
-We can compare this against several different benchmarks.
-The most stark is the trade-war outcome itself, which can also be interpreted as the outcome that would prevail in the absence of executive involvement in trade policy \textit{and} the absence of any effort to make a trade agreement.
-The tariff in the executive-formed trade agreement is about 60$\%$ of the Nash tariff of $0.129$, and the expected level given the probability that the agreement will be broken is 80$\%$ of the Nash level.
-Lobbying expenditures in the agreement-maintenance phase are about 40$\%$ of the those in the Nash game, while expected expenditures in the executive-led trade-agreement scenario are 90$\%$ of what they would be if the legislatures set tariffs unilaterally.
-
-Although the welfare-maximizing governments here are not able to set and maintain tariffs at zero as they would like, they are able to achieve significant reductions in tariff levels through the use of the trade agreement.
-
-Another interesting benchmark is the scenario in which no executives are involved in trade policy, but in which the politically-susceptible legislatures can make use of a trade agreement.
-As in Bagwell and Staiger (2005), I find when maximizing their joint welfare in a legislature-led trade agreement ($W_{\mathit{ML}}(\tau, \tau^*) = W_{\mathit{ML}}^X(\tau) + W_{\mathit{ML}}^Y(\tau^*)$), the cooperative tariff levels will be set at
-\[
-  \tau^L = \frac{4\ga(e,\ve)-4}{25-4\ga(e,\ve)}, \ \tau^{*L} = \frac{4\ga(e^*,\ve)-4}{25-4\ga(e^*,\ve)}
-\]
-The legislatures are able to use the agreement to internalize the terms of trade externality, making political influence more expensive for the lobbies.
-Indeed, I find that lobbying expenditures rise to $0.0027$---60$\%$ higher than in the Nash case---while the agreement tariffs are set at $0.118$, only slightly lower than the Nash tariffs of $0.129$ and still higher than even the expected tariffs from the trade agreement scenario in which the executives choose the tariffs.
-
-If one interprets the addition of a second policy-making actor as consistent with a move from autocracy to democracy, this result helps to explain the extensive empirical evidence that democracies trade more with each other than autocracies.\footnote{cfr.
-Laver and Shepsle (1991); Henisz and Mansfield (2006); Aidt and Gassebner (2010).}
-
-\section{Discussion}
-\label{sec:dis}
-
-\subsection{Separation of Powers}
-\label{sec:sep_powers}
-Consider the model of Section~\ref{sec:main} for the case of $\ve_b=0$---that is, no uncertainty at the break stage (as will be discussed in Section~\ref{sec:uncertainty}, in the case of mean-zero uncertainty, strategic behavior at the trade-war stage is not altered by uncertainty).
-This allows the isolation of the results that derive from the assumption that power over trade policy is shared between the executive and legislative branches.
-Again, we focus on the home lobby.
-
-Now $\ga(e_b)$ is deterministic, so the lobby knows the precise contribution it must make for any given trade agreement tariffs $\bta$ to induce the legislature to break the agreement (see Inequality~\ref{eq:lwcg}).
-Here, the lobby's contribution \textit{increases} in $\bta$ since the higher are the trade agreement tariffs, the larger is the political weight that is required to induce the legislature to find them unsatisfactorily low.
-As long as trade war profits net of the required contribution are greater than trade agreement profits, the lobby will induce a trade war; otherwise, it is not in the lobby's interest to make any contribution.
-Facing this behavior, the executives set the lowest trade agreement tariffs that make it prohibitively expensive for the lobby to have the agreement broken.
-
-I return to the example of Section~\ref{sec:example} to illustrate the effects of changing the environment to one without political uncertainty.
-Recall that in the main example with $\ve$ distributed uniformly on $[-0.25,0.25]$, the expected Nash tariff is $0.129$ and the executive-led trade agreement tariff is $0.078$, with lobbying expenditures of $0.0007$ and a total break probability of $0.505$.
-The expected tariff is then $0.103$.
-
-In this example, if the executives set the trade agreement facing no uncertainty, the tariff level is $0.106$, the lobby exerts zero effort, and the agreement remain in force with probability 1.
-If instead the executives were to set trade-agreement tariffs of $0.105$, the lobby would contribute $0.0025$ and the legislature would break the agreement with probability 1.
-This demonstrates both the agenda-setting power of the executives and the stark discontinuities induced when political uncertainty is not an issue.
-
-Moreover, this case highlights the manipulation of the tariff level to discourage lobbying and therefore the disconnect that can arise between the tariffs that are chosen and the preferences of the governmental actors.
-In the example, the executive is a social welfare maximizer and so prefers zero tariffs, while the legislators with $\ga(e) = 1.25 + e^{.2}$ with $e_b = 0$ prefer $\tau^a = 0.05$.
-However, the trade agreement tariffs are set at $0.106$ precisely to ensure that $e_b =0$ so that we have a relatively free-trading legislature that upholds the agreement.
-
-\subsection{Relation to Grossman and Helpman's `Protection for Sale'}
-\label{sec:gh}
-The legislative welfare function employed in this paper permits flexibility to take account of a wide range of political processes within a non-unitary legislature; alternatively, it can be a model of a unitary policy-maker when special interests face non-constant returns to lobbying activity.
-As noted in Section 2, the non-linear relationship between lobbying effort and tariffs follows Dixit, Grossman and Helpman (1997) and is a departure from the `Protection for Sale' welfare function, which assumes that a unitary government maximizes the sum of contributions and some fraction, $a$, of social welfare.
-
-To see the relationship between the two forms, write the PFS welfare function (replacing $C$ with $e$) as
-\[
-  e + aW = e + a \left[ \mathit{CS}_X(\tau) + \mathit{PS}_X(\tau) + \mathit{CS}_Y(\tau^*) + \mathit{PS}_Y(\tau^*) + \mathit{TR}(\tau) \right]
-\]
-The relationship between $e$ and the weight the legislature places on profits is endogenous; because the PFS government welfare function has such an elegant and completely specified form, it is possible to solve for equilibrium behavior to retrieve this relationship directly.
-
-It is easy to describe the equilibrium behavior and thus how lobbying effort will be related to the weight on profits in equilibrium in the PFS model when there is one lobby who makes a take-it-or-leave-it offer.
-Take the case of no uncertainty and assume that the legislature chooses $\tau=0$ in the absence of lobbying and that $\ga(0)=1$.
-For any $\tau$ that the lobby desires, it must pay according to the  indifference condition $e(\tau) = a \left[ W(0) - W(\tau) \right]$.
-That is, the lobby must pay for the $a$-weighted welfare loss caused by the tariff it requests.
-
-In the model of this paper, this indifference condition, rearranged to solve for $\ga(e)$, is
-		    \[
-		      \ga(e) = \frac{\mathit{CS}_X(0) + \pi_X(0) + \mathit{CS}_Y(0) + \pi_Y(0) + \mathit{TR}(0) - \mathit{CS}_X(\tau) - \mathit{CS}_Y(\tau^*) - \pi_Y(\tau^*) - \mathit{TR}(\tau)}{\pi_X(\tau)}
-		    \] 
-In order to match the form of the PFS framework, we must therefore have
-		    \begin{equation}
-		      \ga(e) = 1 + \frac{e}{a\pi_X(\tau)}
-		      \label{eq:gh}
-		    \end{equation}
-
-Choosing the political weighting function identified in Equation~\ref{eq:gh} aligns the political objective function used in the current model with the PFS framework.
-Although using the $\ga$-weighted government welfare function is a reduced-form approach in the sense that legislative dynamics are not fully modeled, a drawback of the PFS welfare function is that it embodies the restriction that returns to lobbying effort must be constant.
-
-As noted in Dixit, Grossman and Helpman (1997), the linear form also implies transferable utility between the lobby and unitary government; this leads to equilibrium indeterminancies and an inability to speak to distributional questions, prompting them to generalize the PFS form.
-From their results, we have that the same equilibrium analysis holds in the more general environment.
-
-For the trade policy literature, generalizing the linear form in a sense would obviate many of the questions that have been raised by empirical investigations of the PFS model about the relationship of the magnitude of welfare losses relative to the weight placed on social welfare by governments as the $a$ parameter for the constant relative marginal utility provided by social welfare would no longer be present in the model.
-
-The linear PFS form provides for an almost unmatched simplicity and elegance that is able to capture the key cross-industry predictions on import tariffs.
-A perhaps unintended byproduct in the subsequent literature seems to have arisen in the form of a focus on matching fine details of governments' welfare-mindedness using a model that has a very simple unitary government and is thus better suited for other purposes.
-
-It is clear that trade policy is often shaped in a complicated process involving multiple actors (see, for instance, Destler (2005) for the U.S.).
-The model presented herein is a first attempt to more richly model the policy-making process and thus shed light on open questions surrounding how government preferences and the details of political institutions impact trade policy outcomes.
-
-\subsection{The Role of Political Uncertainty}
-\label{sec:uncertainty}
-In this model, if there were no political uncertainty, there would be no lobbying in equilibrium.
-Thus, if nothing else, adding the realism of political uncertainty allows for positive lobbying on the equilibrium path.
-If we examine more closely the assumption that there is significant uncertainty surrounding the legislative lobbying process, additional insights come to light.
-It is most clear in the case of mean-zero uncertainty, so I will assume this throughout this section unless otherwise noted.
-
-In the event of a trade war, the lobbies' behavior, and therefore the expected trade-war tariffs, are not altered by mean-zero uncertainty (see Equation~\ref{eq:lobtw}).
-
-At the earlier stage of lobbying, uncertainty alters optimizing behavior by both the lobby and the executives.
-In contrast to the previous section with no uncertainty, consider adding a very small amount of uncertainty to the previous example: take $\ve$ distributed uniformly on $[-0.01,0.01]$.
-Here, the executives find it optimal to set the tariffs to completely disengage the lobby and ensure the agreement remains in force.
-However, the tariff level will be different because instead of making the decision to lobby according to the certainty condition
-\[\pi(\tau^n) - e_n - \pi(\tau^a) > e_b,
-\]
-the lobby makes this decision according to Condition~\ref{ine:lobint}, which can be re-written as
-\[
-\pi(\tau^n) - e_n - \pi(\tau^a) > \left[\frac{\partial b(0,\bta,\btn(e_n,e_n^*))}{\partial e_b}\right]^{-1}.
-\]
-
-Although the relationship between the tariff under certainty and a small amount of uncertainty will depend on the structure of the economy and the legislative process, in this case, the tariff can be reduced slightly to $0.105$.
-As we have seen above, with significant uncertainty, the tariff in this example is reduced to $0.078$.
-Interestingly, the optimal tariff does not decrease monotonically between these two values as uncertainty increases.
-
-\begin{figure}
-\begin{center}
-\includegraphics[height=3in, width=4in]{lobby_br.jpg}
-\end{center}
-\caption{Lobby's Optimal Expenditure Function (Low and Intermediate Uncertainty)\label{fig:lobby_br}}
-\end{figure}
-
-It is particularly instructive to examine the lobby's reaction function at this very low level of uncertainty as well as the intermediate level when $\ve$ is distributed uniformly on $[-0.14,0.14]$ because it is at the latter that the executives set the highest tariff level of $0.107$.
-For the former, when facing very little uncertainty, the range of tariff levels to which the lobby responds with an intermediate contribution that leaves open the possibility of a break is very small: only those between 0.103 and 0.104.
-The lobby disengages completely at $\bta=0.105$ (see Figure~\ref{fig:lobby_br}).
-
-In contrast, when uncertainty is increased to the interval $[-0.14,0.14]$, the lobby begins to leave open the possibility of a break at $\bta = 0.048$.
-It disengages at $0.107$ and this is the tariff level that the executives find optimal.
-Interestingly, at intermediate uncertainly levels, both portions of the best response curve are steeper and the lobby disengages at slightly lower tariff levels---as low as $0.099$ when $\ve \sim U[-0.06,0.06]$.
-
-The preceding discussion illustrates Part (a) of Result~\ref{res:execsoln}: here the executives maximize their welfare by raising tariffs to the point where the import-competing industry ceases to lobby and thus the agreement remains in force for sure.
-Note, however, that although lobbying expenditures are zero, the potential for lobbying behavior is essential in shaping the trade agreement.
-
-%\begin{figure}
-%\begin{center}
-%\includegraphics[height=3.8in, width=7in]{exec_welfare.jpg}
-%\end{center}
-%\caption{Differences in Executive Welfare when Uncertainty Increases\label{fig:exec_wel}}
-%\end{figure}
-
-When uncertainty rises above this threshold of $\ve \sim U[-0.14,0.14]$, the executives' choices are made according to Part (b) of Result~\ref{res:execsoln}: that is, they trade off reductions in welfare under the agreement with reductions in the probability that the agreement will be abrogated by the legislature.
-%In terms of the lobby's best response function, the executives now choose the optimal point on the downward-sloping portion of the curve instead of the point where it reaches zero.
-In terms of the welfare function for the executives, shown in Figure~\ref{fig:exec_wel}, they are now choosing the point that maximizes the concave portion of the curve instead of the ``spike'' that is created when the lobbies disengage and the chance that the agreement will be abrogated is removed altogether.\footnote{Note that the initial, flat portion of the curve is where the tariff is low enough that the lobby buys a trade war with certainty; we then enter the concave portion where the main results of Section~\ref{sec:main} are applicable; finally, we see the upward spike where the probability of a trade war is reduced to zero and the executives' welfare declines after that because there is no further reduction in break probability as the tariff increases.}
-
-As uncertainty increases further, the general pattern is that tariffs, and therefore lobbying effort, are reduced, while the break probability increases.
-However, this pattern is not smooth or absolute, in contrast to the welfare level achieved by the executives, which increases monotonically in the amount of uncertainty present.
-%It is not clear how general a result this is, but it is particularly interesting in contrast to the non-monotone pattern over the lower-range of uncertainty where executive welfare is minimized at the edges of the range that lead to disengagement of the lobby and maximized in the middle of that range.
-
-Several salient points emerge from this example.
-First, different amounts of uncertainty lead to very different outcomes---both in terms of the tariffs that are set in the trade agreement and the probability of disruption.
-Second, these outcomes vary in nuanced ways with the underlying behavior of the lobbies and the executives.
-The provided tariff levels are strongly influenced by lobbying incentives, but not in the straightforward ways predicted by unitary models.
-
-\begin{comment}
-\section{Extensions}
-\label{sec:ext}
-
-This analysis has been undertaken with the assumption that the import-competing industry is distributed evenly across legislative districts.
-By varying the way in which those profits enter the welfare function of the median legislator (before lobbying takes place), I can provide results about how district composition affects trade policy.
-
-There are also several directions to explore in comparing different government structures. I start by describing the value of a trade agreement in an institutional setting of divided government.
-\begin{result}
-  Lobbying effort is reduced by the presence of executive branches who form a trade agreement (where the counterfactual is legislatures setting unilateral trade policy).
-\end{result}
-\textcolor{blue}{Don't yet know what assumptions I need to make this work} \\
-
-Comparing this model to the unitary-government model is a delicate task.
-The most straightforward approach from the current structure is to disregard the executives and allow the legislative branches to form a trade agreement cooperatively after lobbying takes place.
-Note that we can retain the stochastic nature of the tariff-setting, but there is no vehicle for the agreement to be broken.
-In this case, the trade agreement tariff in home, $\tau_U^a$ will maximize
-\[
-  W_X(\tau,\ga_E) + W_X^*(\tau) = \frac{9}{98} - \frac{5}{49}\tau - \frac{34}{49}\tau^2 +\frac{1}{98}\ga(e,\ve)\left[ 8 + 16\tau + 8\tau^2 \right] + \frac{25}{98} - \frac{3}{49}\tau + \frac{9}{49}(\tau)^2
-\]
-and will thus be set at
-\[
-  \tau_U^a = \frac{4\ga(e,\ve)-4}{25-4\ga(e,\ve)}.
-\]
-$\tau_U^a < \tau^n$ for all $\ga(e,\ve) < 7/4$, while $\ga(e,\ve)=7/4$ results in the prohibitive tariff of $1/6$.
-This is the standard result that a trade agreement internalized the terms of trade externality and results in a lower tariff than unilateral policy making.
-
-\textcolor{blue}{will an executive inserted into this situation set higher or lower tariffs, what will happen to $e$?}
-
-What if we add executive branches with essentially the same preferences as the legislature?
-That is, set $\ga_E = \expect[\ga(e,\ve)]$.
-\textcolor{blue}{It's hard to say how $e$ should be chosen here...still need to figure that out.
-But if we can get around the weird time inconsistency problem...} What agreement tariffs will the executives choose?
-It's easy to see that they will not choose $\tau^a < \tau_U^a$, as $\tau_U^a$ is their ideal point if there were no possibility of the agreement being broken, and lower tariffs only incentivize more lobbying and a higher probability that the agreement will be broken.
-
-Thus, if an interior solution exists \textcolor{blue}{[should have conditions to insert here]}, the executive will exploit the same trade-off explored earlier, raising $\tau^a$ a bit to reduce the probability that the agreement will be abrogated.
-So interestingly, we can see that executives with the same political preferences as the legislature will choose higher trade agreement tariffs than the legislature would by itself due to the possibility that the agreement can be broken and the impact of the tariff choice on lobbying.
-We will also therefore have a reduced level of contributions, and if a trade war were possible, lower trade war tariffs because of it.
-
-Along these lines, allowing for asymmetries in the preferences and political processes between the countries will provide important insights as well.
-
-I also plan to use the richness of this model to parse out how much of the changes in tariffs comes from the fact that the executives place lower weight on producer profits and how much comes from the fact that they make an agreement---that is, compare the effects of structure versus preferences.
-
-\subsection{Asymmetric Trade Agreements}
-\label{sec:asym}
-The assumption that the legislatures do not have the opportunity to break a trade agreement simultaneously serves to simplify the analysis of the interaction between the lobbies and the consequences of relaxing it will be discussed in the following section.
-Maintaining it for now, however, facilitates examination of the assumption that trade agreements must be symmetric.
-This may be quite realistic in some settings where additional political constraints would make ``unfair'' arrangements unpalatable.
-However, it is important to point out that, for some parameter choices, the joint welfare of the executives is in fact maximized by an asymmetric agreement.
-
-The parameterization from the example in Section~\ref{sec:example} is one such instance.
-The optimal unrestricted trade agreement is to set tariffs in one country (say home) at $0.062$ and those in the other (foreign) at 0.
-In this case, the equilibrium break probability in home, if the home legislature is afforded the opportunity, is zero, while the foreign legislature will break the agreement with probability 1 if given the chance.
-With the assumption from the example that each legislature gets the opportunity to repudiate the agreement half the time, this results in the trade agreement being broken with probability $0.50$, just less than the $0.505$ of the optimal symmetric agreement with both tariffs set at $0.078$.
-Because both the trade agreement tariffs and the expected chance that the agreement will remain in force are lower, expected welfare for the executives is higher.
-
-What is behind the rather surprising result that a lower break probability can be achieved when both tariffs are strictly lower than in the symmetric agreement?
-Here, the executives relinquish any chance of passage of the trade agreement in the foreign country (in essence hoping that its legislature will not have a chance to act).
-Once they decide to pursue this course, it is optimal to reduce $\tau^{*a}$ to zero.
-This drastically reduces the home legislature's incentive to break the trade agreement.
-Because of the large direct impact of $\tau^{*a}$ in reducing the break probability, the home tariff is not required to make such a large contribution and can be lowered, which in turn reduces the lobby's incentive to exert effort (see Result~\ref{res:lobby}).
-This effect is strong enough that the executives can set a tariff level lower than that of the symmetric agreement that is sufficient to disengage the home lobby (see Inequality~\ref{ine:lobint}).
-
-It is useful to keep in mind that this result occurs in a completely symmetric environment.
-Recall that, in general, the executives maximize joint welfare in Equation~\ref{eq:jv2}.
-If the chance that the legislatures will have the opportunity to break the agreement ($o$ and $o^*$) are not equal, or if another significant aspect of the economic or political situation is asymmetric, the importance of this result grows in a potentially substantial way.
-In particular, if only one country faces a legislative constraint, an asymmetric agreement seems highly likely as in the large body of literature investigating the Schelling conjecture.
-
-\subsection{Break Opportunities in Both Countries}
-\label{sec:two}
-Although it changes their expected payoffs, extending the model to allow both legislatures the opportunity to break the trade agreement simultaneously does not impact the legislatures' incentives as each decision to break or not break the agreement is independent of the other.
-
-However, the lobbies' incentives are altered because the total probability of a break with which they weight profits changes.
-Now, in addition to their own effort positively influencing their own legislature to break the agreement, the probability of a break in the trade agreement also increases in the effort of the other country's lobby.
-As might be expected, the lobbying effort in each country turns out to be a negative function of that in the trading partner; that is, to some extent, the lobbies have an incentive to free ride.
-
-Without making stronger assumptions, it is not possible to say how much the effort of each lobby will be reduced therefore whether total lobbying effort rises or falls.
-What is clear is that the optimization of the executives becomes significantly more complex as they are now faced with the problem of how to best exploit the free-riding dynamics that occur between the lobbies.
-\end{comment}
-
-\section{Conclusion}
-\label{sec:concl}
-I have shown that the legislature both breaks trade agreements with a higher probability and sets higher trade war tariffs when lobbying activity increases, while the probability with which it breaks agreements (or fails to ratify them) decreases in the domestic trade agreement tariff.
-Because the lobby decreases its effort in response to higher trade agreement tariffs, the executives face a trade-off between the welfare derived while a trade agreement is in force and the probability with which the agreement actually enters into force.
-
-I have also shown that in a policy-making structure in which power is shared, a less politically-motivated executive can utilize an international trade agreement to reduce the political pressure on the ratifying body and therefore increase the probability that the agreement will remain in force.
-Thus, in a model with a richer description of government structure, a political-commitment role for trade agreements can arise.
-
-The executives' incentive to raise tariffs in order to reduce lobbying effort helps to explain the empirical finding in the Protection for Sale literature that levels of protection and associated deadweight losses are too high relative to lobbying expenditure given the high estimates for governments' weighting of social welfare.
-I have shown that this serves to mediate the relationship between preferences for contributions relative to social welfare and the tariffs that are provided.
-The intuition is clear: the observed lobbying expenditure levels may in fact be low \textit{because} tariffs have been raised sufficiently high to prevent political pressure and the increased risk of a costly trade disruption it engenders.
-
-That lobbying and tariff levels are related in systematic ways to the amount of political uncertainty present suggests interesting avenues for future empirical work.
-Several directions for future theoretical work also seem potentially fruitful, including removing the assumption of perfect enforceability and supporting cooperation through repeated interaction and generalizing the model to the case of multiple lobbies.
-
-\section{Appendix}
-\label{sec:appendix}
-\noindent \textbf{\hypertarget{Pr_bincC}{Proof of Result~\ref{res:bincC}}}:\\
-Substituting from Equation~\ref{eq:ml}, Equation~\ref{eq:b} can be re-written as
-\begin{multline}  
-  b(e_b,\bta,\btn) = \Pr [ \mathit{CS}_X(\tau^n) + \ga(e_b,\ve_b) \cdot \mathit{PS}_X(\tau^n) + \mathit{CS}_Y(\tau^{*n}) + \mathit{PS}_Y(\tau^{*n}) + \mathit{TR}(\tau^n) >  \\ \mathit{CS}_X(\tau^a) + \ga(e_b,\ve_b) \cdot \mathit{PS}_X(\tau^a) + \mathit{CS}_Y(\tau^{*a}) + \mathit{PS}_Y(\tau^{*a}) + \mathit{TR}(\tau^a) ]
-\end{multline}
-Rearranging, we have $b(e_b,\bta,\btn) = $
-\begin{multline}  
-  \textstyle \Pr \Big[ \frac{\mathit{CS}_X(\tau^n) + \mathit{PS}_X(\tau^n) + \mathit{CS}_Y(\tau^{*n}) + \mathit{PS}_Y(\tau^{*n}) + \mathit{TR}(\tau^n)  -\mathit{CS}_X(\tau^a) - \mathit{PS}_X(\tau^a) - \mathit{CS}_Y(\tau^{*a}) - \mathit{PS}_Y(\tau^{*a}) - \mathit{TR}(\tau^a)}{\mathit{PS}_X(\tau^a) - \mathit{PS}_X(\tau^n)} + 1 \\ \textstyle < \ga(e_b,\ve_b) \Big]
-  \label{eq:b_ex}
-\end{multline}
-%\noindent The left side of the inequality in Expression~\ref{eq:b_ex} does not depend on $e_b$.
-By Assumption~\ref{as:ga_c}, the right side of the inequality is increasing and concave in $e_b$.
-Thus $b(e_b,\bta,\btn)$ is increasing and concave in $e_b$.
-$\hfill\blacksquare$
-\noindent The left side of the inequality in Expression~\ref{eq:b_ex} does not depend on $e_b$.
-Call it $Z$.
-Thus we have $b(e_b,\bta,\btn) = \Pr [Z < \ga(e_b,\ve_b) ] = 1 - F_{\ga}(\ga=Z)$ where $F_{\ga}(\ga=Z) = F_{\ta}(\ta=h^{-1}(\ga,e))$ by the Change of Variables Theorem and Assumption~\ref{as:ga_ta} with $\ga = h(e,\ta)$ giving the change of variable.
-
-Then $\frac{\partial b}{\partial e} = -\frac{\partial F_{\ta}(\ta=h^{-1}(\ga,e))}{\partial \ta}\frac{\partial h^{-1}(\ga,e)}{\partial e} = -f_\ta(\ta) \frac{\partial h^{-1}(\ga,e)}{\partial e}$ and $\frac{\partial^2 b}{\partial e^2} = -f_\ta(\ta) \frac{\partial^2 h^{-1}(\ga,e)}{\partial e^2}$.
-Because $\ga = h(e,\ta)$ is increasing and concave in $e \ \forall \ta$ by Assumption~\ref{as:ga_c}, its inverse is decreasing and convex $\left(\frac{\partial h^{-1}(\ga,e)}{\partial e}\leq 0; \ \frac{\partial^2 h^{-1}(\ga,e)}{\partial e^2} \geq 0 \right)$ The pdf of $\ta$ is non-negative, so $\frac{\partial b}{\partial e} \geq 0$ and $\frac{\partial^2 b}{\partial e^2} \leq 0$.
-$\hfill\blacksquare$
-
-\vskip.4in
-\noindent \textbf{\hypertarget{Pr_leg_astar}{Proof of Lemma~\ref{lem:leg_astar}}}:\\
-It must be shown that the left hand side of the inequality in Expression~\ref{eq:b_ex} is decreasing in $\tau^{*a}$. The derivative of this quantity with respect to $\tau^{*a}$ is 
-\begin{equation}
-  \frac{-\left(\mathit{PS}_X(\tau^a) - \mathit{PS}_X(\tau^n)\right)\left(\frac{\partial \mathit{CS}_Y(\tau^{*a})}{\partial \tau^{*a}} + \frac{\partial \mathit{PS}_Y(\tau^{*a})}{\partial \tau^{*a}}\right)}{\left(\mathit{PS}_X(\tau^a) - \mathit{PS}_X(\tau^n)\right)^2} = \frac{\frac{\partial \mathit{CS}_Y(\tau^{*a})}{\partial \tau^{*a}} + \frac{\partial \mathit{PS}_Y(\tau^{*a})}{\partial \tau^{*a}}}{\mathit{PS}_X(\tau^n) - \mathit{PS}_X(\tau^a)}.
-  \label{appex:astar}
-\end{equation}
-
-\noindent The price of good $Y$ is decreasing in $\tau^{*a}$, so consumer surplus is increasing and producer surplus is decreasing in $\tau^{*a}$.
-Because $Y$ is being exported, the decrease in producer surplus is larger than the increase in consumer surplus, making the numerator negative.
-Since producer surplus in the denominator is increasing in $\tau$, the expression in Equation~\ref{appex:astar} is negative for all $\tau^{*a}$.
-$\hfill\blacksquare$
-
-\vskip.4in
-\noindent \textbf{\hypertarget{Pr_leg_a}{Proof of Lemma~\ref{res:leg_a}}}:\\
-Using the logic of the proof of Lemma~\ref{lem:leg_astar}, the effect on the break probability is determined by the sign of the derivative of the left hand side of the inequality in Expression~\ref{eq:b_ex} with respect to $\tau^a$; to show that the break probability is decreasing in $\tau^a$, I must demonstrate that this derivative is positive.
-Labeling the numerator of that expression $\left[ W(\btn)-W(\bta)\right]$ (for the change in social welfare), this derivative can be written
-\begin{equation}
-  \frac{\left(\mathit{PS}_X(\tau^n) - \mathit{PS}_X(\tau^a)\right)\left(\frac{\partial \mathit{CS}_X(\tau^{a})}{\partial \tau^{a}} + \frac{\partial \mathit{PS}_X(\tau^a)}{\partial \tau^{a}} + \frac{\partial \mathit{TR}(\tau^{a})}{\partial \tau^{a}}\right) - \left[ W(\btn)-W(\bta)\right] \frac{\partial \mathit{PS}_X(\tau^a)}{\partial \tau^a}}{\left(\mathit{PS}_X(\tau^a) - \mathit{PS}_X(\tau^n)\right)^2}.
-  \label{appex:a}
-\end{equation}
-
-\noindent $\left(\mathit{PS}_X(\tau^n) - \mathit{PS}_X(\tau^a)\right)$ is always positive by Assumption~\ref{as:ga_l_e}.
-Because the optimal unilateral tariff for large welfare-maximizing governments is positive (call it $\tau^O$), $\left(\frac{\partial \mathit{CS}_X(\tau^{a})}{\partial \tau^{a}} + \frac{\partial \mathit{PS}_X(\tau^a)}{\partial \tau^{a}} + \frac{\partial \mathit{TR}(\tau^{a})}{\partial \tau^{a}}\right)$ is increasing up to $\tau^O$ and decreasing above it.
-Thus the first summand is increasing up until $\tau^O$ and decreasing thereafter.
-
-Because total social welfare is maximized at $\tau^a = \tau^{*a} = 0$, $W(\btn)-W(\bta)$ is always negative, whereas producer surplus is increasing in $\tau^a$, so the second summand is positive everywhere.
-With a positive denominator, we thus have that the derivative is positive on $[0,\tau^O]$.
-
-It is also positive over the remaining $(\tau^O,\tau^n)$.
-To see this, add $\left(\tilde{\Gamma} - 1 \right) \frac{\partial \mathit{PS}_X(\tau^a)}{\partial \tau^a} \left(\mathit{PS}_X(\tau^n) - \mathit{PS}_X(\tau^a)\right)$ to the first summand and subtract it from the second.
-For any particular value of $\tilde{\tau}^a$, one can choose the $\tilde{\Gamma}$ weight that would make $\tilde{\tau}^a$ the preferred unilateral tariff; this makes the derivative in the first summand zero.
-Having subtracted the same quantity from the second summand modifies the welfare difference in the second summand to be maximized at $\tilde{\tau}^a$ so that this term is always negative, thus ensuring the result.
-$\hfill\blacksquare$
-
-\vskip.4in
-\noindent \textbf{\hypertarget{Pr_bcomb}{Proof of Lemma~\ref{res:bcomb}}}:\\
-Again, I want to show how the inequality in Expression~\ref{eq:b_ex} changes, now with respect to both $\tau^a$ and $\tau^{*a}$, so I add the derivatives in Expressions~\ref{appex:astar} and \ref{appex:a} to get
-\[
-  \frac{\left(\mathit{PS}_X(\tau^n) - \mathit{PS}_X(\tau^a)\right)\left(\frac{\partial \mathit{W}_X(\bta)}{\partial \tau^{a}} + \frac{\partial \mathit{W}_X(\bta)}{\partial \tau^{*a}}\right) - \left[ W(\btn)-W(\bta)\right] \frac{\partial \mathit{PS}_X(\tau^a)}{\partial \tau^a}}{\left(\mathit{PS}_X(\tau^n) - \mathit{PS}_X(\tau^a)\right)^2}.
-\]
-where $\frac{\partial \mathit{W}_X(\bta)}{\partial \tau^{a}} + \frac{\partial \mathit{W}_X(\bta)}{\partial \tau^{*a}} = \frac{\partial \mathit{CS}_X(\tau^{a})}{\partial \tau^{a}} + \frac{\partial \mathit{PS}_X(\tau^a)}{\partial \tau^{a}} + \frac{\partial \mathit{TR}(\tau^{a})}{\partial \tau^{a}} + \frac{\partial \mathit{CS}_Y(\tau^{*a})}{\partial \tau^{*a}} + \frac{\partial \mathit{PS}_Y(\tau^{*a})}{\partial \tau^{*a}}$ is the total derivative of social welfare.
-Since social welfare is maximized at $\bta = (0,0)$,\footnote{Note, this is identical to the result that joint social welfare is maximized at zero tariffs because of symmetry.}  this is negative $\forall \bm{\tau} \in (\bm{0},\btn]$; note that it is 0 at $\bm{0}$ and vanishingly small for very small tariffs.
-
-Thus the first summand in the numerator is zero at $\bta = \bm{0}$ and increasingly negative as $\bta$ increases.
-The second summand is positive everywhere because social welfare, $W$, is lowest at $\btn$ and producer surplus is increasing everywhere.
-Thus the numerator is positive at $\bm{0}$ and at least for very small $\bta$
-
-It is also positive for all other values of $\bta$ strictly below $\btn$. Just as in the proof of Lemma~\ref{res:leg_a}, one can add
-$\left(\tilde{\Gamma} - 1 \right) \frac{\partial \mathit{PS}_X(\tau^a)}{\partial \tau^a} \left(\mathit{PS}_X(\tau^n) - \mathit{PS}_X(\tau^a)\right)$ to the first summand and subtract it from the second.
-For any particular value of $\tilde{\bta}$, one can choose the $\tilde{\Gamma}$ weight that would make $\tilde{\bta}$ the politically optimal tariff; this makes the derivative in the first summand zero.
-Having subtracted the same quantity from the second summand modifies the welfare difference in the second summand to be maximized at $\tilde{\bta}$ so that this term is always negative, thus ensuring the result.
-
-Because the denominator is positive, the entire expression is positive for all $\bta < \btn$.   $\hfill\blacksquare$
-
-\vskip.4in
-\noindent \textbf{\hypertarget{Pr_lobby}{Proof of Result~\ref{res:lobby}}}:\\
-Proof is via the Implicit Function Theorem using the lobby's first order condition, Equation~\ref{eq:lobbyfoc}, referred to here as $FOC_L$.
-  \[
-		\frac{\partial e_b}{\partial \tau^a} + \frac{\partial e_b}{\partial \tau^{*a}} = - \left[ \frac{\frac{\partial FOC_L}{\partial \tau^a}}{\frac{\partial FOC_L}{\partial e_b}} + \frac{\frac{\partial FOC_L}{\partial \tau^{*a}}}{\frac{\partial FOC_L}{\partial e_b}} \right] = \frac{ \frac{\partial b}{\partial e_b} \frac{\partial \pi(\bta)}{\partial \tau^a} + \frac{\partial b}{\partial e_b} \frac{\partial \pi(\bta)}{\partial \tau^{*a}} - \left\{\frac{\partial^2 b}{\partial e_b \partial \tau^a} + \frac{\partial^2 b}{\partial e_b \partial \tau^{*a}}   \right\}\left[ \pi(\tau^n) - e_n - \pi(\tau^a) \right]}{\frac{\partial^2 b}{\partial e_b{}^2} \left[ \pi(\tau^n) - e_n - \pi(\tau^a) \right]}
-	\]
-Beginning with the denominator: because $\pi(\tau)$ is increasing everywhere, $\left[\pi(\tau^n) - e_n - \pi(\tau^a) \right]$ is positive for all but very large values of $\tau^a$, that is for all $\tau^a$ such that $\pi(\tau^n) - e_n > \pi(\tau^a)$.
-When $\tau^a$ rises above this level, it is no longer in the lobby's interest to ask to have the agreement broken so $e_b=0$ and $\frac{\partial e_b}{\partial \tau^a} = 0$.
-$\frac{\partial^2 b}{\partial e_b{}^2}$ negative by Result~\ref{res:bincC}, so the denominator is negative for all values of $\tau^a$ at which $e_b$ is not constant.
-
-$\frac{\partial b}{\partial e_b}$ is positive, $\frac{\partial \pi(\bta)}{\partial \tau^a}$ is positive by construction, and $\frac{\partial \pi(\bta)}{\partial \tau^{*a}}$ is zero: separability between the sectors implies that profits in the import-competing sector do not depend on $\tau^{*a}$.
-
-We can rewrite $\frac{\partial b(e,\bta)}{\partial  \tau^a} + \frac{\partial b(e,\bta)}{\partial  \tau^{*a}} = -\frac{\partial F_\ga (Z(\bta))}{\partial \tau^a} -\frac{\partial F_\ga (Z(\bta))}{\partial \tau^{*a}} = -\frac{\partial F_\ga (Z(\bta))}{\partial Z(\bta)}\left[\frac{\partial Z(\bta)}{\partial \tau^a} + \frac{\partial Z(\bta)}{\partial \tau^{*a}} \right] = - f_\ga \left[\frac{\partial Z(\bta)}{\partial \tau^a} + \frac{\partial Z(\bta)}{\partial \tau^{*a}} \right]$, where $F_\ga (Z(\bta))$ is the CDF of $\ga$ and $f_\ga (Z(\bta))$ is the pdf of $\ga$ and $Z(\bta)$ represents the left hand side of the inequality in Expression~\ref{eq:b_ex}.
-
-Then $\frac{\partial^2 b}{\partial e_b \partial \tau^a} + \frac{\partial^2 b}{\partial e_b \partial \tau^{*a}} = 	- \frac{\partial}{\partial e} \left( \frac{\partial F_\ga (Z(\bta))}{\partial \tau^a} + \frac{\partial F_\ga (Z(\bta))}{\partial \tau^{*a}} \right) =
-	   - f_\ga \left[\frac{\partial^2 Z(\bta)}{\partial \tau^a \partial e} + \frac{\partial^2 Z(\bta)}{\partial \tau^{*a} \partial e} \right] - \left[\frac{\partial Z(\bta)}{\partial \tau^a} + \frac{\partial Z(\bta)}{\partial \tau^{*a}} \right] \frac{\partial f_\ga}{\partial e} = - \left[\frac{\partial Z(\bta)}{\partial \tau^a} + \frac{\partial Z(\bta)}{\partial \tau^{*a}} \right] \frac{\partial f_\ga}{\partial e}
-$
-where the above equality holds because $Z(\bta)$ does not depend on $e$.
-Lemma~\ref{res:bcomb} shows that $\frac{\partial Z(\bta)}{\partial \tau^a} + \frac{\partial Z(\bta)}{\partial \tau^{*a}} \geq 0$, so $\frac{\partial f_\ga}{\partial e} \geq 0$ ensures that $\left\{\frac{\partial^2 b}{\partial e_b \partial \tau^a} + \frac{\partial^2 b}{\partial e_b \partial \tau^{*a}} \right\} \geq 0$.
-Thus $\frac{\partial e_b}{\partial \tau^a} + \frac{\partial e_b}{\partial \tau^{*a}} \leq 0$.
-$\hfill\blacksquare$
-
-\vskip.4in	  	
-\noindent \textbf{\hypertarget{Pr_bcomB}{Proof of Result~\ref{res:bcomB}}}: \\
-$	\frac{\partial B(\bta)}{\partial \tau^a} + \frac{\partial B(\bta)}{\partial  \tau^{*a}} = \frac{\partial b}{\partial  e_b}\frac{\partial e_b}{\partial \tau^a} + \frac{\partial b}{\partial  e_b}\frac{\partial  e_b}{\partial  \tau^{*a}} + \frac{\partial b}{\partial \tau^a} + \frac{\partial b}{\partial  \tau^{*a}} = \frac{\partial b}{\partial  e_b} \left\{ \frac{\partial e_b}{\partial \tau^a} + \frac{\partial  e_b}{\partial  \tau^{*a}} \right\} + \frac{\partial b}{\partial \tau^a} + \frac{\partial b}{\partial  \tau^{*a}}.$ $\frac{\partial b}{\partial  e_b}$ is positive by Result~\ref{res:bincC}.
-$\left\{ \frac{\partial e_b}{\partial \tau^a} + \frac{\partial  e_b}{\partial  \tau^{*a}} \right\}$ is negative by Result~\ref{res:lobby}.
-Taken together, the final two summands are negative by Lemma~\ref{res:bcomb}.
-Thus the entire expression is negative.
-$\hfill\blacksquare$
-
-\vskip.4in
-\noindent \textbf{\hypertarget{int_soln}{Conditions for Interior Solution to Executives' Problem}} \\
-The first order condition for maximizing joint surplus with respect to $\tau^a$ and $\tau^{*a}$ when the two are constrained to be equal is
-\begin{multline*}
-  \textstyle \left\{ 1 - o \cdot B(\bta) - o^* \cdot B^*(\bta) \right\} \left\{\frac{\partial \bm{W_E}(\bta)}{\partial \tau^a} + \frac{\partial \bm{W_E}(\bta)}{\partial \tau^{*a}} \right\} \\
-     \textstyle + \left\{ o \cdot \left[\frac{\partial B(\bta) }{\partial \tau^a} + \frac{\partial B(\bta) }{\partial \tau^{*a}} \right] + o^* \cdot \left[\frac{\partial B^*(\bta)}{\partial \tau^a} + \frac{\partial B^*(\bta)}{\partial \tau^{*a}} \right] \right\} \left[ \bm{W_E}(\btn)- \bm{W_E}(\bta) \right] = 0
-\end{multline*}
-Because there is no benefit to setting $\bta$ below the executives' preferred level, which I will denote $\bm{\tau^E}$, I will take the choice space to be $[\bte,\btn]$.
-Note that for $\ga_E = 1$, $\bte=0$.
-
-To demonstrate that the executives do not choose $\bta=\bte$, I must show that the left side of the above equation is positive at $\bta=\bte$.
-Assumption~\ref{as:ga_l_e} and Result~\ref{res:bcomB} combined with symmetry ensure that the first term of the second summand is negative.
-That executive welfare is maximized at $\bte$ ensures that the term multiplying it is also negative as well as that $\frac{\partial \bm{W_E}(\bte)}{\partial \tau^a} + \frac{\partial \bm{W_E}(\bte)}{\partial \tau^{*a}}$ is zero.
-Therefore the derivative of joint executive welfare is positive at $\bte$.
-$\hfill\blacksquare$
-
-%\section{Acknowledgements}
-%The author thanks the following people for their insightful comments: Joel Watson, Jim Rauch, Marc Muendler, Lawrence Broz, Peter Cowhey, Kyle Bagwell, Andrei Levchenko, Bob Staiger, two anonymous referees, and seminar participants at UC San Diego, Syracuse, Georgetown, Florida International, Brandeis, BGSU, U Mass Amherst, Swarthmore, SUNY Oswego, Kenyon College, and PEIO 2013.
-All remaining errors are my own.
-
-\section{References}
-\singlespacing
-
-\begin{list}{}{\setlength{\leftmargin}{0.0in}\setlength{\rightmargin}{0.0in}\setlength{\itemindent}{0.0in}\setlength{\itemsep}{0.1in}}
-
-\item Aidt, T., Gassebner, M., 2010. Do Autocratic States Trade Less? The World Bank Economic Review 24, 38-76.
-
-\item Ansolabehere, S., de Figueiredo, J., Snyder Jr., J., 2003. Why Is There so Little Money in U.S. Politics? Journal of Economic Perspectives 17, 105-130.
-
-\item Bagwell, K., Staiger, R., 2005.
-Enforcement, Private Political Pressure, and the General Agreement on Tariffs and Trade/World Trade Organization Escape Clause.
-Journal of Legal Studies 34, 471-513.
-
-\item Baldwin, R.E., 1987. Politically realistic objective functions and trade policy: PROFs and tariffs. Economic Letters 24, 287-90.
-
-\item Beshkar, M., 2010. Trade skirmishes safeguards: A theory of the WTO dispute settlement process. Journal of International Economics 82, 35-48.
-
-\item Bombardini, M., 2008. Firm Heterogeneity and Lobby Participation. Journal of International Economics 75, 329-348.
-
-\item Bombardini, M., Trebbi, F., 2012. Competition and Political Organization: Together or Alone in Lobbying for Trade Policy? Journal of International Economics 87, 18-26.
-
-\item Buzard, K., 2014a. Endogenous Politics and the Design of Trade Agreements. Available at: \url{https://kbuzard.expressions.syr.edu/wp-content/uploads/Endogenous-Politics.pdf}.
-
-\item Buzard, K., 2014b.
-Self-enforcing Trade Agreements, Dispute Settlement and Separation of Powers.
-Available at: \url{https://kbuzard.expressions.syr.edu/wp-content/uploads/2014/04/Self-enforcing-Trade_Agreements.pdf}.
-
-\item Coates, D., Ludema, R., 2001. A Theory of Trade Policy Leadership. Journal of Development Economics 65, 1-29.
-
-\item Dai, X., 2006. Dyadic Myth and Monadic Advantage: Conceptualizing the Effect of Democratic Constraints on Trade. Journal of Theoretical Politics 18, 267-297.
-
-\item Destler, I.M., 2005. American Trade Politics. Institute for International Economics, Washington, DC.
-
-\item Dixit, A., G. Grossman, Helpman, E., 1997. Common Agency and Coordination: General Theory and Application to Government Policy Making. Journal of Political Economy 105, 752-769.
-
-\item Ethier, W., 2002. Unilateralism in a Multilateral World. The Economic Journal 112, 266-292.
-
-\item Ethier, W., 2012. The Political-Support Approach to Protection. Global Journal of Economics 1, 1-14.
-
-\item Feenstra, R., 1992. How Costly is Protectionism? Journal of Economic Perspectives 6, 159-178.
-
-\item Feenstra, R., Lewis, T., 1991. Negotiated Trade Restrictions with Private Political Pressure. Quarterly Journal of Economics 106, 1287-1307.
-
-\item Findlay, R., Wellisz, S., 1982.
-Endogenous Tariffs and the Political Economy of Trade Restrictions and Welfare.
-In Jagdish Bhagwati (ed.) Import Competition and Response, Chicago, IL: University of Chicago, 1982
-
-\item Gawande, K., Bandyopadhyay, U., 2000. Is Protection for Sale? Evidence on the Grossman-Helpman Theory of Endogenous Protection. The Review of Economics and Statistics 82, 139-152.
-
-\item Gawande, K., Hoekman, B., 2006. Lobbying and Agricultural Trade Policy in the United States. International Organization 60, 527-561.
-
-\item Gawande, K., Magee, C., 2012. Free Riding on Protection for Sale. International Studies Quarterly 56, 735-747.
-
-\item Gawande, K., Krishna, P., Robbins, M., 2006. Foreign Lobbies and U.S. Trade Policy. Review of Economics and Statistics 88, 563-571.
-
-\item Gawande, K., Krishna, P., Olarreaga, M., 2009. What Governments Maximize and Why: The View from Trade. International Organization 63, 491-532.
-
-\item Gawande, K., Krishna, P., Olarreaga, M., 2012. Lobbying Competition Over Trade Policy. International Economic Review 53, 115-132.
-
-\item Goldberg, P., Maggi, G., 1999. Protection for Sale: An Empirical Investigation. American Economic Review 89, 1135-1155.
-
-\item Grossman, G., Helpman, E., 1994. Protection for Sale. The American Economic Review 84, 833-850.
-
-\item Grossman, G., Helpman, E., 1995a. The Politics of Free-Trade Agreements. The American Economic Review 85, 667-690.
-
-\item Grossman, G., Helpman, E., 1995b. Trade Wars and Trade Talks. The Journal of Political Economy 103, 675-708.
-
-\item Grossman, G., Helpman, E., 2005. A Protectionist Bias in Majoritarian Politics. The Quarterly Journal of Economics 120, 1239-1282.
-
-\item Henisz, W., Mansfield, E., 2006. Votes and Vetoes: The Political Determinants of Commercial Openness. International Studies Quarterly 50, 189-211.
-
-\item Hillman, A., 1982. Declining Industries and Political-Support Protectionist Motives. The American Economic Review 5, 1180-1187.
-
-\item Iida, K., 1996. Involuntary Defection in Two-Level Games. Public Choice 89, 283-303.
-
-\item Imai, S., Katayama, H., Krishna, K., 2009. Is protection really for sale? A survey and directions for future research. International Review of Economics and Finance 18, 181-191.
-
-%\item Kibris, A., 2012. Uncertainty and Ratification Failure. Public Choice 150, 439-467.
-
-\item Klimenko, M., Ramey, G., Watson, J., 2008. Recurrent Trade Agreements and the Value of External Enforcement. Journal of International Economics 74, 475-499.
-
-%\item Kono, D., 2006. Optimal Obfuscation: Democracy and Trade Policy Transparency. American Political Science Review 100, 369-384.
-
-\item Laver, M., Shepsle, K., 1991. Divided Government: America is Not `Exceptional'. Governance: An International Journal of Policy and Administration 4, 250-269.
-
-\item Le Breton, M., Salanie, F., 2003. Lobbying under political uncertainty. Journal of Public Economics 87, 2589-2610.
-
-\item Le Breton, M., Zaporozhets, V., 2007. Legislative Lobbying under Political Uncertainty. Available at SSRN: \url{http://ssrn.com/abstract=1024686}.
-
-\item Lohmann, S., O'Halloran, S., 1994. Divided Government and US Trade Policy. International Organization 48, 595-632.
-
-\item Maggi, G., Rodr\'{i}guez-Clare, A., 2007. A Political-Economy Theory of Trade Agreements. The American Economic Review 97, 1374-1406.
-
-\item Maggi, G., Staiger, R., 2011. The Role of Dispute Settlement Procedures in International Trade Agreements. Quarterly Journal of Economics 126, 475-515.
-
-\item Mansfield, E., Milner, H., Rosendorff, B.P., 2000. Free to Trade: Democracies, Autocracies, and International Trade. The American Political Science Review 94, 305-321.
-
-%\item Milner, H., Kubota, K., 2005. Why the Move to Free Trade? Democracy and Trade Policy in the Developing Countries. International Organization 59, 107-143.
-
-\item Milner, H., Rosendorff, B.P., 1997.
-Democratic Politics and International trade Negotiations: Elections and Divided Government as Constraints on Trade Liberalization.
-Journal of Conflict Resolution 41, 117-147.
-
-\item Milner, H., Rosendorff, B.P., 2001. The Optimal Design of International Trade Institutions: Uncertainty and Escape. International Organization 55, 829-857.
-
-\item Mitra, D., Thomakos, D., Ulubasoglu, M., 2002. `Protection for Sale' in a Developing Country: Democracy vs. Dictatorship. Review of Economics and Statistics 84, 497-508.
-
-\item Mitra, D., Thomakos, D., Ulubasoglu, M., 2006. Can We Obtain Realistic Parameter Estimates for the Protection for Sale Model? Canadian Journal of Economics 39, 187-210.
-
-%\item Morrow, J., Siverson, R., Tabares, T., 1999. Correction to: ``The Political Determinants of International Trade,' American Political Science Review 94, 931-933.
-
-\item McCalman, P., 2004. Protection for Sale and Trade Liberalization: an Empirical Investigation. Review of International Economics 12, 81-94.
-
-%\item Paltseva, E., 2011. Protection for Sale to Oligopolists. Available at: \url{http://web.econ.ku.dk/okoep/Paltseva_P4S2O_June2011_1.pdf}.
-
-\item Saiegh, S., 2009. Political  Prowess or Lady Luck? Evaluating Chief Executives' Legislative Success Rates. The Journal of Politics 71, 1342-1356.
-
-\item Song, Y., 2008. Protection for Sale: Agenda-Setting and Ratification in the Presence of Lobbying. Korea Institute for International Trade Policy Working Paper Series. Vol. 2008-22, August 2008.
-
-%\item Tarar, A., 2001. International Bargaining with Two-Sided Domestic Constraints. Journal of Conflict Resolution 45, 320-340.
-
-%\item Ward, M., Hoff, P., 2007. Persistent Patterns of International Commerce. Journal of Peace Research 44, 157-175.
-
-\end{list}
+\tikzset{
+% Two node styles for game trees: solid and hollow
+solid node/.style={circle,draw,inner sep=1.5,fill=black},
+hollow node/.style={circle,draw,inner sep=1.5,fill=white}
+%non node/.style={circle,draw,inner sep=.1,fill=black}
+}
+\begin{tikzpicture}[scale=1.5,font=\footnotesize, grow=right]
+% Specify spacing for each level of the tree
+\tikzstyle{level 1}=[level distance=10mm,sibling distance=10mm]
+\tikzstyle{level 2}=[level distance=10mm,sibling distance=10mm]
+% The Tree
+\node(0)[solid node,label=above:{$E,E^*$}]{} 
+child{node(1)[solid node,label=below:{$0$}]{}
+}
+child{[white] node(2)[solid node,xshift=12]{}  %note that you need to adjust the yshift if you change the sibling distance
+child{[black] node(4)[solid node,label=below:{$0$}]{}
+}
+child{[white] node(5)[hollow node,xshift=12]{}  %note that you need to adjust the yshift if you change the sibling distance
+child{[black] node(7)[solid node,label=below:{$0$}]{} 
+}
+child{[white] node(8)[solid node,xshift=12]{}  %note that you need to adjust the yshift if you change the sibling distance
+child{[black] node(10)[solid node,label=below:{}]{} }
+child{[black] node(11)[solid node]{} 
+}
+child{[white] node(12)[solid node,xshift=12]{}  %note that you need to adjust the yshift if you change the sibling distance
+child{[black] node(13)[solid node,label=below:{}]{} }
+child{[black] node(14)[solid node,label=above:{$\infty$}]{} }
+}
+ edge from parent node[black, xshift=23]{$\ta$} %note that you need to adjust the yshift if you change the level distance
+}
+child{[black] node(9)[solid node,label=above:{$\infty$}]{} }
+ edge from parent node[black, xshift=23]{$e_b$} %note that you need to adjust the yshift if you change the level distance
+}
+child{[black] node(6)[solid node,label=above:{$\infty$}]{}}
+ edge from parent node[black, xshift=23]{$\bta$} %note that you need to adjust the yshift if you change the level distance
+}
+child{node(3)[solid node,label=above:{$\infty$}]{}
+}
+;
+% information set
+    \draw[solid,bend right](1)to(3);
+		\draw[solid,bend right](4)to(6);
+		\draw[solid,bend right](7)to(9);
+\end{tikzpicture}
 
 \end{document}
+
+%child{[black] node(4)[solid node,label=below:{$0$}]{} edge from parent node[left]{$C$}}
+%child{[black] node(6)[solid node,label=above:{$\infty$}]{} edge from parent node[right]{$D$}}

### 2016-01-30 game-tree.tex
add game-tree file with finished extensive form

@@
 \node(0)[solid node,label=above:{$E,E^*$}]{} 
-child{node(1)[solid node,label=below:{$0$}]{}
+child{node(1)[solid node,label=below:{$0$}]{}}
+child{[white] node(2)[solid node,xshift=12,label=above:{L}]{}  
+	child{[black] node(4)[solid node,label=below:{$0$}]{}}
+	child{[white] node(5)[hollow node,xshift=12]{} 
+		child{[black] node(7)[solid node,label=below:{$-\infty$}]{}}
+		child{[white] node(8)[solid node,xshift=12]{}  
+			child{[black] node(10)[solid node]{}  }
+			child{[black] node(12)[solid node]{}  
+				child[sibling distance=5mm]{[black] node(13)[solid node,label=below:{$0$}]{}}
+				child{[white] node(15)[solid node,xshift=7]{} 
+					child[sibling distance=5mm]{[black] node(16)[solid node,label=below:{$0$}]{} }
+					child{[white] node(18)[hollow node,xshift=12,label=left:{}]{}  }
+					child[sibling distance=5mm]{[black] node(17)[solid node,label=above:{$\infty$}]{} }
+					edge from parent node[black, xshift=20]{$e_n$} 
 					}
-child{[white] node(2)[solid node,xshift=12]{}  %note that you need to adjust the yshift if you change the sibling distance
-child{[black] node(4)[solid node,label=below:{$0$}]{}
+				child[sibling distance=5mm]{[black] node(14)[solid node,label=above:{$\infty$}]{} }
 				}
-child{[white] node(5)[hollow node,xshift=12]{}  %note that you need to adjust the yshift if you change the sibling distance
-child{[black] node(7)[solid node,label=below:{$0$}]{} 
-}
-child{[white] node(8)[solid node,xshift=12]{}  %note that you need to adjust the yshift if you change the sibling distance
-child{[black] node(10)[solid node,label=below:{}]{} }
-child{[black] node(11)[solid node]{} 
-}
-child{[white] node(12)[solid node,xshift=12]{}  %note that you need to adjust the yshift if you change the sibling distance
-child{[black] node(13)[solid node,label=below:{}]{} }
-child{[black] node(14)[solid node,label=above:{$\infty$}]{} }
-}
- edge from parent node[black, xshift=23]{$\ta$} %note that you need to adjust the yshift if you change the level distance
+			edge from parent node[black, xshift=23]{$\ta$} 
 			}
 		child{[black] node(9)[solid node,label=above:{$\infty$}]{} }
- edge from parent node[black, xshift=23]{$e_b$} %note that you need to adjust the yshift if you change the level distance
+		edge from parent node[black, xshift=23]{$e_b$} 
 		}
 	child{[black] node(6)[solid node,label=above:{$\infty$}]{}}
- edge from parent node[black, xshift=23]{$\bta$} %note that you need to adjust the yshift if you change the level distance
+	edge from parent node[black, xshift=23]{$\bta$} 
 	}
@@
 		\draw[solid,bend right](7)to(9);
+		\draw[solid,bend right](13)to(14);
+		\draw[solid,bend right](16)to(17);
+% labels
+		\node[above,xshift=5,yshift=5]at(2){$L$}; 
+		\node[above,xshift=5,yshift=5]at(5){$N$}; 
+		\node[above,xshift=8,yshift=5]at(8){$Leg$}; 
+		\node[above,xshift=5,yshift=5]at(12){$L$}; 
+		\node[above,xshift=7,yshift=5]at(15){$Leg$}; 
+		\node[left, xshift=-2]at(18){$\tau^n$}; 
+		\node[below left,yshift=-10]at(12){Break}; 
+		\node[below left,yshift=1]at(10){Abide}; 
+		\draw[solid](5)circle(.05cm);
+		\draw[solid](0)circle(.15cm);
+		\node[right]at (10) {$W_E\left(\bta\right),$};
+		\node[right,yshift=-10]at (10) {$\pi\left(\tau^a\right)-e_b,$};
+		\node[right,yshift=-20]at (10) {$W_{ML}\left(\ga(e_b,\ta_b),\bta\right)$};
+		\node[right]at (18) {$W_E\left(\btn\right),$};
+		\node[right,yshift=-10]at (18) {$\pi\left(\tau^n\right)-e_b -e_n,$};
+		\node[right,yshift=-20]at (18) {$W_{ML}\left(\ga(e_n,\ta_n),\btn\right)$};
+
 \end{tikzpicture}

### 2016-01-31 draft.tex
paste extensive form code into draft

--- 2016-01-27 draft.tex
@@
 This economic fact, combined with the assumptions on specific factor ownership, is what motivates political activity.
+
+\begin{figure}
+\begin{center}
+\tikzset{
+% Two node styles for game trees: solid and hollow
+solid node/.style={circle,draw,inner sep=1.5,fill=black},
+hollow node/.style={circle,draw,inner sep=1.5,fill=white}
+%non node/.style={circle,draw,inner sep=.1,fill=black}
+}
+\begin{tikzpicture}[scale=1.5,font=\footnotesize, grow=right]
+% Specify spacing for each level of the tree
+\tikzstyle{level 1}=[level distance=10mm,sibling distance=10mm]
+\tikzstyle{level 2}=[level distance=10mm,sibling distance=10mm]
+% The Tree
+\node(0)[solid node,label=above:{$E,E^*$}]{} 
+child{node(1)[solid node,label=below:{$0$}]{}}
+child{[white] node(2)[solid node,xshift=12,label=above:{L}]{}  
+	child{[black] node(4)[solid node,label=below:{$0$}]{}}
+	child{[white] node(5)[hollow node,xshift=12]{} 
+		child{[black] node(7)[solid node,label=below:{$-\infty$}]{}}
+		child{[white] node(8)[solid node,xshift=12]{}  
+			child{[black] node(10)[solid node]{}  }
+			child{[black] node(12)[solid node]{}  
+				child[sibling distance=5mm]{[black] node(13)[solid node,label=below:{$0$}]{}}
+				child{[white] node(15)[solid node,xshift=7]{} 
+					child[sibling distance=5mm]{[black] node(16)[solid node,label=below:{$0$}]{} }
+					child{[white] node(18)[hollow node,xshift=12,label=left:{}]{}  }
+					child[sibling distance=5mm]{[black] node(17)[solid node,label=above:{$\infty$}]{} }
+					edge from parent node[black, xshift=20]{$e_n$} 
+					}
+				child[sibling distance=5mm]{[black] node(14)[solid node,label=above:{$\infty$}]{} }
+				}
+			edge from parent node[black, xshift=23]{$\ta$} 
+			}
+		child{[black] node(9)[solid node,label=above:{$\infty$}]{} }
+		edge from parent node[black, xshift=23]{$e_b$} 
+		}
+	child{[black] node(6)[solid node,label=above:{$\infty$}]{}}
+	edge from parent node[black, xshift=23]{$\bta$} 
+	}
+child{node(3)[solid node,label=above:{$\infty$}]{}
+}
+;
+% information set
+    \draw[solid,bend right](1)to(3);
+		\draw[solid,bend right](4)to(6);
+		\draw[solid,bend right](7)to(9);
+		\draw[solid,bend right](13)to(14);
+		\draw[solid,bend right](16)to(17);
+% labels
+		\node[above,xshift=5,yshift=5]at(2){$L$}; 
+		\node[above,xshift=5,yshift=5]at(5){$N$}; 
+		\node[above,xshift=8,yshift=5]at(8){$Leg$}; 
+		\node[above,xshift=5,yshift=5]at(12){$L$}; 
+		\node[above,xshift=7,yshift=5]at(15){$Leg$}; 
+		\node[left, xshift=-2]at(18){$\tau^n$}; 
+		\node[below left,yshift=-10]at(12){Break}; 
+		\node[below left,yshift=1]at(10){Abide}; 
+		\draw[solid](5)circle(.05cm);
+		\draw[solid](0)circle(.15cm);
+		\node[right]at (10) {$W_E\left(\bta\right),$};
+		\node[right,yshift=-10]at (10) {$\pi\left(\tau^a\right)-e_b,$};
+		\node[right,yshift=-20]at (10) {$W_{ML}\left(\ga(e_b,\ta_b),\bta\right)$};
+		\node[right]at (18) {$W_E\left(\btn\right),$};
+		\node[right,yshift=-10]at (18) {$\pi\left(\tau^n\right)-e_b -e_n,$};
+		\node[right,yshift=-20]at (18) {$W_{ML}\left(\ga(e_n,\ta_n),\btn\right)$};
+\end{tikzpicture}
+
+\end{center}
+\caption{Extensive Form\label{fig:ext}}
+\end{figure}
 

### 2016-01-31 draft.tex
add subsections to sec:model
move obvious text into sec:basic
replace extensive form code with references to game-tree.tex, comment preamble from game-tree.tex

--- 2016-01-29 draft.tex
@@
 \label{sec:model}
+
+\subsection{The Basic Setup}
+\label{sec:basic}
 \textbf{I begin by describing the basic economic setting within which trade occurs.
@@
 
+\subsection{Preferences}
+\label{sec:pref}
+
+\subsection{Information and Equilibrium Selection}
+\label{sec:info}
+
+\subsection{Modeling Choices}
+\label{sec:choices}
+
 \begin{figure}
 	\begin{center}
-\begin{tikzpicture}[scale=1.5,font=\footnotesize, grow=right]
-% Specify spacing for each level of the tree
-\tikzstyle{level 1}=[level distance=10mm,sibling distance=10mm]
-\tikzstyle{level 2}=[level distance=10mm,sibling distance=10mm]
-% The Tree
-\node(0)[solid node,label=above:{$E,E^*$}]{} 
-child{node(1)[solid node,label=below:{$0$}]{}
-}
-child{[white] node(2)[solid node,xshift=12]{}  %note that you need to adjust the yshift if you change the sibling distance
-child{[black] node(4)[solid node,label=below:{$0$}]{}
-}
-child{[white] node(7)[hollow node,xshift=12]{}  %note that you need to adjust the yshift if you change the sibling distance
-child{[black] node(9)[solid node,label=below:{$0$}]{} }
-child{[black] node(11)[solid node,label=above:{$\infty$}]{} }
- edge from parent node[black, xshift=23]{$e_b$} %note that you need to adjust the yshift if you change the level distance
-}
-child{[black] node(6)[solid node,label=above:{$\infty$}]{}}
- edge from parent node[black, xshift=23]{$\bta$} %note that you need to adjust the yshift if you change the level distance
-}
-child{node(3)[solid node,label=above:{$\infty$}]{}
-}
-;
-% information set
-    \draw[solid,bend right](1)to(3);
-		\draw[solid,bend right](4)to(6);
-		\draw[solid,bend right](9)to(11);
-
-\end{tikzpicture}
-
+		\input{game-tree.tex}
 	\end{center}

### 2016-01-31 draft.tex
move text up to paragraph on timing

@@
 
+{\color{blue} I next describe the politically-relevant actors.
+In order to focus attention on protectionist political forces, I assume that only the import-competing industry in each country is politically-organized and able to lobby and that it is represented by a single lobbying organization (see Footnote~\ref{fn:lobby} for discussion of additional lobbies).
+Each country's government is composed of two branches: an executive who can conclude trade agreements and a legislature that has final say on trade policy.
+In summary, the political process is modeled as involving three players in each country: the lobby, the executive, and the legislature.}
+
+The timing is as follows.
+First, the executives set trade policy cooperatively within the context of an international agreement.
+After the trade agreement is concluded, the lobbies attempt to persuade the legislators in their respective countries to vote ``no'' on the ratification of the agreement; or alternatively, if there is no ratification hurdle, to break the trade agreement.\footnote{As will be seen shortly, these two interpretations are equivalent in this setting because tariffs in a ``trade war'' are assumed to be set non-cooperatively in the same manner as if there were no trade agreement.
+For simplicity, I refer to this as the ``break'' decision throughout.} Next, the uncertainty about the the median legislator's identity is resolved and the legislatures decide whether to abide by the agreement or to provoke a trade war.
+In the event that the trade agreement does not enter into or remain in force, there is a final stage of lobbying, resolution of the uncertainty surrounding this decision, and voting to set the trade-war tariffs.\footnote{Alternatively, one can model the lobbying for the ratification decision as determining the potential non-cooperative tariff; this changes behavior slightly depending on the relationship between the lobby and the legislature's preferences, but does not change the fundamental dynamics of the model.
+Yet another possibility is to replace the trade-war phase with the ex-ante status quo, which would most naturally be the non-cooperative equilibrium.
+This is precisely the outcome that obtains in the trade-war phase, so the results of the model are unaltered.} Once all political decisions are taken, producers and consumers make their decisions.
+
 \subsection{Preferences}
@@
 \label{sec:info}
+
+Although uncertainty is present, information about it is symmetric, so the appropriate solution concept is subgame perfect Nash equilibrium.
+As this game is solved by backward induction, it is intuitive to start by describing the incentives of the legislatures, whose decisions I model as being taken by a median legislator.
+As the economy is fully separable and the economic and political structures are symmetric, I focus here on the home country and the $X$-sector.
+The details are analogous for $Y$ and foreign.\footnote{\textbf{These are the relevant sectors for analysis because they are the sectors in which there are lobbies.}}
+
+I assume that an external authority can ensure enforcement of the agreement in the one-shot game analyzed here.\footnote{Buzard (2014b) extends the analysis to a repeated-game context and can therefore remove this restriction.}
 
@@
 \end{figure}
-
-I next describe the politically-relevant actors.
-In order to focus attention on protectionist political forces, I assume that only the import-competing industry in each country is politically-organized and able to lobby and that it is represented by a single lobbying organization (see Footnote~\ref{fn:lobby} for discussion of additional lobbies).
-Each country's government is composed of two branches: an executive who can conclude trade agreements and a legislature that has final say on trade policy.
-In summary, the political process is modeled as involving three players in each country: the lobby, the executive, and the legislature.
-
-The timing is as follows.
-First, the executives set trade policy cooperatively within the context of an international agreement.
-I assume that an external authority can ensure enforcement of the agreement in the one-shot game analyzed here.\footnote{Buzard (2014b) extends the analysis to a repeated-game context and can therefore remove this restriction.} After the trade agreement is concluded, the lobbies attempt to persuade the legislators in their respective countries to vote ``no'' on the ratification of the agreement; or alternatively, if there is no ratification hurdle, to break the trade agreement.\footnote{As will be seen shortly, these two interpretations are equivalent in this setting because tariffs in a ``trade war'' are assumed to be set non-cooperatively in the same manner as if there were no trade agreement.
-For simplicity, I refer to this as the ``break'' decision throughout.} Next, the uncertainty about the the median legislator's identity is resolved and the legislatures decide whether to abide by the agreement or to provoke a trade war.
-In the event that the trade agreement does not enter into or remain in force, there is a final stage of lobbying, resolution of the uncertainty surrounding this decision, and voting to set the trade-war tariffs.\footnote{Alternatively, one can model the lobbying for the ratification decision as determining the potential non-cooperative tariff; this changes behavior slightly depending on the relationship between the lobby and the legislature's preferences, but does not change the fundamental dynamics of the model.
-Yet another possibility is to replace the trade-war phase with the ex-ante status quo, which would most naturally be the non-cooperative equilibrium.
-This is precisely the outcome that obtains in the trade-war phase, so the results of the model are unaltered.} Once all political decisions are taken, producers and consumers make their decisions.
-
-Although uncertainty is present, information about it is symmetric, so the appropriate solution concept is subgame perfect Nash equilibrium.
-As this game is solved by backward induction, it is intuitive to start by describing the incentives of the legislatures, whose decisions I model as being taken by a median legislator.
-As the economy is fully separable and the economic and political structures are symmetric, I focus here on the home country and the $X$-sector.
-The details are analogous for $Y$ and foreign.\footnote{\textbf{These are the relevant sectors for analysis because they are the sectors in which there are lobbies.}}
 

### 2016-01-31 game-tree.tex
replace extensive form code with references to game-tree.tex, comment preamble from game-tree.tex
fix game-tree for uncertainty at trade war stage

@@
-\documentclass[10pt]{article}
+%\documentclass[10pt]{article}
 
-\addtolength{\textwidth}{1.3in}
-\addtolength{\oddsidemargin}{-.65in} %left margin
-\addtolength{\evensidemargin}{-.65in}
-\setlength{\textheight}{9in}
-\setlength{\topmargin}{-.5in}
-\setlength{\headheight}{0.0in}
-\setlength{\footskip}{.375in}
-\renewcommand{\baselinestretch}{1.0}
-%\linespread{1.5}
-\usepackage{setspace}
-\doublespacing
+%\addtolength{\textwidth}{1.3in}
+%\addtolength{\oddsidemargin}{-.65in} %left margin
+%\addtolength{\evensidemargin}{-.65in}
+%\setlength{\textheight}{9in}
+%\setlength{\topmargin}{-.5in}
+%\setlength{\headheight}{0.0in}
+%\setlength{\footskip}{.375in}
+%\renewcommand{\baselinestretch}{1.0}
+%\usepackage{setspace}
+%\doublespacing
 
-\usepackage{titling}
+%\usepackage{titling}
 
-\usepackage[pdftex]{graphicx}
+%\usepackage[pdftex]{graphicx}
 
-\usepackage[usenames,dvipsnames]{color}
-\usepackage{cite}
-\usepackage{times, verbatim,bm,pifont,pdfsync}
-\usepackage{caption}
-\usepackage{subcaption}
+%\usepackage[usenames,dvipsnames]{color}
+%\usepackage{cite}
+%\usepackage{times, verbatim,bm,pifont,pdfsync}
+%\usepackage{caption}
+%\usepackage{subcaption}
 
-\usepackage{tikz}
+%\usepackage{tikz}
 
@@
 
-\usepackage{amsbsy,amssymb, amsmath, amsthm, MnSymbol,bbding}
-\usepackage[hang,flushmargin]{footmisc}
+%\usepackage{amsbsy,amssymb, amsmath, amsthm, MnSymbol,bbding}
+%\usepackage[hang,flushmargin]{footmisc}
 
-\usepackage[pdftex,
-bookmarks=true,
-bookmarksnumbered=false,
-pdfview=fitH,
-bookmarksopen=true]{hyperref}
+%\usepackage[pdftex,
+%bookmarks=true,
+%bookmarksnumbered=false,
+%pdfview=fitH,
+%bookmarksopen=true]{hyperref}
 
-\newtheorem{definition}{Definition}
-\newtheorem{theorem}{Theorem}
-\newtheorem{lemma}{Lemma}
-\newtheorem*{lemma*}{Lemma}
-\newtheorem{corollary}{Corollary}
-\newtheorem{assumption}{Assumption}
-\newtheorem{fact}{Fact}
-\newtheorem{result}{Result}
+%\newcommand{\ta}{\theta}
+%\newcommand{\al}{\alpha}
+%\newcommand{\Ta}{\Theta}
+%\newcommand{\expect}{\mathbb{E}}
+%\newcommand{\Bt}{B(\bm{\tau^a})}
+%\newcommand{\bta}{\bm{\tau^a}}
+%\newcommand{\bte}{\bm{\tau^E}}
+%\newcommand{\btn}{\bm{\tau^n}}
+%\newcommand{\ga}{\gamma}
 
-\newcommand{\ve}{\theta}
-\newcommand{\ta}{\theta}
-\newcommand{\ov}{\overline}
-\newcommand{\un}{\underline}
-\newcommand{\al}{\alpha}
-\newcommand{\Ta}{\Theta}
-\newcommand{\expect}{\mathbb{E}}
-\newcommand{\Bt}{B(\bm{\tau^a})}
-\newcommand{\bta}{\bm{\tau^a}}
-\newcommand{\bte}{\bm{\tau^E}}
-\newcommand{\btn}{\bm{\tau^n}}
-\newcommand{\ga}{\gamma}
-
-\begin{document}
+%\begin{document}
 
@@
 				child{[white] node(15)[solid node,xshift=7]{} 
+					child[sibling distance=5mm]{[black] node(19)[solid node,label=below:{$-\infty$}]{}}
+					child{[white] node(21)[solid node,xshift=7]{} 
 						child[sibling distance=5mm]{[black] node(16)[solid node,label=below:{$0$}]{} }
@@
 						child[sibling distance=5mm]{[black] node(17)[solid node,label=above:{$\infty$}]{} }
+						edge from parent node[black, xshift=20]{$\ta_n$} 
+						}
+					child[sibling distance=5mm]{[black] node(21)[solid node,label=above:{$\infty$}]{} }
 					edge from parent node[black, xshift=20]{$e_n$} 
@@
 				}
-			edge from parent node[black, xshift=23]{$\ta$} 
+			edge from parent node[black, xshift=23]{$\ta_b$} 
 			}
@@
 		\draw[solid,bend right](16)to(17);
+		\draw[solid,bend right](19)to(21);
 % labels
@@
 		\node[above,xshift=5,yshift=5]at(12){$L$};
-		\node[above,xshift=7,yshift=5]at(15){$Leg$}; 
+		\node[above,xshift=15,yshift=-17]at(21){$Leg$}; 
+		\node[above,xshift=7,yshift=5]at(15){$N$}; 
 		\node[left, xshift=-2]at(18){$\tau^n$}; 
@@
 
-\end{document}
-
-%child{[black] node(4)[solid node,label=below:{$0$}]{} edge from parent node[left]{$C$}}
-%child{[black] node(6)[solid node,label=above:{$\infty$}]{} edge from parent node[right]{$D$}}
+%\end{document}

### 2016-01-31 draft.tex
fix game-tree for uncertainty at trade war stage
move preference defns to new sec:pref

@@
 
-The timing is as follows.
+{\color{blue} The timing is as follows.
 First, the executives set trade policy cooperatively within the context of an international agreement.
@@
 This is precisely the outcome that obtains in the trade-war phase, so the results of the model are unaltered.} Once all political decisions are taken, producers and consumers make their decisions.
-
-\subsection{Preferences}
-\label{sec:pref}
-
-\subsection{Information and Equilibrium Selection}
-\label{sec:info}
-
-Although uncertainty is present, information about it is symmetric, so the appropriate solution concept is subgame perfect Nash equilibrium.
-As this game is solved by backward induction, it is intuitive to start by describing the incentives of the legislatures, whose decisions I model as being taken by a median legislator.
-As the economy is fully separable and the economic and political structures are symmetric, I focus here on the home country and the $X$-sector.
-The details are analogous for $Y$ and foreign.\footnote{\textbf{These are the relevant sectors for analysis because they are the sectors in which there are lobbies.}}
-
-I assume that an external authority can ensure enforcement of the agreement in the one-shot game analyzed here.\footnote{Buzard (2014b) extends the analysis to a repeated-game context and can therefore remove this restriction.}
-
-\subsection{Modeling Choices}
-\label{sec:choices}
+}
 
@@
 
-In modeling the objective of the legislature, I introduce a variant of the general political support function introduced by Hillman (1982).
-While Hillman is credited with first proposing that policy makers choose tariffs in order to maximize their own political support, the PFS model contributed the explicit specification that lobbying contributions enter the political support function with constant marginal utility as $C + aW$.
-
-Dixit, Grossman and Helpman (1997) point out that linearity in contributions prevents complete analysis of distributional questions; the PFS form also restricts the returns to lobbying activity to be constant.\footnote{See Section \ref{sec:gh} for further discussion of this point and an in-depth discussion of the relationship between the functional form presented herein and the PFS form.
-Note that the assumption of diminishing returns to lobbying effort has been present in the literature going back at least to Findlay and Wellisz (1982).} They show that the PFS equilibrium analysis can be extended to a more general setting where, among other features, returns to lobbying are not restricted to be constant.
-
-For the case of trade policy, I specify the welfare function of the home legislature as
+\subsection{Preferences}
+\label{sec:pref}
+I specify the welfare function of the home legislature as
 \begin{equation}
@@
 where $\mathit{CS}$ is consumer surplus, $\mathit{PS}$ is producer surplus, $\ga(e,\ta)$ is the weight placed on producer surplus (profits) in the import-competing industry, $e$ is lobbying effort, $\ta$ is a random variable, and $\mathit{TR}$ is tariff revenue.\footnote{\textbf{Labor income $l$ could also be included in both median legislator and executive welfare.
-This alters none of the results and so would only serve to complicate the exposition.}}
-
-Notice that, aside from the endogeneity of the weight the legislature places on the lobbying industry's profits, this is precisely the \textit{deus ex machina} government objective function popularized by Baldwin (1987) that is commonly employed in the literature on the political economy of trade agreements.
-Since trade policies are overwhelmingly determined within the context of trade agreements, it is useful to have a framework to bring together the endogenous political pressure of PFS-style modeling with the trade agreements approach; the formulation in Equation~\ref{eq:ml} is intended as a bridge between the two.
-Buzard (2014a) for an example of how this form can be used to integrate the analysis of endogenous political activity and exogenous shocks to the political welfare of governments.
-
-In this formulation, the weight the median legislator places on the profits of the import-competing industry, $\ga(e,\ve)$, is affected by the level of lobbying effort and a random variable $\ta$.
-This allows for the analysis of endogenous lobbying to be integrated with the standard political economy shocks in the trade agreements literature.
-Alternatively, as motivated below, it can represent reduced-form modeling of political uncertainty arising from a complex decision-making process within a legislature.
-Modeling the objective function so closely on the standard in the trade agreements literature allows for direct comparisons to the large extant body of work that studies exogenous shocks only, revealing cleanly the effects of the addition of endogenous lobbying---and if one wants---this different kind of political uncertainty.
-
-I allow for the possibility of this political uncertainty, which can be interpreted as uncertainty over the identity of the median legislator, in order to model the idea that the lobbying and voting ``game'' that goes on within legislatures is often complex enough that none of the actors know precisely what the outcome will be---that is, no one can predict exactly which legislator will be the median and therefore what weight will be placed on import-industry profits at the time of the vote.
-Thus one way we can think about $\ta$ is as an additional influence on the legislature's preferences about which all parties---executives, legislators and lobbies alike---are uncertain.
-It is this interpretation of $\ta$ I take throughout the paper.
-
-One can conceptualize this uncertainty as being a result of the lobby's strategic choice to engage with only some key members of the legislature who wield significant influence but who cannot deliver policy decisions without the cooperation of others within the legislature, which is in turn uncertain.
-Different assumptions on $\ga(\cdot,\cdot)$ correspond to different institutional features and will affect optimal lobbying- and tariff-setting behavior.
-While a structural model of the legislative process might be more satisfying, the reduced form can deliver the key insights without distracting from the central policy choice and lobbying dynamics that take place at the level of the executives and lobbies, to which we turn next.
+This alters none of the results and so would only serve to complicate the exposition.}} The weight the median legislator places on the profits of the import-competing industry, $\ga(e,\ve)$, is affected by the level of lobbying effort and a random variable $\ta$.
 
@@
 
-I assume the lobby's contribution is not observable to the foreign legislature.
-The implication is that the lobby can directly influence only the home legislature, and so the influence of one country's lobby on the other country's legislature occurs only through the tariffs selected.\footnote{cfr.
-Grossman and Helpman (1995b), page 685.}
-
 In the first stage, the executives choose the trade agreement tariffs $\bta=\left(\tau^a,\tau^{*a} \right)$ via a negotiating process that I assume to be efficient.
@@
 This construction permits me to integrate the influence of the import-competing industry on negotiations prior to the formation of the trade agreement while also reflecting the idea that the mechanisms through which political considerations influence the executives' preferences appear to be quite different from those at the legislative level (the electoral college and primary/caucus systems for instance).
+
+\subsection{Information and Equilibrium Selection}
+\label{sec:info}
+
+Although uncertainty is present, information about it is symmetric, so the appropriate solution concept is subgame perfect Nash equilibrium.
+As this game is solved by backward induction, it is intuitive to start by describing the incentives of the legislatures, whose decisions I model as being taken by a median legislator.
+As the economy is fully separable and the economic and political structures are symmetric, I focus here on the home country and the $X$-sector.
+The details are analogous for $Y$ and foreign.\footnote{\textbf{These are the relevant sectors for analysis because they are the sectors in which there are lobbies.}}
+
+I assume that an external authority can ensure enforcement of the agreement in the one-shot game analyzed here.\footnote{Buzard (2014b) extends the analysis to a repeated-game context and can therefore remove this restriction.}
+
+I assume the lobby's contribution is not observable to the foreign legislature.
+The implication is that the lobby can directly influence only the home legislature, and so the influence of one country's lobby on the other country's legislature occurs only through the tariffs selected.\footnote{cfr.
+Grossman and Helpman (1995b), page 685.}
+
+\subsection{Modeling Choices}
+\label{sec:choices}
+
+In modeling the objective of the legislature, I introduce a variant of the general political support function introduced by Hillman (1982).
+While Hillman is credited with first proposing that policy makers choose tariffs in order to maximize their own political support, the PFS model contributed the explicit specification that lobbying contributions enter the political support function with constant marginal utility as $C + aW$.
+
+Dixit, Grossman and Helpman (1997) point out that linearity in contributions prevents complete analysis of distributional questions; the PFS form also restricts the returns to lobbying activity to be constant.\footnote{See Section \ref{sec:gh} for further discussion of this point and an in-depth discussion of the relationship between the functional form presented herein and the PFS form.
+Note that the assumption of diminishing returns to lobbying effort has been present in the literature going back at least to Findlay and Wellisz (1982).} They show that the PFS equilibrium analysis can be extended to a more general setting where, among other features, returns to lobbying are not restricted to be constant.
+
+For the case of trade policy, ...
+
+Notice that, aside from the endogeneity of the weight the legislature places on the lobbying industry's profits, this is precisely the \textit{deus ex machina} government objective function popularized by Baldwin (1987) that is commonly employed in the literature on the political economy of trade agreements.
+Since trade policies are overwhelmingly determined within the context of trade agreements, it is useful to have a framework to bring together the endogenous political pressure of PFS-style modeling with the trade agreements approach; the formulation in Equation~\ref{eq:ml} is intended as a bridge between the two.
+Buzard (2014a) for an example of how this form can be used to integrate the analysis of endogenous political activity and exogenous shocks to the political welfare of governments.
+
+the weight the median legislator places on the profits of the import-competing industry, $\ga(e,\ve)$, is affected by the level of lobbying effort and a random variable $\ta$.
+This allows for the analysis of endogenous lobbying to be integrated with the standard political economy shocks in the trade agreements literature.
+Alternatively, as motivated below, it can represent reduced-form modeling of political uncertainty arising from a complex decision-making process within a legislature.
+Modeling the objective function so closely on the standard in the trade agreements literature allows for direct comparisons to the large extant body of work that studies exogenous shocks only, revealing cleanly the effects of the addition of endogenous lobbying---and if one wants---this different kind of political uncertainty.
+
+I allow for the possibility of this political uncertainty, which can be interpreted as uncertainty over the identity of the median legislator, in order to model the idea that the lobbying and voting ``game'' that goes on within legislatures is often complex enough that none of the actors know precisely what the outcome will be---that is, no one can predict exactly which legislator will be the median and therefore what weight will be placed on import-industry profits at the time of the vote.
+Thus one way we can think about $\ta$ is as an additional influence on the legislature's preferences about which all parties---executives, legislators and lobbies alike---are uncertain.
+It is this interpretation of $\ta$ I take throughout the paper.
+
+One can conceptualize this uncertainty as being a result of the lobby's strategic choice to engage with only some key members of the legislature who wield significant influence but who cannot deliver policy decisions without the cooperation of others within the legislature, which is in turn uncertain.
+Different assumptions on $\ga(\cdot,\cdot)$ correspond to different institutional features and will affect optimal lobbying- and tariff-setting behavior.
+While a structural model of the legislative process might be more satisfying, the reduced form can deliver the key insights without distracting from the central policy choice and lobbying dynamics that take place at the level of the executives and lobbies, to which we turn next.
 

### 2016-01-31 draft.tex
proof sec:pref

@@
 where $\mathit{CS}$ is consumer surplus, $\mathit{PS}$ is producer surplus, $\ga(e,\ta)$ is the weight placed on producer surplus (profits) in the import-competing industry, $e$ is lobbying effort, $\ta$ is a random variable, and $\mathit{TR}$ is tariff revenue.\footnote{\textbf{Labor income $l$ could also be included in both median legislator and executive welfare.
-This alters none of the results and so would only serve to complicate the exposition.}} The weight the median legislator places on the profits of the import-competing industry, $\ga(e,\ve)$, is affected by the level of lobbying effort and a random variable $\ta$.
+I omit it because its inclusion alters none of the results and only serves to complicate the exposition.}} Notice that the weight the median legislator places on the profits of the import-competing industry, $\ga(e,\ve)$, is affected by the level of lobbying effort and a random variable $\ta$.
 
@@
 
-In the first stage, the executives choose the trade agreement tariffs $\bta=\left(\tau^a,\tau^{*a} \right)$ via a negotiating process that I assume to be efficient.
+In the first stage of the game, the executives choose the trade agreement tariffs $\bta=\left(\tau^a,\tau^{*a} \right)$ via a negotiating process that I assume to be efficient.
 In this symmetric environment, it therefore maximizes the expected joint payoffs:
@@
 Note that this is identical to the welfare function for the legislature aside from the weight $\ga_E$ the executive places on the profits of the import industry, which is not a function of lobbying effort.
-This construction permits me to integrate the influence of the import-competing industry on negotiations prior to the formation of the trade agreement while also reflecting the idea that the mechanisms through which political considerations influence the executives' preferences appear to be quite different from those at the legislative level (the electoral college and primary/caucus systems for instance).
 
@@
 While a structural model of the legislative process might be more satisfying, the reduced form can deliver the key insights without distracting from the central policy choice and lobbying dynamics that take place at the level of the executives and lobbies, to which we turn next.
+
+This construction (exec vs.
+leg weight on X profits) permits me to integrate the influence of the import-competing industry on negotiations prior to the formation of the trade agreement while also reflecting the idea that the mechanisms through which political considerations influence the executives' preferences appear to be quite different from those at the legislative level (the electoral college and primary/caucus systems for instance).
 

### 2016-01-31 draft.tex
re-write paragraph(s) on timing

@@
 As profits and producer surplus (identical in this model) in a sector are increasing in the price of its good, profits in the import-competing sector are also increasing in the domestic tariff.
-This economic fact, combined with the assumptions on specific factor ownership, is what motivates political activity.
-
-{\color{blue} I next describe the politically-relevant actors.
+This economic fact, combined with the assumptions on specific factor ownership, is what motivates political activity by import-competing lobbies.
+
+I now turn to describing the strategic players in the game.
 In order to focus attention on protectionist political forces, I assume that only the import-competing industry in each country is politically-organized and able to lobby and that it is represented by a single lobbying organization (see Footnote~\ref{fn:lobby} for discussion of additional lobbies).
 Each country's government is composed of two branches: an executive who can conclude trade agreements and a legislature that has final say on trade policy.
-In summary, the political process is modeled as involving three players in each country: the lobby, the executive, and the legislature.}
-
-{\color{blue} The timing is as follows.
-First, the executives set trade policy cooperatively within the context of an international agreement.
+To summarize, the political process is modeled as involving three players in each country: the lobby, the executive, and the legislature.
+
+Figure~\ref{fig:ext} illustrates the timing of the game from the perspective of the home country.
+First, the executives set trade policy cooperatively within the context of an international agreement, that is they choose the trade agreement tariffs $\bta = \left(\tau^a,\tau^{*a}\right)$.
 After the trade agreement is concluded, the lobbies attempt to persuade the legislators in their respective countries to vote ``no'' on the ratification of the agreement; or alternatively, if there is no ratification hurdle, to break the trade agreement.\footnote{As will be seen shortly, these two interpretations are equivalent in this setting because tariffs in a ``trade war'' are assumed to be set non-cooperatively in the same manner as if there were no trade agreement.
-For simplicity, I refer to this as the ``break'' decision throughout.} Next, the uncertainty about the the median legislator's identity is resolved and the legislatures decide whether to abide by the agreement or to provoke a trade war.
+For simplicity, I refer to this as the ``break'' decision throughout.} They do this by choice of lobbying effort $e_b$, which can take on any non-negative value.
+Next, the uncertainty about the the median legislator's identity is resolved.
+This is modeled as the realization of a random variable $\ta$, the value of which all players observe simultaneously.
+The ``break stage'' concludes with the legislature making a choice to abide by the agreement or to break the agreement.
+
 In the event that the trade agreement does not enter into or remain in force, there is a final stage of lobbying, resolution of the uncertainty surrounding this decision, and voting to set the trade-war tariffs.\footnote{Alternatively, one can model the lobbying for the ratification decision as determining the potential non-cooperative tariff; this changes behavior slightly depending on the relationship between the lobby and the legislature's preferences, but does not change the fundamental dynamics of the model.
@@
 This is precisely the outcome that obtains in the trade-war phase, so the results of the model are unaltered.} Once all political decisions are taken, producers and consumers make their decisions.
-}
 

### 2016-01-31 letter.tex
add first part of note on E6 to letter (had forgotten on 1/25)
add paragraph on E4 to letter

@@
 
+Editor's point $\#$3: Thank you for the suggestion to add an extensive form and to define the game more clearly and formally.
+I think the addition of the normal form in Figure 1 made this portion of the paper, in the new subsections 2.1 (The Basic Setup) and 2.2 (Preferences) significantly more readable.
+I hope you will agree.
+
 Editor's point $\#$4: I have added a traded numeraire good to address the concern about balanced trade.
@@
 
+Editor's point $\#$6: I have removed the duplicate statement of Assumption 1. [{\color{blue} much more to do here}]
+
 Editor's point $\#$7: Thank you for pointing out that the second order condition holds everywhere and that the proof of what was formerly Lemma 1 can be drastically simplified.

### 2016-02-01 draft.tex
clean up sec:choices a bit, fix ref to pe_shocks, add transition after sec:pref

@@
 Note that this is identical to the welfare function for the legislature aside from the weight $\ga_E$ the executive places on the profits of the import industry, which is not a function of lobbying effort.
+\textbf{Section~\ref{sec:choices} discusses these modeling choices in depth.}
 
@@
 
-In modeling the objective of the legislature, I introduce a variant of the general political support function introduced by Hillman (1982).
+In modeling the objective of the legislature \textbf{(see Equation~\ref{eq:ml}), I have introduced} a variant of the general political support function introduced by Hillman (1982).
 While Hillman is credited with first proposing that policy makers choose tariffs in order to maximize their own political support, the PFS model contributed the explicit specification that lobbying contributions enter the political support function with constant marginal utility as $C + aW$.
@@
 
-For the case of trade policy, ...
-
 Notice that, aside from the endogeneity of the weight the legislature places on the lobbying industry's profits, this is precisely the \textit{deus ex machina} government objective function popularized by Baldwin (1987) that is commonly employed in the literature on the political economy of trade agreements.
 Since trade policies are overwhelmingly determined within the context of trade agreements, it is useful to have a framework to bring together the endogenous political pressure of PFS-style modeling with the trade agreements approach; the formulation in Equation~\ref{eq:ml} is intended as a bridge between the two.
-Buzard (2014a) for an example of how this form can be used to integrate the analysis of endogenous political activity and exogenous shocks to the political welfare of governments.
+Buzard (2016) for an example of how this form can be used to integrate the analysis of endogenous political activity and exogenous shocks to the political welfare of governments.
 
@@
 
-\item Buzard, K., 2014a. Endogenous Politics and the Design of Trade Agreements. Available at: \url{https://kbuzard.expressions.syr.edu/wp-content/uploads/Endogenous-Politics.pdf}.
+\item Buzard, K., 2016. Endogenous Politics and the Design of Trade Agreements. Available at: \url{https://kbuzard.expressions.syr.edu/wp-content/uploads/Endogenous-Politics.pdf}.
 

### 2016-02-01 draft.tex
fix ref to sop_repeated

@@
 
-I assume that an external authority can ensure enforcement of the agreement in the one-shot game analyzed here.\footnote{Buzard (2014b) extends the analysis to a repeated-game context and can therefore remove this restriction.}
+I assume that an external authority can ensure enforcement of the agreement in the one-shot game analyzed here.\footnote{Buzard (2015) extends the analysis to a repeated-game context and can therefore remove this restriction.}
 
@@
 
+\item Buzard, K., 2015. Self-enforcing Trade Agreements, Dispute Settlement and Separation of Powers. Available at: \url{http://papers.ssrn.com/sol3/papers.cfm?abstract_id=2333728}.
+
 \item Buzard, K., 2016. Endogenous Politics and the Design of Trade Agreements. Available at: \url{https://kbuzard.expressions.syr.edu/wp-content/uploads/Endogenous-Politics.pdf}.
-
-\item Buzard, K., 2014b.
-Self-enforcing Trade Agreements, Dispute Settlement and Separation of Powers.
-Available at: \url{https://kbuzard.expressions.syr.edu/wp-content/uploads/2014/04/Self-enforcing-Trade_Agreements.pdf}.
 

### 2016-02-01 draft.tex
rewrite leg part of sec:choices

@@
 
-Notice that, aside from the endogeneity of the weight the legislature places on the lobbying industry's profits, this is precisely the \textit{deus ex machina} government objective function popularized by Baldwin (1987) that is commonly employed in the literature on the political economy of trade agreements.
-Since trade policies are overwhelmingly determined within the context of trade agreements, it is useful to have a framework to bring together the endogenous political pressure of PFS-style modeling with the trade agreements approach; the formulation in Equation~\ref{eq:ml} is intended as a bridge between the two.
-Buzard (2016) for an example of how this form can be used to integrate the analysis of endogenous political activity and exogenous shocks to the political welfare of governments.
-
-the weight the median legislator places on the profits of the import-competing industry, $\ga(e,\ve)$, is affected by the level of lobbying effort and a random variable $\ta$.
-This allows for the analysis of endogenous lobbying to be integrated with the standard political economy shocks in the trade agreements literature.
-Alternatively, as motivated below, it can represent reduced-form modeling of political uncertainty arising from a complex decision-making process within a legislature.
+Notice that, aside from the endogeneity of the weight the legislature places on the lobbying industry's profits, Equation~\ref{eq:ml} is precisely the \textit{deus ex machina} government objective function popularized by Baldwin (1987) that is commonly employed in the literature on the political economy of trade agreements.
+Instead of the weight on profits being constant as in the exogenous case or necessarily linear in contributions as in the PFS case, this form allows for both endogeneity and non-linearity.
+
+Since trade policies are overwhelmingly determined within the context of trade agreements, it is useful to have a framework to bring together the endogenous political pressure of PFS-style modeling with the trade agreements approach.
+The formulation in Equation~\ref{eq:ml} is intended as a bridge between the two.
+See Buzard (2016) for an example of how this form can be used to integrate the analysis of endogenous political activity and exogenous shocks to the political welfare of governments.
+
+In addition, letting the weight the median legislator places on the profits of the import-competing industry $\ga(e,\ve)$ be affected by the level of lobbying effort as well a random variable $\ta$ allows for the analysis of endogenous lobbying to be integrated with the standard political economy shocks of the trade agreements literature.
 Modeling the objective function so closely on the standard in the trade agreements literature allows for direct comparisons to the large extant body of work that studies exogenous shocks only, revealing cleanly the effects of the addition of endogenous lobbying---and if one wants---this different kind of political uncertainty.
 
-I allow for the possibility of this political uncertainty, which can be interpreted as uncertainty over the identity of the median legislator, in order to model the idea that the lobbying and voting ``game'' that goes on within legislatures is often complex enough that none of the actors know precisely what the outcome will be---that is, no one can predict exactly which legislator will be the median and therefore what weight will be placed on import-industry profits at the time of the vote.
+Alternatively, the integration of the random variable $\ta$ can represent reduced-form modeling of political uncertainty arising from a complex decision-making process within a legislature.
+One can interpret $\ta$ as measuring uncertainty over the identity of the median legislator in order to represent the idea that the lobbying and voting ``game'' that goes on within legislatures is often complex enough that none of the actors know precisely what the outcome will be---that is, no one can predict exactly which legislator will be the median and therefore what weight will be placed on import-industry profits at the time of the vote.
 Thus one way we can think about $\ta$ is as an additional influence on the legislature's preferences about which all parties---executives, legislators and lobbies alike---are uncertain.
-It is this interpretation of $\ta$ I take throughout the paper.
+It is this interpretation of $\ta$ I take throughout this paper.
 

### 2016-02-01 draft.tex
rewrite sec:info

@@
 \label{sec:pref}
+As the economy is fully separable and the economic and political structures are symmetric, I focus throughout the paper on the home country and the $X$-sector.
+The details are analogous for $Y$ and foreign.\footnote{\textbf{These are the relevant sectors for analysis because they are the sectors in which there are lobbies.}}
+
 I specify the welfare function of the home legislature as
@@
 where $\mathit{CS}$ is consumer surplus, $\mathit{PS}$ is producer surplus, $\ga(e,\ta)$ is the weight placed on producer surplus (profits) in the import-competing industry, $e$ is lobbying effort, $\ta$ is a random variable, and $\mathit{TR}$ is tariff revenue.\footnote{\textbf{Labor income $l$ could also be included in both median legislator and executive welfare.
-I omit it because its inclusion alters none of the results and only serves to complicate the exposition.}} Notice that the weight the median legislator places on the profits of the import-competing industry, $\ga(e,\ve)$, is affected by the level of lobbying effort and a random variable $\ta$.
+I omit it because its inclusion alters none of the results and only serves to complicate the exposition.}} I model the decisions of the legislature as being taken by a median legislator (more on this in Section~\ref{sec:choices}.
+Notice that the weight the median legislator places on the profits of the import-competing industry, $\ga(e,\ve)$, is affected by the level of lobbying effort and a random variable $\ta$.
 
@@
 \label{sec:info}
-
-Although uncertainty is present, information about it is symmetric, so the appropriate solution concept is subgame perfect Nash equilibrium.
-As this game is solved by backward induction, it is intuitive to start by describing the incentives of the legislatures, whose decisions I model as being taken by a median legislator.
-As the economy is fully separable and the economic and political structures are symmetric, I focus here on the home country and the $X$-sector.
-The details are analogous for $Y$ and foreign.\footnote{\textbf{These are the relevant sectors for analysis because they are the sectors in which there are lobbies.}}
-
-I assume that an external authority can ensure enforcement of the agreement in the one-shot game analyzed here.\footnote{Buzard (2015) extends the analysis to a repeated-game context and can therefore remove this restriction.}
-
-I assume the lobby's contribution is not observable to the foreign legislature.
+I examine a simple class of equilibria that have the following features.
+
+First, on information.
+Although uncertainty is present, information about it is symmetric.
+However, in line with the literature, I assume the lobby's contribution is not observable to the foreign legislature.
 The implication is that the lobby can directly influence only the home legislature, and so the influence of one country's lobby on the other country's legislature occurs only through the tariffs selected.\footnote{cfr.
-Grossman and Helpman (1995b), page 685.}
+Grossman and Helpman (1995b), page 685.} Given the game's structure and the assumption that lobbying effort is not observable across international borders, players in the same country can take advantage of more information than those who are in different countries.
+That is, the solution concept employed here is perfect public equilibrium (PPE).
+
+Second, whenever there is a possibility of multiple equilibria, I will focus on the one that maximizes the welfare of the executives.
+
+Finally, I assume that an external authority can ensure enforcement of the agreement in the one-shot game analyzed here.\footnote{Buzard (2015) extends the analysis to a repeated-game context and can therefore remove this restriction.}
 

### 2016-02-01 letter.tex
add paragraph to letter on E2; add more to E3 para

@@
 
+Editor's point $\#$2: I have created four subsections within Section 2 to structure the exposition of the model. I think this is much clearer than the previous draft.
+
 Editor's point $\#$3: Thank you for the suggestion to add an extensive form and to define the game more clearly and formally.
-I think the addition of the normal form in Figure 1 made this portion of the paper, in the new subsections 2.1 (The Basic Setup) and 2.2 (Preferences) significantly more readable.
+I think the addition of the normal form in Figure 1 made this portion of the paper significantly more readable.
 I hope you will agree.
+The discussion of the players, timing, and the presentation of the extensive form are at the end of Section 2.1 (The Basic Setup).
+I largely allow the extensive form to do the work of laying out the strategy spaces so as not to clutter the text.
+The payoffs are gathered together in their own section (2.2: Preferences).
 

### 2016-02-01 game-tree.tex
fix game-tree for opportunity to break

--- 2016-01-30 game-tree.tex
@@
 \renewcommand{\baselinestretch}{1.0}
-%\linespread{1.5}
 \usepackage{setspace}
@@
 % disables chapter, section and subsection numbering
-%\setcounter{secnumdepth}{-1}
+\setcounter{secnumdepth}{-1}
 
@@
 
-\newtheorem{definition}{Definition}
-\newtheorem{theorem}{Theorem}
-\newtheorem{lemma}{Lemma}
-\newtheorem*{lemma*}{Lemma}
-\newtheorem{corollary}{Corollary}
-\newtheorem{assumption}{Assumption}
-\newtheorem{fact}{Fact}
-\newtheorem{result}{Result}
-
-\newcommand{\ve}{\theta}
 \newcommand{\ta}{\theta}
-\newcommand{\ov}{\overline}
-\newcommand{\un}{\underline}
 \newcommand{\al}{\alpha}
@@
 % Specify spacing for each level of the tree
-\tikzstyle{level 1}=[level distance=10mm,sibling distance=10mm]
-\tikzstyle{level 2}=[level distance=10mm,sibling distance=10mm]
+\tikzstyle{level 1}=[level distance=10mm,sibling distance=8mm]
+\tikzstyle{level 2}=[level distance=8mm,sibling distance=8mm]
 % The Tree
@@
 	child{[white] node(5)[hollow node,xshift=12]{} 
-		child{[black] node(7)[solid node,label=below:{$-\infty$}]{}}
-		child{[white] node(8)[solid node,xshift=12]{}  
-			child{[black] node(10)[solid node]{}  }
-			child{[black] node(12)[solid node]{}  
+	 child[sibling distance=11mm]{[black] node(22)[solid node]{}  }
+	 child[sibling distance=11mm]{[black] node(23)[hollow node]{}  
+		child[sibling distance=8mm]{[black] node(7)[solid node,label=below:{$-\infty$}]{}}
+		child[sibling distance=8mm]{[white] node(8)[solid node,xshift=12]{}  
+			child[sibling distance=10mm]{[black] node(10)[solid node]{}  }
+			child[sibling distance=10mm]{[black] node(12)[solid node]{}  
 				child[sibling distance=5mm]{[black] node(13)[solid node,label=below:{$0$}]{}}
 				child{[white] node(15)[solid node,xshift=7]{} 
+					child[sibling distance=5mm]{[black] node(19)[solid node,label=below:{$-\infty$}]{}}
+					child{[white] node(21)[solid node,xshift=7]{} 
 						child[sibling distance=5mm]{[black] node(16)[solid node,label=below:{$0$}]{} }
@@
 						child[sibling distance=5mm]{[black] node(17)[solid node,label=above:{$\infty$}]{} }
-					edge from parent node[black, xshift=20]{$e_n$} 
+						edge from parent node[black, xshift=16]{$\ta_n$} 
+						}
+					child[sibling distance=5mm]{[black] node(21)[solid node,label=above:{$\infty$}]{} }
+					edge from parent node[black, xshift=15]{$e_n$} 
 				}
@@
 				}
-			edge from parent node[black, xshift=23]{$\ta$} 
+			edge from parent node[black, xshift=20]{$\ta_b$} 
 			}
-		child{[black] node(9)[solid node,label=above:{$\infty$}]{} }
-		edge from parent node[black, xshift=23]{$e_b$} 
+		child[sibling distance=8mm]{[black] node(9)[solid node,label=above:{$\infty$}]{} }
+		}
+	 edge from parent node[black, xshift=20]{$e_b$} 	
 	 }
@@
 		\draw[solid,bend right](16)to(17);
+		\draw[solid,bend right](19)to(21);
 % labels
-		\node[above,xshift=5,yshift=5]at(2){$L$}; 
-		\node[above,xshift=5,yshift=5]at(5){$N$}; 
-		\node[above,xshift=8,yshift=5]at(8){$Leg$}; 
+		\node[above,xshift=3,yshift=5]at(2){$L$}; 
+		\node[above,xshift=1,yshift=5]at(23){$N$}; 
+		\node[above,xshift=3,yshift=5]at(5){$N$}; 
+		\node[above,xshift=6,yshift=5]at(8){$Leg$}; 
 		\node[above,xshift=5,yshift=5]at(12){$L$};
-		\node[above,xshift=7,yshift=5]at(15){$Leg$}; 
+		\node[above,xshift=13,yshift=-17]at(21){$Leg$}; 
+		\node[above,xshift=7,yshift=5]at(15){$N$}; 
 		\node[left, xshift=-2]at(18){$\tau^n$}; 
-		\node[below left,yshift=-10]at(12){Break}; 
-		\node[below left,yshift=1]at(10){Abide}; 
+		\node[below left,yshift=-12]at(12){Break}; 
+		\node[below left,yshift=4]at(10){Abide}; 
 		\draw[solid](5)circle(.05cm);
@@
 		\node[right,yshift=-20]at (18) {$W_{ML}\left(\ga(e_n,\ta_n),\btn\right)$};
-
+		\node[right,yshift=-5]at (22) {$W_E\left(\bta\right),$};
+		\node[right,yshift=-15]at (22) {$\pi\left(\tau^a\right)-e_b,$};
+		\node[right,yshift=-25]at (22) {$W_{ML}\left(\ga(e_b,\ta_b),\bta\right)$};
+		\node[below left,xshift=-4,yshift=8]at(23){Opp.}; 
+		\node[below left,xshift=-4,yshift=-10]at(23){(o)}; 
+		\node[below left,xshift=-3,yshift=7]at(22){No opp.}; 
+		\node[below left,xshift=6,yshift=20]at(22){(1-o)}; 
 \end{tikzpicture}
@@
 \end{document}
-
-%child{[black] node(4)[solid node,label=below:{$0$}]{} edge from parent node[left]{$C$}}
-%child{[black] node(6)[solid node,label=above:{$\infty$}]{} edge from parent node[right]{$D$}}

### 2016-02-01 draft.tex
compile draft with updated game-tree
revise text for straightforward presentation of one-sided break

@@
 
-Figure~\ref{fig:ext} illustrates the timing of the game from the perspective of the home country.
-First, the executives set trade policy cooperatively within the context of an international agreement, that is they choose the trade agreement tariffs $\bta = \left(\tau^a,\tau^{*a}\right)$.
+Figure~\ref{fig:ext} illustrates the timing of the game from the perspective of the home country.\footnote{\textbf{This extensive form is for the case where the foreign legislature does not break the trade agreement.}} First, the executives set trade policy cooperatively within the context of an international agreement, that is they choose the trade agreement tariffs $\bta = \left(\tau^a,\tau^{*a}\right)$.
 After the trade agreement is concluded, the lobbies attempt to persuade the legislators in their respective countries to vote ``no'' on the ratification of the agreement; or alternatively, if there is no ratification hurdle, to break the trade agreement.\footnote{As will be seen shortly, these two interpretations are equivalent in this setting because tariffs in a ``trade war'' are assumed to be set non-cooperatively in the same manner as if there were no trade agreement.
 For simplicity, I refer to this as the ``break'' decision throughout.} They do this by choice of lobbying effort $e_b$, which can take on any non-negative value.
-Next, the uncertainty about the the median legislator's identity is resolved.
+
+Next, nature determines with probabilities $o$ and $o^*$ whether each legislature will have the opportunity to take a vote to break the agreement.
+I assume that these probabilities are less than one-half and mutually exclusive.\footnote{\textbf{While this assumption is made for purposes of exposition and tractability, one can think of $o$ and $o^*$ as being determined by events beyond the executives' or legislatures' control that determine whether or not the legislatures will be willing to consider the lobby's request; for example, they would be affected by the occurrence of an economic crisis that diverts the legislature's attention from less-pressing matters.}} This allows me to focus on the key interaction between the domestic actors and abstract from the strategic interaction between the lobbies in the two countries.\footnote{\textbf{An incentive for each lobby to free ride on the effort of the other makes the optimization problem of the executives more complex as they are faced with the problem of how best to exploit the free-riding dynamic.}}
+
+After this, the uncertainty about the the median legislator's identity is resolved.
 This is modeled as the realization of a random variable $\ta$, the value of which all players observe simultaneously.
@@
 The realization of uncertainty about the median legislator's identity will similarly be denoted $\ve_b$ so as to distinguish it from $\ve_n$ at the trade-war stage.
-
-I also assume that each legislature will have the opportunity to take a vote to break the agreement with probability less than one-half, and that the probability that both legislatures have the opportunity is vanishingly small.
-This allows me to focus here on the key interaction between the domestic actors and abstract from the strategic interaction between the lobbies in the two countries.\footnote{\textbf{An incentive for each lobby to free ride on the effort of the other makes the optimization problem of the executives more complex as they are faced with the problem of how best to exploit the free-riding dynamic.}}
 
@@
 Here $o$ and $o^*$ are the probabilities that the home and foreign legislatures, respectively, will have the opportunity to break the trade agreement; these are assumed to be mutually exclusive events that are realized after the conclusion of the trade agreement.
-One can think of $o$ and $o^*$ as being determined by events beyond the executives' or legislatures' control that determine whether or not the legislatures will be willing to consider the lobby's request; for example, they would be affected by the occurrence of an economic crisis that diverts the legislature's attention from less-pressing matters.
 

### 2016-02-01 letter.tex
add paragraph on one-sided break (E9) to letter

@@
 
+Editor's point $\#9$: I have added the move of nature that determines whether the legislature gets the opportunity to the extensive form and to the discussion of the timing in the initial discussion of the model (Section 2.1: The Basic Setup).
+I have gathered all relevant discussion in this section as suggested.
+
 Editor's point $\#10$: I have removed the extensions (Section 6) per your suggestion. {\color{blue}ADD discussion about fixing exposition around one break at a time}

### 2016-02-01 draft.tex
add e to as:ga_l_e and adjust text, part of #6

@@
 \begin{assumption}
-  $\ga(e,\ve) \geq \ga_E \geq 1 \ \forall \ve$.
+  $\ga(e,\ta) \geq \ga_E \geq 1 \ \forall \bm{e},\ta$.
   \label{as:ga_l_e}
@@
 
-That is, even for the least favorable outcome of the lobbying process, the legislature will be at least slightly more protectionist than the executive.
+That is, even for the least favorable outcome of the lobbying process, the legislature will be at least slightly more protectionist than the executive \textbf{regardless of the lobby's effort choice}.
 This assumption is not essential, but it matches well the empirical findings that politicians with larger constituencies are less sensitive to special interests, and it simplifies the analysis considerably, in particular ensuring that trade agreement tariffs are less than trade war tariffs \textbf{(See Destler (2005) and footnote~\ref{fn:ga_l_e} above)}.
@@
 \subsection{To Break or Not to Break?}
+\label{sec:break}
 With trade-war decisions specified, we can proceed to analyze the incentives regarding the legislature's decision to uphold or break the trade agreement.

### 2016-02-01 draft.tex
rewrite paragraphs on exec / lobbying / PFS (#6)

@@
 
-This construction (exec vs.
-leg weight on X profits) permits me to integrate the influence of the import-competing industry on negotiations prior to the formation of the trade agreement while also reflecting the idea that the mechanisms through which political considerations influence the executives' preferences appear to be quite different from those at the legislative level (the electoral college and primary/caucus systems for instance).
-
-This does \textit{not} require that the executives are not lobbied; only that their preferences are not directly altered in a significant way by lobbying over trade---that they do not sell protection in order to finance their re-election campaigns.
+The executives' objective function is taken to be identical to the legislatures' except that profits in the import-competing industry receive a fixed, exogenous weight.
+This construction permits me to integrate the influence of the import-competing industry on negotiations prior to the formation of the trade agreement while also reflecting the idea that the mechanisms through which political considerations influence the executives' preferences appear to be quite different from those at the legislative level---the electoral college and primary/caucus systems for instance.
+
+This setup does not strictly require that the executives are not lobbied.
+The assumption that the weight the executives place on import-competing profits is constant only requires that the extent to which they favor this industry cannot be \textit{directly} influenced by lobbying effort.
+The assumption is innocuous as long as the executives' preferences are not directly altered in a significant way by lobbying over trade.
+
 In the case of the post-war United States, where the Congress has consistently been significantly more protectionist than the President, this seems to reasonably reflect the political reality.

### 2016-02-01 letter.tex
add paragraph on E6 to letter

@@
 
-Editor's point $\#$6: I have removed the duplicate statement of Assumption 1. [{\color{blue} much more to do here}]
+Editor's point $\#$6: I have rewritten the passage about the influence of lobbying on the executive to read
+\begin{quote}
+  This setup does not strictly require that the executives are not lobbied.
+The assumption that the weight the executives place on import-competing profits is constant only requires that the extent to which they favor this industry cannot be \textit{directly} influenced by lobbying effort.
+The assumption is innocuous as long as the executives' preferences are not directly altered in a significant way by lobbying over trade.
+\end{quote}
+I think this is a softer claim than what was in the earlier draft and avoids getting into an interpretation of exactly what `Protection for Sale' means in this context since this is not central---what is important is to defend the modeling assumption directly.
+
+I have also removed the duplicate statement of Assumption 1 and combined redundant text surrounding the two statements. I have added $\forall e$ to the statement of Assumption 1.
 

### 2016-02-01 draft.tex
add paragraph on applicability to parliamentary regimes #10

@@
 Although the political process here most closely matches that of the United States in the post-war era,\footnote{In particular, the binary decision by the legislature about whether to abide by or break the trade agreement is modeled on the ``Fast Track Authority'' that the U.S.
-Congress granted to the Executive branch almost continuously from 1974-1994 and then again as ``Trade Promotion Authority'' from 2002-2007.} the model is broadly applicable because the central dynamic that emerges---that trade agreements are formed in the \textit{shadow} of lobbying pressure instead of in direct response to it---appears to be present in a wide range of countries in which the formation and maintenance of trade policy requires ratification by a politically susceptible body either at home or in a trading partner.
+Congress granted to the Executive branch almost continuously from 1974-1994 and then again as ``Trade Promotion Authority'' from 2002-2007.} the model is broadly applicable because the central dynamic that emerges---that trade agreements are formed in the \textit{shadow} of lobbying pressure instead of in direct response to it---appears to be present in a wide range of countries in which the formation and maintenance of trade policy requires \textit{approval} by a politically susceptible body either at home or in a trading partner.
+
+The results apply beyond separation-of-powers presidential systems because the presence of veto points is what is essential for the dynamic posited here to occur.
+Although the number of formal veto points tends to be lower in, for instance, parliamentary systems, veto points are typically not absent altogether (Henisz and Mansfield 2006).
+In reality, constraints on the executive are almost always present in democratic regimes, although they may operate in different ways, e.g.
+over questions of government dissolution and formation in parliamentary regimes.
+It is important to consider that constraints may be formal or informal, and they may come from legislative, regulatory, judicial, sub-national or even other executive bodies.
 

### 2016-02-01 letter.tex
add paragraph for referee concern about parliamentary regimes

@@
 
+Referee's point $\#1$ (``Introduction is derailing at the start''):
+
+Referee's point $\#2$ (``Intuition in the Introduction about political uncertainty''):
+
+Referee's point $\#3$ (``Parliamentary democracies''): I have added a paragraph at the very end of Section 2 to address this concern.
+I would also note, as a point of interest, that the number of veto points does not always correlate particularly well with the political system.
+I was not easily able to find empirical work on the number of veto points on trade issues for Turkey, but Sparkes, Bump and Reich (2015) were somewhat surprisingly able to identify \textit{seven} veto points for health issues in Turkey.\footnote{Sparkes, S.P., Bump, J.
+B.
+and M.
+R.
+Reich (2015), ``Political Strategies for Health Reform in Turkey: Extending Veto Point Theory,'' \textit{Health Systems and Reform}, 1(4), 263-275.}
+
+Referee's point $\#4$ (``Strong language regarding contribution''):
+
 \end{document}

### 2016-02-01 draft.tex
fix typos in sec:main, remove assumption 2 content from res:lobby

@@
 Next, nature determines with probabilities $o$ and $o^*$ whether each legislature will have the opportunity to take a vote to break the agreement.
-I assume that these probabilities are less than one-half and mutually exclusive.\footnote{\textbf{While this assumption is made for purposes of exposition and tractability, one can think of $o$ and $o^*$ as being determined by events beyond the executives' or legislatures' control that determine whether or not the legislatures will be willing to consider the lobby's request; for example, they would be affected by the occurrence of an economic crisis that diverts the legislature's attention from less-pressing matters.}} This allows me to focus on the key interaction between the domestic actors and abstract from the strategic interaction between the lobbies in the two countries.\footnote{\textbf{An incentive for each lobby to free ride on the effort of the other makes the optimization problem of the executives more complex as they are faced with the problem of how best to exploit the free-riding dynamic.}}
+I assume that these probabilities are less than one-half and mutually exclusive.\footnote{\textbf{While this assumption is made for purposes of exposition and tractability, one can think of $o$ and $o^*$ as being determined by events beyond the executives' or legislatures' control that determine whether or not the legislatures will be willing to consider the lobby's request; for example, they would be affected by the occurrence of an economic crisis that diverts the legislature's attention from less-pressing matters.}} This allows me to focus on the key interaction between the domestic actors and abstract from the strategic interaction between the lobbies in the two countries.\footnote{\textbf{When the legislatures in both countries have an opportunity to break the agreement simultaneously, an incentive for each lobby to free ride on the effort of the other makes the optimization problem of the executives more complex as they are faced with the problem of how best to exploit the free-riding dynamic.}}
 
@@
 With trade-war decisions specified, we can proceed to analyze the incentives regarding the legislature's decision to uphold or break the trade agreement.
-I assume that the form of the political weighting function is the same in both cases, but the model could be extended to allow the two to differ.
+I assume that the form of the political weighting function is the same \textbf{regardless of the break decision}, but the model could be extended to allow the two to differ.
 What is in principle different are the lobbying efforts toward influencing the break decision, denoted $e_b$ at this stage to distinguish them from those that are made to influence the Nash tariff, $e_n$.
@@
 
-The legislature's decisions are stochastic, so the outcome of the vote on whether or not to break the trade agreement, as well as the level at which the tariff will be set in the case of a disruption, is not known to \textit{any} player until the uncertainty over the identity of the median legislator is resolved at each stage---that is, until the moment the vote takes place.\footnote{We might think of the trade-war tariff as being set in the legislation before a vote takes place in the case of administered protection for instance; the results are robust to this variant.} I represent the probability that the home legislature breaks the trade agreement and sets the tariff at $\tau^n$ as:
+The legislature's decisions are stochastic, so the outcome of the vote on whether or not to break the trade agreement, as well as the level at which the tariff will be set in the case of a disruption, is not known to \textit{any} player until the uncertainty over the identity of the median legislator is resolved at the moment the vote takes place on each issue.\footnote{We might think of the trade-war tariff as being set in the legislation before a vote takes place in the case of administered protection for instance; the results are robust to this variant.} I represent the probability that the home legislature breaks the trade agreement and sets the tariff at $\tau^n$ as:
 \begin{multline}
@@
 
-Note that the legislature's best response functions (both for whether to break the agreement, and at what level to set the tariff if it does) will be affected by the shape of $\ga(e,\ve)$.
+Note that the legislature's best response functions---both for whether to break the agreement, and at what level to set the tariff if it does---will be affected by the shape of $\ga(e,\ve)$.
 I make the following further assumptions on the weight on import-industry profits:
@@
 I will henceforth suppress the Nash tariffs in the expression of the break probability to simplify notation.\footnote{Since the decisions over the Nash tariffs are taken in the final period, they do not vary from the point of view of earlier stages.
-Similarly, because the lobbying effort made in that period does not interact with other decision variables in earlier periods, I will suppress the dependence of the Nash tariffs on $e_n$.} Recalling that if the lobby has the chance to act, there is no chance for the foreign legislature to break the agreement, the lobby's decision is
+Similarly, because the lobbying effort made in that period does not interact with other decision variables in earlier periods, I suppress the dependence of the Nash tariffs on $e_n$.} Recalling that if the lobby has the chance to act, there is no chance for the foreign legislature to break the agreement, the lobby's decision is
 \[
@@
 \begin{result}
-  When the trade agreement remains in force with positive probability and the distribution of $\ga(e,\ta)$ is weakly increasing in $e$, lobbying effort is weakly decreasing in trade agreement tariffs.\footnote{See Section~\ref{sec:uncertainty} for a detailed account of behavior when the lobby is able to induce a trade war with probability 1.}
+  When the trade agreement remains in force with positive probability, lobbying effort is weakly decreasing in trade agreement tariffs.\footnote{See Section~\ref{sec:uncertainty} for a detailed account of behavior when the lobby is able to induce a trade war with probability 1.}
   \label{res:lobby}
@@
 Note that the symmetric increase in the foreign tariff does not impact profits in the import-competing industry, but it can impact lobbying incentives through the break probability.
-However, there is an analogous effect on the break probability from raising $\tau^a$ and the net effect is negative given the assumption on the probability distribution function of $\ga(e,\ta)$, which ensures that the effect of lobbying effort on the break probability does not increase as $\bta$ increases.
-This assumption can be significantly weakened: this effect cd be positive as long as it is smaller in absolute value than the negative effect on trade agreement profits.
+However, there is an analogous effect on the break probability from raising $\tau^a$ and the net effect is negative given Assumption~\ref{as:ga_c} on the probability distribution function of $\ga(e,\ta)$, which ensures that the effect of lobbying effort on the break probability does not increase as $\bta$ increases.
+This assumption can be significantly weakened: this effect can be positive as long as it is smaller in absolute value than the negative effect on trade agreement profits.
 

### 2016-02-02 draft.tex
capitalized a reference title
write two new opening paragraphs for intro

@@
 
-Empirical investigations of Grossman and Helpman's (1994) `Protection for Sale' model (henceforth `PFS') broadly support the predictions of the theory qualitatively, yet their quantitative estimates present several interesting puzzles.\footnote{cfr.
+Much of trade policy is determined in the context of international agreements that must be ratified by some domestic body.
+This paper seeks to shed light on such a policy-making process by incorporating Protection-for-Sale (Grossman and Helpman 1994, hereafter PFS) style lobbying into a model of trade agreements in which the executive branches of the governments set tariff levels in anticipation of political pressure upon a ratifying body, assumed here to be a legislature.
+This political pressure may be interpreted either as lobbying against ratification of the trade agreement or lobbying to break the agreement once in force.
+Important realism is achieved not only through the addition of the ratification constraint but also by the incorporation of political uncertainty into the model.
+
+Assuming for now that lobbies represent only import-competing interests, lobbies in this environment would like to prevent or derail a trade agreement that would reduce their tariff protection.
+They work toward this end by exerting effort to convince their legislatures to vote against the trade agreement but face some uncertainty about whether their efforts will result in a voting outcome consistent with their interests.
+I show that the probability that the trade agreement is derailed increases in lobbying effort and decreases in the trade agreement tariff.
+Further, the lobby decreases its effort as the tariff it would face under the trade agreement rises: it has less to lose the higher is the trade agreement tariff.
+When forming the trade agreement tariff, the executives trade off the welfare they receive under the agreement with the probability that the agreement actually remains in force.
+Ultimately, I find that observed lobbying effort may often be low because the executive branches of the governments set trade agreement tariffs sufficiently high to reduce political pressure on the legislatures.
+
+\vskip.2in
+I find that the separation-of-powers government structure can explain the empirical puzzle surrounding governments' welfare weights and that the addition of political uncertainty adds important realism so that the model presented here can speak to important trade policy-making events such as the four-year delay in the ratification of the U.S.-Korea Free Trade Agreement.
+
+Empirical investigations of Grossman and Helpman's (1994) PFS model broadly support the predictions of the theory qualitatively, yet their quantitative estimates present several interesting puzzles.\footnote{cfr.
 Goldberg and Maggi (1999); Gawande and Bandyopadhyay (2000); Mitra, Thomakos and Ulubasoglu (2002); McCalman (2004).\label{fn:gh_lobby}} These studies have consistently found the weight governments place on social welfare to be many times that which they place on lobbying effort, while estimates indicate that the deadweight losses caused by trade distortions are several orders of magnitude larger than lobbying expenditures.\footnote{Feenstra (1992) assembles estimates from the mid-1980s that place a floor of $\$$8 billion per year on the deadweight losses from protection, whereas Bombardini and Trebbi (2012) find that total annual lobbying expenditure on trade issues in 1999-2001 when this data first became available was about $\$$200 million.} This raises the question: how could governments value social welfare so highly, yet grant these quantities of protection at such a low price?
@@
 Gawande, Krishna and Robbins (2006); Mitra, Thomakos and Ulubasoglu (2006); Bombardini (2008); and Gawande, Krishna and Olarreaga (2012) among others.} Taking a different approach, Gawande and Hoekman (2006) suggest that one can reconcile the empirical results with the model by acknowledging the complexity of the policy-making process and the uncertainty that arises from it.
-
-Since so much of trade policy is determined in the context of international agreements that must be ratified by domestic legislative bodies, I incorporate PFS-style lobbying into a model of trade agreements in which the executive branches of the governments set tariff levels in anticipation of political pressure upon the legislatures.
-This political pressure may be interpreted either as lobbying against ratification of the trade agreement or lobbying to break the agreement once in force.
-I find that the separation-of-powers government structure can explain the empirical puzzle surrounding governments' welfare weights and that the addition of political uncertainty adds important realism so that the model presented here can speak to important trade policy-making events such as the four-year delay in the ratification of the U.S.-Korea Free Trade Agreement.
 
@@
 
-\item Le Breton, M., Salanie, F., 2003. Lobbying under political uncertainty. Journal of Public Economics 87, 2589-2610.
+\item Le Breton, M., Salanie, F., 2003. Lobbying under Political Uncertainty. Journal of Public Economics 87, 2589-2610.
 

### 2016-02-02 draft.tex
move GH puzzle text to end of intro

@@
 \vskip.2in
-I find that the separation-of-powers government structure can explain the empirical puzzle surrounding governments' welfare weights and that the addition of political uncertainty adds important realism so that the model presented here can speak to important trade policy-making events such as the four-year delay in the ratification of the U.S.-Korea Free Trade Agreement.
-
-Empirical investigations of Grossman and Helpman's (1994) PFS model broadly support the predictions of the theory qualitatively, yet their quantitative estimates present several interesting puzzles.\footnote{cfr.
-Goldberg and Maggi (1999); Gawande and Bandyopadhyay (2000); Mitra, Thomakos and Ulubasoglu (2002); McCalman (2004).\label{fn:gh_lobby}} These studies have consistently found the weight governments place on social welfare to be many times that which they place on lobbying effort, while estimates indicate that the deadweight losses caused by trade distortions are several orders of magnitude larger than lobbying expenditures.\footnote{Feenstra (1992) assembles estimates from the mid-1980s that place a floor of $\$$8 billion per year on the deadweight losses from protection, whereas Bombardini and Trebbi (2012) find that total annual lobbying expenditure on trade issues in 1999-2001 when this data first became available was about $\$$200 million.} This raises the question: how could governments value social welfare so highly, yet grant these quantities of protection at such a low price?
-
-Attempts to more fully model the lobbying process have successfully reduced the parameter estimates for the government's welfare considerations somewhat, but the smallest estimates still indicate that the U.S.
-government, for instance, values social welfare about twenty times as much as contributions.\footnote{cfr.
-Gawande, Krishna and Robbins (2006); Mitra, Thomakos and Ulubasoglu (2006); Bombardini (2008); and Gawande, Krishna and Olarreaga (2012) among others.} Taking a different approach, Gawande and Hoekman (2006) suggest that one can reconcile the empirical results with the model by acknowledging the complexity of the policy-making process and the uncertainty that arises from it.
-
 Consider first the case of political certainty, and assume that the executive is more pro-social than the legislature, as has been the case throughout the post-war period in the United States.\footnote{The standard assumption that the executive is less protectionist than the legislature is a special case of the finding that susceptibility to special interests generally declines with constituency size.
@@
 Ultimately, it is the addition of a rich structure of government, tailored carefully to the trade policy making process, that resolves the main empirical puzzle surrounding the PFS model: in this framework, even a welfare-maximizing executive and lobbying effort at zero can be consistent with high tariff levels.
+
+\vskip.2in
+I find that the separation-of-powers government structure can explain the empirical puzzle surrounding governments' welfare weights and that the addition of political uncertainty adds important realism so that the model presented here can speak to important trade policy-making events such as the four-year delay in the ratification of the U.S.-Korea Free Trade Agreement.
+
+Empirical investigations of Grossman and Helpman's (1994) PFS model broadly support the predictions of the theory qualitatively, yet their quantitative estimates present several interesting puzzles.\footnote{cfr.
+Goldberg and Maggi (1999); Gawande and Bandyopadhyay (2000); Mitra, Thomakos and Ulubasoglu (2002); McCalman (2004).\label{fn:gh_lobby}} These studies have consistently found the weight governments place on social welfare to be many times that which they place on lobbying effort, while estimates indicate that the deadweight losses caused by trade distortions are several orders of magnitude larger than lobbying expenditures.\footnote{Feenstra (1992) assembles estimates from the mid-1980s that place a floor of $\$$8 billion per year on the deadweight losses from protection, whereas Bombardini and Trebbi (2012) find that total annual lobbying expenditure on trade issues in 1999-2001 when this data first became available was about $\$$200 million.} This raises the question: how could governments value social welfare so highly, yet grant these quantities of protection at such a low price?
+
+Attempts to more fully model the lobbying process have successfully reduced the parameter estimates for the government's welfare considerations somewhat, but the smallest estimates still indicate that the U.S.
+government, for instance, values social welfare about twenty times as much as contributions.\footnote{cfr.
+Gawande, Krishna and Robbins (2006); Mitra, Thomakos and Ulubasoglu (2006); Bombardini (2008); and Gawande, Krishna and Olarreaga (2012) among others.} Taking a different approach, Gawande and Hoekman (2006) suggest that one can reconcile the empirical results with the model by acknowledging the complexity of the policy-making process and the uncertainty that arises from it.
 

### 2016-02-02 draft.tex
revise 'how it achieves' and contributions' parts of intro

--- 2016-02-02 draft.tex
@@
 
-\vskip.2in
-I find that the separation-of-powers government structure can explain the empirical puzzle surrounding governments' welfare weights and that the addition of political uncertainty adds important realism so that the model presented here can speak to important trade policy-making events such as the four-year delay in the ratification of the U.S.-Korea Free Trade Agreement.
-
-Empirical investigations of Grossman and Helpman's (1994) PFS model broadly support the predictions of the theory qualitatively, yet their quantitative estimates present several interesting puzzles.\footnote{cfr.
-Goldberg and Maggi (1999); Gawande and Bandyopadhyay (2000); Mitra, Thomakos and Ulubasoglu (2002); McCalman (2004).\label{fn:gh_lobby}} These studies have consistently found the weight governments place on social welfare to be many times that which they place on lobbying effort, while estimates indicate that the deadweight losses caused by trade distortions are several orders of magnitude larger than lobbying expenditures.\footnote{Feenstra (1992) assembles estimates from the mid-1980s that place a floor of $\$$8 billion per year on the deadweight losses from protection, whereas Bombardini and Trebbi (2012) find that total annual lobbying expenditure on trade issues in 1999-2001 when this data first became available was about $\$$200 million.} This raises the question: how could governments value social welfare so highly, yet grant these quantities of protection at such a low price?
-
-Attempts to more fully model the lobbying process have successfully reduced the parameter estimates for the government's welfare considerations somewhat, but the smallest estimates still indicate that the U.S.
-government, for instance, values social welfare about twenty times as much as contributions.\footnote{cfr.
-Gawande, Krishna and Robbins (2006); Mitra, Thomakos and Ulubasoglu (2006); Bombardini (2008); and Gawande, Krishna and Olarreaga (2012) among others.} Taking a different approach, Gawande and Hoekman (2006) suggest that one can reconcile the empirical results with the model by acknowledging the complexity of the policy-making process and the uncertainty that arises from it.
-
-Consider first the case of political certainty, and assume that the executive is more pro-social than the legislature, as has been the case throughout the post-war period in the United States.\footnote{The standard assumption that the executive is less protectionist than the legislature is a special case of the finding that susceptibility to special interests generally declines with constituency size.
+To see how this works, consider first the case of political certainty, and assume that the executive is more pro-social than the legislature, as has been the case throughout the post-war period in the United States.\footnote{The standard assumption that the executive is less protectionist than the legislature is a special case of the finding that susceptibility to special interests generally declines with constituency size.
 One simple illustration from the realm of trade policy is the following: a legislator whose district has a large concentration of a particular industry does not take into account the impact of tariffs on the welfare of consumers in other districts, while the executive, whose constituency encompasses the whole country, will internalize these diffuse consumption effects.
@@
 Thus, any trade agreement under political certainty involves a tariff high enough to disengage the lobby completely; that is, even welfare-maximizing executives set high tariffs and induce zero contributions in equilibrium because they set policy in the shadow of a ratification decision influenced by lobbying.
-Careful modeling of the political process demonstrates that, while tariff-setting behavior is closely linked to deadweight-loss calculations, the empirical estimates of the responsiveness of protective measures to import penetration and the elasticity of import demand do not necessarily map directly onto the PFS parameter for welfare-mindedness.
-
-In the non-unitary government model, these estimates and the earlier-cited stylized facts become consistent with each other.
-The ``case of the missing contributions,'' as Gawande and Hoekman (2006) refer to it, turns out to be well explained by an equilibrium phenomenon arising from the agenda-setting power of the executive branch.
-In particular, this model can explain the approximately 15$\%$ of sectors that receive protection in spite of apparently putting forth no lobbying effort.
+Careful modeling of the political process demonstrates that \textbf{the relationship between deadweight loss calculations and the tariffs that are provided in a trade agreement may be mediated in important ways by the ratification constraint.}
 
@@
 One contribution of this model, therefore, is the idea that trade agreements can act as a kind of domestic political commitment device.
-Rather than a time inconsistency such as that in Maggi and Rodr\'{i}guez-Clare (2007) where the welfare of the unitary policymaker is improved via commitment to a trade agreement by changing the investment decisions of firms, here we have an inconsistency between the preferences of the executive and legislative branches.\footnote{Indeed, depending on the modeling choices and parameters, the utility of the legislative branch taken as a whole may be lower under the trade agreement scenario.} Although the legislature has the final word both with and without the trade agreement, the executives are able to use commitment to the trade agreement to reduce lobbying effort, thus muting the legislature's protectionist bias and bringing the outcome more in line with their relatively liberal preferences.
+Rather than a time inconsistency such as that in Maggi and Rodr\'{i}guez-Clare (2007) where the welfare of the unitary policymaker is improved via commitment to a trade agreement by changing the investment decisions of firms, here we have an inconsistency between the preferences of the executive and legislative branches.\footnote{Indeed, depending on the modeling choices and parameters, the utility of the legislative branch taken as a whole may be lower under the trade agreement scenario.} Although the legislature has the final word both with and without the trade agreement, the executives are able to use the trade agreement to reduce lobbying effort, thus muting the legislature's protectionist bias and bringing the outcome more in line with their relatively liberal preferences.
 
@@
 Ultimately, it is the addition of a rich structure of government, tailored carefully to the trade policy making process, that resolves the main empirical puzzle surrounding the PFS model: in this framework, even a welfare-maximizing executive and lobbying effort at zero can be consistent with high tariff levels.
+
+\vskip.2in
+I find that the separation-of-powers government structure can explain the empirical puzzle surrounding governments' welfare weights and that the addition of political uncertainty adds important realism so that the model presented here can speak to important trade policy-making events such as the four-year delay in the ratification of the U.S.-Korea Free Trade Agreement.
+
+Empirical investigations of Grossman and Helpman's (1994) PFS model broadly support the predictions of the theory qualitatively, yet their quantitative estimates present several interesting puzzles.\footnote{cfr.
+Goldberg and Maggi (1999); Gawande and Bandyopadhyay (2000); Mitra, Thomakos and Ulubasoglu (2002); McCalman (2004).\label{fn:gh_lobby}} These studies have consistently found the weight governments place on social welfare to be many times that which they place on lobbying effort, while estimates indicate that the deadweight losses caused by trade distortions are several orders of magnitude larger than lobbying expenditures.\footnote{Feenstra (1992) assembles estimates from the mid-1980s that place a floor of $\$$8 billion per year on the deadweight losses from protection, whereas Bombardini and Trebbi (2012) find that total annual lobbying expenditure on trade issues in 1999-2001 when this data first became available was about $\$$200 million.} This raises the question: how could governments value social welfare so highly, yet grant these quantities of protection at such a low price?
+
+Attempts to more fully model the lobbying process have successfully reduced the parameter estimates for the government's welfare considerations somewhat, but the smallest estimates still indicate that the U.S.
+government, for instance, values social welfare about twenty times as much as contributions.\footnote{cfr.
+Gawande, Krishna and Robbins (2006); Mitra, Thomakos and Ulubasoglu (2006); Bombardini (2008); and Gawande, Krishna and Olarreaga (2012) among others.} Taking a different approach, Gawande and Hoekman (2006) suggest that one can reconcile the empirical results with the model by acknowledging the complexity of the policy-making process and the uncertainty that arises from it.
+
+In the non-unitary government model, these estimates and the earlier-cited stylized facts become consistent with each other.
+The ``case of the missing contributions,'' as Gawande and Hoekman (2006) refer to it, turns out to be well explained by an equilibrium phenomenon arising from the agenda-setting power of the executive branch.
+In particular, this model can explain the approximately 15$\%$ of sectors that receive protection in spite of apparently putting forth no lobbying effort.
 

### 2016-02-02 draft.tex
revise abstract, end of intro on PFS puzzle

@@
 Thus trade agreements act as a domestic political commitment device: executives set relatively high tariffs to discourage lobbying and increase the chance that the agreement will be ratified.
-This reconciles the result from tests of Grossman and Helpman's model that protection levels are high relative to contributions given estimates of governments' social-welfare weights.
-
+This finding sheds light on the empirical puzzle surrounding governments' welfare weights in the Grossman and Helpman (1994) model, while the addition of political uncertainty adds important realism so that the model presented here can speak to important trade policy-making events such as the four-year delay in the ratification of the U.S.-Korea Free Trade Agreement.
 \textit{JEL classification:} D72, D80, F13, F53, F55 \\
@@
 Ultimately, I find that observed lobbying effort may often be low because the executive branches of the governments set trade agreement tariffs sufficiently high to reduce political pressure on the legislatures.
+I also show that other examples of realism are gained when political uncertainty is added to the model: for instance, the model rationalizes the four-year delay and eventual ratification of the U.S.-Korea Free Trade Agreement.
 
@@
 
-Ultimately, it is the addition of a rich structure of government, tailored carefully to the trade policy making process, that resolves the main empirical puzzle surrounding the PFS model: in this framework, even a welfare-maximizing executive and lobbying effort at zero can be consistent with high tariff levels.
-
-\vskip.2in
-I find that the separation-of-powers government structure can explain the empirical puzzle surrounding governments' welfare weights and that the addition of political uncertainty adds important realism so that the model presented here can speak to important trade policy-making events such as the four-year delay in the ratification of the U.S.-Korea Free Trade Agreement.
-
-Empirical investigations of Grossman and Helpman's (1994) PFS model broadly support the predictions of the theory qualitatively, yet their quantitative estimates present several interesting puzzles.\footnote{cfr.
-Goldberg and Maggi (1999); Gawande and Bandyopadhyay (2000); Mitra, Thomakos and Ulubasoglu (2002); McCalman (2004).\label{fn:gh_lobby}} These studies have consistently found the weight governments place on social welfare to be many times that which they place on lobbying effort, while estimates indicate that the deadweight losses caused by trade distortions are several orders of magnitude larger than lobbying expenditures.\footnote{Feenstra (1992) assembles estimates from the mid-1980s that place a floor of $\$$8 billion per year on the deadweight losses from protection, whereas Bombardini and Trebbi (2012) find that total annual lobbying expenditure on trade issues in 1999-2001 when this data first became available was about $\$$200 million.} This raises the question: how could governments value social welfare so highly, yet grant these quantities of protection at such a low price?
+The addition of a rich structure of government, tailored carefully to the trade policy making process, helps shed light on the empirical puzzle surrounding the PFS model.
+Empirical tests of the model\footnote{cfr.
+Goldberg and Maggi (1999); Gawande and Bandyopadhyay (2000); Mitra, Thomakos and Ulubasoglu (2002); McCalman (2004).\label{fn:gh_lobby}} have consistently found the weight governments place on social welfare to be many times that which they place on lobbying effort, while estimates indicate that the deadweight losses caused by trade distortions are several orders of magnitude larger than lobbying expenditures.\footnote{Feenstra (1992) assembles estimates from the mid-1980s that place a floor of $\$$8 billion per year on the deadweight losses from protection, whereas Bombardini and Trebbi (2012) find that total annual lobbying expenditure on trade issues in 1999-2001 when this data first became available was about $\$$200 million.} This raises the question: how could governments value social welfare so highly, yet grant these quantities of protection at such a low price?
 
@@
 The ``case of the missing contributions,'' as Gawande and Hoekman (2006) refer to it, turns out to be well explained by an equilibrium phenomenon arising from the agenda-setting power of the executive branch.
-In particular, this model can explain the approximately 15$\%$ of sectors that receive protection in spite of apparently putting forth no lobbying effort.
+In particular, in this model, there is inconsistency for the approximately 15$\%$ of sectors that receive protection but apparently put forth no lobbying effort.
+In this framework, even a welfare-maximizing executive and lobbying effort at zero can be reconciled with high tariff levels.
 

### 2016-02-02 draft.tex
add more intuition on uncertainty to intro

@@
 The addition of political uncertainty provides the missing realism, smooths the results and delivers additional intuition.
-I assume all actors are uncertain about the weight the median legislator in a non-unitary legislature places on the profits of the lobbying industry and that this is attributable to the complexity of the legislative process; say, for instance, because a small number of lobbied legislators cannot deliver the votes of their non-lobbied colleagues with certainty.
-In this case, I show that the legislatures disrupt trade agreements with a higher probability and set higher trade war tariffs when lobbying effort increases.
+I assume all actors are uncertain about the weight the median legislator in a non-unitary legislature places on the profits of the lobbying industry.
+\textbf{This is meant to represent the idea that the lobbying and voting ``game'' that goes on within legislatures is often complex enough that none of the actors know precisely what the outcome will be---that is, no one can predict exactly which legislator will be the median and therefore what weight will be placed on import-industry profits at the time of the vote.
+For instance, perhaps} a small number of lobbied legislators cannot deliver the votes of their non-lobbied colleagues with certainty.
+\textbf{One can think of this uncertainty as an additional influence on the legislature's preferences about which all parties---executives, legislators and lobbies alike---are uncertain.}
+
+In this case \textbf{of political uncertainty}, I show that the legislatures disrupt trade agreements with a higher probability and set higher trade war tariffs when lobbying effort increases.
 Because the lobbies respond to higher trade agreement tariffs by decreasing effort (and therefore the probability the agreement will fail to be ratified or will be broken), the executives must trade off the level of social welfare derived while a trade agreement is in force---since they prefer low tariffs---and the chances that the agreement will remain in force.

### 2016-02-02 letter.tex
fix typos in sec:main, remove assumption 2 content from res:lobby
add paragraph to letter on R2

@@
 
-Referee's point $\#2$ (``Intuition in the Introduction about political uncertainty''):
+Referee's point $\#2$ (``Intuition in the Introduction about political uncertainty''): There are several kinds of intuition about political uncertainty, and I struggled a bit to guess which kind was meant.
+Because I already give some intuition about the effects of political uncertainty on the results, I went with the interpretation that this meant intuition/interpretation about what I mean by political uncertainty itself.
+I have thus added a few sentences in the fifth paragraph of the introduction to this end.
+I would be happy to do something different/additional if I have interpreted this suggestion incorrectly.
 
@@
 
+\vskip.5in
+Other changes
+\begin{itemize}
+	\item I have removed the statement ``and the distribution of $\gamma(e,\theta)$ is weakly increasing in $e$'' from Result 2 because this is the content of Assumption 2 and does not need to be repeated.
+\end{itemize}
+
 \end{document}

### 2016-02-02 letter.tex
add paragraphs to letter for E1,E8,R1,R4

@@
 
+Editor's point $\#$1: I have revised the introduction according to the formula you suggested.
+I have toned down considerably the claims about the PFS puzzle as both you and the referee suggest; they are now at the very end of the introduction and in a soft way in the abstract.
+
 Editor's point $\#$2: I have created four subsections within Section 2 to structure the exposition of the model. I think this is much clearer than the previous draft.
@@
 
+Editor's point $\#8$: I agree that my model does not nest PFS.
+The discussion in the new Section 2.4, footnote 17 and Section 5.2 is meant to highlight the similarities and differences.
+I hope that I have addressed your concerns about over-selling the contribution relative to the PFS puzzle so that this is no longer an issue.
+
 Editor's point $\#9$: I have added the move of nature that determines whether the legislature gets the opportunity to the extensive form and to the discussion of the timing in the initial discussion of the model (Section 2.1: The Basic Setup).
@@
 
-Referee's point $\#1$ (``Introduction is derailing at the start''):
+Referee's point $\#1$ (``Introduction is derailing at the start''): I agree and I thank you for your detailed suggestions about how to reformulate the introduction and incorporate a toned-down discussion about the PFS puzzle.
+Please see respone to Editor's point $\#$1 for logistics.
 
@@
 
-Referee's point $\#4$ (``Strong language regarding contribution''):
+Referee's point $\#4$ (``Strong language regarding contribution''): I agree. Please see response to Editor's point $\#$1.
 

### 2016-02-02 draft.tex
update lit review with uncertainty lit

@@
 
+Another related literature considers the impact of exogenously-determined political uncertainty on the potential for trade cooperation.
+These studies (e.g.
+Feenstra and Lewis (1991); Milner and Rosendorff (2001); Bagwell and Staiger (2005); Beshkar (2010); Horn, Maggi and Staiger (2010); Beshkar and Bond (2012); Amador and Bagwell (2013)) derive various implications for the design of international trade agreements using a Baldwin (1987)-style government welfare function with exogenous shocks to the political-pressure parameter.
+As explained in the next section, the model presented here is intended to be a bridge between this literature and that discussed above that focuses on endogenously-determined political economy weights.
+To preview, this model is closest to that of Bagwell and Staiger (2005), with two main changes: in place of their unitary government signing a trade agreement and having different preferences ex-ante and ex-post, this model has two branches of government with differing preferences, where---in the case of the legislature---the political economy weights are determined endogenously.
+
+A very recent literature has begun to study the effects of trade agreements for reducing the uncertainty that firms (Handley 2014 and Handley and Lim\~{a}o 2015) and governments (Lim\~{a}o and Maggi 2015) face.
+
+Several papers have studied the impact of executive/legislative interactions on international agreements in the case of exogenously-determined preferences.
+Mansfield, Milner and Rosendorff (2000) construct a model in which democracies trade more with each other because the domestic legislature's influence requires the foreign government to make deeper tariff cuts to ensure its cooperation.
+Dai (2006) typifies a second approach, which, in line with the Schelling conjecture, argues that a country's legislative constraint assists its own executive in maintaining higher barriers.
+Ethier (2002) treats the effect of a different separation of powers on the international trade system: that between trade negotiators and the government officials who administer protection.
+
+The above work is patterned on the model of Milner and Rosendorff (1997), which explores how uncertainty can affect trade policy and the probability of ratification failure when political preferences, and therefore uncertainty, are exogenous;  Iida (1996) presents a similar model.
+Related is Le Breton and Salanie (2003), which studies lobbying when the lobby is uncertain about the preferences of a unitary decision maker.
+Le Breton and Zaporozhets (2007) go a step further and replace the unitary decision maker with a legislature with multiple actors.
+Song (2008) presents a model in which policy-making power is shared between an executive and a unitary legislature where lobbies are able to endogenously influence the political preferences of the legislature in a context of unilateral policy making with no uncertainty, and Coates and Ludema (2001) study trade policy leadership in a model with endogenous lobbying in the presence of political uncertainty with imperfect monitoring.
+
 The extensive literature that explores the empirical implications of the PFS model is well-summarized by Gawande and Magee (2012).
@@
 Indeed, Imai, Katayama and Krishna (2009) argue that many of the classification schemes present serious challenges for the validity of results.
-
-Another related literature considers the impact of exogenously-determined political uncertainty on the potential for trade cooperation.
-These studies (e.g.
-Feenstra and Lewis (1991); Milner and Rosendorff (2001); Bagwell and Staiger (2005); Beshkar (2010)) derive various implications for the design of international trade agreements using a Baldwin (1987)-style government welfare function with exogenous shocks to the political-pressure parameter.
-As explained in the next section, the model presented here is intended to be a bridge between this literature and that discussed above that focuses on endogenously-determined political economy weights.
-To preview, this model is closest to that of Bagwell and Staiger (2005), with two main changes: in place of their unitary government signing a trade agreement and having different preferences ex-ante and ex-post, this model has two branches of government with differing preferences, where---in the case of the legislature---the political economy weights are determined endogenously.
-
-Several papers have studied the impact of executive/legislative interactions on international agreements in the case of exogenously-determined preferences.
-Mansfield, Milner and Rosendorff (2000) construct a model in which democracies trade more with each other because the domestic legislature's influence requires the foreign government to make deeper tariff cuts to ensure its cooperation.
-Dai (2006) typifies a second approach, which, in line with the Schelling conjecture, argues that a country's legislative constraint assists its own executive in maintaining higher barriers.
-Ethier (2002) treats the effect of a different separation of powers on the international trade system: that between trade negotiators and the government officials who administer protection.
-
-The above work is patterned on the model of Milner and Rosendorff (1997), which explores how uncertainty can affect trade policy and the probability of ratification failure when political preferences, and therefore uncertainty, are exogenous;  Iida (1996) presents a similar model.
-Related is Le Breton and Salanie (2003), which studies lobbying when the lobby is uncertain about the preferences of a unitary decision maker.
-Le Breton and Zaporozhets (2007) go a step further and replace the unitary decision maker with a legislature with multiple actors.
-Song (2008) presents a model in which policy-making power is shared between an executive and a unitary legislature where lobbies are able to endogenously influence the political preferences of the legislature in a context of unilateral policy making with no uncertainty, and Coates and Ludema (2001) study trade policy leadership in a model with endogenous lobbying in the presence of political uncertainty with imperfect monitoring.
 
@@
 
+\item Amador, M., and Bagwell, K., 2013. The Theory of Optimal Delegation with an Application to Tariff Caps. Econometrica 81, 1541-1599.
+
 \item Ansolabehere, S., de Figueiredo, J., Snyder Jr., J., 2003. Why Is There so Little Money in U.S. Politics? Journal of Economic Perspectives 17, 105-130.
@@
 
+\item Beshkar, M. and Bond, E., 2012. Cap and Escape in Trade Agreements. Available at \url{http://papers.ssrn.com/sol3/papers.cfm?abstract_id=2145679}.
+
 \item Bombardini, M., 2008. Firm Heterogeneity and Lobby Participation. Journal of International Economics 75, 329-348.
@@
 
+\item Handley, K., 2014. Exporting under Trade Policy Uncertainty: Theory and Evidence. Journal of international Economics 94, 50-66.
+
+\item Handley, K., Lim\~{a}o, N., 2015. Trade and Investment under Policy Uncertainty: Theory and Firm Evidence. American Economic Journal: Economic Policy 7, 189-222.
+
 \item Henisz, W., Mansfield, E., 2006. Votes and Vetoes: The Political Determinants of Commercial Openness. International Studies Quarterly 50, 189-211.
@@
 \item Le Breton, M., Zaporozhets, V., 2007. Legislative Lobbying under Political Uncertainty. Available at SSRN: \url{http://ssrn.com/abstract=1024686}.
+
+\item Lim\~{a}o, N., Maggi, G., 2015. Uncertainty and Trade Agreements. American Economic Journal: Microeconomics 7, 1-42.
 

### 2016-02-02 draft.tex
move assumptions above sec:twt

@@
 Different assumptions on $\ga(\cdot,\cdot)$ correspond to different institutional features and will affect optimal lobbying- and tariff-setting behavior.
+I make the following two basic assumptions on the weight on import-industry profits:
+
+\begin{assumption}
+  $\ga(e,\ta)$ is increasing and concave in $e$ for every $\ta \in \Theta$.
+  \label{as:ga_c}
+\end{assumption}
+
+\begin{assumption}
+  $\ga(e,\ta)$ is increasing in $\ta$.
+  \label{as:ga_ta}
+\end{assumption}
+
+Assumption~\ref{as:ga_c} formalizes the intuition that the legislature favors the import-competing industry more the higher is its lobbying effort, but that there are diminishing returns to lobbying activity.\footnote{The diminishing returns here take the form of declining increments to the lobby's influence as effort increases; in Ethier (2012), the returns to lobbying decline as the level of protection increases.} It allows the political uncertainty inherent in $\ta$ to be correlated in interesting ways with the level of $e$, ruling out only (i) that higher effort levels make lower $\ga$-weights more likely and (ii) that effort and uncertainty are correlated in such a way that increasing lobbying effort changes the structure of uncertainty so that higher $\ga$-weights become more likely at an accelerating pace.
+Assumption~\ref{as:ga_ta} formalizes the intuition that larger realization of the uncertainty variable $\ta$ increase the value of the political economy weight.
+
 While a structural model of the legislative process might be more satisfying, the reduced form can deliver the key insights without distracting from the central policy choice and lobbying dynamics that take place at the level of the executives and lobbies, to which we turn next.
@@
 
-Therefore, in line with both the theoretical and empirical literature, I make the following assumption:
+Therefore, in line with both the theoretical and empirical literature, I make the following, final assumption on $\ga(e,\ta)$:
 
@@
 Note that the legislature's best response functions---both for whether to break the agreement, and at what level to set the tariff if it does---will be affected by the shape of $\ga(e,\ve)$.
-I make the following further assumptions on the weight on import-industry profits:
-
-\begin{assumption}
-  $\ga(e,\ta)$ is increasing and concave in $e$ for every $\ta \in \Theta$.
-  \label{as:ga_c}
-\end{assumption}
-
-\begin{assumption}
-  $\ga(e,\ta)$ is increasing in $\ta$.
-  \label{as:ga_ta}
-\end{assumption}
-
-Assumption~\ref{as:ga_c} formalizes the intuition that the legislature favors the import-competing industry more the higher is its lobbying effort, but that there are diminishing returns to lobbying activity.\footnote{The diminishing returns here take the form of declining increments to the lobby's influence as effort increases; in Ethier (2012), the returns to lobbying decline as the level of protection increases.} It allows the political uncertainty inherent in $\ta$ to be correlated in interesting ways with the level of $e$, ruling out only (i) that higher effort levels make lower $\ga$-weights more likely and (ii) that effort and uncertainty are correlated in such a way that increasing lobbying effort changes the structure of uncertainty so that higher $\ga$-weights become more likely at an accelerating pace.
-
-Assumption~\ref{as:ga_ta} formalizes the intuition that larger realization of the uncertainty variable $\ta$ increase the value of the political economy weight.
+Recall that we have made three basic assumptions on $\ga(e,\ve)$ in Section~\ref{sec:choices}.
+More restrictions could be added to deliver further results.
 

### 2016-02-02 letter.tex
add note on literature update ot letter
explain moving as:ga_c and as:ga_ta in letter

@@
 
-Editor's point $\#$2: I have created four subsections within Section 2 to structure the exposition of the model. I think this is much clearer than the previous draft.
+Editor's point $\#$2: I have created four subsections within Section 2 to structure the exposition of the model.
+I think this is much clearer than the previous draft.
+Note that I have moved (former) Assumptions 2 and 3 from Section 3 into Section 2.4.
+They are now labeled Assumptions 1 and 2, while the former Assumption 1 is now Assumption 3 because of this rearrangement.
 
@@
 
-I have also removed the duplicate statement of Assumption 1 and combined redundant text surrounding the two statements. I have added $\forall e$ to the statement of Assumption 1.
+I have also removed the duplicate statement of Assumption 1 and combined redundant text surrounding the two statements.
+I have added $\forall e$ to the statement of Assumption 1, which is now relabeled Assumption 3 because I have moved the former Assumptions 2 and 3 earlier in Section 2.4 as part of the restructuring discussed in point $\#2$.
 
@@
 	\item I have removed the statement ``and the distribution of $\gamma(e,\theta)$ is weakly increasing in $e$'' from Result 2 because this is the content of Assumption 2 and does not need to be repeated.
+	\item I have added some newer references to the uncertainty and trade agreements literature to the literature review and updated working papers where possible.
 \end{itemize}

### 2016-02-02 draft.tex
boldchanges since Jan. 28, fix a few typos

--- 2016-02-02 draft.tex
@@
 Thus trade agreements act as a domestic political commitment device: executives set relatively high tariffs to discourage lobbying and increase the chance that the agreement will be ratified.
-This finding sheds light on the empirical puzzle surrounding governments' welfare weights in the Grossman and Helpman (1994) model, while the addition of political uncertainty adds important realism so that the model presented here can speak to important trade policy-making events such as the four-year delay in the ratification of the U.S.-Korea Free Trade Agreement.
+\textbf{This finding sheds light on the empirical puzzle surrounding governments' welfare weights in the Grossman and Helpman (1994) model, while the addition of political uncertainty adds important realism so that the model presented here can speak to important trade policy-making events such as the four-year delay in the ratification of the U.S.-Korea Free Trade Agreement.}
+
 \textit{JEL classification:} D72, D80, F13, F53, F55 \\
@@
 
-Much of trade policy is determined in the context of international agreements that must be ratified by some domestic body.
+\textbf{Much of trade policy is determined in the context of international agreements that must be ratified by some domestic body.
 This paper seeks to shed light on such a policy-making process by incorporating Protection-for-Sale (Grossman and Helpman 1994, hereafter PFS) style lobbying into a model of trade agreements in which the executive branches of the governments set tariff levels in anticipation of political pressure upon a ratifying body, assumed here to be a legislature.
@@
 Important realism is achieved not only through the addition of the ratification constraint but also by the incorporation of political uncertainty into the model.
-
-Assuming for now that lobbies represent only import-competing interests, lobbies in this environment would like to prevent or derail a trade agreement that would reduce their tariff protection.
+}
+\textbf{Assuming for now that lobbies represent only import-competing interests, lobbies in this environment would like to prevent or derail a trade agreement that would reduce their tariff protection.
 They work toward this end by exerting effort to convince their legislatures to vote against the trade agreement but face some uncertainty about whether their efforts will result in a voting outcome consistent with their interests.
@@
 Ultimately, I find that observed lobbying effort may often be low because the executive branches of the governments set trade agreement tariffs sufficiently high to reduce political pressure on the legislatures.
-I also show that other examples of realism are gained when political uncertainty is added to the model: for instance, the model rationalizes the four-year delay and eventual ratification of the U.S.-Korea Free Trade Agreement.
-
-To see how this works, consider first the case of political certainty, and assume that the executive is more pro-social than the legislature, as has been the case throughout the post-war period in the United States.\footnote{The standard assumption that the executive is less protectionist than the legislature is a special case of the finding that susceptibility to special interests generally declines with constituency size.
+I also show that other examples of realism are gained when political uncertainty is added to the model: for instance, the model rationalizes the four-year delay and eventual ratification of the U.S.-Korea Free Trade Agreement.}
+
+\textbf{To see how this works,} consider first the case of political certainty, and assume that the executive is more pro-social than the legislature, as has been the case throughout the post-war period in the United States.\footnote{The standard assumption that the executive is less protectionist than the legislature is a special case of the finding that susceptibility to special interests generally declines with constituency size.
 One simple illustration from the realm of trade policy is the following: a legislator whose district has a large concentration of a particular industry does not take into account the impact of tariffs on the welfare of consumers in other districts, while the executive, whose constituency encompasses the whole country, will internalize these diffuse consumption effects.
@@
 
-The addition of a rich structure of government, tailored carefully to the trade policy making process, helps shed light on the empirical puzzle surrounding the PFS model.
+\textbf{The addition of a rich structure of government}, tailored carefully to the trade policy making process, helps shed light on the empirical puzzle surrounding the PFS model.
 Empirical tests of the model\footnote{cfr.
@@
 The ``case of the missing contributions,'' as Gawande and Hoekman (2006) refer to it, turns out to be well explained by an equilibrium phenomenon arising from the agenda-setting power of the executive branch.
-In particular, in this model, there is inconsistency for the approximately 15$\%$ of sectors that receive protection but apparently put forth no lobbying effort.
-In this framework, even a welfare-maximizing executive and lobbying effort at zero can be reconciled with high tariff levels.
+In particular, \textbf{in this model, there is inconsistency for the approximately 15$\%$ of sectors that receive protection but apparently put forth no lobbying effort.
+In this framework, even a welfare-maximizing executive and lobbying effort at zero can be reconciled with high tariff levels.}
 
@@
 
+Another related literature considers the impact of exogenously-determined political uncertainty on the potential for trade cooperation.
+These studies (e.g.
+Feenstra and Lewis (1991); Milner and Rosendorff (2001); Bagwell and Staiger (2005); Beshkar (2010); \textbf{Horn, Maggi and Staiger (2010); Beshkar and Bond (2012); Amador and Bagwell (2013)}) derive various implications for the design of international trade agreements using a Baldwin (1987)-style government welfare function with exogenous shocks to the political-pressure parameter.
+As explained in the next section, the model presented here is intended to be a bridge between this literature and that discussed above that focuses on endogenously-determined political economy weights.
+To preview, this model is closest to that of Bagwell and Staiger (2005), with two main changes: in place of their unitary government signing a trade agreement and having different preferences ex-ante and ex-post, this model has two branches of government with differing preferences, where---in the case of the legislature---the political economy weights are determined endogenously.
+
+\textbf{A very recent literature has begun to study the effects of trade agreements for reducing the uncertainty that firms (Handley 2014 and Handley and Lim\~{a}o 2015) and governments (Lim\~{a}o and Maggi 2015) face.}
+
+Several papers have studied the impact of executive/legislative interactions on international agreements in the case of exogenously-determined preferences.
+Mansfield, Milner and Rosendorff (2000) construct a model in which democracies trade more with each other because the domestic legislature's influence requires the foreign government to make deeper tariff cuts to ensure its cooperation.
+Dai (2006) typifies a second approach, which, in line with the Schelling conjecture, argues that a country's legislative constraint assists its own executive in maintaining higher barriers.
+Ethier (2002) treats the effect of a different separation of powers on the international trade system: that between trade negotiators and the government officials who administer protection.
+
+The above work is patterned on the model of Milner and Rosendorff (1997), which explores how uncertainty can affect trade policy and the probability of ratification failure when political preferences, and therefore uncertainty, are exogenous;  Iida (1996) presents a similar model.
+Related is Le Breton and Salanie (2003), which studies lobbying when the lobby is uncertain about the preferences of a unitary decision maker.
+Le Breton and Zaporozhets (2007) go a step further and replace the unitary decision maker with a legislature with multiple actors.
+Song (2008) presents a model in which policy-making power is shared between an executive and a unitary legislature where lobbies are able to endogenously influence the political preferences of the legislature in a context of unilateral policy making with no uncertainty, and Coates and Ludema (2001) study trade policy leadership in a model with endogenous lobbying in the presence of political uncertainty with imperfect monitoring.
+
 The extensive literature that explores the empirical implications of the PFS model is well-summarized by Gawande and Magee (2012).
@@
 As profits and producer surplus (identical in this model) in a sector are increasing in the price of its good, profits in the import-competing sector are also increasing in the domestic tariff.
-This economic fact, combined with the assumptions on specific factor ownership, is what motivates political activity by import-competing lobbies.
-
-I now turn to describing the strategic players in the game.
-In order to focus attention on protectionist political forces, I assume that only the import-competing industry in each country is politically-organized and able to lobby and that it is represented by a single lobbying organization (see Footnote~\ref{fn:lobby} for discussion of additional lobbies).
+This economic fact, combined with the assumptions on specific factor ownership, is what motivates political activity \textbf{by import-competing lobbies}.
+
+\textbf{I now turn to describing the strategic players in the game.} In order to focus attention on protectionist political forces, I assume that only the import-competing industry in each country is politically-organized and able to lobby and that it is represented by a single lobbying organization (see Footnote~\ref{fn:lobby} for discussion of additional lobbies).
 Each country's government is composed of two branches: an executive who can conclude trade agreements and a legislature that has final say on trade policy.
-To summarize, the political process is modeled as involving three players in each country: the lobby, the executive, and the legislature.
-
-Figure~\ref{fig:ext} illustrates the timing of the game from the perspective of the home country.\footnote{\textbf{This extensive form is for the case where the foreign legislature does not break the trade agreement.}} First, the executives set trade policy cooperatively within the context of an international agreement, that is they choose the trade agreement tariffs $\bta = \left(\tau^a,\tau^{*a}\right)$.
+\textbf{To summarize}, the political process is modeled as involving three players in each country: the lobby, the executive, and the legislature.
+
+\textbf{Figure~\ref{fig:ext} illustrates the timing of the game from the perspective of the home country.}\footnote{\textbf{This extensive form is for the case where the foreign legislature does not break the trade agreement.}} First, the executives set trade policy cooperatively within the context of an international agreement, \textbf{that is they choose the trade agreement tariffs $\bta = \left(\tau^a,\tau^{*a}\right)$}.
 After the trade agreement is concluded, the lobbies attempt to persuade the legislators in their respective countries to vote ``no'' on the ratification of the agreement; or alternatively, if there is no ratification hurdle, to break the trade agreement.\footnote{As will be seen shortly, these two interpretations are equivalent in this setting because tariffs in a ``trade war'' are assumed to be set non-cooperatively in the same manner as if there were no trade agreement.
-For simplicity, I refer to this as the ``break'' decision throughout.} They do this by choice of lobbying effort $e_b$, which can take on any non-negative value.
-
-Next, nature determines with probabilities $o$ and $o^*$ whether each legislature will have the opportunity to take a vote to break the agreement.
-I assume that these probabilities are less than one-half and mutually exclusive.\footnote{\textbf{While this assumption is made for purposes of exposition and tractability, one can think of $o$ and $o^*$ as being determined by events beyond the executives' or legislatures' control that determine whether or not the legislatures will be willing to consider the lobby's request; for example, they would be affected by the occurrence of an economic crisis that diverts the legislature's attention from less-pressing matters.}} This allows me to focus on the key interaction between the domestic actors and abstract from the strategic interaction between the lobbies in the two countries.\footnote{\textbf{When the legislatures in both countries have an opportunity to break the agreement simultaneously, an incentive for each lobby to free ride on the effort of the other makes the optimization problem of the executives more complex as they are faced with the problem of how best to exploit the free-riding dynamic.}}
-
-After this, the uncertainty about the the median legislator's identity is resolved.
+For simplicity, I refer to this as the ``break'' decision throughout.} \textbf{They do this by choice of lobbying effort $e_b$, which can take on any non-negative value.}
+
+\textbf{Next, nature determines with probabilities $o$ and $o^*$ whether each legislature will have the opportunity to take a vote to break the agreement.
+I assume that these probabilities are less than one-half and mutually exclusive.}\footnote{\textbf{While this assumption is made for purposes of exposition and tractability, one can think of $o$ and $o^*$ as being determined by events beyond the executives' or legislatures' control that determine whether or not the legislatures will be willing to consider the lobby's request; for example, they would be affected by the occurrence of an economic crisis that diverts the legislature's attention from less-pressing matters.}} \textbf{This allows me to focus on the key interaction between the domestic actors and abstract from the strategic interaction between the lobbies in the two countries.}\footnote{\textbf{When the legislatures in both countries have an opportunity to break the agreement simultaneously, an incentive for each lobby to free ride on the effort of the other makes the optimization problem of the executives more complex as they are faced with the problem of how best to exploit the free-riding dynamic.}}
+
+\textbf{After this}, the uncertainty about the the median legislator's identity is resolved.
 This is modeled as the realization of a random variable $\ta$, the value of which all players observe simultaneously.
@@
 	\end{center}
-	\caption{Extensive Form\label{fig:ext}}
+	\caption{\textbf{Extensive Form\label{fig:ext}}}
 \end{figure}
@@
 where $\mathit{CS}$ is consumer surplus, $\mathit{PS}$ is producer surplus, $\ga(e,\ta)$ is the weight placed on producer surplus (profits) in the import-competing industry, $e$ is lobbying effort, $\ta$ is a random variable, and $\mathit{TR}$ is tariff revenue.\footnote{\textbf{Labor income $l$ could also be included in both median legislator and executive welfare.
-I omit it because its inclusion alters none of the results and only serves to complicate the exposition.}} I model the decisions of the legislature as being taken by a median legislator (more on this in Section~\ref{sec:choices}.
-Notice that the weight the median legislator places on the profits of the import-competing industry, $\ga(e,\ve)$, is affected by the level of lobbying effort and a random variable $\ta$.
+I omit it because its inclusion alters none of the results and only serves to complicate the exposition.}} I model the decisions of the legislature as being taken by a median legislator \textbf{(more on this in Section~\ref{sec:choices}).
+Notice that} the weight the median legislator places on the profits of the import-competing industry, $\ga(e,\ve)$, is affected by the level of lobbying effort and a random variable $\ta$.
 
@@
 
-In the first stage of the game, the executives choose the trade agreement tariffs $\bta=\left(\tau^a,\tau^{*a} \right)$ via a negotiating process that I assume to be efficient.
+In the first stage \textbf{of the game}, the executives choose the trade agreement tariffs $\bta=\left(\tau^a,\tau^{*a} \right)$ via a negotiating process that I assume to be efficient.
 In this symmetric environment, it therefore maximizes the expected joint payoffs:
@@
 \label{sec:info}
-I examine a simple class of equilibria that have the following features.
-
-First, on information.
-Although uncertainty is present, information about it is symmetric.
+\textbf{I examine a simple class of equilibria that have the following features.}
+
+\textbf{First, on information.} Although uncertainty is present, information about it is symmetric.
 However, in line with the literature, I assume the lobby's contribution is not observable to the foreign legislature.
 The implication is that the lobby can directly influence only the home legislature, and so the influence of one country's lobby on the other country's legislature occurs only through the tariffs selected.\footnote{cfr.
-Grossman and Helpman (1995b), page 685.} Given the game's structure and the assumption that lobbying effort is not observable across international borders, players in the same country can take advantage of more information than those who are in different countries.
-That is, the solution concept employed here is perfect public equilibrium (PPE).
-
-Second, whenever there is a possibility of multiple equilibria, I will focus on the one that maximizes the welfare of the executives.
-
-Finally, I assume that an external authority can ensure enforcement of the agreement in the one-shot game analyzed here.\footnote{Buzard (2015) extends the analysis to a repeated-game context and can therefore remove this restriction.}
+Grossman and Helpman (1995b), page 685.} \textbf{Given the game's structure and the assumption that lobbying effort is not observable across international borders, players in the same country can take advantage of more information than those who are in different countries.
+That is, the solution concept employed here is perfect public equilibrium (PPE).}
+
+\textbf{Second, whenever there is a possibility of multiple equilibria, I will focus on the one that maximizes the welfare of the executives. }
+
+\textbf{Finally, I assume that an external authority can ensure enforcement of the agreement in the one-shot game analyzed here.}\footnote{Buzard (\textbf{2015}) extends the analysis to a repeated-game context and can therefore remove this restriction.}
 
@@
 
-Notice that, aside from the endogeneity of the weight the legislature places on the lobbying industry's profits, Equation~\ref{eq:ml} is precisely the \textit{deus ex machina} government objective function popularized by Baldwin (1987) that is commonly employed in the literature on the political economy of trade agreements.
-Instead of the weight on profits being constant as in the exogenous case or necessarily linear in contributions as in the PFS case, this form allows for both endogeneity and non-linearity.
+Notice that, aside from the endogeneity of the weight the legislature places on the lobbying industry's profits, \textbf{Equation~\ref{eq:ml}} is precisely the \textit{deus ex machina} government objective function popularized by Baldwin (1987) that is commonly employed in the literature on the political economy of trade agreements.
+\textbf{Instead of the weight on profits being constant as in the exogenous case or necessarily linear in contributions as in the PFS case, this form allows for both endogeneity and non-linearity.}
 
@@
 The formulation in Equation~\ref{eq:ml} is intended as a bridge between the two.
-See Buzard (2016) for an example of how this form can be used to integrate the analysis of endogenous political activity and exogenous shocks to the political welfare of governments.
-
-In addition, letting the weight the median legislator places on the profits of the import-competing industry $\ga(e,\ve)$ be affected by the level of lobbying effort as well a random variable $\ta$ allows for the analysis of endogenous lobbying to be integrated with the standard political economy shocks of the trade agreements literature.
+See Buzard (\textbf{2016}) for an example of how this form can be used to integrate the analysis of endogenous political activity and exogenous shocks to the political welfare of governments.
+
+\textbf{In addition, letting the weight the median legislator places on the profits of the import-competing industry $\ga(e,\ve)$ be affected by the level of lobbying effort as well a random variable $\ta$} allows for the analysis of endogenous lobbying to be integrated with the standard political economy shocks of the trade agreements literature.
 Modeling the objective function so closely on the standard in the trade agreements literature allows for direct comparisons to the large extant body of work that studies exogenous shocks only, revealing cleanly the effects of the addition of endogenous lobbying---and if one wants---this different kind of political uncertainty.
@@
 Alternatively, the integration of the random variable $\ta$ can represent reduced-form modeling of political uncertainty arising from a complex decision-making process within a legislature.
-One can interpret $\ta$ as measuring uncertainty over the identity of the median legislator in order to represent the idea that the lobbying and voting ``game'' that goes on within legislatures is often complex enough that none of the actors know precisely what the outcome will be---that is, no one can predict exactly which legislator will be the median and therefore what weight will be placed on import-industry profits at the time of the vote.
+\textbf{One can interpret $\ta$ as measuring uncertainty over} the identity of the median legislator in order to represent the idea that the lobbying and voting ``game'' that goes on within legislatures is often complex enough that none of the actors know precisely what the outcome will be---that is, no one can predict exactly which legislator will be the median and therefore what weight will be placed on import-industry profits at the time of the vote.
 Thus one way we can think about $\ta$ is as an additional influence on the legislature's preferences about which all parties---executives, legislators and lobbies alike---are uncertain.
@@
 Different assumptions on $\ga(\cdot,\cdot)$ correspond to different institutional features and will affect optimal lobbying- and tariff-setting behavior.
+\textbf{I make the following two basic assumptions on the weight on import-industry profits:}
+
+\begin{assumption}
+  $\ga(e,\ta)$ is increasing and concave in $e$ for every $\ta \in \Theta$.
+  \label{as:ga_c}
+\end{assumption}
+
+\begin{assumption}
+  $\ga(e,\ta)$ is increasing in $\ta$.
+  \label{as:ga_ta}
+\end{assumption}
+
+Assumption~\ref{as:ga_c} formalizes the intuition that the legislature favors the import-competing industry more the higher is its lobbying effort, but that there are diminishing returns to lobbying activity.\footnote{The diminishing returns here take the form of declining increments to the lobby's influence as effort increases; in Ethier (2012), the returns to lobbying decline as the level of protection increases.} It allows the political uncertainty inherent in $\ta$ to be correlated in interesting ways with the level of $e$, ruling out only (i) that higher effort levels make lower $\ga$-weights more likely and (ii) that effort and uncertainty are correlated in such a way that increasing lobbying effort changes the structure of uncertainty so that higher $\ga$-weights become more likely at an accelerating pace.
+Assumption~\ref{as:ga_ta} formalizes the intuition that larger realization of the uncertainty variable $\ta$ increase the value of the political economy weight.
+
 While a structural model of the legislative process might be more satisfying, the reduced form can deliver the key insights without distracting from the central policy choice and lobbying dynamics that take place at the level of the executives and lobbies, to which we turn next.
 
-The executives' objective function is taken to be identical to the legislatures' except that profits in the import-competing industry receive a fixed, exogenous weight.
-This construction permits me to integrate the influence of the import-competing industry on negotiations prior to the formation of the trade agreement while also reflecting the idea that the mechanisms through which political considerations influence the executives' preferences appear to be quite different from those at the legislative level---the electoral college and primary/caucus systems for instance.
-
-This setup does not strictly require that the executives are not lobbied.
+\textbf{The executives' objective function is taken to be identical to the legislatures' except that profits in the import-competing industry receive a fixed, exogenous weight.} This construction permits me to integrate the influence of the import-competing industry on negotiations prior to the formation of the trade agreement while also reflecting the idea that the mechanisms through which political considerations influence the executives' preferences appear to be quite different from those at the legislative level---the electoral college and primary/caucus systems for instance.
+
+\textbf{This setup does not strictly require that the executives are not lobbied.
 The assumption that the weight the executives place on import-competing profits is constant only requires that the extent to which they favor this industry cannot be \textit{directly} influenced by lobbying effort.
 The assumption is innocuous as long as the executives' preferences are not directly altered in a significant way by lobbying over trade.
-
+}
 In the case of the post-war United States, where the Congress has consistently been significantly more protectionist than the President, this seems to reasonably reflect the political reality.
@@
 
-Therefore, in line with both the theoretical and empirical literature, I make the following assumption:
+Therefore, in line with both the theoretical and empirical literature, I make the following, \textbf{final} assumption \textbf{on $\ga(e,\ta)$}:
 
@@
 
-The results apply beyond separation-of-powers presidential systems because the presence of veto points is what is essential for the dynamic posited here to occur.
+\textbf{The results apply beyond separation-of-powers presidential systems because the presence of veto points is what is essential for the dynamic posited here to occur.
 Although the number of formal veto points tends to be lower in, for instance, parliamentary systems, veto points are typically not absent altogether (Henisz and Mansfield 2006).
@@
 It is important to consider that constraints may be formal or informal, and they may come from legislative, regulatory, judicial, sub-national or even other executive bodies.
+}
 
@@
 
-The legislature's decisions are stochastic, so the outcome of the vote on whether or not to break the trade agreement, as well as the level at which the tariff will be set in the case of a disruption, is not known to \textit{any} player until the uncertainty over the identity of the median legislator is resolved at the moment the vote takes place on each issue.\footnote{We might think of the trade-war tariff as being set in the legislation before a vote takes place in the case of administered protection for instance; the results are robust to this variant.} I represent the probability that the home legislature breaks the trade agreement and sets the tariff at $\tau^n$ as:
+The legislature's decisions are stochastic, so the outcome of the vote on whether or not to break the trade agreement, as well as the level at which the tariff will be set in the case of a disruption, is not known to \textit{any} player until the uncertainty over the identity of the median legislator is resolved at the moment the vote takes place \textbf{on each issue}.\footnote{We might think of the trade-war tariff as being set in the legislation before a vote takes place in the case of administered protection for instance; the results are robust to this variant.} I represent the probability that the home legislature breaks the trade agreement and sets the tariff at $\tau^n$ as:
 \begin{multline}
@@
 Note that the legislature's best response functions---both for whether to break the agreement, and at what level to set the tariff if it does---will be affected by the shape of $\ga(e,\ve)$.
-I make the following further assumptions on the weight on import-industry profits:
-
-\begin{assumption}
-  $\ga(e,\ta)$ is increasing and concave in $e$ for every $\ta \in \Theta$.
-  \label{as:ga_c}
-\end{assumption}
-
-\begin{assumption}
-  $\ga(e,\ta)$ is increasing in $\ta$.
-  \label{as:ga_ta}
-\end{assumption}
-
-Assumption~\ref{as:ga_c} formalizes the intuition that the legislature favors the import-competing industry more the higher is its lobbying effort, but that there are diminishing returns to lobbying activity.\footnote{The diminishing returns here take the form of declining increments to the lobby's influence as effort increases; in Ethier (2012), the returns to lobbying decline as the level of protection increases.} It allows the political uncertainty inherent in $\ta$ to be correlated in interesting ways with the level of $e$, ruling out only (i) that higher effort levels make lower $\ga$-weights more likely and (ii) that effort and uncertainty are correlated in such a way that increasing lobbying effort changes the structure of uncertainty so that higher $\ga$-weights become more likely at an accelerating pace.
-
-Assumption~\ref{as:ga_ta} formalizes the intuition that larger realization of the uncertainty variable $\ta$ increase the value of the political economy weight.
+\textbf{Recall that we have made three basic assumptions on $\ga(e,\ve)$ in Section~\ref{sec:choices}.
+More restrictions could be added to deliver further results.}
 
@@
 Note that the symmetric increase in the foreign tariff does not impact profits in the import-competing industry, but it can impact lobbying incentives through the break probability.
-However, there is an analogous effect on the break probability from raising $\tau^a$ and the net effect is negative given Assumption~\ref{as:ga_c} on the probability distribution function of $\ga(e,\ta)$, which ensures that the effect of lobbying effort on the break probability does not increase as $\bta$ increases.
+However, there is an analogous effect on the break probability from raising $\tau^a$ and the net effect is negative given \textbf{Assumption~\ref{as:ga_c}} on the probability distribution function of $\ga(e,\ta)$, which ensures that the effect of lobbying effort on the break probability does not increase as $\bta$ increases.
 This assumption can be significantly weakened: this effect can be positive as long as it is smaller in absolute value than the negative effect on trade agreement profits.
@@
 
+\item \textbf{Amador, M., and Bagwell, K., 2013. The Theory of Optimal Delegation with an Application to Tariff Caps. Econometrica 81, 1541-1599.}
+
 \item Ansolabehere, S., de Figueiredo, J., Snyder Jr., J., 2003. Why Is There so Little Money in U.S. Politics? Journal of Economic Perspectives 17, 105-130.
@@
 
+\item \textbf{Beshkar, M. and Bond, E., 2012. Cap and Escape in Trade Agreements. Available at \url{http://papers.ssrn.com/sol3/papers.cfm?abstract_id=2145679}.}
+
 \item Bombardini, M., 2008. Firm Heterogeneity and Lobby Participation. Journal of International Economics 75, 329-348.
@@
 
-\item Buzard, K., 2015. Self-enforcing Trade Agreements, Dispute Settlement and Separation of Powers. Available at: \url{http://papers.ssrn.com/sol3/papers.cfm?abstract_id=2333728}.
-
-\item Buzard, K., 2016. Endogenous Politics and the Design of Trade Agreements. Available at: \url{https://kbuzard.expressions.syr.edu/wp-content/uploads/Endogenous-Politics.pdf}.
+\item Buzard, K., \textbf{2015}. Self-enforcing Trade Agreements, Dispute Settlement and Separation of Powers. Available at: \url{http://papers.ssrn.com/sol3/papers.cfm?abstract_id=2333728}.
+
+\item Buzard, K., \textbf{2016}. Endogenous Politics and the Design of Trade Agreements. Available at: \url{https://kbuzard.expressions.syr.edu/wp-content/uploads/Endogenous-Politics.pdf}.
 
@@
 
+\item \textbf{Handley, K., 2014. Exporting under Trade Policy Uncertainty: Theory and Evidence. Journal of international Economics 94, 50-66.}
+
+\item \textbf{Handley, K., Lim\~{a}o, N., 2015. Trade and Investment under Policy Uncertainty: Theory and Firm Evidence. American Economic Journal: Economic Policy 7, 189-222.}
+
 \item Henisz, W., Mansfield, E., 2006. Votes and Vetoes: The Political Determinants of Commercial Openness. International Studies Quarterly 50, 189-211.
@@
 
+\item \textbf{Horn, H., Maggi, G., Staiger, R. W., 2010. Trade Agreements as Endogenously Incomplete Contracts. American Economic Review 100, 394-419.}
+
 \item Iida, K., 1996. Involuntary Defection in Two-Level Games. Public Choice 89, 283-303.
@@
 \item Le Breton, M., Zaporozhets, V., 2007. Legislative Lobbying under Political Uncertainty. Available at SSRN: \url{http://ssrn.com/abstract=1024686}.
+
+\item \textbf{Lim\~{a}o, N., Maggi, G., 2015. Uncertainty and Trade Agreements. American Economic Journal: Microeconomics 7, 1-42.}
 

### 2016-02-02 draft.tex
revise letter to editor/referee (final)

@@
 This political pressure may be interpreted either as lobbying against ratification of the trade agreement or lobbying to break the agreement once in force.
-Important realism is achieved not only through the addition of the ratification constraint but also by the incorporation of political uncertainty into the model.
-}
+Important realism is achieved not only through the addition of the ratification constraint but also by the incorporation of political uncertainty into the model.}
+
 \textbf{Assuming for now that lobbies represent only import-competing interests, lobbies in this environment would like to prevent or derail a trade agreement that would reduce their tariff protection.
@@
 Of central concern is how the break probabilities behave as a function of the trade agreement tariffs given the lobby's optimal response.
-Recalling the definition of the total break probability, $B(\bta)=b(e_b(\bta),\bta)$, we must combine th direct effect of tariffs on the legislature's break decision with the indirect effect through the impact on the lobby's choice of $e_b$.
+Recalling the definition of the total break probability, $B(\bta)=b(e_b(\bta),\bta)$, we must combine the direct effect of tariffs on the legislature's break decision with the indirect effect through the impact on the lobby's choice of $e_b$.
 
@@
 
-As noted in Dixit, Grossman and Helpman (1997), the linear form also implies transferable utility between the lobby and unitary government; this leads to equilibrium indeterminancies and an inability to speak to distributional questions, prompting them to generalize the PFS form.
+As noted in Dixit, Grossman and Helpman (1997), the linear form also implies transferable utility between the lobby and unitary government; this leads to equilibrium indeterminacies and an inability to speak to distributional questions, prompting them to generalize the PFS form.
 From their results, we have that the same equilibrium analysis holds in the more general environment.

### 2016-02-02 letter.tex
revise letter to editor/referee (final)

@@
-\documentclass[12pt]{article}
+\documentclass[12pt]{report}
 
@@
 \addtolength{\evensidemargin}{-.5in}
+\setlength{\parindent}{0pt}
 \setlength{\textheight}{8.5in}
@@
 
-Editor's point $\#$1: I have revised the introduction according to the formula you suggested.
+\vskip.2in
+I would like to begin by thanking you both for very thoughtful and helpful suggestions on the previous draft of the paper.
+I believe that each one contributed to improving the current draft.
+In particularly, following your comments, the first half of the paper has been almost entirely re-written.
+\\
+
+To make the review process as efficient as possible, I have bolded the changes made as requested.
+Note that I have only bolded text when it was newly-written or changed, not when it was moved, as were substantial portions of Sections 1 and 2.
+This letter details all changes beyond correction of typos and minor expository changes.
+I have tried as much as possible to list the responses here in the order the changes appear in the text.
+\\
+
+\textbf{Editor's point $\#$1 / Referee's point $\#$4}: I have revised the introduction according to the formula you suggested.
 I have toned down considerably the claims about the PFS puzzle as both you and the referee suggest; they are now at the very end of the introduction and in a soft way in the abstract.
+\\
 
-Editor's point $\#$2: I have created four subsections within Section 2 to structure the exposition of the model.
+\textbf{Referee's point $\#$1} (``Introduction is derailing at the start''): I agree and I thank you for your detailed suggestions about how to reformulate the introduction and incorporate a toned-down discussion about the PFS puzzle.
+Please see response to Editor's point $\#$1 for logistics.
+\\
+
+\textbf{Referee's point $\#$2} (``Intuition in the Introduction about political uncertainty''): There are several kinds of intuition about political uncertainty, and I struggled a bit to guess which kind was meant.
+Because I already give some intuition about the effects of political uncertainty on the results, I went with the interpretation that this meant intuition/interpretation about what I mean by political uncertainty itself.
+I have thus added a few sentences in the fifth paragraph of the introduction to this end.
+I would be happy to do something different/additional if I have interpreted this suggestion incorrectly.
+\\
+
+\textbf{Editor's point $\#$8}: I agree that my model does not nest PFS.
+The discussion in the new Section 2.4, footnote 17 and Section 5.2 is meant to highlight the similarities and differences between my model and the PFS model.
+I hope that I have addressed your concerns about over-selling the contribution relative to the PFS puzzle so that this is no longer an issue.
+\\
+
+\textbf{Editor's point $\#$4}: I have added a traded numeraire good to address the concern about balanced trade.
+As all the results hinge on the incentives of the lobby, the executive and the median legislator---none of which are impacted by the addition of the numeraire good---this does not change any of the results.
+In carefully examining this issue, I realized that it would be more complete to include labor income in the welfare functions of the executive and lobby (again, this would change no results, as they all involve either differencing two welfare terms or taking derivatives with respect to a tariff).
+However, since much of the literature I'm following abstracts from a balanced trade condition and a numeraire sector with labor input, this would complicate the exposition and comparison to the extant literature.
+I have therefore chosen to insert Footnote 14 to this effect.
+It would be a simple matter to put labor income into the welfare functions if that is judged preferable.
+\\
+
+\textbf{Editor's point $\#$2}: I have created four subsections within Section 2 to structure the exposition of the model.
 I think this is much clearer than the previous draft.
@@
 They are now labeled Assumptions 1 and 2, while the former Assumption 1 is now Assumption 3 because of this rearrangement.
+\\
 
-Editor's point $\#$3: Thank you for the suggestion to add an extensive form and to define the game more clearly and formally.
-I think the addition of the normal form in Figure 1 made this portion of the paper significantly more readable.
+\textbf{Editor's point $\#$3}: Thank you for the suggestion to add an extensive form and to define the game more clearly and formally.
+I think the addition of the normal form as Figure 1 made this portion of the paper significantly more readable.
 I hope you will agree.
@@
 The payoffs are gathered together in their own section (2.2: Preferences).
+\\
 
-Editor's point $\#$4: I have added a traded numeraire good to address the concern about balanced trade.
-As all the results hinge on the incentives of the lobby, the executive and the median legislator---none of which are impacted by the addition of the numeraire good---this does not change any of the results.
-In carefully examining this issue, I realized that it would be more complete to include labor income in the welfare functions of the executive and lobby (again, this would change no results, as they all involve either differencing two welfare terms or taking derivatives with respect to a tariff).
-However, since much of the literature I'm following abstracts from a balanced trade condition and a numeraire sector with labor input, this would complicate the exposition and comparison to the extant literature.
-I have therefore chosen to insert Footnote [{\color{blue}ADD FOOTNOTE REFERENCE}] to this effect.
-It would be a simple matter to put labor income into the welfare functions if that is judged preferable.
+\textbf{Editor's point $\#$9}: I have added the move of nature that determines whether the legislature gets the opportunity to break the agreement to the extensive form as well as to the discussion of the timing in the initial discussion of the model (Section 2.1: The Basic Setup).
+I have gathered all relevant discussion in this section as suggested.
+\\
 
-Editor's point $\#$5: I have added expectations operators to Equations (3) and (4) [{\color{blue}CHECK EQ:JV AND EQ:LV}] to address the redundant notation.
-Thank you for pointing out this needed clarification.
+\textbf{Editor's point $\#$5}: I have added expectations operators to Equations (2) and (3) to address the redundant notation. Thank you for pointing out this needed clarification. \\
 
-Editor's point $\#$6: I have rewritten the passage about the influence of lobbying on the executive to read
+\textbf{Editor's point $\#$6}: I have rewritten the passage about the influence of lobbying on the executive to read
 \begin{quote}
@@
 I think this is a softer claim than what was in the earlier draft and avoids getting into an interpretation of exactly what `Protection for Sale' means in this context since this is not central---what is important is to defend the modeling assumption directly.
-
 I have also removed the duplicate statement of Assumption 1 and combined redundant text surrounding the two statements.
 I have added $\forall e$ to the statement of Assumption 1, which is now relabeled Assumption 3 because I have moved the former Assumptions 2 and 3 earlier in Section 2.4 as part of the restructuring discussed in point $\#2$.
+\\
 
-Editor's point $\#$7: Thank you for pointing out that the second order condition holds everywhere and that the proof of what was formerly Lemma 1 can be drastically simplified.
-I moved the discussion of the second order condition into the text and integrated it with the simple result on the increasingness of trade war tariff in the political economy weight.
-As this did not need to be referenced elsewhere and no longer requires a proof in the appendix, it is no longer labeled Lemma 1.
-Likewise, Lemma 0 is no longer necessary and has been removed.
-Note that this advances the numbering of the lemmas in the revised draft relative to the previous draft.
-
-Editor's point $\#8$: I agree that my model does not nest PFS.
-The discussion in the new Section 2.4, footnote 17 and Section 5.2 is meant to highlight the similarities and differences.
-I hope that I have addressed your concerns about over-selling the contribution relative to the PFS puzzle so that this is no longer an issue.
-
-Editor's point $\#9$: I have added the move of nature that determines whether the legislature gets the opportunity to the extensive form and to the discussion of the timing in the initial discussion of the model (Section 2.1: The Basic Setup).
-I have gathered all relevant discussion in this section as suggested.
-
-Editor's point $\#10$: I have removed the extensions (Section 6) per your suggestion. {\color{blue}ADD discussion about fixing exposition around one break at a time}
-
-Referee's point $\#1$ (``Introduction is derailing at the start''): I agree and I thank you for your detailed suggestions about how to reformulate the introduction and incorporate a toned-down discussion about the PFS puzzle.
-Please see respone to Editor's point $\#$1 for logistics.
-
-Referee's point $\#2$ (``Intuition in the Introduction about political uncertainty''): There are several kinds of intuition about political uncertainty, and I struggled a bit to guess which kind was meant.
-Because I already give some intuition about the effects of political uncertainty on the results, I went with the interpretation that this meant intuition/interpretation about what I mean by political uncertainty itself.
-I have thus added a few sentences in the fifth paragraph of the introduction to this end.
-I would be happy to do something different/additional if I have interpreted this suggestion incorrectly.
-
-Referee's point $\#3$ (``Parliamentary democracies''): I have added a paragraph at the very end of Section 2 to address this concern.
+\textbf{Referee's point $\#$3} (``Parliamentary democracies''): I have added a paragraph at the very end of Section 2 to address this concern.
 I would also note, as a point of interest, that the number of veto points does not always correlate particularly well with the political system.
@@
 R.
-Reich (2015), ``Political Strategies for Health Reform in Turkey: Extending Veto Point Theory,'' \textit{Health Systems and Reform}, 1(4), 263-275.}
+Reich (2015), ``Political Strategies for Health Reform in Turkey: Extending Veto Point Theory,'' \textit{Health Systems and Reform}, 1(4), 263-275.} \\
 
-Referee's point $\#4$ (``Strong language regarding contribution''): I agree. Please see response to Editor's point $\#$1.
+\textbf{Editor's point $\#$7}: Thank you for pointing out that the second order condition holds everywhere and that the proof of what was formerly Lemma 1 can be drastically simplified.
+I moved the discussion of the second order condition into the text (Section 3.1) and integrated it with the simple result on the increasingness of trade war tariff in the political economy weight.
+As this did not need to be referenced elsewhere and no longer requires a proof in the appendix, it is no longer labeled Lemma 1.
+Likewise, Lemma 0 is no longer necessary and has been removed.
+Note that this advances the numbering of the lemmas in the revised draft relative to the previous draft; this should not cause too much trouble as there were no changes to any of the results aside from the statement of Result 2 (see below).
+\\
 
-\vskip.5in
-Other changes
+\textbf{Editor's point $\#$10}: I have removed the extensions (Section 6) per your suggestion.
+Some of the discussion about the assumption that the legislatures do not take simultaneous decisions to break the agreement was moved into Section 2.1.
+\\
+
+I made two other small changes:
 \begin{itemize}
 	\item I have removed the statement ``and the distribution of $\gamma(e,\theta)$ is weakly increasing in $e$'' from Result 2 because this is the content of Assumption 2 and does not need to be repeated.
-	\item I have added some newer references to the uncertainty and trade agreements literature to the literature review and updated working papers where possible.
+	\item I have added some newer references to the uncertainty and trade agreements literature to the literature review and updated working papers where possible.\\
 \end{itemize}
 
+Again, I thank you for your thoughtful and thorough feedback and look forward to your further comments. \\
+
+Best wishes,\\
+\\
+\\
+Kristy Buzard
+
 \end{document}

### 2016-02-02 draft.tex
final proofread before submission

@@
 Thus trade agreements act as a domestic political commitment device: executives set relatively high tariffs to discourage lobbying and increase the chance that the agreement will be ratified.
-\textbf{This finding sheds light on the empirical puzzle surrounding governments' welfare weights in the Grossman and Helpman (1994) model, while the addition of political uncertainty adds important realism so that the model presented here can speak to important trade policy-making events such as the four-year delay in the ratification of the U.S.-Korea Free Trade Agreement.}
+\textbf{The model sheds light on the empirical puzzle surrounding governments' welfare weights in the Grossman and Helpman (1994) model and can speak to important trade policy-making events such as the four-year delay in the ratification of the U.S.-Korea Free Trade Agreement.}
 
@@
 
-\textbf{Assuming for now that lobbies represent only import-competing interests, lobbies in this environment would like to prevent or derail a trade agreement that would reduce their tariff protection.
+\textbf{Assuming that lobbies represent only import-competing interests, lobbies in this environment would like to prevent or derail a trade agreement that would reduce their tariff protection.
 They work toward this end by exerting effort to convince their legislatures to vote against the trade agreement but face some uncertainty about whether their efforts will result in a voting outcome consistent with their interests.
@@
 Further, the lobby decreases its effort as the tariff it would face under the trade agreement rises: it has less to lose the higher is the trade agreement tariff.
-When forming the trade agreement tariff, the executives trade off the welfare they receive under the agreement with the probability that the agreement actually remains in force.
+When forming the trade agreement tariff, the executives therefore trade off the welfare they receive under the agreement with the probability that the agreement actually remains in force.
 Ultimately, I find that observed lobbying effort may often be low because the executive branches of the governments set trade agreement tariffs sufficiently high to reduce political pressure on the legislatures.
-I also show that other examples of realism are gained when political uncertainty is added to the model: for instance, the model rationalizes the four-year delay and eventual ratification of the U.S.-Korea Free Trade Agreement.}
+There are other ways in which realism is gained when political uncertainty is added to the model: for instance, the model rationalizes the four-year delay and eventual ratification of the U.S.-Korea Free Trade Agreement.}
 
@@
 automakers, after which they removed their objections and the agreement was ratified by Congress in 2011 and entered into force in 2012.
-In a certain world, this ratification failure should not have occurred, and it was remedied by making the trade agreement more appealing to a key lobby and thus reducing its lobbying effort.
-\footnote{See \url{http://www.ustr.gov/trade-agreements/free-trade-agreements/korus-fta}.
+In a certain world, this ratification failure should not have occurred, and it was remedied by making the trade agreement more appealing to a key lobby and thus reducing its lobbying effort.\footnote{See \url{http://www.ustr.gov/trade-agreements/free-trade-agreements/korus-fta}.
 In this case, automakers are both importers and exporters, and the concessions were concentrated on market access for auto exports in South Korea.
@@
 
-\textbf{The addition of a rich structure of government}, tailored carefully to the trade policy making process, helps shed light on the empirical puzzle surrounding the PFS model.
+\textbf{The addition of a rich structure of government}, tailored carefully to the trade policy making process, also helps shed light on the empirical puzzle surrounding the PFS model.
 Empirical tests of the model\footnote{cfr.
@@
 The ``case of the missing contributions,'' as Gawande and Hoekman (2006) refer to it, turns out to be well explained by an equilibrium phenomenon arising from the agenda-setting power of the executive branch.
-In particular, \textbf{in this model, there is inconsistency for the approximately 15$\%$ of sectors that receive protection but apparently put forth no lobbying effort.
+In particular, \textbf{in this model, there is no inconsistency for the approximately 15$\%$ of sectors that receive protection but apparently put forth no lobbying effort.
 In this framework, even a welfare-maximizing executive and lobbying effort at zero can be reconciled with high tariff levels.}
@@
 \textbf{After this}, the uncertainty about the the median legislator's identity is resolved.
-This is modeled as the realization of a random variable $\ta$, the value of which all players observe simultaneously.
+This is modeled as the realization of a random variable $\ta_b$, the value of which all players observe simultaneously.
 The ``break stage'' concludes with the legislature making a choice to abide by the agreement or to break the agreement.
@@
 Note that this is identical to the welfare function for the legislature aside from the weight $\ga_E$ the executive places on the profits of the import industry, which is not a function of lobbying effort.
-\textbf{Section~\ref{sec:choices} discusses these modeling choices in depth.}
+\textbf{Section~\ref{sec:choices} contains an in-depth discussion of these modeling choices.}
 
@@
 
-\textbf{Finally, I assume that an external authority can ensure enforcement of the agreement in the one-shot game analyzed here.}\footnote{Buzard (\textbf{2015}) extends the analysis to a repeated-game context and can therefore remove this restriction.}
+\textbf{Finally, I assume that an external authority can ensure enforcement of the agreement in the one-shot game analyzed here.}\footnote{Buzard (\textbf{2015}) considers a similar model in a repeated-game context and can therefore remove this restriction.}
 
@@
 
-Notice that, aside from the endogeneity of the weight the legislature places on the lobbying industry's profits, \textbf{Equation~\ref{eq:ml}} is precisely the \textit{deus ex machina} government objective function popularized by Baldwin (1987) that is commonly employed in the literature on the political economy of trade agreements.
+Now, notice that, aside from the endogeneity of the weight the legislature places on the lobbying industry's profits, \textbf{Equation~\ref{eq:ml}} is precisely the \textit{deus ex machina} government objective function popularized by Baldwin (1987) that is commonly employed in the literature on the political economy of trade agreements.
 \textbf{Instead of the weight on profits being constant as in the exogenous case or necessarily linear in contributions as in the PFS case, this form allows for both endogeneity and non-linearity.}
@@
 The formulation in Equation~\ref{eq:ml} is intended as a bridge between the two.
-See Buzard (\textbf{2016}) for an example of how this form can be used to integrate the analysis of endogenous political activity and exogenous shocks to the political welfare of governments.
-
-\textbf{In addition, letting the weight the median legislator places on the profits of the import-competing industry $\ga(e,\ve)$ be affected by the level of lobbying effort as well a random variable $\ta$} allows for the analysis of endogenous lobbying to be integrated with the standard political economy shocks of the trade agreements literature.
+\textbf{See Buzard (\textbf{2016}) for an example of how letting the weight the median legislator places on the profits of the import-competing industry $\ga(e,\ve)$ be affected by the level of lobbying effort as well a random variable $\ta$ can integrate both endogenous political activity and exogenous shocks into the analysis of questions regarding optimal trade agreement design and the political welfare of governments.}
+
 Modeling the objective function so closely on the standard in the trade agreements literature allows for direct comparisons to the large extant body of work that studies exogenous shocks only, revealing cleanly the effects of the addition of endogenous lobbying---and if one wants---this different kind of political uncertainty.
@@
 
-Therefore, in line with both the theoretical and empirical literature, I make the following, \textbf{final} assumption \textbf{on $\ga(e,\ta)$}:
+Therefore, in line with both the theoretical and empirical literature, I make \textbf{one more} assumption \textbf{on $\ga(e,\ta)$}:
 
@@
 
-\textbf{The results apply beyond separation-of-powers presidential systems because the presence of veto points is what is essential for the dynamic posited here to occur.
-Although the number of formal veto points tends to be lower in, for instance, parliamentary systems, veto points are typically not absent altogether (Henisz and Mansfield 2006).
+\textbf{We can see that the results apply beyond separation-of-powers presidential systems because the presence of veto points is what is essential for the dynamic posited here to occur.
+Although the number of formal veto points tends to be lower in, for instance, parliamentary systems, veto points are typically not absent altogether, at least in democracies (Henisz and Mansfield 2006).
 In reality, constraints on the executive are almost always present in democratic regimes, although they may operate in different ways, e.g.
@@
 To understand how the executives optimally structure trade agreements, we must first examine the incentives of the lobbies and how the legislatures make decisions regarding breach of the trade agreement, including how trade-war tariffs are set.
-The symmetric structure of the model permits restriction of attention to the home country.
+The symmetric, separable structure of the model permits restriction of attention to the home country.
 
@@
 That is, at this stage, the lobby chooses the level of effort that equates its expected marginal increase in profits with its marginal payment.
-Further details about the political weighting function $\ga$ and assumed structure of uncertainty are provided in the next section.
 
@@
 \label{sec:break}
-With trade-war decisions specified, we can proceed to analyze the incentives regarding the legislature's decision to uphold or break the trade agreement.
+With trade-war decisions specified, we can proceed by backward induction to analyze the incentives regarding the legislature's decision to uphold or break the trade agreement.
 I assume that the form of the political weighting function is the same \textbf{regardless of the break decision}, but the model could be extended to allow the two to differ.
@@
 
-The executives' incentive to raise tariffs in order to reduce lobbying effort helps to explain the empirical finding in the Protection for Sale literature that levels of protection and associated deadweight losses are too high relative to lobbying expenditure given the high estimates for governments' weighting of social welfare.
+The executives' incentive to raise tariffs in order to reduce lobbying effort \textbf{provides insight into the empirical puzzle} in the Protection for Sale literature that levels of protection and associated deadweight losses are too high relative to lobbying expenditure given the high estimates for governments' weighting of social welfare.
 I have shown that this serves to mediate the relationship between preferences for contributions relative to social welfare and the tariffs that are provided.

### 2016-02-22 to-do2.tex
finalize list of editor comments

\documentclass[12pt]{article}

\addtolength{\textwidth}{1.4in}
\addtolength{\oddsidemargin}{-.7in} %left margin
\addtolength{\evensidemargin}{-.7in}
\setlength{\textheight}{8.5in}
\setlength{\topmargin}{0.0in}
\setlength{\headsep}{0.0in}
\setlength{\headheight}{0.0in}
\setlength{\footskip}{.5in}
\renewcommand{\baselinestretch}{1.0}
\setlength{\parindent}{0pt}
\linespread{1.1}

\usepackage{amssymb, amsmath, amsthm, bm}
\usepackage{graphicx,csquotes,verbatim}
\usepackage[backend=biber,block=space,style=authoryear]{biblatex}
\setlength{\bibitemsep}{\baselineskip}
\usepackage[american]{babel}
%dell laptop
\addbibresource{C:/Users/Kristy/Dropbox/Research/xBibs/tradeagreements.bib}
%\addbibresource{C:/Users/Kristy/Documents/Dropbox/Research/xBibs/tradeagreements.bib}
\renewcommand{\newunitpunct}{,}
\renewbibmacro{in:}{}

\DeclareMathOperator*{\argmax}{arg\,max}
\usepackage{xcolor}
\hbadness=10000

\newcommand{\ve}{\varepsilon}
\newcommand{\ov}{\overline}
\newcommand{\un}{\underline}
\newcommand{\ta}{\theta}
\newcommand{\al}{\alpha}
\newcommand{\Ta}{\Theta}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\Bt}{B(\bm{\tau^a})}
\newcommand{\bta}{\bm{\tau^a}}
\newcommand{\btn}{\bm{\tau^n}}
\newcommand{\ga}{\gamma}
\newcommand{\Ga}{\Gamma}
\newcommand{\de}{\delta}

\begin{document}
\begin{center}
RIE R$\&$R of SOP, 2nd Round\\
February 22, 2016
\end{center}

\vskip.1in

\begin{enumerate}
  \item What I would like you to do is write a much more focused, crispier version of the paper. The idea is sufficiently straightforward that it does not need 28 pages of exposition. Please reduce the length of the paper by half.
	\item Avoid changes of notation as you go along. This is totally unhelpful to the reader. Please pick one and stick to it throughout the paper.
	\item Avoid digressions, lengthy discussions, and overreach: you stretch your model far too much. Stick to the main story. The paper still degenerates far too often into a series of discussions that bring little additional insight.
	\item Section 2.4 either belongs to section 1.1 or should move as section 1.2.
	\item Section 2.3 is far too long.
	\item Your recall stuff far too often. Not all readers suffer from memory losses.
	\item Assuming that lobbies represent only import-competing interests is a strong assumption and it requires theoretical and/or empirical justification.
	\item Your model is static, so claiming that it rationalizes the four-year delay and eventual ratification of the US-SK FTA is a stretch. In particular, the interactions with automakers suggests a that a bargaining game under incomplete information would be more appropriate.
	\item Your assumption that the identity of the median voter is not known a priori is standard in probability voting models, so you may want to stress the analogy between your setting and such models.
  \item You claim that trade agreements act as a domestic political commitment device. To make this point formally, it would be nice to establish conditions under which, in equilibrium, “fast track authority” dominates cherry picking ex ante (in a version of your model with several import competing sectors).
	\item I don’t see how the move of Nature (which determines the probabilities o and o* is consistent with the idea that there is uncertainty about who is the median voter (which was part of your motivation). Assuming that o and o* are mutually exclusive is not great. It would be better to assume some timing (rather than have a simultaneous choice). Since the game is symmetric, assuming that Home’s legislature moves before Foreign’s would be without further loss of generality.
	\item “the uncertainty about the median legislator’s identity is resolved.” What is this uncertainty really about?
	\item The equilibrium selection criterion requires justification and motivation.
	\item Assumption 2 does not formalize any intuition: it is just a labelling.
	\item I feel uneasy about Assumption 3. I would rather have it as an outcome and, if it is not the only possibly one, a discussion of the conditions under which this arises.
	\item The full separability of sectors implies very minimal interactions already. Assuming a fully symmetric model is then highly restrictive. How difficult would it be to relax the latter assumption?
	\item Expression 6 establishes that the necessary second order condition holds by assumption 1, does not it?
	\item Writing that lobbying expenditures are equal to 0.0007 is meaningless absent units. Are these cents? Trillions of dollars? It would be much more informative if these were expressed as a fraction of equilibrium profits.
	\item The last paragraph of section 4 is quite a stretch; is there any evidence in the political science literature that the separation of power is correlated with trade openness among democracies?
\end{enumerate}
\end{document}

### 2016-02-22 letter-round2.tex
create letter for 2nd round revision

--- 2016-02-02 letter.tex
@@
 \large Response to Editor and Referee \\
-\normalsize February 2, 2016
+\normalsize Month ???, 2016
 \end{center}
 
-Editor's point $\#$1: I have revised the introduction according to the formula you suggested.
-I have toned down considerably the claims about the PFS puzzle as both you and the referee suggest; they are now at the very end of the introduction and in a soft way in the abstract.
+\vskip.2in
+Thank you. \\
 
-Editor's point $\#$2: I have created four subsections within Section 2 to structure the exposition of the model. I think this is much clearer than the previous draft.
+Logistics. Points were not numbered, so pasting them here\\
 
-Editor's point $\#$3: Thank you for the suggestion to add an extensive form and to define the game more clearly and formally.
-I think the addition of the normal form in Figure 1 made this portion of the paper significantly more readable.
-I hope you will agree.
-The discussion of the players, timing, and the presentation of the extensive form are at the end of Section 2.1 (The Basic Setup).
-I largely allow the extensive form to do the work of laying out the strategy spaces so as not to clutter the text.
-The payoffs are gathered together in their own section (2.2: Preferences).
-
-Editor's point $\#$4: I have added a traded numeraire good to address the concern about balanced trade.
-As all the results hinge on the incentives of the lobby, the executive and the median legislator---none of which are impacted by the addition of the numeraire good---this does not change any of the results.
-In carefully examining this issue, I realized that it would be more complete to include labor income in the welfare functions of the executive and lobby (again, this would change no results, as they all involve either differencing two welfare terms or taking derivatives with respect to a tariff).
-However, since much of the literature I'm following abstracts from a balanced trade condition and a numeraire sector with labor input, this would complicate the exposition and comparison to the extant literature.
-I have therefore chosen to insert Footnote [{\color{blue}ADD FOOTNOTE REFERENCE}] to this effect.
-It would be a simple matter to put labor income into the welfare functions if that is judged preferable.
-
-Editor's point $\#$5: I have added expectations operators to Equations (3) and (4) [{\color{blue}CHECK EQ:JV AND EQ:LV}] to address the redundant notation.
-Thank you for pointing out this needed clarification.
-
-Editor's point $\#$6: I have rewritten the passage about the influence of lobbying on the executive to read
-\begin{quote}
-  This setup does not strictly require that the executives are not lobbied.
-The assumption that the weight the executives place on import-competing profits is constant only requires that the extent to which they favor this industry cannot be \textit{directly} influenced by lobbying effort.
-The assumption is innocuous as long as the executives' preferences are not directly altered in a significant way by lobbying over trade.
-\end{quote}
-I think this is a softer claim than what was in the earlier draft and avoids getting into an interpretation of exactly what `Protection for Sale' means in this context since this is not central---what is important is to defend the modeling assumption directly.
-
-I have also removed the duplicate statement of Assumption 1 and combined redundant text surrounding the two statements. I have added $\forall e$ to the statement of Assumption 1.
-
-Editor's point $\#$7: Thank you for pointing out that the second order condition holds everywhere and that the proof of what was formerly Lemma 1 can be drastically simplified.
-I moved the discussion of the second order condition into the text and integrated it with the simple result on the increasingness of trade war tariff in the political economy weight.
-As this did not need to be referenced elsewhere and no longer requires a proof in the appendix, it is no longer labeled Lemma 1.
-Likewise, Lemma 0 is no longer necessary and has been removed.
-Note that this advances the numbering of the lemmas in the revised draft relative to the previous draft.
-
-Editor's point $\#8$: I agree that my model does not nest PFS.
-The discussion in the new Section 2.4, footnote 17 and Section 5.2 is meant to highlight the similarities and differences.
-I hope that I have addressed your concerns about over-selling the contribution relative to the PFS puzzle so that this is no longer an issue.
-
-Editor's point $\#9$: I have added the move of nature that determines whether the legislature gets the opportunity to the extensive form and to the discussion of the timing in the initial discussion of the model (Section 2.1: The Basic Setup).
-I have gathered all relevant discussion in this section as suggested.
-
-Editor's point $\#10$: I have removed the extensions (Section 6) per your suggestion. {\color{blue}ADD discussion about fixing exposition around one break at a time}
-
-Referee's point $\#1$ (``Introduction is derailing at the start''): I agree and I thank you for your detailed suggestions about how to reformulate the introduction and incorporate a toned-down discussion about the PFS puzzle.
-Please see respone to Editor's point $\#$1 for logistics.
-
-Referee's point $\#2$ (``Intuition in the Introduction about political uncertainty''): There are several kinds of intuition about political uncertainty, and I struggled a bit to guess which kind was meant.
-Because I already give some intuition about the effects of political uncertainty on the results, I went with the interpretation that this meant intuition/interpretation about what I mean by political uncertainty itself.
-I have thus added a few sentences in the fifth paragraph of the introduction to this end.
-I would be happy to do something different/additional if I have interpreted this suggestion incorrectly.
-
-Referee's point $\#3$ (``Parliamentary democracies''): I have added a paragraph at the very end of Section 2 to address this concern.
-I would also note, as a point of interest, that the number of veto points does not always correlate particularly well with the political system.
-I was not easily able to find empirical work on the number of veto points on trade issues for Turkey, but Sparkes, Bump and Reich (2015) were somewhat surprisingly able to identify \textit{seven} veto points for health issues in Turkey.\footnote{Sparkes, S.P., Bump, J.
-B.
-and M.
-R.
-Reich (2015), ``Political Strategies for Health Reform in Turkey: Extending Veto Point Theory,'' \textit{Health Systems and Reform}, 1(4), 263-275.}
-
-Referee's point $\#4$ (``Strong language regarding contribution''): I agree. Please see response to Editor's point $\#$1.
-
-\vskip.5in
-Other changes
-\begin{itemize}
-	\item I have removed the statement ``and the distribution of $\gamma(e,\theta)$ is weakly increasing in $e$'' from Result 2 because this is the content of Assumption 2 and does not need to be repeated.
-\end{itemize}
+Best wishes,\\
+\\
+\\
+Kristy Buzard
 

### 2016-02-22 letter-round2.tex
paste text of editor's comments into letter

@@
 
+\textit{The main issue is still point 2 of my previous report: the paper is fundamentally quite interesting but poorly executed.
+What I would like you to do is write a much more focused, crispier version of the paper.
+The idea is sufficiently straightforward that it does not need 28 pages of exposition.
+Please reduce the length of the paper by half.}
+
+\vskip.3in
+Stylistic advice:
+
+\begin{itemize}
+	\item Avoid changes of notation as you go along. This is totally unhelpful to the reader. Please pick one and stick to it throughout the paper.
+	\item Avoid digressions, lengthy discussions, and overreach: you stretch your model far too much.
+Stick to the main story.
+The paper still degenerates far too often into a series of discussions that bring little additional insight.
+	\item Section 2.4 either belongs to section 1.1 or should move as section 1.2. Section 2.3 is far too long.
+	\item Your recall stuff far too often. Not all readers suffer from memory losses.
+\end{itemize}
+
+\vskip.3in
+Some comments on substance, chronologically:
+\begin{itemize}
+	\item Assuming that lobbies represent only import-competing interests is a strong assumption and it requires theoretical and/or empirical justification.
+	\item Your model is static, so claiming that it rationalizes the four-year delay and eventual ratification of the US-SK FTA is a stretch.
+In particular, the interactions with automakers suggests a that a bargaining game under incomplete information would be more appropriate.
+  \item Your assumption that the identity of the median voter is not known a priori is standard in probability voting models, so you may want to stress the analogy between your setting and such models.
+  \item You claim that trade agreements act as a domestic political commitment device.
+To make this point formally, it would be nice to establish conditions under which, in equilibrium, “fast track authority” dominates cherry picking ex ante (in a version of your model with several import competing sectors).
+  \item I don’t see how the move of Nature (which determines the probabilities o and o* is consistent with the idea that there is uncertainty about who is the median voter (which was part of your motivation).
+Assuming that o and o* are mutually exclusive is not great.
+It would be better to assume some timing (rather than have a simultaneous choice).
+Since the game is symmetric, assuming that Home’s legislature moves before Foreign’s would be without further loss of generality.
+  \item “the uncertainty about the median legislator’s identity is resolved.” What is this uncertainty really about?
+  \item The equilibrium selection criterion requires justification and motivation.
+  \item Assumption 2 does not formalize any intuition: it is just a labelling.
+  \item I feel uneasy about Assumption 3. I would rather have it as an outcome and, if it is not the only possibly one, a discussion of the conditions under which this arises.
+  \item The full separability of sectors implies very minimal interactions already.
+Assuming a fully symmetric model is then highly restrictive.
+How difficult would it be to relax the latter assumption?
+  \item Expression 6 establishes that the necessary second order condition holds by assumption 1, does not it?
+  \item Writing that lobbying expenditures are equal to 0.0007 is meaningless absent units.
+Are these cents?
+Trillions of dollars?
+It would be much more informative if these were expressed as a fraction of equilibrium profits.
+  \item The last paragraph of section 4 is quite a stretch; is there any evidence in the political science literature that the separation of power is correlated with trade openness among democracies?
+\end{itemize}
+
 Best wishes,\\

### 2016-02-22 to-do2.tex
number to do list to correspond to issue numbers

@@
 \begin{enumerate}
-  \item What I would like you to do is write a much more focused, crispier version of the paper.
+  \item[11.] What I would like you to do is write a much more focused, crispier version of the paper.
 The idea is sufficiently straightforward that it does not need 28 pages of exposition.
 Please reduce the length of the paper by half.
-	\item Avoid changes of notation as you go along. This is totally unhelpful to the reader. Please pick one and stick to it throughout the paper.
-	\item Avoid digressions, lengthy discussions, and overreach: you stretch your model far too much.
+	\item[12.] Avoid changes of notation as you go along. This is totally unhelpful to the reader. Please pick one and stick to it throughout the paper.
+	\item[13.] Avoid digressions, lengthy discussions, and overreach: you stretch your model far too much.
 Stick to the main story.
 The paper still degenerates far too often into a series of discussions that bring little additional insight.
-	\item Section 2.4 either belongs to section 1.1 or should move as section 1.2.
-	\item Section 2.3 is far too long.
-	\item Your recall stuff far too often. Not all readers suffer from memory losses.
-	\item Assuming that lobbies represent only import-competing interests is a strong assumption and it requires theoretical and/or empirical justification.
-	\item Your model is static, so claiming that it rationalizes the four-year delay and eventual ratification of the US-SK FTA is a stretch.
+	\item[14.] Section 2.4 either belongs to section 1.1 or should move as section 1.2.
+	\item[15.] Section 2.3 is far too long.
+	\item[16.] Your recall stuff far too often. Not all readers suffer from memory losses.
+	\item[17.] Assuming that lobbies represent only import-competing interests is a strong assumption and it requires theoretical and/or empirical justification.
+	\item[18.] Your model is static, so claiming that it rationalizes the four-year delay and eventual ratification of the US-SK FTA is a stretch.
 In particular, the interactions with automakers suggests a that a bargaining game under incomplete information would be more appropriate.
-	\item Your assumption that the identity of the median voter is not known a priori is standard in probability voting models, so you may want to stress the analogy between your setting and such models.
-  \item You claim that trade agreements act as a domestic political commitment device.
+	\item[19.] Your assumption that the identity of the median voter is not known a priori is standard in probability voting models, so you may want to stress the analogy between your setting and such models.
+  \item[20.] You claim that trade agreements act as a domestic political commitment device.
 To make this point formally, it would be nice to establish conditions under which, in equilibrium, “fast track authority” dominates cherry picking ex ante (in a version of your model with several import competing sectors).
-	\item I don’t see how the move of Nature (which determines the probabilities o and o* is consistent with the idea that there is uncertainty about who is the median voter (which was part of your motivation).
+	\item[21.] I don’t see how the move of Nature (which determines the probabilities o and o* is consistent with the idea that there is uncertainty about who is the median voter (which was part of your motivation).
 Assuming that o and o* are mutually exclusive is not great.
@@
 Since the game is symmetric, assuming that Home’s legislature moves before Foreign’s would be without further loss of generality.
-	\item “the uncertainty about the median legislator’s identity is resolved.” What is this uncertainty really about?
-	\item The equilibrium selection criterion requires justification and motivation.
-	\item Assumption 2 does not formalize any intuition: it is just a labelling.
-	\item I feel uneasy about Assumption 3. I would rather have it as an outcome and, if it is not the only possibly one, a discussion of the conditions under which this arises.
-	\item The full separability of sectors implies very minimal interactions already.
+	\item[22.] “the uncertainty about the median legislator’s identity is resolved.” What is this uncertainty really about?
+	\item[23.] The equilibrium selection criterion requires justification and motivation.
+	\item[24.] Assumption 2 does not formalize any intuition: it is just a labelling.
+	\item[25.] I feel uneasy about Assumption 3. I would rather have it as an outcome and, if it is not the only possibly one, a discussion of the conditions under which this arises.
+	\item[26.] The full separability of sectors implies very minimal interactions already.
 Assuming a fully symmetric model is then highly restrictive.
 How difficult would it be to relax the latter assumption?
-	\item Expression 6 establishes that the necessary second order condition holds by assumption 1, does not it?
-	\item Writing that lobbying expenditures are equal to 0.0007 is meaningless absent units.
+	\item[27.] Expression 6 establishes that the necessary second order condition holds by assumption 1, does not it?
+	\item[28.] Writing that lobbying expenditures are equal to 0.0007 is meaningless absent units.
 Are these cents?
@@
 It would be much more informative if these were expressed as a fraction of equilibrium profits.
-	\item The last paragraph of section 4 is quite a stretch; is there any evidence in the political science literature that the separation of power is correlated with trade openness among democracies?
+	\item[29.] The last paragraph of section 4 is quite a stretch; is there any evidence in the political science literature that the separation of power is correlated with trade openness among democracies?
 \end{enumerate}

### 2016-02-23 draft.tex
remove bold / color commands from first round revision

@@
 Thus trade agreements act as a domestic political commitment device: executives set relatively high tariffs to discourage lobbying and increase the chance that the agreement will be ratified.
-\textbf{The model sheds light on the empirical puzzle surrounding governments' welfare weights in the Grossman and Helpman (1994) model and can speak to important trade policy-making events such as the four-year delay in the ratification of the U.S.-Korea Free Trade Agreement.}
+The model sheds light on the empirical puzzle surrounding governments' welfare weights in the Grossman and Helpman (1994) model and can speak to important trade policy-making events such as the four-year delay in the ratification of the U.S.-Korea Free Trade Agreement.
 
@@
 
-\textbf{Much of trade policy is determined in the context of international agreements that must be ratified by some domestic body.
+Much of trade policy is determined in the context of international agreements that must be ratified by some domestic body.
 This paper seeks to shed light on such a policy-making process by incorporating Protection-for-Sale (Grossman and Helpman 1994, hereafter PFS) style lobbying into a model of trade agreements in which the executive branches of the governments set tariff levels in anticipation of political pressure upon a ratifying body, assumed here to be a legislature.
 This political pressure may be interpreted either as lobbying against ratification of the trade agreement or lobbying to break the agreement once in force.
-Important realism is achieved not only through the addition of the ratification constraint but also by the incorporation of political uncertainty into the model.}
-
-\textbf{Assuming that lobbies represent only import-competing interests, lobbies in this environment would like to prevent or derail a trade agreement that would reduce their tariff protection.
+Important realism is achieved not only through the addition of the ratification constraint but also by the incorporation of political uncertainty into the model.
+
+Assuming that lobbies represent only import-competing interests, lobbies in this environment would like to prevent or derail a trade agreement that would reduce their tariff protection.
 They work toward this end by exerting effort to convince their legislatures to vote against the trade agreement but face some uncertainty about whether their efforts will result in a voting outcome consistent with their interests.
@@
 Ultimately, I find that observed lobbying effort may often be low because the executive branches of the governments set trade agreement tariffs sufficiently high to reduce political pressure on the legislatures.
-There are other ways in which realism is gained when political uncertainty is added to the model: for instance, the model rationalizes the four-year delay and eventual ratification of the U.S.-Korea Free Trade Agreement.}
-
-\textbf{To see how this works,} consider first the case of political certainty, and assume that the executive is more pro-social than the legislature, as has been the case throughout the post-war period in the United States.\footnote{The standard assumption that the executive is less protectionist than the legislature is a special case of the finding that susceptibility to special interests generally declines with constituency size.
+There are other ways in which realism is gained when political uncertainty is added to the model: for instance, the model rationalizes the four-year delay and eventual ratification of the U.S.-Korea Free Trade Agreement.
+
+To see how this works, consider first the case of political certainty, and assume that the executive is more pro-social than the legislature, as has been the case throughout the post-war period in the United States.\footnote{The standard assumption that the executive is less protectionist than the legislature is a special case of the finding that susceptibility to special interests generally declines with constituency size.
 One simple illustration from the realm of trade policy is the following: a legislator whose district has a large concentration of a particular industry does not take into account the impact of tariffs on the welfare of consumers in other districts, while the executive, whose constituency encompasses the whole country, will internalize these diffuse consumption effects.
@@
 Thus, any trade agreement under political certainty involves a tariff high enough to disengage the lobby completely; that is, even welfare-maximizing executives set high tariffs and induce zero contributions in equilibrium because they set policy in the shadow of a ratification decision influenced by lobbying.
-Careful modeling of the political process demonstrates that \textbf{the relationship between deadweight loss calculations and the tariffs that are provided in a trade agreement may be mediated in important ways by the ratification constraint.}
-
-The fact that \textit{zero} contributions are predicted in equilibrium is not particularly satisfying.
+Careful modeling of the political process demonstrates that the relationship between deadweight loss calculations and the tariffs that are provided in a trade agreement may be mediated in important ways by the ratification constraint.
+
+The fact that zero contributions are predicted in equilibrium is not particularly satisfying.
 The addition of political uncertainty provides the missing realism, smooths the results and delivers additional intuition.
 I assume all actors are uncertain about the weight the median legislator in a non-unitary legislature places on the profits of the lobbying industry.
-\textbf{This is meant to represent the idea that the lobbying and voting ``game'' that goes on within legislatures is often complex enough that none of the actors know precisely what the outcome will be---that is, no one can predict exactly which legislator will be the median and therefore what weight will be placed on import-industry profits at the time of the vote.
-For instance, perhaps} a small number of lobbied legislators cannot deliver the votes of their non-lobbied colleagues with certainty.
-\textbf{One can think of this uncertainty as an additional influence on the legislature's preferences about which all parties---executives, legislators and lobbies alike---are uncertain.}
-
-In this case \textbf{of political uncertainty}, I show that the legislatures disrupt trade agreements with a higher probability and set higher trade war tariffs when lobbying effort increases.
+This is meant to represent the idea that the lobbying and voting ``game'' that goes on within legislatures is often complex enough that none of the actors know precisely what the outcome will be---that is, no one can predict exactly which legislator will be the median and therefore what weight will be placed on import-industry profits at the time of the vote.
+For instance, perhaps a small number of lobbied legislators cannot deliver the votes of their non-lobbied colleagues with certainty.
+One can think of this uncertainty as an additional influence on the legislature's preferences about which all parties---executives, legislators and lobbies alike---are uncertain.
+
+In this case of political uncertainty, I show that the legislatures disrupt trade agreements with a higher probability and set higher trade war tariffs when lobbying effort increases.
 Because the lobbies respond to higher trade agreement tariffs by decreasing effort (and therefore the probability the agreement will fail to be ratified or will be broken), the executives must trade off the level of social welfare derived while a trade agreement is in force---since they prefer low tariffs---and the chances that the agreement will remain in force.
@@
 
-\textbf{The addition of a rich structure of government}, tailored carefully to the trade policy making process, also helps shed light on the empirical puzzle surrounding the PFS model.
+The addition of a rich structure of government, tailored carefully to the trade policy making process, also helps shed light on the empirical puzzle surrounding the PFS model.
 Empirical tests of the model\footnote{cfr.
@@
 The ``case of the missing contributions,'' as Gawande and Hoekman (2006) refer to it, turns out to be well explained by an equilibrium phenomenon arising from the agenda-setting power of the executive branch.
-In particular, \textbf{in this model, there is no inconsistency for the approximately 15$\%$ of sectors that receive protection but apparently put forth no lobbying effort.
-In this framework, even a welfare-maximizing executive and lobbying effort at zero can be reconciled with high tariff levels.}
+In particular, in this model, there is no inconsistency for the approximately 15$\%$ of sectors that receive protection but apparently put forth no lobbying effort.
+In this framework, even a welfare-maximizing executive and lobbying effort at zero can be reconciled with high tariff levels.
 
@@
 These studies (e.g.
-Feenstra and Lewis (1991); Milner and Rosendorff (2001); Bagwell and Staiger (2005); Beshkar (2010); \textbf{Horn, Maggi and Staiger (2010); Beshkar and Bond (2012); Amador and Bagwell (2013)}) derive various implications for the design of international trade agreements using a Baldwin (1987)-style government welfare function with exogenous shocks to the political-pressure parameter.
+Feenstra and Lewis (1991); Milner and Rosendorff (2001); Bagwell and Staiger (2005); Beshkar (2010); Horn, Maggi and Staiger (2010); Beshkar and Bond (2012); Amador and Bagwell (2013)) derive various implications for the design of international trade agreements using a Baldwin (1987)-style government welfare function with exogenous shocks to the political-pressure parameter.
 As explained in the next section, the model presented here is intended to be a bridge between this literature and that discussed above that focuses on endogenously-determined political economy weights.
@@
 
-\textbf{A very recent literature has begun to study the effects of trade agreements for reducing the uncertainty that firms (Handley 2014 and Handley and Lim\~{a}o 2015) and governments (Lim\~{a}o and Maggi 2015) face.}
+A very recent literature has begun to study the effects of trade agreements for reducing the uncertainty that firms (Handley 2014 and Handley and Lim\~{a}o 2015) and governments (Lim\~{a}o and Maggi 2015) face.
 
@@
 \label{sec:basic}
-\textbf{I begin by describing the basic economic setting within which trade occurs.
+I begin by describing the basic economic setting within which trade occurs.
 It is a three-good model with two countries: home (no asterisk) and foreign (asterisk).
@@
 Further, the demand functions for $X$ and $Y$ are taken to be identical and denoted $D(P_i)$ in home and $D(P_i^*)$ in foreign.
-$P_i$ denotes the home price of good $i \in \left\{N,X,Y\right\}$ and $P_i^*$ denotes the foreign price of good $i$.}
-
-\textbf{Good $N$ is produced with labor alone with constant returns to scale technology so that $Q_N = l_N$.
-I assume the aggregate labor supply $l$ is large enough to ensure that the output of good $N$ is large enough to guarantee balanced trade.} The supply functions for good $X$ are $Q_X(P_X)$ and $Q_X^*(P_X^*)$ and are assumed strictly increasing and twice continuously differentiable for all prices that elicit positive supply.
+$P_i$ denotes the home price of good $i \in \left\{N,X,Y\right\}$ and $P_i^*$ denotes the foreign price of good $i$.
+
+Good $N$ is produced with labor alone with constant returns to scale technology so that $Q_N = l_N$.
+I assume the aggregate labor supply $l$ is large enough to ensure that the output of good $N$ is large enough to guarantee balanced trade.
+The supply functions for good $X$ are $Q_X(P_X)$ and $Q_X^*(P_X^*)$ and are assumed strictly increasing and twice continuously differentiable for all prices that elicit positive supply.
 I also assume $Q_X^*(P_X) > Q_X(P_X)$ for any such $P_X$ so that the home country is a net importer of good $X$.
 The production structure for good $Y$ is taken to be symmetric, with both demand and supply such that the economy is separable in goods $X$ and $Y$.
-It is assumed that the production of goods $X$ and $Y$ requires, \textbf{in addition to labor,} the possession of a sector-specific factor that is available in inelastic supply and is non-tradable so that the income of owners of the specific factors is tied to the price of the good in whose production their factor is used.
+It is assumed that the production of goods $X$ and $Y$ requires, in addition to labor, the possession of a sector-specific factor that is available in inelastic supply and is non-tradable so that the income of owners of the specific factors is tied to the price of the good in whose production their factor is used.
 
@@
 where $M_X$ are home-county imports and $E_X^*$ are foreign exports of good $X$ and $E_Y$ are home-county exports and $M_Y^*$ are foreign imports of good $Y$.
-\textbf{Given the production technology and no trade taxes, the price of the numeraire is equal to one in both countries and on the world market.}
+Given the production technology and no trade taxes, the price of the numeraire is equal to one in both countries and on the world market.
 
@@
 As profits and producer surplus (identical in this model) in a sector are increasing in the price of its good, profits in the import-competing sector are also increasing in the domestic tariff.
-This economic fact, combined with the assumptions on specific factor ownership, is what motivates political activity \textbf{by import-competing lobbies}.
-
-\textbf{I now turn to describing the strategic players in the game.} In order to focus attention on protectionist political forces, I assume that only the import-competing industry in each country is politically-organized and able to lobby and that it is represented by a single lobbying organization (see Footnote~\ref{fn:lobby} for discussion of additional lobbies).
+This economic fact, combined with the assumptions on specific factor ownership, is what motivates political activity by import-competing lobbies.
+
+I now turn to describing the strategic players in the game.
+In order to focus attention on protectionist political forces, I assume that only the import-competing industry in each country is politically-organized and able to lobby and that it is represented by a single lobbying organization (see Footnote~\ref{fn:lobby} for discussion of additional lobbies).
 Each country's government is composed of two branches: an executive who can conclude trade agreements and a legislature that has final say on trade policy.
-\textbf{To summarize}, the political process is modeled as involving three players in each country: the lobby, the executive, and the legislature.
-
-\textbf{Figure~\ref{fig:ext} illustrates the timing of the game from the perspective of the home country.}\footnote{\textbf{This extensive form is for the case where the foreign legislature does not break the trade agreement.}} First, the executives set trade policy cooperatively within the context of an international agreement, \textbf{that is they choose the trade agreement tariffs $\bta = \left(\tau^a,\tau^{*a}\right)$}.
+To summarize, the political process is modeled as involving three players in each country: the lobby, the executive, and the legislature.
+
+Figure~\ref{fig:ext} illustrates the timing of the game from the perspective of the home country.\footnote{This extensive form is for the case where the foreign legislature does not break the trade agreement.} First, the executives set trade policy cooperatively within the context of an international agreement, that is they choose the trade agreement tariffs $\bta = \left(\tau^a,\tau^{*a}\right)$.
 After the trade agreement is concluded, the lobbies attempt to persuade the legislators in their respective countries to vote ``no'' on the ratification of the agreement; or alternatively, if there is no ratification hurdle, to break the trade agreement.\footnote{As will be seen shortly, these two interpretations are equivalent in this setting because tariffs in a ``trade war'' are assumed to be set non-cooperatively in the same manner as if there were no trade agreement.
-For simplicity, I refer to this as the ``break'' decision throughout.} \textbf{They do this by choice of lobbying effort $e_b$, which can take on any non-negative value.}
-
-\textbf{Next, nature determines with probabilities $o$ and $o^*$ whether each legislature will have the opportunity to take a vote to break the agreement.
-I assume that these probabilities are less than one-half and mutually exclusive.}\footnote{\textbf{While this assumption is made for purposes of exposition and tractability, one can think of $o$ and $o^*$ as being determined by events beyond the executives' or legislatures' control that determine whether or not the legislatures will be willing to consider the lobby's request; for example, they would be affected by the occurrence of an economic crisis that diverts the legislature's attention from less-pressing matters.}} \textbf{This allows me to focus on the key interaction between the domestic actors and abstract from the strategic interaction between the lobbies in the two countries.}\footnote{\textbf{When the legislatures in both countries have an opportunity to break the agreement simultaneously, an incentive for each lobby to free ride on the effort of the other makes the optimization problem of the executives more complex as they are faced with the problem of how best to exploit the free-riding dynamic.}}
-
-\textbf{After this}, the uncertainty about the the median legislator's identity is resolved.
+For simplicity, I refer to this as the ``break'' decision throughout.} They do this by choice of lobbying effort $e_b$, which can take on any non-negative value.
+
+Next, nature determines with probabilities $o$ and $o^*$ whether each legislature will have the opportunity to take a vote to break the agreement.
+I assume that these probabilities are less than one-half and mutually exclusive.\footnote{While this assumption is made for purposes of exposition and tractability, one can think of $o$ and $o^*$ as being determined by events beyond the executives' or legislatures' control that determine whether or not the legislatures will be willing to consider the lobby's request; for example, they would be affected by the occurrence of an economic crisis that diverts the legislature's attention from less-pressing matters.} This allows me to focus on the key interaction between the domestic actors and abstract from the strategic interaction between the lobbies in the two countries.\footnote{When the legislatures in both countries have an opportunity to break the agreement simultaneously, an incentive for each lobby to free ride on the effort of the other makes the optimization problem of the executives more complex as they are faced with the problem of how best to exploit the free-riding dynamic.}
+
+After this, the uncertainty about the the median legislator's identity is resolved.
 This is modeled as the realization of a random variable $\ta_b$, the value of which all players observe simultaneously.
@@
 	\end{center}
-	\caption{\textbf{Extensive Form\label{fig:ext}}}
+	\caption{Extensive Form\label{fig:ext}}
 \end{figure}
@@
 As the economy is fully separable and the economic and political structures are symmetric, I focus throughout the paper on the home country and the $X$-sector.
-The details are analogous for $Y$ and foreign.\footnote{\textbf{These are the relevant sectors for analysis because they are the sectors in which there are lobbies.}}
+The details are analogous for $Y$ and foreign.\footnote{These are the relevant sectors for analysis because they are the sectors in which there are lobbies.}
 
@@
 \end{equation}
-where $\mathit{CS}$ is consumer surplus, $\mathit{PS}$ is producer surplus, $\ga(e,\ta)$ is the weight placed on producer surplus (profits) in the import-competing industry, $e$ is lobbying effort, $\ta$ is a random variable, and $\mathit{TR}$ is tariff revenue.\footnote{\textbf{Labor income $l$ could also be included in both median legislator and executive welfare.
-I omit it because its inclusion alters none of the results and only serves to complicate the exposition.}} I model the decisions of the legislature as being taken by a median legislator \textbf{(more on this in Section~\ref{sec:choices}).
-Notice that} the weight the median legislator places on the profits of the import-competing industry, $\ga(e,\ve)$, is affected by the level of lobbying effort and a random variable $\ta$.
+where $\mathit{CS}$ is consumer surplus, $\mathit{PS}$ is producer surplus, $\ga(e,\ta)$ is the weight placed on producer surplus (profits) in the import-competing industry, $e$ is lobbying effort, $\ta$ is a random variable, and $\mathit{TR}$ is tariff revenue.\footnote{Labor income $l$ could also be included in both median legislator and executive welfare.
+I omit it because its inclusion alters none of the results and only serves to complicate the exposition.} I model the decisions of the legislature as being taken by a median legislator (more on this in Section~\ref{sec:choices}).
+Notice that the weight the median legislator places on the profits of the import-competing industry, $\ga(e,\ve)$, is affected by the level of lobbying effort and a random variable $\ta$.
 
@@
 \begin{equation}
-  {\color{blue} \expect \left[U_L \right]} = \Pr\left[ \text{TradeWar}(\bta,e_b,e_n) \right] \left[ \pi(\tau^{\mathit{n}}) - e_n \right] + \Pr\left[ \text{TradeAgreement}(\bta,e_b,e_n) \right] \pi(\tau^a) - e_b
+  \expect \left[U_L \right] = \Pr\left[ \text{TradeWar}(\bta,e_b,e_n) \right] \left[ \pi(\tau^{\mathit{n}}) - e_n \right] + \Pr\left[ \text{TradeAgreement}(\bta,e_b,e_n) \right] \pi(\tau^a) - e_b
   \label{eq:lv}
@@
 
-In the first stage \textbf{of the game}, the executives choose the trade agreement tariffs $\bta=\left(\tau^a,\tau^{*a} \right)$ via a negotiating process that I assume to be efficient.
+In the first stage of the game, the executives choose the trade agreement tariffs $\bta=\left(\tau^a,\tau^{*a} \right)$ via a negotiating process that I assume to be efficient.
 In this symmetric environment, it therefore maximizes the expected joint payoffs:
 \begin{multline}
-  {\color{blue}\expect \left[W_E(\bta) \right] + \expect \left[ W_E^{*}(\bta) \right]} = \Pr\left[ \text{TradeWar}(\bta) \right] \left[W_E(\bm{\tau}^{\mathit{n}}) + W_E^*(\bm{\tau}^{\mathit{n}}) \right] + \\ \Pr\left[ \text{TradeAgreement}(\bta) \right] \left[W_E(\bta) + W_E^*(\bta) \right]
+  \expect \left[W_E(\bta) \right] + \expect \left[ W_E^{*}(\bta) \right] = \Pr\left[ \text{TradeWar}(\bta) \right] \left[W_E(\bm{\tau}^{\mathit{n}}) + W_E^*(\bm{\tau}^{\mathit{n}}) \right] + \\ \Pr\left[ \text{TradeAgreement}(\bta) \right] \left[W_E(\bta) + W_E^*(\bta) \right]
   \label{eq:jv}
@@
 Note that this is identical to the welfare function for the legislature aside from the weight $\ga_E$ the executive places on the profits of the import industry, which is not a function of lobbying effort.
-\textbf{Section~\ref{sec:choices} contains an in-depth discussion of these modeling choices.}
+Section~\ref{sec:choices} contains an in-depth discussion of these modeling choices.
 
@@
 \label{sec:info}
-\textbf{I examine a simple class of equilibria that have the following features.}
-
-\textbf{First, on information.} Although uncertainty is present, information about it is symmetric.
+I examine a simple class of equilibria that have the following features.
+
+First, on information.
+Although uncertainty is present, information about it is symmetric.
 However, in line with the literature, I assume the lobby's contribution is not observable to the foreign legislature.
 The implication is that the lobby can directly influence only the home legislature, and so the influence of one country's lobby on the other country's legislature occurs only through the tariffs selected.\footnote{cfr.
-Grossman and Helpman (1995b), page 685.} \textbf{Given the game's structure and the assumption that lobbying effort is not observable across international borders, players in the same country can take advantage of more information than those who are in different countries.
-That is, the solution concept employed here is perfect public equilibrium (PPE).}
-
-\textbf{Second, whenever there is a possibility of multiple equilibria, I will focus on the one that maximizes the welfare of the executives. }
-
-\textbf{Finally, I assume that an external authority can ensure enforcement of the agreement in the one-shot game analyzed here.}\footnote{Buzard (\textbf{2015}) considers a similar model in a repeated-game context and can therefore remove this restriction.}
+Grossman and Helpman (1995b), page 685.} Given the game's structure and the assumption that lobbying effort is not observable across international borders, players in the same country can take advantage of more information than those who are in different countries.
+That is, the solution concept employed here is perfect public equilibrium (PPE).
+
+Second, whenever there is a possibility of multiple equilibria, I will focus on the one that maximizes the welfare of the executives.
+
+Finally, I assume that an external authority can ensure enforcement of the agreement in the one-shot game analyzed here.\footnote{Buzard (2015) considers a similar model in a repeated-game context and can therefore remove this restriction.}
 
@@
 
-In modeling the objective of the legislature \textbf{(see Equation~\ref{eq:ml}), I have introduced} a variant of the general political support function introduced by Hillman (1982).
+In modeling the objective of the legislature (see Equation~\ref{eq:ml}), I have introduced a variant of the general political support function introduced by Hillman (1982).
 While Hillman is credited with first proposing that policy makers choose tariffs in order to maximize their own political support, the PFS model contributed the explicit specification that lobbying contributions enter the political support function with constant marginal utility as $C + aW$.
@@
 
-Now, notice that, aside from the endogeneity of the weight the legislature places on the lobbying industry's profits, \textbf{Equation~\ref{eq:ml}} is precisely the \textit{deus ex machina} government objective function popularized by Baldwin (1987) that is commonly employed in the literature on the political economy of trade agreements.
-\textbf{Instead of the weight on profits being constant as in the exogenous case or necessarily linear in contributions as in the PFS case, this form allows for both endogeneity and non-linearity.}
+Now, notice that, aside from the endogeneity of the weight the legislature places on the lobbying industry's profits, Equation~\ref{eq:ml} is precisely the \textit{deus ex machina} government objective function popularized by Baldwin (1987) that is commonly employed in the literature on the political economy of trade agreements.
+Instead of the weight on profits being constant as in the exogenous case or necessarily linear in contributions as in the PFS case, this form allows for both endogeneity and non-linearity.
 
@@
 The formulation in Equation~\ref{eq:ml} is intended as a bridge between the two.
-\textbf{See Buzard (\textbf{2016}) for an example of how letting the weight the median legislator places on the profits of the import-competing industry $\ga(e,\ve)$ be affected by the level of lobbying effort as well a random variable $\ta$ can integrate both endogenous political activity and exogenous shocks into the analysis of questions regarding optimal trade agreement design and the political welfare of governments.}
+See Buzard (2016) for an example of how letting the weight the median legislator places on the profits of the import-competing industry $\ga(e,\ve)$ be affected by the level of lobbying effort as well a random variable $\ta$ can integrate both endogenous political activity and exogenous shocks into the analysis of questions regarding optimal trade agreement design and the political welfare of governments.
 
@@
 Alternatively, the integration of the random variable $\ta$ can represent reduced-form modeling of political uncertainty arising from a complex decision-making process within a legislature.
-\textbf{One can interpret $\ta$ as measuring uncertainty over} the identity of the median legislator in order to represent the idea that the lobbying and voting ``game'' that goes on within legislatures is often complex enough that none of the actors know precisely what the outcome will be---that is, no one can predict exactly which legislator will be the median and therefore what weight will be placed on import-industry profits at the time of the vote.
+One can interpret $\ta$ as measuring uncertainty over the identity of the median legislator in order to represent the idea that the lobbying and voting ``game'' that goes on within legislatures is often complex enough that none of the actors know precisely what the outcome will be---that is, no one can predict exactly which legislator will be the median and therefore what weight will be placed on import-industry profits at the time of the vote.
 Thus one way we can think about $\ta$ is as an additional influence on the legislature's preferences about which all parties---executives, legislators and lobbies alike---are uncertain.
@@
 Different assumptions on $\ga(\cdot,\cdot)$ correspond to different institutional features and will affect optimal lobbying- and tariff-setting behavior.
-\textbf{I make the following two basic assumptions on the weight on import-industry profits:}
+I make the following two basic assumptions on the weight on import-industry profits:
 
@@
 
-\textbf{The executives' objective function is taken to be identical to the legislatures' except that profits in the import-competing industry receive a fixed, exogenous weight.} This construction permits me to integrate the influence of the import-competing industry on negotiations prior to the formation of the trade agreement while also reflecting the idea that the mechanisms through which political considerations influence the executives' preferences appear to be quite different from those at the legislative level---the electoral college and primary/caucus systems for instance.
-
-\textbf{This setup does not strictly require that the executives are not lobbied.
+The executives' objective function is taken to be identical to the legislatures' except that profits in the import-competing industry receive a fixed, exogenous weight.
+This construction permits me to integrate the influence of the import-competing industry on negotiations prior to the formation of the trade agreement while also reflecting the idea that the mechanisms through which political considerations influence the executives' preferences appear to be quite different from those at the legislative level---the electoral college and primary/caucus systems for instance.
+
+This setup does not strictly require that the executives are not lobbied.
 The assumption that the weight the executives place on import-competing profits is constant only requires that the extent to which they favor this industry cannot be \textit{directly} influenced by lobbying effort.
 The assumption is innocuous as long as the executives' preferences are not directly altered in a significant way by lobbying over trade.
-}
+
 In the case of the post-war United States, where the Congress has consistently been significantly more protectionist than the President, this seems to reasonably reflect the political reality.
@@
 
-Therefore, in line with both the theoretical and empirical literature, I make \textbf{one more} assumption \textbf{on $\ga(e,\ta)$}:
+Therefore, in line with both the theoretical and empirical literature, I make one more assumption on $\ga(e,\ta)$:
 
 \begin{assumption}
-  $\ga(e,\ta) \geq \ga_E \geq 1 \ \forall \bm{e},\ta$.
+  $\ga(e,\ta) \geq \ga_E \geq 1 \ \forall e,\ta$.
   \label{as:ga_l_e}
@@
 
-That is, even for the least favorable outcome of the lobbying process, the legislature will be at least slightly more protectionist than the executive \textbf{regardless of the lobby's effort choice}.
-This assumption is not essential, but it matches well the empirical findings that politicians with larger constituencies are less sensitive to special interests, and it simplifies the analysis considerably, in particular ensuring that trade agreement tariffs are less than trade war tariffs \textbf{(See Destler (2005) and footnote~\ref{fn:ga_l_e} above)}.
+That is, even for the least favorable outcome of the lobbying process, the legislature will be at least slightly more protectionist than the executive regardless of the lobby's effort choice.
+This assumption is not essential, but it matches well the empirical findings that politicians with larger constituencies are less sensitive to special interests, and it simplifies the analysis considerably, in particular ensuring that trade agreement tariffs are less than trade war tariffs (See Destler (2005) and footnote~\ref{fn:ga_l_e} above).
 
@@
 
-\textbf{We can see that the results apply beyond separation-of-powers presidential systems because the presence of veto points is what is essential for the dynamic posited here to occur.
+We can see that the results apply beyond separation-of-powers presidential systems because the presence of veto points is what is essential for the dynamic posited here to occur.
 Although the number of formal veto points tends to be lower in, for instance, parliamentary systems, veto points are typically not absent altogether, at least in democracies (Henisz and Mansfield 2006).
@@
 It is important to consider that constraints may be formal or informal, and they may come from legislative, regulatory, judicial, sub-national or even other executive bodies.
-}
 
@@
 (Note this is Equation~\ref{eq:lv} simplified by the resolution of uncertainty over the legislature's decision on the trade agreement).
-This implies a first order condition \textbf{for the lobby} of
+This implies a first order condition for the lobby of
 \begin{equation}
-	\bm{\frac{\partial \pi(\tau^n)}{\partial \tau^n}\frac{\partial \tau^n}{\partial \ga} \frac{\partial \ga}{\partial e_n} = 1}
+	\frac{\partial \pi(\tau^n)}{\partial \tau^n}\frac{\partial \tau^n}{\partial \ga} \frac{\partial \ga}{\partial e_n} = 1
   \label{eq:lobtw}
@@
 
-\textbf{In accordance with intuition, because profits are increasing in the tariff, trade war tariffs are increasing in the weight attached to the profits of the import-competing industry.
+In accordance with intuition, because profits are increasing in the tariff, trade war tariffs are increasing in the weight attached to the profits of the import-competing industry.
 This can be seen by rearranging Equation~\ref{eq:lobtw} as
@@
 By the implicit function theorem, $\frac{\partial \tau^n}{\partial \ga} = -\frac{\frac{\partial \text{Equation}~\ref{eq:legfoc}}{\partial \ga}}{\frac{\partial \text{Equation}~\ref{eq:legfoc}}{\partial \tau^n}} = -\frac{\frac{\partial \pi(\tau^n)}{\partial \tau^n}}{SOC}$ where ``SOC'' stands for `Second Order Condition', the second derivative of the legislature's objective function which should be everywhere negative.
-Setting this expression equal to Equation~\ref{eq:lem1}, the second order condition must be satisfied since $\frac{\partial \ga}{\partial e_n}$ is positive by Assumption~\ref{as:ga_c}.}
+Setting this expression equal to Equation~\ref{eq:lem1}, the second order condition must be satisfied since $\frac{\partial \ga}{\partial e_n}$ is positive by Assumption~\ref{as:ga_c}.
 
@@
 With trade-war decisions specified, we can proceed by backward induction to analyze the incentives regarding the legislature's decision to uphold or break the trade agreement.
-I assume that the form of the political weighting function is the same \textbf{regardless of the break decision}, but the model could be extended to allow the two to differ.
+I assume that the form of the political weighting function is the same regardless of the break decision, but the model could be extended to allow the two to differ.
 What is in principle different are the lobbying efforts toward influencing the break decision, denoted $e_b$ at this stage to distinguish them from those that are made to influence the Nash tariff, $e_n$.
@@
 
-The legislature's decisions are stochastic, so the outcome of the vote on whether or not to break the trade agreement, as well as the level at which the tariff will be set in the case of a disruption, is not known to \textit{any} player until the uncertainty over the identity of the median legislator is resolved at the moment the vote takes place \textbf{on each issue}.\footnote{We might think of the trade-war tariff as being set in the legislation before a vote takes place in the case of administered protection for instance; the results are robust to this variant.} I represent the probability that the home legislature breaks the trade agreement and sets the tariff at $\tau^n$ as:
+The legislature's decisions are stochastic, so the outcome of the vote on whether or not to break the trade agreement, as well as the level at which the tariff will be set in the case of a disruption, is not known to \textit{any} player until the uncertainty over the identity of the median legislator is resolved at the moment the vote takes place on each issue.\footnote{We might think of the trade-war tariff as being set in the legislation before a vote takes place in the case of administered protection for instance; the results are robust to this variant.} I represent the probability that the home legislature breaks the trade agreement and sets the tariff at $\tau^n$ as:
 \begin{multline}
@@
 Note that the legislature's best response functions---both for whether to break the agreement, and at what level to set the tariff if it does---will be affected by the shape of $\ga(e,\ve)$.
-\textbf{Recall that we have made three basic assumptions on $\ga(e,\ve)$ in Section~\ref{sec:choices}.
-More restrictions could be added to deliver further results.}
+Recall that we have made three basic assumptions on $\ga(e,\ve)$ in Section~\ref{sec:choices}.
+More restrictions could be added to deliver further results.
 
@@
 
-\noindent \textbf{All proofs are in the Appendix.}
+\noindent All proofs are in the Appendix.
 
@@
 Note that the symmetric increase in the foreign tariff does not impact profits in the import-competing industry, but it can impact lobbying incentives through the break probability.
-However, there is an analogous effect on the break probability from raising $\tau^a$ and the net effect is negative given \textbf{Assumption~\ref{as:ga_c}} on the probability distribution function of $\ga(e,\ta)$, which ensures that the effect of lobbying effort on the break probability does not increase as $\bta$ increases.
+However, there is an analogous effect on the break probability from raising $\tau^a$ and the net effect is negative given Assumption~\ref{as:ga_c} on the probability distribution function of $\ga(e,\ta)$, which ensures that the effect of lobbying effort on the break probability does not increase as $\bta$ increases.
 This assumption can be significantly weakened: this effect can be positive as long as it is smaller in absolute value than the negative effect on trade agreement profits.
@@
 
-The executives' incentive to raise tariffs in order to reduce lobbying effort \textbf{provides insight into the empirical puzzle} in the Protection for Sale literature that levels of protection and associated deadweight losses are too high relative to lobbying expenditure given the high estimates for governments' weighting of social welfare.
+The executives' incentive to raise tariffs in order to reduce lobbying effort provides insight into the empirical puzzle in the Protection for Sale literature that levels of protection and associated deadweight losses are too high relative to lobbying expenditure given the high estimates for governments' weighting of social welfare.
 I have shown that this serves to mediate the relationship between preferences for contributions relative to social welfare and the tariffs that are provided.
@@
 
-\item \textbf{Amador, M., and Bagwell, K., 2013. The Theory of Optimal Delegation with an Application to Tariff Caps. Econometrica 81, 1541-1599.}
+\item Amador, M., and Bagwell, K., 2013. The Theory of Optimal Delegation with an Application to Tariff Caps. Econometrica 81, 1541-1599.
 
@@
 
-\item \textbf{Beshkar, M. and Bond, E., 2012. Cap and Escape in Trade Agreements. Available at \url{http://papers.ssrn.com/sol3/papers.cfm?abstract_id=2145679}.}
+\item Beshkar, M. and Bond, E., 2012. Cap and Escape in Trade Agreements. Available at \url{http://papers.ssrn.com/sol3/papers.cfm?abstract_id=2145679}.
 
@@
 
-\item Buzard, K., \textbf{2015}. Self-enforcing Trade Agreements, Dispute Settlement and Separation of Powers. Available at: \url{http://papers.ssrn.com/sol3/papers.cfm?abstract_id=2333728}.
-
-\item Buzard, K., \textbf{2016}. Endogenous Politics and the Design of Trade Agreements. Available at: \url{https://kbuzard.expressions.syr.edu/wp-content/uploads/Endogenous-Politics.pdf}.
+\item Buzard, K., 2015. Self-enforcing Trade Agreements, Dispute Settlement and Separation of Powers. Available at: \url{http://papers.ssrn.com/sol3/papers.cfm?abstract_id=2333728}.
+
+\item Buzard, K., 2016. Endogenous Politics and the Design of Trade Agreements. Available at: \url{https://kbuzard.expressions.syr.edu/wp-content/uploads/Endogenous-Politics.pdf}.
 
@@
 
-\item \textbf{Handley, K., 2014. Exporting under Trade Policy Uncertainty: Theory and Evidence. Journal of international Economics 94, 50-66.}
-
-\item \textbf{Handley, K., Lim\~{a}o, N., 2015. Trade and Investment under Policy Uncertainty: Theory and Firm Evidence. American Economic Journal: Economic Policy 7, 189-222.}
+\item Handley, K., 2014. Exporting under Trade Policy Uncertainty: Theory and Evidence. Journal of international Economics 94, 50-66.
+
+\item Handley, K., Lim\~{a}o, N., 2015. Trade and Investment under Policy Uncertainty: Theory and Firm Evidence. American Economic Journal: Economic Policy 7, 189-222.
 
@@
 
-\item \textbf{Horn, H., Maggi, G., Staiger, R. W., 2010. Trade Agreements as Endogenously Incomplete Contracts. American Economic Review 100, 394-419.}
+\item Horn, H., Maggi, G., Staiger, R. W., 2010. Trade Agreements as Endogenously Incomplete Contracts. American Economic Review 100, 394-419.
 
@@
 
-\item \textbf{Lim\~{a}o, N., Maggi, G., 2015. Uncertainty and Trade Agreements. American Economic Journal: Microeconomics 7, 1-42.}
+\item Lim\~{a}o, N., Maggi, G., 2015. Uncertainty and Trade Agreements. American Economic Journal: Microeconomics 7, 1-42.
 

### 2016-02-23 draft.tex
fix description of as:ga_ta, note in letter; closes #24

@@
 Assumption~\ref{as:ga_c} formalizes the intuition that the legislature favors the import-competing industry more the higher is its lobbying effort, but that there are diminishing returns to lobbying activity.\footnote{The diminishing returns here take the form of declining increments to the lobby's influence as effort increases; in Ethier (2012), the returns to lobbying decline as the level of protection increases.} It allows the political uncertainty inherent in $\ta$ to be correlated in interesting ways with the level of $e$, ruling out only (i) that higher effort levels make lower $\ga$-weights more likely and (ii) that effort and uncertainty are correlated in such a way that increasing lobbying effort changes the structure of uncertainty so that higher $\ga$-weights become more likely at an accelerating pace.
-Assumption~\ref{as:ga_ta} formalizes the intuition that larger realization of the uncertainty variable $\ta$ increase the value of the political economy weight.
+Assumption~\ref{as:ga_ta} \textbf{simply provides for an intuitive labelling so that} larger realization of the uncertainty variable $\ta$ increase the value of the political economy weight.
 

### 2016-02-23 letter-round2.tex
fix description of as:ga_ta, note in letter; closes #24
reply to SOC issue in letter; closes #27

@@
 \begin{itemize}
+\renewcommand\labelitemi{-}
 	\item Assuming that lobbies represent only import-competing interests is a strong assumption and it requires theoretical and/or empirical justification.
@@
   \item The equilibrium selection criterion requires justification and motivation.
-  \item Assumption 2 does not formalize any intuition: it is just a labelling.
+  \item \textit{Assumption 2 does not formalize any intuition: it is just a labelling.} \\
+	Quite right. The text has been modified appropriately.
   \item I feel uneasy about Assumption 3. I would rather have it as an outcome and, if it is not the only possibly one, a discussion of the conditions under which this arises.
@@
 How difficult would it be to relax the latter assumption?
-  \item Expression 6 establishes that the necessary second order condition holds by assumption 1, does not it?
+  \item \textit{Expression 6 establishes that the necessary second order condition holds by assumption 1, does not it?} \\
+	I agree that this is the case.
+As you so helpfully pointed this out in your last set of comments, I have included an argument to this effect in the two full sentences following Expression 6.
+If there is a more direct argument, I would be happy to replace the current one.
   \item Writing that lobbying expenditures are equal to 0.0007 is meaningless absent units.

### 2016-02-25 draft.tex
expand comments on autocracy vs. democracy at end of sec:example; addressed in letter; closes #29

@@
 
-If one interprets the addition of a second policy-making actor as consistent with a move from autocracy to democracy, this result helps to explain the extensive empirical evidence that democracies trade more with each other than autocracies.\footnote{cfr.
-Laver and Shepsle (1991); Henisz and Mansfield (2006); Aidt and Gassebner (2010).}
+If one interprets the addition of a second policy-making actor as consistent with a move from autocracy to democracy, this result helps to explain the extensive empirical evidence that democracies trade more with each other than autocracies (Laver and Shepsle 1991, Aidt and Gassebner 2010).
+In particular, Frye and Mansfield (2003) argue that the reduction in trade barriers becomes more likely as the number of veto points increases as countries move from autarky to democracy as exemplified by the transition of Communist regimes in the post-Cold-War era.
+Here, a more open trading stance emerged as groups with a liberal stance were integrated into the political process following the fall of the Berlin Wall.
+
+Crucially, Henisz and Mansfield (2006) demonstrate that the impact of adding veto players depends on the status quo policy.
+In the case of these Communist regimes, Cold War policy was near autarkic.
+The additional actors had an interest in liberalization.
+Henisz and Mansfield's careful empirical work shows that, among democracies, systems with more veto players as in a separation-of-powers structure will be more robust to retreats from liberal policies in the face of protectionist pressures than are those with fewer veto players.
+Conversely, it stands to reason that protectionist separation-of-power regimes would also be less susceptible to liberalization than unitary democractic regimes, \textit{ceteris paribus}.
 
@@
 
+\item Frye, T., Mansfield, E., 2003. Fragmenting Protection: The Political Economy of Trade Policy in the Post-Communist World. British Journal of Political Science 33, 633-657.
+
 \item Gawande, K., Bandyopadhyay, U., 2000. Is Protection for Sale? Evidence on the Grossman-Helpman Theory of Endogenous Protection. The Review of Economics and Statistics 82, 139-152.

### 2016-02-25 letter-round2.tex
expand comments on autocracy vs. democracy at end of sec:example; addressed in letter; closes #29

@@
 It would be much more informative if these were expressed as a fraction of equilibrium profits.
-  \item The last paragraph of section 4 is quite a stretch; is there any evidence in the political science literature that the separation of power is correlated with trade openness among democracies?
+  \item \textit{The last paragraph of section 4 is quite a stretch; is there any evidence in the political science literature that the separation of power is correlated with trade openness among democracies?} \\
+	The issue is quite a nuanced, in particular among democracies.
+I have discussed the empirical evidence on democracies---separate from the claim that I made about moving from autocracy to democracy---from Henisz and Mansfield (2006) in a new paragraph.
+I have also added a reference to Frye and Mansfield (2003) to support the original claim.
+I don't consider any of this essential and would be happy to remove either or both paragraphs.
 \end{itemize}

### 2016-02-26 to-do2.tex
reorder to do list for ease of implementation
move completed items to end of to-do list
add notes on timing change to to-do list

@@
 \begin{enumerate}
+	\item[14.] Section 2.4 either belongs to section 1.1 or should move as section 1.2.
+	\item[15.] Section 2.3 is far too long.
+	\item[17.] Assuming that lobbies represent only import-competing interests is a strong assumption and it requires theoretical and/or empirical justification.
+	\item[19.] Your assumption that the identity of the median voter is not known a priori is standard in probability voting models, so you may want to stress the analogy between your setting and such models.
+	\item[22.] “the uncertainty about the median legislator’s identity is resolved.” What is this uncertainty really about?
+	\item[23.] The equilibrium selection criterion requires justification and motivation.
+	\item[25.] I feel uneasy about Assumption 3. I would rather have it as an outcome and, if it is not the only possibly one, a discussion of the conditions under which this arises.
+	\item[28.] Writing that lobbying expenditures are equal to 0.0007 is meaningless absent units.
+Are these cents?
+Trillions of dollars?
+It would be much more informative if these were expressed as a fraction of equilibrium profits.
 	\item[11.] What I would like you to do is write a much more focused, crispier version of the paper.
@@
 The paper still degenerates far too often into a series of discussions that bring little additional insight.
-	\item[14.] Section 2.4 either belongs to section 1.1 or should move as section 1.2.
-	\item[15.] Section 2.3 is far too long.
 	\item[16.] Your recall stuff far too often. Not all readers suffer from memory losses.
-	\item[17.] Assuming that lobbies represent only import-competing interests is a strong assumption and it requires theoretical and/or empirical justification.
 	\item[18.] Your model is static, so claiming that it rationalizes the four-year delay and eventual ratification of the US-SK FTA is a stretch.
 In particular, the interactions with automakers suggests a that a bargaining game under incomplete information would be more appropriate.
-	\item[19.] Your assumption that the identity of the median voter is not known a priori is standard in probability voting models, so you may want to stress the analogy between your setting and such models.
   \item[20.] You claim that trade agreements act as a domestic political commitment device.
@@
 Since the game is symmetric, assuming that Home’s legislature moves before Foreign’s would be without further loss of generality.
-	\item[22.] “the uncertainty about the median legislator’s identity is resolved.” What is this uncertainty really about?
-	\item[23.] The equilibrium selection criterion requires justification and motivation.
-	\item[24.] Assumption 2 does not formalize any intuition: it is just a labelling.
-	\item[25.] I feel uneasy about Assumption 3. I would rather have it as an outcome and, if it is not the only possibly one, a discussion of the conditions under which this arises.
+		\begin{itemize}
+			\item From the point of view of existence of the executives' problem, I think this is feasible. The objective function changes in the following way:
+				\begin{enumerate}
+					\item Home (first mover) gets a break in one of two ways: getting its home legislature to break, or getting the foreign to do so: $Br + (1-Br)\cdot Br^* = Br + Br^* - Br \cdot Br^*$
+					\item There is no break when $(1-Br)(1-Br^*) = 1 - Br - Br^* + Br \cdot Br^*$
+				\end{enumerate}
+			\item Only difference in proof of existence is addition of $- Br \cdot Br^*$ term.
+Looks like I just distribute it out into the two earlier terms and have $(1 - Br)$ times the derivative, for example, which doesn't change the sign.
+			\item Current results should hold for second mover (foreign, WLOG).
+Either gets a chance to move, or is agreement is already broken and does nothing.
+But if it moves, decisions aren't conditioned on choice variables of home lobby (only legislature through break decision)
+				\begin{itemize}
+					\item I don't see why the legislature should make a different choice just because the foreign lobby/legislature may do something in the future
+				\end{itemize}
+			\item Need to explore for first mover (home)
+		\end{itemize}
 	\item[26.] The full separability of sectors implies very minimal interactions already.
@@
 How difficult would it be to relax the latter assumption?
+\end{enumerate}
+
+\vskip.5in
+Done
+\begin{enumerate}
+	\item[24.] Assumption 2 does not formalize any intuition: it is just a labelling.
 	\item[27.] Expression 6 establishes that the necessary second order condition holds by assumption 1, does not it?
-	\item[28.] Writing that lobbying expenditures are equal to 0.0007 is meaningless absent units.
-Are these cents?
-Trillions of dollars?
-It would be much more informative if these were expressed as a fraction of equilibrium profits.
 	\item[29.] The last paragraph of section 4 is quite a stretch; is there any evidence in the political science literature that the separation of power is correlated with trade openness among democracies?
 \end{enumerate}
+
 \end{document}

### 2016-02-27 to-do2.tex
add more notes on sequential timing to to-do

@@
 But if it moves, decisions aren't conditioned on choice variables of home lobby (only legislature through break decision)
+			\item Results need to be modified for first mover (home)
 				\begin{itemize}
-					\item I don't see why the legislature should make a different choice just because the foreign lobby/legislature may do something in the future
+					\item Legislature (equations 7 and 8 in the draft):
+						\begin{equation}
+							W_{ML}(\btn,\ga(e_b,\ve_b)) > W_{ML}\left(\bta,\ga(e_b,\ve_b)\right)
+							\label{eq:lwcg}
+						\end{equation}
+						becomes
+						\[ 
+						  W_{ML}(\btn,\ga(e_b,\ve_b)) > b^* W_{ML}\left(\btn,\ga(e_b,\ve_b)\right) + (1-b^*) W_{ML}\left(\bta,\ga(e_b,\ve_b)\right)
+						\]
+						This can be arranged as 
+						\[ 
+						  W_{ML}(\btn,\ga(e_b,\ve_b)) - W_{ML}(\bta,\ga(e_b,\ve_b)) > b^* \left[ W_{ML}\left(\btn,\ga(e_b,\ve_b)\right) - W_{ML}\left(\bta,\ga(e_b,\ve_b)\right) \right]
+						\]
+						This has the interesting structure that it is always true when (\ref{eq:lwcg}) is true, and is false whenever (\ref{eq:lwcg}) is false.
+So all the same results go through after modifying the objective function for the legislature
+					\item Lobby: objective function was
+					  \[
+							\max_{e_b} b(e_b,\bta) \left[\pi(\tau^n) - e_n \right] + [1 - b(e_b,\bta)] \pi(\tau^a) - e_b
+						\]
+						Now it's 
+						\begin{multline*}
+							\max_{e_b} \left\{ b(e_b,\bta) + b^*(1-b(e_b,\bta)) \right\} \left[\pi(\tau^n) - e_n \right] + \\
+							[1 - b(e_b,\bta) - b^* + b(e_b,\bta) b^*] \pi(\tau^a) - e_b
+						\end{multline*}
+						So the first order condition becomes 
+						\begin{equation}
+							\frac{\partial b(e_b,\bta)}{\partial e_b} \left[1-b^*(e_b^*,\bta)\right] \left[ \pi(\tau^n) -e_n - \pi(\tau^a) \right] = 1 
+						\end{equation}
+						I just need to check that the extra $\bta$ will not mess up any results.
 				\end{itemize}
-			\item Need to explore for first mover (home)
 		\end{itemize}

### 2016-02-29 to-do2.tex
intuition for why result 2 holds in sequential model

@@
 						\end{equation}
-						I just need to check that the extra $\bta$ will not mess up any results.
+						I believe that only Result 2 is in peril.
+A new term is added in the numerator (the denominator just gets a $(1-b^*)$; by the product rule, there are two terms after differentiating.
+The first is negative, the second is positive but can be combined with the pre-existing second term which just becomes less negative now.
+Should write all this up.
 				\end{itemize}

### 2016-02-29 to-do2.tex
add language for letter to editor on sequential model

@@
 				\end{itemize}
+			\item As long as there is a move of nature that occurs \textit{after} the trade agreement tariffs are set that determines which country's legislature moves first, at a first pass the results seem to go through with little additional complication for the sequential model you suggest.
+
+			The current results would apply for the second mover and a new set of results would need to be added for the actors in the country that is the first mover since their objective functions are altered by the future chance that the second moving legislature will break the agreement if the first moving legislature does not.
+This adds a nice new result that was only mentioned in passing in the previous draft about the first-mover lobby free-riding on the potential effort of the second-mover lobby.
+This is also why preserving the current results would require that the executives are behind a veil of ignorance about which country's legislature moves first: if they knew, they would like to exploit this information to reduce the tariff in the country where the lobby will suffer from the free-riding incentive.
+While there are probably some countries in which an executive can control the timing of a vote, it's probably fairly realistic to think that executives do know know for sure in which of their countries a vote will first take place.
+I can handle the case of asymmetric tariffs, but some of the propositions rely on the symmetry in tariffs for simplicity and I believe an extra condition would have to be invoked in several cases to bound magnitudes.
+I would like to avoid this especially if complicating the model further.
+
+			I think that this modification of the model adds something beneficial, but the tradeoff is both in terms of complexity and, necessarily, length.
+Thus I write to ask your opinion given your request to cut the length of the paper in half.
+I am working toward this end, but I do not see that this is possible if I were to add this complication and new results that go along with it.
+
 		\end{itemize}

### 2016-03-01 to-do2.tex
add response to fast-track request to to-do list

@@
 To make this point formally, it would be nice to establish conditions under which, in equilibrium, “fast track authority” dominates cherry picking ex ante (in a version of your model with several import competing sectors).
+		\begin{itemize}
+			\item I believe that I have already made this point formally in Result 3.
+The proof invokes Result 2 and Lemma 3 as it combines both dynamics that are explained intuitively in the text.
+Perhaps I should make this a bit more explicit by referring to those results in the text.
+
+			I agree that the idea you propose concerning fast-track authority / TPA is an important one and it is an idea that a colleague and I are exploring in a separate project as I believe it is beyond the scope of the current paper.
+The extension of the model to one with multiple sectors and lobbies is one that I believe is worthwhile for many purposes.
+However, the essential idea of the current paper that is, as you have judged, `sufficiently straightforward' that in my view the complications involved in the multi-sector model are not warranted.
+
+		\end{itemize}
 	\item[21.] I don’t see how the move of Nature (which determines the probabilities o and o* is consistent with the idea that there is uncertainty about who is the median voter (which was part of your motivation).

### 2016-04-26 letter-round2.tex
fix reference to sec 2.4 being too long after confirmation from editor

@@
 The paper still degenerates far too often into a series of discussions that bring little additional insight.
-	\item Section 2.4 either belongs to section 1.1 or should move as section 1.2. Section 2.3 is far too long.
+	\item Section 2.4 either belongs to section 1.1 or should move as section 1.2. Section 2.4 is far too long.
 	\item Your recall stuff far too often. Not all readers suffer from memory losses.
@@
   \item Your assumption that the identity of the median voter is not known a priori is standard in probability voting models, so you may want to stress the analogy between your setting and such models.
-  \item You claim that trade agreements act as a domestic political commitment device.
-To make this point formally, it would be nice to establish conditions under which, in equilibrium, “fast track authority” dominates cherry picking ex ante (in a version of your model with several import competing sectors).
+  \item \textit{You claim that trade agreements act as a domestic political commitment device.
+To make this point formally, it would be nice to establish conditions under which, in equilibrium, “fast track authority” dominates cherry picking ex ante (in a version of your model with several import competing sectors).}
+
   \item I don’t see how the move of Nature (which determines the probabilities o and o* is consistent with the idea that there is uncertainty about who is the median voter (which was part of your motivation).

### 2016-04-26 letter-round2.tex
format letter to editor

@@
 \begin{itemize}
-	\item Avoid changes of notation as you go along. This is totally unhelpful to the reader. Please pick one and stick to it throughout the paper.
-	\item Avoid digressions, lengthy discussions, and overreach: you stretch your model far too much.
+\renewcommand\labelitemi{-}
+	\item \textit{Avoid changes of notation as you go along. This is totally unhelpful to the reader. Please pick one and stick to it throughout the paper.}
+	\item \textit{Avoid digressions, lengthy discussions, and overreach: you stretch your model far too much.
 Stick to the main story.
-The paper still degenerates far too often into a series of discussions that bring little additional insight.
-	\item Section 2.4 either belongs to section 1.1 or should move as section 1.2. Section 2.4 is far too long.
-	\item Your recall stuff far too often. Not all readers suffer from memory losses.
+The paper still degenerates far too often into a series of discussions that bring little additional insight.}
+	\item \textit{Section 2.4 either belongs to section 1.1 or should move as section 1.2. Section 2.4 is far too long.}
+	\item \textit{Your recall stuff far too often. Not all readers suffer from memory losses.}
 \end{itemize}
@@
 \renewcommand\labelitemi{-}
-	\item Assuming that lobbies represent only import-competing interests is a strong assumption and it requires theoretical and/or empirical justification.
-	\item Your model is static, so claiming that it rationalizes the four-year delay and eventual ratification of the US-SK FTA is a stretch.
-In particular, the interactions with automakers suggests a that a bargaining game under incomplete information would be more appropriate.
-  \item Your assumption that the identity of the median voter is not known a priori is standard in probability voting models, so you may want to stress the analogy between your setting and such models.
+	\item \textit{Assuming that lobbies represent only import-competing interests is a strong assumption and it requires theoretical and/or empirical justification.}
+	\item \textit{Your model is static, so claiming that it rationalizes the four-year delay and eventual ratification of the US-SK FTA is a stretch.
+In particular, the interactions with automakers suggests a that a bargaining game under incomplete information would be more appropriate.}
+  \item \textit{Your assumption that the identity of the median voter is not known a priori is standard in probability voting models, so you may want to stress the analogy between your setting and such models.}
   \item \textit{You claim that trade agreements act as a domestic political commitment device.
@@
 
-  \item I don’t see how the move of Nature (which determines the probabilities o and o* is consistent with the idea that there is uncertainty about who is the median voter (which was part of your motivation).
+  \item \textit{I don’t see how the move of Nature (which determines the probabilities o and o* is consistent with the idea that there is uncertainty about who is the median voter (which was part of your motivation).
 Assuming that o and o* are mutually exclusive is not great.
 It would be better to assume some timing (rather than have a simultaneous choice).
-Since the game is symmetric, assuming that Home’s legislature moves before Foreign’s would be without further loss of generality.
-  \item “the uncertainty about the median legislator’s identity is resolved.” What is this uncertainty really about?
-  \item The equilibrium selection criterion requires justification and motivation.
+Since the game is symmetric, assuming that Home’s legislature moves before Foreign’s would be without further loss of generality.}
+  \item \textit{“the uncertainty about the median legislator’s identity is resolved.” What is this uncertainty really about?}
+  \item \textit{The equilibrium selection criterion requires justification and motivation.}
   \item \textit{Assumption 2 does not formalize any intuition: it is just a labelling.} \\
 	Quite right. The text has been modified appropriately.
-  \item I feel uneasy about Assumption 3. I would rather have it as an outcome and, if it is not the only possibly one, a discussion of the conditions under which this arises.
-  \item The full separability of sectors implies very minimal interactions already.
+  \item \textit{I feel uneasy about Assumption 3. I would rather have it as an outcome and, if it is not the only possibly one, a discussion of the conditions under which this arises.}
+  \item \textit{The full separability of sectors implies very minimal interactions already.
 Assuming a fully symmetric model is then highly restrictive.
-How difficult would it be to relax the latter assumption?
+How difficult would it be to relax the latter assumption?}
   \item \textit{Expression 6 establishes that the necessary second order condition holds by assumption 1, does not it?} \\
@@
 If there is a more direct argument, I would be happy to replace the current one.
-  \item Writing that lobbying expenditures are equal to 0.0007 is meaningless absent units.
+  \item \textit{Writing that lobbying expenditures are equal to 0.0007 is meaningless absent units.
 Are these cents?
 Trillions of dollars?
-It would be much more informative if these were expressed as a fraction of equilibrium profits.
+It would be much more informative if these were expressed as a fraction of equilibrium profits.}
   \item \textit{The last paragraph of section 4 is quite a stretch; is there any evidence in the political science literature that the separation of power is correlated with trade openness among democracies?} \\

### 2016-04-26 draft.tex
make formality of commitment device argument clear in text; response in letter; closes #20

@@
 \noindent That is, when the executives raise the trade agreement tariffs the legislature becomes less likely to abrogate the agreement, for three reasons.
-First, the legislature prefers a higher domestic tariff; second, the higher tariff discourages lobbying; finally, the lower lobbying effort reduces the legislature's preferred tariff further.
+First, the legislature prefers a higher domestic tariff (Lemma~\ref{res:leg_a}); second, the higher tariff discourages lobbying (Result~\ref{res:lobby}); and finally, the lower lobbying effort reduces the legislature's preferred tariff further (Lemma~\ref{res:bcomb}).
 Thus, beyond promising a lower tariff from the trading partner, we can think of the trade agreement as a sort of political commitment device that can be used to reduce political pressure and therefore get the legislature to maintain a lower tariff than it otherwise would.

### 2016-04-26 letter-round2.tex
make formality of commitment device argument clear in text; response in letter; closes #20

@@
   \item \textit{You claim that trade agreements act as a domestic political commitment device.
-To make this point formally, it would be nice to establish conditions under which, in equilibrium, “fast track authority” dominates cherry picking ex ante (in a version of your model with several import competing sectors).}
+To make this point formally, it would be nice to establish conditions under which, in equilibrium, “fast track authority” dominates cherry picking ex ante (in a version of your model with several import competing sectors).} \\
+		I believe that I have already made this point formally in Result 3.
+The proof invokes Result 2 and Lemma 3 as it combines both dynamics that are explained intuitively in the text.
+I have made this more explicit by referring to those results in the text.
+
+		I agree that the idea you propose concerning fast-track authority / TPA is an important one and it is an idea that a colleague and I are exploring in a separate project as I believe it is beyond the scope of the current paper.
+The extension of the model to one with multiple sectors and lobbies is one that I believe is worthwhile for many purposes.
+However, the essential idea of the current paper is, as you have judged, `sufficiently straightforward' that in my view the complications involved in the multi-sector model are not warranted.
 

### 2016-04-26 to-do2.tex
fix reference to sec 2.4 being too long after confirmation from editor
make formality of commitment device argument clear in text; response in letter; closes #20

@@
 	\item[14.] Section 2.4 either belongs to section 1.1 or should move as section 1.2.
-	\item[15.] Section 2.3 is far too long.
+	\item[15.] Section 2.4 is far too long.
 	\item[17.] Assuming that lobbies represent only import-competing interests is a strong assumption and it requires theoretical and/or empirical justification.
@@
 In particular, the interactions with automakers suggests a that a bargaining game under incomplete information would be more appropriate.
-  \item[20.] You claim that trade agreements act as a domestic political commitment device.
-To make this point formally, it would be nice to establish conditions under which, in equilibrium, “fast track authority” dominates cherry picking ex ante (in a version of your model with several import competing sectors).
-		\begin{itemize}
-			\item I believe that I have already made this point formally in Result 3.
-The proof invokes Result 2 and Lemma 3 as it combines both dynamics that are explained intuitively in the text.
-Perhaps I should make this a bit more explicit by referring to those results in the text.
 
-			I agree that the idea you propose concerning fast-track authority / TPA is an important one and it is an idea that a colleague and I are exploring in a separate project as I believe it is beyond the scope of the current paper.
-The extension of the model to one with multiple sectors and lobbies is one that I believe is worthwhile for many purposes.
-However, the essential idea of the current paper that is, as you have judged, `sufficiently straightforward' that in my view the complications involved in the multi-sector model are not warranted.
-
-		\end{itemize}
 	\item[21.] I don’t see how the move of Nature (which determines the probabilities o and o* is consistent with the idea that there is uncertainty about who is the median voter (which was part of your motivation).
@@
 \begin{enumerate}
+	\item[20.] You claim that trade agreements act as a domestic political commitment device.
+To make this point formally, it would be nice to establish conditions under which, in equilibrium, “fast track authority” dominates cherry picking ex ante (in a version of your model with several import competing sectors).
 	\item[24.] Assumption 2 does not formalize any intuition: it is just a labelling.

### 2016-04-26 letter-round2.tex
explain choice of symmetry in letter; closes #26

@@
 Assuming a fully symmetric model is then highly restrictive.
-How difficult would it be to relax the latter assumption?}
+How difficult would it be to relax the latter assumption?} \\
+	It's not very hard to relax the symmetry of the \textit{equilibrium}, but it requires adding some restrictions so that the impact on the break probability from $\tau^{*a}$ in Lemma 2 does not outweigh that of $\tau^a$ in Lemma 1.
+I chose symmetry because it's prevalent in the literature (along with separability) and avoids such seemingly ad-hoc assumptions.
+I also take advantage of the symmetry for the argument establishing existence of the executives' problem and I'm not completely sure the most elegant way with the least additional restrictions to get that without symmetry.
+
+	How hard it would be to make the underlying model asymmetric depends on the particular form of asymmetry.
+The one such avenue that is regularly followed in this literature is to have political pressure in only one country.
+That's a simplification so would easily go through.
+The only other asymmetry I've seen with any regularity is in the size of the economies (one large, one small).
+I think I might need a restriction on the size of the asymmetry in this case.
+I'd also have to generalize the bargaining solution in the case of any model asymmetry.
+
   \item \textit{Expression 6 establishes that the necessary second order condition holds by assumption 1, does not it?} \\

### 2016-04-26 draft.tex
bold changes for issue #20

--- 2016-02-25 draft.tex
@@
 \noindent That is, when the executives raise the trade agreement tariffs the legislature becomes less likely to abrogate the agreement, for three reasons.
-First, the legislature prefers a higher domestic tariff; second, the higher tariff discourages lobbying; finally, the lower lobbying effort reduces the legislature's preferred tariff further.
+First, the legislature prefers a higher domestic tariff \textbf{(Lemma~\ref{res:leg_a})}; second, the higher tariff discourages lobbying \textbf{(Result~\ref{res:lobby})}; and finally, the lower lobbying effort reduces the legislature's preferred tariff further \textbf{(Lemma~\ref{res:bcomb})}.
 Thus, beyond promising a lower tariff from the trading partner, we can think of the trade agreement as a sort of political commitment device that can be used to reduce political pressure and therefore get the legislature to maintain a lower tariff than it otherwise would.

### 2016-04-26 draft.tex
add theoretical justification for only one lobby, note in letter; closes #17

@@
 I now turn to describing the strategic players in the game.
-In order to focus attention on protectionist political forces, I assume that only the import-competing industry in each country is politically-organized and able to lobby and that it is represented by a single lobbying organization (see Footnote~\ref{fn:lobby} for discussion of additional lobbies).
 Each country's government is composed of two branches: an executive who can conclude trade agreements and a legislature that has final say on trade policy.
-To summarize, the political process is modeled as involving three players in each country: the lobby, the executive, and the legislature.
+In order to focus attention on protectionist political forces\textbf{, highlight the model's central mechanisms and minimize the number of actors that must be modeled}, I assume that only the import-competing industry in each country is politically-organized.
+Further, I assume that this industry is represented by a single lobbying organization (see Footnote~\ref{fn:lobby} for discussion of additional lobbies).
 

### 2016-04-26 letter-round2.tex
add theoretical justification for only one lobby, note in letter; closes #17

@@
 \renewcommand\labelitemi{-}
-	\item \textit{Assuming that lobbies represent only import-competing interests is a strong assumption and it requires theoretical and/or empirical justification.}
+	\item \textit{Assuming that lobbies represent only import-competing interests is a strong assumption and it requires theoretical and/or empirical justification.} \\
+	Done, on page [8]
 	\item \textit{Your model is static, so claiming that it rationalizes the four-year delay and eventual ratification of the US-SK FTA is a stretch.

### 2016-04-27 letter-round2.tex
add notes about not repeating explanation of uncertainty; closes #22

@@
 Since the game is symmetric, assuming that Home’s legislature moves before Foreign’s would be without further loss of generality.}
-  \item \textit{“the uncertainty about the median legislator’s identity is resolved.” What is this uncertainty really about?}
+  \item \textit{“the uncertainty about the median legislator’s identity is resolved.” What is this uncertainty really about?} \\
+	This is explained in Section [currently 2.4, but will move], so I did not repeat it or reference it here in line with the earlier stylistic advice.
+
   \item \textit{The equilibrium selection criterion requires justification and motivation.}

### 2016-04-28 draft.tex
add notes about not repeating explanation of uncertainty; closes #22
add justification for eqm selection criterion; closes #23

@@
 
-Second, whenever there is a possibility of multiple equilibria, I will focus on the one that maximizes the welfare of the executives.
+Second, whenever there is a possibility of multiple equilibria, I focus on the one that maximizes the welfare of the executives.
+Given the assumption that the executives are social welfare maximizers, this selection criterion puts results in terms of the maximal level of trade policy cooperation that is possible.
+It also allows the question of whether governments use trade agreements as political commitment devices to be answered in a straightforward way.
 
@@
 It is this interpretation of $\ta$ I take throughout this paper.
+%need to keep this paragraph and the next, or at least parts of it, to satisfy #22
 
@@
 Different assumptions on $\ga(\cdot,\cdot)$ correspond to different institutional features and will affect optimal lobbying- and tariff-setting behavior.
-I make the following two basic assumptions on the weight on import-industry profits:
+I make the following two basic assumptions on the weight on import-industry profits: %need to keep this paragraph and the previous, or at least parts of them, to satisfy #22
 

### 2016-04-29 draft.tex
add net profits and e/pi to example; closes #28

@@
 Facing this specification of the political process, the lobby maximizes its objective function at $e_n = 0.00166$, which produces an expected Nash tariff of $0.129$.
+Here net profits are $0.10239$ in terms of the numeraire and lobbying effort is $1.60\%$ of gross profits.
 
@@
 Assuming that each legislature has the opportunity to vote to break the agreement with probability $\frac{1}{2}$ and that the executives are social-welfare maximizers (i.e.
-$\ga_E = \ga_E^* = 1$), they will set trade agreements tariffs of $\tau^a = \tau^{*a} = 0.078$, with lobbying expenditures of $0.0007$ and a total break probability of $0.505$.
-The expected tariff is then $0.103$.
+$\ga_E = \ga_E^* = 1$), they will set trade agreements tariffs of $\tau^a = \tau^{*a} = 0.078$, with lobbying expenditures at the trade maintence phase of $0.0007$ and a total break probability of $0.505$.
+Total lobbying expenditure is $1.54\%$ of the expected profits of $0.98591$.
+The expected tariff is $0.103$.
 
@@
 The legislatures are able to use the agreement to internalize the terms of trade externality, making political influence more expensive for the lobbies.
-Indeed, I find that lobbying expenditures rise to $0.0027$---60$\%$ higher than in the Nash case---while the agreement tariffs are set at $0.118$, only slightly lower than the Nash tariffs of $0.129$ and still higher than even the expected tariffs from the trade agreement scenario in which the executives choose the tariffs.
+Indeed, I find that lobbying expenditures rise to $0.0027$---60$\%$ higher than in the Nash case and $2.64\%$ of the gross profits of $0.10204$---while the agreement tariffs are set at $0.118$, only slightly lower than the Nash tariffs of $0.129$ and still higher than even the expected tariffs from the trade agreement scenario in which the executives choose the tariffs.
+Comparing net (expected profits) across the three scenarios, they are highest in the Nash game ($0.1024$), lowest under the trade agreement ($0.9789$) and intermediate when the legislatures make policy directly ($0.0993$).
 

### 2016-04-29 letter-round2.tex
add justification for eqm selection criterion; closes #23
add net profits and e/pi to example; closes #28

@@
 
-  \item \textit{The equilibrium selection criterion requires justification and motivation.}
+  \item \textit{The equilibrium selection criterion requires justification and motivation.} \\
+	Done. See page [10]
+
   \item \textit{Assumption 2 does not formalize any intuition: it is just a labelling.} \\
@@
 Trillions of dollars?
-It would be much more informative if these were expressed as a fraction of equilibrium profits.}
+It would be much more informative if these were expressed as a fraction of equilibrium profits.} \\
+	I have added the numbers both for expenditure as a fraction of profits and for net profits.
+
   \item \textit{The last paragraph of section 4 is quite a stretch; is there any evidence in the political science literature that the separation of power is correlated with trade openness among democracies?} \\

### 2016-05-03 draft.tex
remove claim that I speak to KORUS delay, move fn:lobby; closes #18

@@
 Thus trade agreements act as a domestic political commitment device: executives set relatively high tariffs to discourage lobbying and increase the chance that the agreement will be ratified.
-The model sheds light on the empirical puzzle surrounding governments' welfare weights in the Grossman and Helpman (1994) model and can speak to important trade policy-making events such as the four-year delay in the ratification of the U.S.-Korea Free Trade Agreement.
+The model sheds light on the empirical puzzle surrounding governments' welfare weights in the Grossman and Helpman (1994) model and \textbf{provides a new explanation for failures to ratify trade agreements}.
 
@@
 Ultimately, I find that observed lobbying effort may often be low because the executive branches of the governments set trade agreement tariffs sufficiently high to reduce political pressure on the legislatures.
-There are other ways in which realism is gained when political uncertainty is added to the model: for instance, the model rationalizes the four-year delay and eventual ratification of the U.S.-Korea Free Trade Agreement.
 
@@
 Careful consideration of the political process seems to be important for understanding why governments so often fail to cooperate.
-The recent ratification of the U.S.-Korea Free Trade Agreement is a case in point.
-The agreement, as originally negotiated in 2007, was not able to be ratified in the United States until President Obama gained concessions for U.S.
-automakers, after which they removed their objections and the agreement was ratified by Congress in 2011 and entered into force in 2012.
-In a certain world, this ratification failure should not have occurred, and it was remedied by making the trade agreement more appealing to a key lobby and thus reducing its lobbying effort.\footnote{See \url{http://www.ustr.gov/trade-agreements/free-trade-agreements/korus-fta}.
-In this case, automakers are both importers and exporters, and the concessions were concentrated on market access for auto exports in South Korea.
-For simplicity, the model presented herein focuses on lobbying by an import-competing industry in a model of interindustry trade; the central insights do not depend on this structure and can be extended to the case of multiple lobbies and/or intraindustry trade.
-In particular, adding an export lobby in the simplest case of no uncertainty predicts that import lobbies will remain disengaged while the executive will be able to reduce tariffs by recruiting support from exporters.
-The Panama Trade Promotion Act seems to follow this pattern, where there was virtually no lobbying against, significant lobbying in favor, and numerous industries receiving import tariff protection.\label{fn:lobby}}
 
@@
 Each country's government is composed of two branches: an executive who can conclude trade agreements and a legislature that has final say on trade policy.
-In order to focus attention on protectionist political forces\textbf{, highlight the model's central mechanisms and minimize the number of actors that must be modeled}, I assume that only the import-competing industry in each country is politically-organized.
-Further, I assume that this industry is represented by a single lobbying organization (see Footnote~\ref{fn:lobby} for discussion of additional lobbies).
+In order to focus attention on protectionist political forces\textbf{, highlight the model's central mechanisms and minimize the number of actors that must be modeled}, I assume that only the import-competing industry in each country is politically-organized.\footnote{The model's central insights do not depend on this structure and can be extended to the case of multiple lobbies.
+In particular, adding an export lobby in the simplest case of no uncertainty predicts that import lobbies will remain disengaged while the executive will be able to reduce tariffs by recruiting support from exporters.
+The Panama Trade Promotion Act seems to follow this pattern, where there was virtually no lobbying against, significant lobbying in favor, and numerous industries receiving import tariff protection.\label{fn:lobby}} Further, I assume that this industry is represented by a single lobbying organization.
 

### 2016-05-03 letter-round2.tex
remove claim that I speak to KORUS delay, move fn:lobby; closes #18

@@
 	\item \textit{Your model is static, so claiming that it rationalizes the four-year delay and eventual ratification of the US-SK FTA is a stretch.
-In particular, the interactions with automakers suggests a that a bargaining game under incomplete information would be more appropriate.}
+In particular, the interactions with automakers suggests a that a bargaining game under incomplete information would be more appropriate.} \\
+	I have removed this claim.
   \item \textit{Your assumption that the identity of the median voter is not known a priori is standard in probability voting models, so you may want to stress the analogy between your setting and such models.}

### 2016-05-04 letter-round2.tex
explain assumption 3 in letter

@@
   \item \textit{You claim that trade agreements act as a domestic political commitment device.
-To make this point formally, it would be nice to establish conditions under which, in equilibrium, “fast track authority” dominates cherry picking ex ante (in a version of your model with several import competing sectors).} \\
+To make this point formally, it would be nice to establish conditions under which, in equilibrium, ``fast track authority'' dominates cherry picking ex ante (in a version of your model with several import competing sectors).} \\
 		I believe that I have already made this point formally in Result 3.
@@
 
-  \item \textit{I don’t see how the move of Nature (which determines the probabilities o and o* is consistent with the idea that there is uncertainty about who is the median voter (which was part of your motivation).
+  \item \textit{I don't see how the move of Nature (which determines the probabilities o and o* is consistent with the idea that there is uncertainty about who is the median voter (which was part of your motivation).
 Assuming that o and o* are mutually exclusive is not great.
 It would be better to assume some timing (rather than have a simultaneous choice).
-Since the game is symmetric, assuming that Home’s legislature moves before Foreign’s would be without further loss of generality.}
-  \item \textit{“the uncertainty about the median legislator’s identity is resolved.” What is this uncertainty really about?} \\
+Since the game is symmetric, assuming that Home’s legislature moves before Foreign's would be without further loss of generality.}
+  \item \textit{``the uncertainty about the median legislator's identity is resolved.'' What is this uncertainty really about?} \\
 	This is explained in Section [currently 2.4, but will move], so I did not repeat it or reference it here in line with the earlier stylistic advice.
@@
 	Quite right. The text has been modified appropriately.
-  \item \textit{I feel uneasy about Assumption 3. I would rather have it as an outcome and, if it is not the only possibly one, a discussion of the conditions under which this arises.}
+  \item \textit{I feel uneasy about Assumption 3. I would rather have it as an outcome and, if it is not the only possibly one, a discussion of the conditions under which this arises.} \\
+	I can remove Assumption 3, but the following happens: we get a second case for most of the results for when there are extreme realizations of $\theta$ such that the median legislator is more pro-trade than the executive at the trade-war phase so that the trade war tariffs are below the trade agreement tariffs.
+In this case, the results of the paper are uninteresting.
+So removing this assumption complicates the analysis and results without adding anything interesting or particularly realistic.
+
+	Alternatively, I can make an assumption on the support of $\theta$ and the functional form of $\ga$ so that Assumption 3 becomes a result, but this seems overly restrictive to make the points of the paper.
+
   \item \textit{The full separability of sectors implies very minimal interactions already.

### 2016-05-05 draft.tex
add reference to Roemer 1997 about probabilistic voting; closes #19

@@
 Alternatively, the integration of the random variable $\ta$ can represent reduced-form modeling of political uncertainty arising from a complex decision-making process within a legislature.
-One can interpret $\ta$ as measuring uncertainty over the identity of the median legislator in order to represent the idea that the lobbying and voting ``game'' that goes on within legislatures is often complex enough that none of the actors know precisely what the outcome will be---that is, no one can predict exactly which legislator will be the median and therefore what weight will be placed on import-industry profits at the time of the vote.
+One can interpret $\ta$ as measuring uncertainty over the identity of the median legislator.
+\textbf{In this sense, the idea is similar to that of the probabilistic voting models that incorporate policy motivation such as Roemer (1997).} Here, it is meant to represent the idea that the lobbying and voting ``game'' that goes on within legislatures is often complex enough that none of the actors know precisely what the outcome will be---that is, no one can predict exactly which legislator will be the median and therefore what weight will be placed on import-industry profits at the time of the vote.
 Thus one way we can think about $\ta$ is as an additional influence on the legislature's preferences about which all parties---executives, legislators and lobbies alike---are uncertain.
@@
 
+\item \textbf{Roemer, J., 1997. Political-economic Equilibrium when Parties Represent Constituents: The Unidimensional Case. Social Choice and Welfare 14, 479-502.}
+
 \item Saiegh, S., 2009. Political  Prowess or Lady Luck? Evaluating Chief Executives' Legislative Success Rates. The Journal of Politics 71, 1342-1356.

### 2016-05-05 letter-round2.tex
add reference to Roemer 1997 about probabilistic voting; closes #19

@@
 	I have removed this claim.
-  \item \textit{Your assumption that the identity of the median voter is not known a priori is standard in probability voting models, so you may want to stress the analogy between your setting and such models.}
+  \item \textit{Your assumption that the identity of the median voter is not known a priori is standard in probability voting models, so you may want to stress the analogy between your setting and such models.} \\
+	Thank you for this suggestion. I have incorporated a short reference on page [ ].
+
   \item \textit{You claim that trade agreements act as a domestic political commitment device.
@@
 
-	Alternatively, I can make an assumption on the support of $\theta$ and the functional form of $\ga$ so that Assumption 3 becomes a result, but this seems overly restrictive to make the points of the paper.
+	Alternatively, I can make an assumption on the support of $\theta$ and the functional form of $\gamma$ so that Assumption 3 becomes a result, but this seems overly restrictive to make the points of the paper.
 

### 2016-05-11 draft.tex
fix typo on proof of res:lobby

@@
 where the above equality holds because $Z(\bta)$ does not depend on $e$.
-Lemma~\ref{res:bcomb} shows that $\frac{\partial Z(\bta)}{\partial \tau^a} + \frac{\partial Z(\bta)}{\partial \tau^{*a}} \geq 0$, so $\frac{\partial f_\ga}{\partial e} \geq 0$ ensures that $\left\{\frac{\partial^2 b}{\partial e_b \partial \tau^a} + \frac{\partial^2 b}{\partial e_b \partial \tau^{*a}} \right\} \geq 0$.
+Lemma~\ref{res:bcomb} shows that $\frac{\partial Z(\bta)}{\partial \tau^a} + \frac{\partial Z(\bta)}{\partial \tau^{*a}} \geq 0$, so $\frac{\partial f_\ga}{\partial e} \geq 0$ ensures that $\left\{\frac{\partial^2 b}{\partial e_b \partial \tau^a} + \frac{\partial^2 b}{\partial e_b \partial \tau^{*a}} \right\} \leq 0$.
 Thus $\frac{\partial e_b}{\partial \tau^a} + \frac{\partial e_b}{\partial \tau^{*a}} \leq 0$.

### 2016-05-11 round2_calc.R
add R program w/new calculations for example

@@ -1,0 +1,16 @@
+e = .0027
+ta = .078
+pi_ta = .25*((4+4*ta)/7)^2
+
+t = .118
+pi = .25*((4+4*t)/7)^2
+
+pi - e
+
+efrac = e / pi * 100
+
+epi = .505 * pi_ta + .495 * (pi - e)
+
+efrac2 = .0007 / epi * 100
+
+epi - .0007

### 2016-06-09 draft.tex
create new section for second mover (home), begin modifying text

@@
 \label{sec:break}
-With trade-war decisions specified, we can proceed by backward induction to analyze the incentives regarding the legislature's decision to uphold or break the trade agreement.
+With trade-war decisions specified, we can proceed by backward induction to analyze the incentives regarding the legislatures' decisions to uphold or break the trade agreement.
 I assume that the form of the political weighting function is the same regardless of the break decision, but the model could be extended to allow the two to differ.
-What is in principle different are the lobbying efforts toward influencing the break decision, denoted $e_b$ at this stage to distinguish them from those that are made to influence the Nash tariff, $e_n$.
-The realization of uncertainty about the median legislator's identity will similarly be denoted $\ve_b$ so as to distinguish it from $\ve_n$ at the trade-war stage.
-
-The legislature will break the agreement and set the tariff at $\tau^n$ if the median legislator's utility from the Nash tariffs is higher than his utility from the trade agreement tariffs, i.e. if
+What is in principle different are the lobbying efforts toward influencing the break decision, denoted $e_b$ and $e_b^*$ at this stage to distinguish them from those that are made to influence the Nash tariff, $e_n$ and $e_n^*$.
+The realization of uncertainty about the identity of the median legislator in each country will similarly be denoted $\ve_b$ and $\ve_b^*$ so as to distinguish them from $\ve_n$ and $\ve_n^*$ at the trade-war stage.
+
+\subsubsection{Home Legislature (Second Mover)}
+
+I begin with the second-moving home legislature.
+The home legislature's decision only has consequence if the foreign legislature does not break the trade agreement.
+I assume that there is no political activity in home if the foreign legislature breaks the trade agreement.
+
+The home legislature will break the agreement and set the tariff at $\tau^n$ if the home median legislator's utility from the Nash tariffs is higher than his utility from the trade agreement tariffs, i.e.
+if
 \begin{equation}
@@
 
-The legislature's decisions are stochastic, so the outcome of the vote on whether or not to break the trade agreement, as well as the level at which the tariff will be set in the case of a disruption, is not known to \textit{any} player until the uncertainty over the identity of the median legislator is resolved at the moment the vote takes place on each issue.\footnote{We might think of the trade-war tariff as being set in the legislation before a vote takes place in the case of administered protection for instance; the results are robust to this variant.} I represent the probability that the home legislature breaks the trade agreement and sets the tariff at $\tau^n$ as:
+Each legislature's decisions are stochastic, so the outcome of the vote on whether or not to break the trade agreement, as well as the level at which the tariff will be set in the case of a disruption, is not known to \textit{any} player until the uncertainty over the identity of the median legislator is resolved at the moment the vote takes place on each issue.\footnote{We might think of the trade-war tariff as being set in the legislation before a vote takes place in the case of administered protection for instance; the results are robust to this variant.} I represent the probability that the home legislature breaks the trade agreement and sets the tariff at $\tau^n$ as:
 \begin{multline}
@@
 
-Note that the legislature's best response functions---both for whether to break the agreement, and at what level to set the tariff if it does---will be affected by the shape of $\ga(e,\ve)$.
-Recall that we have made three basic assumptions on $\ga(e,\ve)$ in Section~\ref{sec:choices}.
-More restrictions could be added to deliver further results.
-
-%Assumption~\ref{as:ga_l_e} ensures that $\tau^a < \tau^n$, and more generally, that the legislature's incentives are more closely aligned with the lobby's than are those of the executive.
-This is not essential but simplifies the analysis and matches well the empirical findings that politicians with larger constituencies are less sensitive to special interests (See Destler (2005) and footnote~\ref{fn:ga_l_e} above).
-
-We are now in a position to examine the legislature's decision more closely. Of central concern is how the probability that the legislature will break the trade agreement varies with lobbying effort:
+Note that the legislatures' best response functions---both for whether to break the agreement, and at what level to set the tariff if it does---will be affected by the shape of $\ga(e,\ve)$.
+
+We are now in a position to examine the home legislature's decision more closely.
+Of central concern is how the probability that the legislature will break the trade agreement varies with lobbying effort:
 
 \begin{result}
-  The probability that the legislature breaks a trade agreement is increasing and concave in $e_b$.
+  The probability that the home legislature breaks a trade agreement is increasing and concave in $e_b$.
   \label{res:bincC}

### 2016-06-09 draft.tex
finish modifying second mover section (old legislature section)

@@
 
-Turning to the effects of the trade-agreement tariffs on the probability that the agreement will be abrogated, it is straightforward that the legislature always prefers lower levels of the foreign tariff:
+Turning to the effects of the trade-agreement tariffs on the probability that the agreement will be abrogated, it is straightforward that the home legislature always prefers lower levels of the foreign tariff:
 \begin{lemma}
@@
 Because the home country is an exporter of good $Y$, the lower world price for $Y$ has a negative effect on producer surplus that is larger than the positive effect on consumer surplus.
-Thus the net effect of an increase in foreign trade-agreement tariffs on legislative (and home country) welfare is negative, leading to an increased probability that the trade agreement will be broken.
-
-The relationship between the home country's trade-agreement tariff $\tau^a$ and the legislature's probability of breaking the agreement is slightly more complex because $\tau^a$ interacts directly with $\ga(e_b,\ta_b)$.
+Thus the net effect of an increase in foreign trade-agreement tariffs on both home legislative and home total social welfare is negative, leading to an increased probability that the trade agreement will be broken.
+
+The relationship between the home country's trade-agreement tariff $\tau^a$ and the home legislature's probability of breaking the agreement is slightly more complex because $\tau^a$ interacts directly with $\ga(e_b,\ta_b)$.
 The intuition is as follows.
@@
 That is, the break probability decreases as $\tau^a$ increases.
-Even if the trade agreement tariff is so high that it is disliked by most, or even all, potential median legislators, it is preferable to the --- even higher --- trade war tariff, so the probability of breaking either becomes zero or continues to decrease in $\tau^a$.
+Even if the trade agreement tariff is so high that it is disliked by most, or even all, potential home median legislators, it is preferable to the --- even higher --- trade war tariff, so the probability of breaking either becomes zero or continues to decrease in $\tau^a$.
 
 \begin{lemma}
-  Holding lobbying effort constant, the probability the legislature breaks a trade agreement is weakly decreasing in $\tau^a$.
+  Holding lobbying effort constant, the probability the home legislature breaks a trade agreement is weakly decreasing in $\tau^a$.
   \label{res:leg_a}
@@
 \begin{lemma}
-	Holding lobbying effort constant, the probability the legislature breaks a trade agreement is weakly decreasing in $\bta$ (i.e.
+	Holding lobbying effort constant, the probability the home legislature breaks a trade agreement is weakly decreasing in $\bta$ (i.e.
 $\frac{\partial b(e,\bta)}{\partial  \tau^a} + \frac{\partial b(e,\bta)}{\partial  \tau^{*a}} \leq 0$).
@@
 
-\noindent When any change in one of the trade agreement tariffs must be met with an equal change in the other (as in a symmetric agreement), the negative impact on the legislature's propensity to break the trade agreement of raising the home tariff always outweighs the positive impact of raising the foreign tariff because of the legislature's bias toward the import-competing industry.
+\noindent When any change in one of the trade agreement tariffs must be met with an equal change in the other (as in a symmetric agreement), the negative impact on the home legislature's propensity to break the trade agreement of raising the home tariff always outweighs the positive impact of raising the foreign tariff because of the home legislature's bias toward its import-competing industry.
+
+\subsubsection{Foreign Legislature (First Mover)}
 

### 2016-06-09 draft.tex
move lobbying sec as first mover lobbying sec and modify

@@
 In the event of a trade war, the lobby chooses its effort $e_n$ given this tariff-setting behavior by maximizing its profits net of effort: $\pi\left(\tau^n\left(\ga\left(e_n,\ve_n\right)\right)\right) - e_n$.
-(Note this is Equation~\ref{eq:lv} simplified by the resolution of uncertainty over the legislature's decision on the trade agreement).
 This implies a first order condition for the lobby of
@@
 \subsubsection{Home Legislature (Second Mover)}
-
 I begin with the second-moving home legislature.
@@
 Note that the legislatures' best response functions---both for whether to break the agreement, and at what level to set the tariff if it does---will be affected by the shape of $\ga(e,\ve)$.
+I suppress the Nash tariffs in the expression of the break probability since they do not vary from the point of view of earlier stages.
 
@@
 
-\subsubsection{Foreign Legislature (First Mover)}
-
-\subsubsection{Lobby}
+\subsubsection{Second-Mover Lobby}
 \label{sec:lob_un}
-The lobby chooses its level of effort $e_b$ as a function of $\bta$, given the implications of that choice on the legislature's probability of breaking the agreement, $b(e_b,\bta,\btn)$.
-I will henceforth suppress the Nash tariffs in the expression of the break probability to simplify notation.\footnote{Since the decisions over the Nash tariffs are taken in the final period, they do not vary from the point of view of earlier stages.
-Similarly, because the lobbying effort made in that period does not interact with other decision variables in earlier periods, I suppress the dependence of the Nash tariffs on $e_n$.} Recalling that if the lobby has the chance to act, there is no chance for the foreign legislature to break the agreement, the lobby's decision is
+The home lobby chooses its level of effort $e_b$ as a function of $\bta$, given the implications of that choice on the home legislature's probability of breaking the agreement.
+The lobby's decision can be written as
 \[
@@
 \begin{result}
-  When the trade agreement remains in force with positive probability, lobbying effort is weakly decreasing in trade agreement tariffs.\footnote{See Section~\ref{sec:uncertainty} for a detailed account of behavior when the lobby is able to induce a trade war with probability 1.}
+  When the trade agreement remains in force with positive probability, home lobbying effort is weakly decreasing in trade agreement tariffs.\footnote{See Section~\ref{sec:uncertainty} for a detailed account of behavior when the lobby is able to induce a trade war with probability 1.}
   \label{res:lobby}
@@
 
-Note that the symmetric increase in the foreign tariff does not impact profits in the import-competing industry, but it can impact lobbying incentives through the break probability.
+\textit{Note that the symmetric increase in the foreign tariff does not impact profits in the import-competing industry, but it can impact lobbying incentives through the break probability.
 However, there is an analogous effect on the break probability from raising $\tau^a$ and the net effect is negative given Assumption~\ref{as:ga_c} on the probability distribution function of $\ga(e,\ta)$, which ensures that the effect of lobbying effort on the break probability does not increase as $\bta$ increases.
-This assumption can be significantly weakened: this effect can be positive as long as it is smaller in absolute value than the negative effect on trade agreement profits.
+This assumption can be significantly weakened: this effect can be positive as long as it is smaller in absolute value than the negative effect on trade agreement profits.}
+
+\subsubsection{Foreign Legislature (First Mover)}
+The only strategic difference between the first mover and the second mover is that the first mover faces uncertainty about whether the other legislature will break the trade agreement.
+
+				\begin{itemize}
+					\item Legislature (equations 7 and 8 in the draft):
+						\begin{equation}
+							W_{ML}(\btn,\ga(e_b,\ve_b)) > W_{ML}\left(\bta,\ga(e_b,\ve_b)\right)
+							\label{eq:lwcg}
+						\end{equation}
+						becomes
+						\[ 
+						  W_{ML}(\btn,\ga(e_b,\ve_b)) > b^* W_{ML}\left(\btn,\ga(e_b,\ve_b)\right) + (1-b^*) W_{ML}\left(\bta,\ga(e_b,\ve_b)\right)
+						\]
+						This can be arranged as 
+						\[ 
+						  W_{ML}(\btn,\ga(e_b,\ve_b)) - W_{ML}(\bta,\ga(e_b,\ve_b)) > b^* \left[ W_{ML}\left(\btn,\ga(e_b,\ve_b)\right) - W_{ML}\left(\bta,\ga(e_b,\ve_b)\right) \right]
+						\]
+						This has the interesting structure that it is always true when (\ref{eq:lwcg}) is true, and is false whenever (\ref{eq:lwcg}) is false.
+So all the same results go through after modifying the objective function for the legislature
+					\item Lobby: objective function was
+					  \[
+							\max_{e_b} b(e_b,\bta) \left[\pi(\tau^n) - e_n \right] + [1 - b(e_b,\bta)] \pi(\tau^a) - e_b
+						\]
+						Now it's 
+						\begin{multline*}
+							\max_{e_b} \left\{ b(e_b,\bta) + b^*(1-b(e_b,\bta)) \right\} \left[\pi(\tau^n) - e_n \right] + \\
+							[1 - b(e_b,\bta) - b^* + b(e_b,\bta) b^*] \pi(\tau^a) - e_b
+						\end{multline*}
+						So the first order condition becomes 
+						\begin{equation}
+							\frac{\partial b(e_b,\bta)}{\partial e_b} \left[1-b^*(e_b^*,\bta)\right] \left[ \pi(\tau^n) -e_n - \pi(\tau^a) \right] = 1 
+						\end{equation}
+						I believe that only Result 2 is in peril.
+A new term is added in the numerator (the denominator just gets a $(1-b^*)$; by the product rule, there are two terms after differentiating.
+The first is negative, the second is positive but can be combined with the pre-existing second term which just becomes less negative now.
+Should write all this up.
+				\end{itemize}
 

### 2016-06-09 draft.tex
begin section on first moving leg

@@
 \subsubsection{Foreign Legislature (First Mover)}
-The only strategic difference between the first mover and the second mover is that the first mover faces uncertainty about whether the other legislature will break the trade agreement.
-
+The only strategic difference between the first-moving legislature and the second moving legislature is that the first mover faces uncertainty about whether the other legislature will break the trade agreement.
+In constructing the break probability for the foreign legislature $b^*(e_b^*,\bta)$, Expression~\ref{eq:lwcg} becomes
+		\[ 
+		  W_{ML}^*(\btn,\ga(e_b^*,\ve_b^*)) > b(e_b,\bta) W_{ML}^*\left(\btn,\ga(e_b^*,\ve_b^*)\right) + \left[1-b(e_b,\bta)\right] W_{ML}^*\left(\bta,\ga(e_b^*,\ve_b^*)\right)
+		\]
+	This inequality determining when the foreign legislature prefers to break the trade agreement can be arranged as 
+		\begin{equation}
+						  W_{ML}^*(\btn,\ga(e_b^*,\ve_b^*)) - W_{ML}^*(\bta,\ga(e_b^*,\ve_b^*)) > b(e_b,\bta) \left[ W_{ML}^*\left(\btn,\ga(e_b^*,\ve_b^*)\right) - W_{ML}^*\left(\bta,\ga(e_b^*,\ve_b^*)\right) \right]
+			\label{eq:lwcg_for}
+		\end{equation}
+	Given symmetry, Expression~\ref{eq:lwcg_for} is true precisely when Expression~\ref{eq:lwcg} is true and false whenever Expression~\ref{eq:lwcg}) is false. 
 				\begin{itemize}
-					\item Legislature (equations 7 and 8 in the draft):
-						\begin{equation}
-							W_{ML}(\btn,\ga(e_b,\ve_b)) > W_{ML}\left(\bta,\ga(e_b,\ve_b)\right)
-							\label{eq:lwcg}
-						\end{equation}
-						becomes
-						\[ 
-						  W_{ML}(\btn,\ga(e_b,\ve_b)) > b^* W_{ML}\left(\btn,\ga(e_b,\ve_b)\right) + (1-b^*) W_{ML}\left(\bta,\ga(e_b,\ve_b)\right)
-						\]
-						This can be arranged as 
-						\[ 
-						  W_{ML}(\btn,\ga(e_b,\ve_b)) - W_{ML}(\bta,\ga(e_b,\ve_b)) > b^* \left[ W_{ML}\left(\btn,\ga(e_b,\ve_b)\right) - W_{ML}\left(\bta,\ga(e_b,\ve_b)\right) \right]
-						\]
-						This has the interesting structure that it is always true when (\ref{eq:lwcg}) is true, and is false whenever (\ref{eq:lwcg}) is false.
-So all the same results go through after modifying the objective function for the legislature
+
+					\item	So all the same results go through after modifying the objective function for the legislature
 					\item Lobby: objective function was

### 2016-06-10 draft.tex
almost completed first mover, decided not to go forward with it

@@
 The lobby's decision can be written as
-\[
+\begin{equation}
   \max_{e_b} b(e_b,\bta) \left[\pi(\tau^n) - e_n \right] + [1 - b(e_b,\bta)] \pi(\tau^a) - e_b
-\]
+	\label{eq:lobby}
+\end{equation}
 The lobby simply maximizes probability-weighted profits net of effort.
@@
 		\]
-	This inequality determining when the foreign legislature prefers to break the trade agreement can be arranged as 
+	This inequality determining when the foreign legislature prefers to break the trade agreement can be rearranged as 
 		\begin{equation}
@@
 		\end{equation}
-	Given symmetry, Expression~\ref{eq:lwcg_for} is true precisely when Expression~\ref{eq:lwcg} is true and false whenever Expression~\ref{eq:lwcg}) is false. 
-				\begin{itemize}
-
-					\item	So all the same results go through after modifying the objective function for the legislature
-					\item Lobby: objective function was
-					  \[
-							\max_{e_b} b(e_b,\bta) \left[\pi(\tau^n) - e_n \right] + [1 - b(e_b,\bta)] \pi(\tau^a) - e_b
-						\]
-						Now it's 
+	This is equivalent to 	
+		\[
+			W_{ML}^*(\btn,\ga(e_b^*,\ve_b^*)) > W_{ML}^*\left(\bta,\ga(e_b^*,\ve_b^*)\right)
+		\]
+	Then given symmetry, the foreign legislature's incentives to break the trade agreement are identical to the home legislature's incentives to break the trade agreement given in Expression~\ref{eq:lwcg}.
+We can therefore state propositions analogous to Result~\ref{res:bincC} and Lemma~\ref{res:bcomb} for the foreign legislature.
+The proofs are identical modulo symmetry.
+
+\begin{result}
+  The probability that the foreign legislature breaks a trade agreement is increasing and concave in $e_b^*$.
+  \label{res:bincCforeign}
+\end{result}
+
+\begin{lemma}
+	Holding lobbying effort constant, the probability the foreign legislature breaks a trade agreement is weakly decreasing in $\bta$ (i.e.
+$\frac{\partial b^*(e^*,\bta)}{\partial  \tau^a} + \frac{\partial b^*(e^*,\bta)}{\partial  \tau^{*a}} \leq 0$).
+	\label{res:bcombforeign}
+\end{lemma}
+
+\subsubsection{First Mover Lobby}
+The lobby in the foreign country likewise has a more complex decision. The second-moving lobby's objective function in Expression~\ref{eq:lobby} becomes
   \begin{multline*}
 		\max_{e_b} \left\{ b(e_b,\bta) + b^*(1-b(e_b,\bta)) \right\} \left[\pi(\tau^n) - e_n \right] + \\
-							[1 - b(e_b,\bta) - b^* + b(e_b,\bta) b^*] \pi(\tau^a) - e_b
+							[1 - b(e_b,\bta) - b^* + b(e_b,\bta) b^*] \pi(\tau^a) - e_b.
 	\end{multline*}
-						So the first order condition becomes 
+The first order condition is 	
   \begin{equation}
@@
 	\end{equation}
+
+\begin{result}
+  When the trade agreement remains in force with positive probability, home lobbying effort is weakly decreasing in trade agreement tariffs.
+  \label{res:lobbyforeign}
+\end{result}
+
 						I believe that only Result 2 is in peril.
@@
 Should write all this up.
-				\end{itemize}
 
@@
 \vskip.4in
+\noindent \textbf{\hypertarget{Pr_lobbyforeign}{Proof of Result~\ref{res:lobbyforeign}}}:\\
+Proof is via the Implicit Function Theorem using the lobby's first order condition, Equation~\ref{eq:lobbyfoc}, referred to here as $FOC_L$.
+  \[
+		\frac{\partial e_b}{\partial \tau^a} + \frac{\partial e_b}{\partial \tau^{*a}} = - \left[ \frac{\frac{\partial FOC_L}{\partial \tau^a}}{\frac{\partial FOC_L}{\partial e_b}} + \frac{\frac{\partial FOC_L}{\partial \tau^{*a}}}{\frac{\partial FOC_L}{\partial e_b}} \right] = \frac{ \frac{\partial b}{\partial e_b} \frac{\partial \pi(\bta)}{\partial \tau^a} + \frac{\partial b}{\partial e_b} \frac{\partial \pi(\bta)}{\partial \tau^{*a}} - \left\{\frac{\partial^2 b}{\partial e_b \partial \tau^a} + \frac{\partial^2 b}{\partial e_b \partial \tau^{*a}}   \right\}\left[ \pi(\tau^n) - e_n - \pi(\tau^a) \right]}{\frac{\partial^2 b}{\partial e_b{}^2} \left[ \pi(\tau^n) - e_n - \pi(\tau^a) \right]}
+	\]
+Beginning with the denominator: because $\pi(\tau)$ is increasing everywhere, $\left[\pi(\tau^n) - e_n - \pi(\tau^a) \right]$ is positive for all but very large values of $\tau^a$, that is for all $\tau^a$ such that $\pi(\tau^n) - e_n > \pi(\tau^a)$.
+When $\tau^a$ rises above this level, it is no longer in the lobby's interest to ask to have the agreement broken so $e_b=0$ and $\frac{\partial e_b}{\partial \tau^a} = 0$.
+$\frac{\partial^2 b}{\partial e_b{}^2}$ negative by Result~\ref{res:bincC}, so the denominator is negative for all values of $\tau^a$ at which $e_b$ is not constant.
+
+$\frac{\partial b}{\partial e_b}$ is positive, $\frac{\partial \pi(\bta)}{\partial \tau^a}$ is positive by construction, and $\frac{\partial \pi(\bta)}{\partial \tau^{*a}}$ is zero: separability between the sectors implies that profits in the import-competing sector do not depend on $\tau^{*a}$.
+
+We can rewrite $\frac{\partial b(e,\bta)}{\partial  \tau^a} + \frac{\partial b(e,\bta)}{\partial  \tau^{*a}} = -\frac{\partial F_\ga (Z(\bta))}{\partial \tau^a} -\frac{\partial F_\ga (Z(\bta))}{\partial \tau^{*a}} = -\frac{\partial F_\ga (Z(\bta))}{\partial Z(\bta)}\left[\frac{\partial Z(\bta)}{\partial \tau^a} + \frac{\partial Z(\bta)}{\partial \tau^{*a}} \right] = - f_\ga \left[\frac{\partial Z(\bta)}{\partial \tau^a} + \frac{\partial Z(\bta)}{\partial \tau^{*a}} \right]$, where $F_\ga (Z(\bta))$ is the CDF of $\ga$ and $f_\ga (Z(\bta))$ is the pdf of $\ga$ and $Z(\bta)$ represents the left hand side of the inequality in Expression~\ref{eq:b_ex}.
+
+Then $\frac{\partial^2 b}{\partial e_b \partial \tau^a} + \frac{\partial^2 b}{\partial e_b \partial \tau^{*a}} = 	- \frac{\partial}{\partial e} \left( \frac{\partial F_\ga (Z(\bta))}{\partial \tau^a} + \frac{\partial F_\ga (Z(\bta))}{\partial \tau^{*a}} \right) =
+	   - f_\ga \left[\frac{\partial^2 Z(\bta)}{\partial \tau^a \partial e} + \frac{\partial^2 Z(\bta)}{\partial \tau^{*a} \partial e} \right] - \left[\frac{\partial Z(\bta)}{\partial \tau^a} + \frac{\partial Z(\bta)}{\partial \tau^{*a}} \right] \frac{\partial f_\ga}{\partial e} = - \left[\frac{\partial Z(\bta)}{\partial \tau^a} + \frac{\partial Z(\bta)}{\partial \tau^{*a}} \right] \frac{\partial f_\ga}{\partial e}
+$
+where the above equality holds because $Z(\bta)$ does not depend on $e$.
+Lemma~\ref{res:bcomb} shows that $\frac{\partial Z(\bta)}{\partial \tau^a} + \frac{\partial Z(\bta)}{\partial \tau^{*a}} \geq 0$, so $\frac{\partial f_\ga}{\partial e} \geq 0$ ensures that $\left\{\frac{\partial^2 b}{\partial e_b \partial \tau^a} + \frac{\partial^2 b}{\partial e_b \partial \tau^{*a}} \right\} \leq 0$.
+Thus $\frac{\partial e_b}{\partial \tau^a} + \frac{\partial e_b}{\partial \tau^{*a}} \leq 0$.
+$\hfill\blacksquare$
+
+\vskip.4in	  	
 \noindent \textbf{\hypertarget{Pr_bcomB}{Proof of Result~\ref{res:bcomB}}}: \\

### 2016-06-10 draft.tex
small revisions throughout sec:break

--- 2016-06-09 draft.tex
@@
 \label{sec:break}
-With trade-war decisions specified, we can proceed by backward induction to analyze the incentives regarding the legislatures' decisions to uphold or break the trade agreement.
+With trade-war decisions specified, we can proceed by backward induction to analyze the incentives regarding the legislature's decision to uphold or break the trade agreement.
 I assume that the form of the political weighting function is the same regardless of the break decision, but the model could be extended to allow the two to differ.
-What is in principle different are the lobbying efforts toward influencing the break decision, denoted $e_b$ and $e_b^*$ at this stage to distinguish them from those that are made to influence the Nash tariff, $e_n$ and $e_n^*$.
-The realization of uncertainty about the identity of the median legislator in each country will similarly be denoted $\ve_b$ and $\ve_b^*$ so as to distinguish them from $\ve_n$ and $\ve_n^*$ at the trade-war stage.
-
-\subsubsection{Home Legislature (Second Mover)}
-I begin with the second-moving home legislature.
-The home legislature's decision only has consequence if the foreign legislature does not break the trade agreement.
-I assume that there is no political activity in home if the foreign legislature breaks the trade agreement.
-
-The home legislature will break the agreement and set the tariff at $\tau^n$ if the home median legislator's utility from the Nash tariffs is higher than his utility from the trade agreement tariffs, i.e.
-if
+What is in principle different are the lobbying efforts toward influencing the break decision, denoted $e_b$ at this stage to distinguish them from those that are made to influence the Nash tariff, $e_n$.
+The realization of uncertainty about the median legislator's identity will similarly be denoted $\ve_b$ so as to distinguish it from $\ve_n$ at the trade-war stage.
+
+The legislature will break the agreement and set the tariff at $\tau^n$ if the median legislator's utility from the Nash tariffs is higher than his utility from the trade agreement tariffs, i.e. if
 \begin{equation}
@@
 
-Each legislature's decisions are stochastic, so the outcome of the vote on whether or not to break the trade agreement, as well as the level at which the tariff will be set in the case of a disruption, is not known to \textit{any} player until the uncertainty over the identity of the median legislator is resolved at the moment the vote takes place on each issue.\footnote{We might think of the trade-war tariff as being set in the legislation before a vote takes place in the case of administered protection for instance; the results are robust to this variant.} I represent the probability that the home legislature breaks the trade agreement and sets the tariff at $\tau^n$ as:
+The legislature's decisions are stochastic, so the outcome of the vote on whether or not to break the trade agreement, as well as the level at which the tariff will be set in the case of a disruption, is not known to \textit{any} player until the uncertainty over the identity of the median legislator is resolved at the moment the vote takes place on each issue.\footnote{We might think of the trade-war tariff as being set in the legislation before a vote takes place in the case of administered protection for instance; the results are robust to this variant.} I represent the probability that the home legislature breaks the trade agreement and sets the tariff at $\tau^n$ as:
 \begin{multline}
@@
 
-Note that the legislatures' best response functions---both for whether to break the agreement, and at what level to set the tariff if it does---will be affected by the shape of $\ga(e,\ve)$.
+Note that the legislature's best response functions---both for whether to break the agreement, and at what level to set the tariff if it does---will be affected by the shape of $\ga(e,\ve)$.
 I suppress the Nash tariffs in the expression of the break probability since they do not vary from the point of view of earlier stages.
 
-We are now in a position to examine the home legislature's decision more closely.
-Of central concern is how the probability that the legislature will break the trade agreement varies with lobbying effort:
+We are now in a position to examine the legislature's decision more closely. Of central concern is how the probability that the legislature will break the trade agreement varies with lobbying effort:
 
 \begin{result}
-  The probability that the home legislature breaks a trade agreement is increasing and concave in $e_b$.
+  The probability that the legislature breaks a trade agreement is increasing and concave in $e_b$.
   \label{res:bincC}
@@
 
-Turning to the effects of the trade-agreement tariffs on the probability that the agreement will be abrogated, it is straightforward that the home legislature always prefers lower levels of the foreign tariff:
+Turning to the effects of the trade-agreement tariffs on the probability that the agreement will be abrogated, it is straightforward that the legislature always prefers lower levels of the foreign tariff:
 \begin{lemma}
@@
 
-The relationship between the home country's trade-agreement tariff $\tau^a$ and the home legislature's probability of breaking the agreement is slightly more complex because $\tau^a$ interacts directly with $\ga(e_b,\ta_b)$.
+The relationship between the home country's trade-agreement tariff $\tau^a$ and the legislature's probability of breaking the agreement is slightly more complex because $\tau^a$ interacts directly with $\ga(e_b,\ta_b)$.
 The intuition is as follows.
@@
 That is, the break probability decreases as $\tau^a$ increases.
-Even if the trade agreement tariff is so high that it is disliked by most, or even all, potential home median legislators, it is preferable to the --- even higher --- trade war tariff, so the probability of breaking either becomes zero or continues to decrease in $\tau^a$.
+Even if the trade agreement tariff is so high that it is disliked by most, or even all, potential median legislators, it is preferable to the --- even higher --- trade war tariff, so the probability of breaking either becomes zero or continues to decrease in $\tau^a$.
 
 \begin{lemma}
-  Holding lobbying effort constant, the probability the home legislature breaks a trade agreement is weakly decreasing in $\tau^a$.
+  Holding lobbying effort constant, the probability the legislature breaks a trade agreement is weakly decreasing in $\tau^a$.
   \label{res:leg_a}
@@
 \begin{lemma}
-	Holding lobbying effort constant, the probability the home legislature breaks a trade agreement is weakly decreasing in $\bta$ (i.e.
+	Holding lobbying effort constant, the probability the legislature breaks a trade agreement is weakly decreasing in $\bta$ (i.e.
 $\frac{\partial b(e,\bta)}{\partial  \tau^a} + \frac{\partial b(e,\bta)}{\partial  \tau^{*a}} \leq 0$).
@@
 
-\noindent When any change in one of the trade agreement tariffs must be met with an equal change in the other (as in a symmetric agreement), the negative impact on the home legislature's propensity to break the trade agreement of raising the home tariff always outweighs the positive impact of raising the foreign tariff because of the home legislature's bias toward its import-competing industry.
-
-\subsubsection{Second-Mover Lobby}
+\noindent When any change in one of the trade agreement tariffs must be met with an equal change in the other (as in a symmetric agreement), the negative impact on the legislature's propensity to break the trade agreement of raising the home tariff always outweighs the positive impact of raising the foreign tariff because of the legislature's bias toward the import-competing industry.
+
+\subsubsection{Lobby}
 \label{sec:lob_un}
-The home lobby chooses its level of effort $e_b$ as a function of $\bta$, given the implications of that choice on the home legislature's probability of breaking the agreement.
-The lobby's decision can be written as
+The lobby chooses its level of effort $e_b$ as a function of $\bta$, given the implications of that choice on the legislature's probability of breaking the agreement.
+The lobby's decision is taken according to
 \[
@@
 \begin{result}
-  When the trade agreement remains in force with positive probability, home lobbying effort is weakly decreasing in trade agreement tariffs.\footnote{See Section~\ref{sec:uncertainty} for a detailed account of behavior when the lobby is able to induce a trade war with probability 1.}
+  When the trade agreement remains in force with positive probability, lobbying effort is weakly decreasing in trade agreement tariffs.\footnote{See Section~\ref{sec:uncertainty} for a detailed account of behavior when the lobby is able to induce a trade war with probability 1.}
   \label{res:lobby}
@@
 
-\textit{Note that the symmetric increase in the foreign tariff does not impact profits in the import-competing industry, but it can impact lobbying incentives through the break probability.
+Note that the symmetric increase in the foreign tariff does not impact profits in the import-competing industry, but it can impact lobbying incentives through the break probability.
 However, there is an analogous effect on the break probability from raising $\tau^a$ and the net effect is negative given Assumption~\ref{as:ga_c} on the probability distribution function of $\ga(e,\ta)$, which ensures that the effect of lobbying effort on the break probability does not increase as $\bta$ increases.
-This assumption can be significantly weakened: this effect can be positive as long as it is smaller in absolute value than the negative effect on trade agreement profits.}
-
-\subsubsection{Foreign Legislature (First Mover)}
-The only strategic difference between the first mover and the second mover is that the first mover faces uncertainty about whether the other legislature will break the trade agreement.
-
-				\begin{itemize}
-					\item Legislature (equations 7 and 8 in the draft):
-						\begin{equation}
-							W_{ML}(\btn,\ga(e_b,\ve_b)) > W_{ML}\left(\bta,\ga(e_b,\ve_b)\right)
-							\label{eq:lwcg}
-						\end{equation}
-						becomes
-						\[ 
-						  W_{ML}(\btn,\ga(e_b,\ve_b)) > b^* W_{ML}\left(\btn,\ga(e_b,\ve_b)\right) + (1-b^*) W_{ML}\left(\bta,\ga(e_b,\ve_b)\right)
-						\]
-						This can be arranged as 
-						\[ 
-						  W_{ML}(\btn,\ga(e_b,\ve_b)) - W_{ML}(\bta,\ga(e_b,\ve_b)) > b^* \left[ W_{ML}\left(\btn,\ga(e_b,\ve_b)\right) - W_{ML}\left(\bta,\ga(e_b,\ve_b)\right) \right]
-						\]
-						This has the interesting structure that it is always true when (\ref{eq:lwcg}) is true, and is false whenever (\ref{eq:lwcg}) is false.
-So all the same results go through after modifying the objective function for the legislature
-					\item Lobby: objective function was
-					  \[
-							\max_{e_b} b(e_b,\bta) \left[\pi(\tau^n) - e_n \right] + [1 - b(e_b,\bta)] \pi(\tau^a) - e_b
-						\]
-						Now it's 
-						\begin{multline*}
-							\max_{e_b} \left\{ b(e_b,\bta) + b^*(1-b(e_b,\bta)) \right\} \left[\pi(\tau^n) - e_n \right] + \\
-							[1 - b(e_b,\bta) - b^* + b(e_b,\bta) b^*] \pi(\tau^a) - e_b
-						\end{multline*}
-						So the first order condition becomes 
-						\begin{equation}
-							\frac{\partial b(e_b,\bta)}{\partial e_b} \left[1-b^*(e_b^*,\bta)\right] \left[ \pi(\tau^n) -e_n - \pi(\tau^a) \right] = 1 
-						\end{equation}
-						I believe that only Result 2 is in peril.
-A new term is added in the numerator (the denominator just gets a $(1-b^*)$; by the product rule, there are two terms after differentiating.
-The first is negative, the second is positive but can be combined with the pre-existing second term which just becomes less negative now.
-Should write all this up.
-				\end{itemize}
+This assumption can be significantly weakened: this effect can be positive as long as it is smaller in absolute value than the negative effect on trade agreement profits.
 

### 2016-06-10 draft.tex
explain no timing change, add footnote to text; closes #21

@@
 Next, nature determines with probabilities $o$ and $o^*$ whether each legislature will have the opportunity to take a vote to break the agreement.
-I assume that these probabilities are less than one-half and mutually exclusive.\footnote{While this assumption is made for purposes of exposition and tractability, one can think of $o$ and $o^*$ as being determined by events beyond the executives' or legislatures' control that determine whether or not the legislatures will be willing to consider the lobby's request; for example, they would be affected by the occurrence of an economic crisis that diverts the legislature's attention from less-pressing matters.} This allows me to focus on the key interaction between the domestic actors and abstract from the strategic interaction between the lobbies in the two countries.\footnote{When the legislatures in both countries have an opportunity to break the agreement simultaneously, an incentive for each lobby to free ride on the effort of the other makes the optimization problem of the executives more complex as they are faced with the problem of how best to exploit the free-riding dynamic.}
+I assume that these probabilities are less than one-half and mutually exclusive.\footnote{While this assumption is made for purposes of exposition and tractability, one can think of $o$ and $o^*$ as being determined by events beyond the executives' or legislatures' control that determine whether or not the legislatures will be willing to consider the lobby's request; for example, they would be affected by the occurrence of an economic crisis that diverts the legislature's attention from less-pressing matters.} This allows me to focus on the key interaction between the domestic actors and abstract from the strategic interaction between the lobbies in the two countries.\footnote{When the legislatures in both countries have an opportunity to break the agreement simultaneously, an incentive for each lobby to free ride on the effort of the other makes the optimization problem of the executives more complex as they are faced with the problem of how best to exploit the free-riding dynamic.
+\textbf{Alternatively, one can take the break decisions in the two countries to be sequential with the executives behind a veil of ignorance about which legislature will move first.
+In that set up, the end results are the same as in the current model but additional intermediate results are needed.}}
 

### 2016-06-10 letter-round2.tex
explain no timing change, add footnote to text; closes #21

@@
 It would be better to assume some timing (rather than have a simultaneous choice).
-Since the game is symmetric, assuming that Home’s legislature moves before Foreign's would be without further loss of generality.}
+Since the game is symmetric, assuming that Home’s legislature moves before Foreign's would be without further loss of generality.} \\
+	I view the move of nature that determines the probabilities $o$ and $o^*$ as distinct from the uncertainty about the identity of the median legislator.
+This is really just a theoretical device to simplify the model.
+
+	The sequential model you suggested is very attractive.
+I worked it out and wrote it into the draft.
+All the results go through (with one small additional assumption on magnitudes: it must be ensured that the lobby's direct response to an increase in trade agreement tariffs is stronger than a new effect wherein the lobby might increase effort to compensate for the trading partner's legislature reducing its break probability in response to that same tariff increase).
+However, the model is significantly complicated as two actors have to be modeled and results added since they can no longer be accounted for by symmetry.
+This added considerable length without compensating insight, so I opted to add a footnote explaining the near-equivalence of the models in line with your desire to simplify the presentation and not distract from the central point.
+The new text reads as follows can can be found in \textbf{Footnote ???
+on Page???}:
+	\begin{quote}
+		Alternatively, one can take the break decisions in the two countries to be sequential with the executives behind a veil of ignorance about which legislature will move first.
+In that set up, the end results are the same as in the current model but additional intermediate results are needed.
+	\end{quote}
+
   \item \textit{``the uncertainty about the median legislator's identity is resolved.'' What is this uncertainty really about?} \\

### 2016-06-10 draft.tex
move uncertainty text from sec:choices to intro

@@
 
+\newpage
 The fact that zero contributions are predicted in equilibrium is not particularly satisfying.
@@
 I assume all actors are uncertain about the weight the median legislator in a non-unitary legislature places on the profits of the lobbying industry.
-This is meant to represent the idea that the lobbying and voting ``game'' that goes on within legislatures is often complex enough that none of the actors know precisely what the outcome will be---that is, no one can predict exactly which legislator will be the median and therefore what weight will be placed on import-industry profits at the time of the vote.
-For instance, perhaps a small number of lobbied legislators cannot deliver the votes of their non-lobbied colleagues with certainty.
+\textbf{Similar to probabilistic voting models that incorporate policy motivation such as Roemer (1997),} this represents the idea that the lobbying and voting ``game'' that goes on within legislatures is often complex enough that none of the actors know precisely what the outcome will be---that is, no one can predict exactly which legislator will be the median and therefore what weight will be placed on import-industry profits at the time of the vote.
+
+\textbf{One can conceptualize this uncertainty as being a result of the lobby's strategic choice to engage with only some key members of the legislature who wield significant influence but who cannot deliver policy decisions without the cooperation of others within the legislature, which is in turn uncertain.} For instance, perhaps a small number of lobbied legislators cannot deliver the votes of their non-lobbied colleagues with certainty.
 One can think of this uncertainty as an additional influence on the legislature's preferences about which all parties---executives, legislators and lobbies alike---are uncertain.
@@
 Modeling the objective function so closely on the standard in the trade agreements literature allows for direct comparisons to the large extant body of work that studies exogenous shocks only, revealing cleanly the effects of the addition of endogenous lobbying---and if one wants---this different kind of political uncertainty.
-
-Alternatively, the integration of the random variable $\ta$ can represent reduced-form modeling of political uncertainty arising from a complex decision-making process within a legislature.
-One can interpret $\ta$ as measuring uncertainty over the identity of the median legislator.
-\textbf{In this sense, the idea is similar to that of the probabilistic voting models that incorporate policy motivation such as Roemer (1997).} Here, it is meant to represent the idea that the lobbying and voting ``game'' that goes on within legislatures is often complex enough that none of the actors know precisely what the outcome will be---that is, no one can predict exactly which legislator will be the median and therefore what weight will be placed on import-industry profits at the time of the vote.
-Thus one way we can think about $\ta$ is as an additional influence on the legislature's preferences about which all parties---executives, legislators and lobbies alike---are uncertain.
-It is this interpretation of $\ta$ I take throughout this paper.
-%need to keep this paragraph and the next, or at least parts of it, to satisfy #22
-
-One can conceptualize this uncertainty as being a result of the lobby's strategic choice to engage with only some key members of the legislature who wield significant influence but who cannot deliver policy decisions without the cooperation of others within the legislature, which is in turn uncertain.
-Different assumptions on $\ga(\cdot,\cdot)$ correspond to different institutional features and will affect optimal lobbying- and tariff-setting behavior.
-I make the following two basic assumptions on the weight on import-industry profits: %need to keep this paragraph and the previous, or at least parts of them, to satisfy #22
 

### 2016-06-10 draft.tex
condense and move sec 2.4 text; closes #14; closes #15

@@
 
-\newpage
 The fact that zero contributions are predicted in equilibrium is not particularly satisfying.
@@
 
+\textbf{Although the political process here most closely matches that of the United States in the post-war era,\footnote{In particular, the binary decision by the legislature about whether to abide by or break the trade agreement is modeled on the ``Fast Track Authority'' that the U.S.
+Congress granted to the Executive branch almost continuously from 1974-1994 and then again as ``Trade Promotion Authority'' from 2002-2007 and 2015 to present.} the model is broadly applicable because the central dynamic that emerges---that trade agreements are formed in the \textit{shadow} of lobbying pressure instead of in direct response to it---appears to be present in a wide range of countries in which the formation and maintenance of trade policy requires approval by a politically susceptible body either at home or in a trading partner.}
+
+\textbf{The results apply beyond separation-of-powers presidential systems because the presence of veto points is what is essential for the dynamic posited here to occur.
+Although the number of formal veto points tends to be lower in, for instance, parliamentary systems, veto points are typically not absent altogether, at least in democracies (Henisz and Mansfield 2006).
+In reality, constraints on the executive are almost always present in democratic regimes, although they may operate in different ways.}
+
 \subsection{Related Literature}
-The foundations of this work rest on Grossman and Helpman's (1994) `Protection for Sale.' Two related papers `Trade Wars and Trade Talks' (Grossman and Helpman (1995b)) and `The Politics of Free-Trade Agreements' (Grossman and Helpman (1995a)) first employed the PFS approach in the study of trade agreements.
+The foundations of this work rest on the `political support function' introduced by Hillman (1982) and Grossman and Helpman's (1994) `Protection for Sale.' Two related papers `Trade Wars and Trade Talks' (Grossman and Helpman (1995b)) and `The Politics of Free-Trade Agreements' (Grossman and Helpman (1995a)) first employed the PFS approach in the study of trade agreements.
 The first considers ``Trade Wars'' and ``Trade Talks'' separately, whereas my approach allows the ``Trade Talks'' to be influenced by the desire to avoid a ``Trade War'' and thus views the ``Trade War'' as a crucial subgame.
@@
 Feenstra and Lewis (1991); Milner and Rosendorff (2001); Bagwell and Staiger (2005); Beshkar (2010); Horn, Maggi and Staiger (2010); Beshkar and Bond (2012); Amador and Bagwell (2013)) derive various implications for the design of international trade agreements using a Baldwin (1987)-style government welfare function with exogenous shocks to the political-pressure parameter.
-As explained in the next section, the model presented here is intended to be a bridge between this literature and that discussed above that focuses on endogenously-determined political economy weights.
-To preview, this model is closest to that of Bagwell and Staiger (2005), with two main changes: in place of their unitary government signing a trade agreement and having different preferences ex-ante and ex-post, this model has two branches of government with differing preferences, where---in the case of the legislature---the political economy weights are determined endogenously.
+
+\textbf{Since trade policies are overwhelmingly determined within the context of trade agreements, it is useful to have a framework to bring together the endogenous political pressure of PFS-style modeling with the trade agreements approach.
+The government objective function introduced in this paper is intended to be a bridge between the two.
+Buzard (2016) is an example of how this modeling approach can be used to integrate  endogenous political activity and exogenous shocks into the analysis of questions regarding optimal trade agreement design.}
+
+\textbf{The objective function I introduce starts from the \textit{deus ex machina} government objective function popularized by Baldwin (1987).
+It is more flexible in that the weight the legislature places on profits is endogenous (PFS) as well as potentially non-linear (Findlay and Wellisz 1982); Dixit, Grossman and Helpman 1997) and stochastic.}
+
+\textbf{Modeling the objective function so closely on the standard in the trade agreements literature allows for direct comparisons to the large extant body of work that studies exogenous shocks only, revealing cleanly the effects of the addition of endogenous lobbying---and if one wants---this different kind of political uncertainty.} To preview, this model is closest to that of Bagwell and Staiger (2005), with two main changes: in place of their unitary government signing a trade agreement and having different preferences ex-ante and ex-post, this model has two branches of government with differing preferences, where---in the case of the legislature---the political economy weights are determined endogenously.
 
@@
 
-In the event that the trade agreement does not enter into or remain in force, there is a final stage of lobbying, resolution of the uncertainty surrounding this decision, and voting to set the trade-war tariffs.\footnote{Alternatively, one can model the lobbying for the ratification decision as determining the potential non-cooperative tariff; this changes behavior slightly depending on the relationship between the lobby and the legislature's preferences, but does not change the fundamental dynamics of the model.
+In the event that the trade agreement does not enter into or remain in force, there is a final stage of lobbying, resolution of the uncertainty surrounding this decision, and voting to set the trade-war tariffs.\footnote{Alternatively, one can model the lobbying for the ratification decision as determining the potential non-cooperative tariff; this changes behavior slightly but does not change the fundamental dynamics.
 Yet another possibility is to replace the trade-war phase with the ex-ante status quo, which would most naturally be the non-cooperative equilibrium.
@@
 where $\mathit{CS}$ is consumer surplus, $\mathit{PS}$ is producer surplus, $\ga(e,\ta)$ is the weight placed on producer surplus (profits) in the import-competing industry, $e$ is lobbying effort, $\ta$ is a random variable, and $\mathit{TR}$ is tariff revenue.\footnote{Labor income $l$ could also be included in both median legislator and executive welfare.
-I omit it because its inclusion alters none of the results and only serves to complicate the exposition.} I model the decisions of the legislature as being taken by a median legislator (more on this in Section~\ref{sec:choices}).
+I omit it because its inclusion alters none of the results and only serves to complicate the exposition.} I model the decisions of the legislature as being taken by a median legislator.
 Notice that the weight the median legislator places on the profits of the import-competing industry, $\ga(e,\ve)$, is affected by the level of lobbying effort and a random variable $\ta$.
+I make the following assumptions on $\ga(e,\ta)$:
+
+\begin{assumption}
+  $\ga(e,\ta)$ is increasing and concave in $e$ for every $\ta \in \Theta$.
+  \label{as:ga_c}
+\end{assumption}
+
+\begin{assumption}
+  $\ga(e,\ta)$ is increasing in $\ta$.
+  \label{as:ga_ta}
+\end{assumption}
+
+Assumption~\ref{as:ga_c} formalizes the intuition that the legislature favors the import-competing industry more the higher is its lobbying effort, but that there are diminishing returns to lobbying activity.\footnote{The diminishing returns here take the form of declining increments to the lobby's influence as effort increases; in Ethier (2012), the returns to lobbying decline as the level of protection increases.} It allows the political uncertainty inherent in $\ta$ to be correlated in interesting ways with the level of $e$, ruling out only (i) that higher effort levels make lower $\ga$-weights more likely and (ii) that effort and uncertainty are correlated in such a way that increasing lobbying effort changes the structure of uncertainty so that higher $\ga$-weights become more likely at an accelerating pace.
+Assumption~\ref{as:ga_ta} \textbf{simply provides for an intuitive labelling so that} larger realization of the uncertainty variable $\ta$ increase the value of the political economy weight.
 
@@
 Note that this is identical to the welfare function for the legislature aside from the weight $\ga_E$ the executive places on the profits of the import industry, which is not a function of lobbying effort.
-Section~\ref{sec:choices} contains an in-depth discussion of these modeling choices.
+\textbf{This setup does not strictly require that the executives are not lobbied, only that the extent to which the executives favor this industry cannot be \textit{directly} influenced by lobbying effort.
+The assumption is innocuous as long as the executives' preferences are not directly altered in a significant way by lobbying over trade.}
+
+\textbf{I make one more assumption on $\ga(e,\ta)$:}
+
+\begin{assumption}
+  $\ga(e,\ta) \geq \ga_E \geq 1 \ \forall e,\ta$.
+  \label{as:ga_l_e}
+\end{assumption}
+
+\textbf{That is, even for the least favorable outcome of the lobbying process, the legislature will be at least slightly more protectionist than the executive regardless of the lobby's effort choice.
+This assumption is not essential, but it simplifies the analysis considerably, ensuring that trade agreement tariffs are less than trade war tariffs.
+It matches the empirical findings that politicians with larger constituencies are less sensitive to special interests and seems to represent particularly well the post-war U.S.
+where Congress has consistently been significantly more protectionist than the President (See Destler (2005) and footnote~\ref{fn:ga_l_e} above).}
 
@@
 Finally, I assume that an external authority can ensure enforcement of the agreement in the one-shot game analyzed here.\footnote{Buzard (2015) considers a similar model in a repeated-game context and can therefore remove this restriction.}
-
-\subsection{Modeling Choices}
-\label{sec:choices}
-
-In modeling the objective of the legislature (see Equation~\ref{eq:ml}), I have introduced a variant of the general political support function introduced by Hillman (1982).
-While Hillman is credited with first proposing that policy makers choose tariffs in order to maximize their own political support, the PFS model contributed the explicit specification that lobbying contributions enter the political support function with constant marginal utility as $C + aW$.
-
-Dixit, Grossman and Helpman (1997) point out that linearity in contributions prevents complete analysis of distributional questions; the PFS form also restricts the returns to lobbying activity to be constant.\footnote{See Section \ref{sec:gh} for further discussion of this point and an in-depth discussion of the relationship between the functional form presented herein and the PFS form.
-Note that the assumption of diminishing returns to lobbying effort has been present in the literature going back at least to Findlay and Wellisz (1982).} They show that the PFS equilibrium analysis can be extended to a more general setting where, among other features, returns to lobbying are not restricted to be constant.
-
-Now, notice that, aside from the endogeneity of the weight the legislature places on the lobbying industry's profits, Equation~\ref{eq:ml} is precisely the \textit{deus ex machina} government objective function popularized by Baldwin (1987) that is commonly employed in the literature on the political economy of trade agreements.
-Instead of the weight on profits being constant as in the exogenous case or necessarily linear in contributions as in the PFS case, this form allows for both endogeneity and non-linearity.
-
-Since trade policies are overwhelmingly determined within the context of trade agreements, it is useful to have a framework to bring together the endogenous political pressure of PFS-style modeling with the trade agreements approach.
-The formulation in Equation~\ref{eq:ml} is intended as a bridge between the two.
-See Buzard (2016) for an example of how letting the weight the median legislator places on the profits of the import-competing industry $\ga(e,\ve)$ be affected by the level of lobbying effort as well a random variable $\ta$ can integrate both endogenous political activity and exogenous shocks into the analysis of questions regarding optimal trade agreement design and the political welfare of governments.
-
-Modeling the objective function so closely on the standard in the trade agreements literature allows for direct comparisons to the large extant body of work that studies exogenous shocks only, revealing cleanly the effects of the addition of endogenous lobbying---and if one wants---this different kind of political uncertainty.
-
-\begin{assumption}
-  $\ga(e,\ta)$ is increasing and concave in $e$ for every $\ta \in \Theta$.
-  \label{as:ga_c}
-\end{assumption}
-
-\begin{assumption}
-  $\ga(e,\ta)$ is increasing in $\ta$.
-  \label{as:ga_ta}
-\end{assumption}
-
-Assumption~\ref{as:ga_c} formalizes the intuition that the legislature favors the import-competing industry more the higher is its lobbying effort, but that there are diminishing returns to lobbying activity.\footnote{The diminishing returns here take the form of declining increments to the lobby's influence as effort increases; in Ethier (2012), the returns to lobbying decline as the level of protection increases.} It allows the political uncertainty inherent in $\ta$ to be correlated in interesting ways with the level of $e$, ruling out only (i) that higher effort levels make lower $\ga$-weights more likely and (ii) that effort and uncertainty are correlated in such a way that increasing lobbying effort changes the structure of uncertainty so that higher $\ga$-weights become more likely at an accelerating pace.
-Assumption~\ref{as:ga_ta} \textbf{simply provides for an intuitive labelling so that} larger realization of the uncertainty variable $\ta$ increase the value of the political economy weight.
-
-While a structural model of the legislative process might be more satisfying, the reduced form can deliver the key insights without distracting from the central policy choice and lobbying dynamics that take place at the level of the executives and lobbies, to which we turn next.
-
-The executives' objective function is taken to be identical to the legislatures' except that profits in the import-competing industry receive a fixed, exogenous weight.
-This construction permits me to integrate the influence of the import-competing industry on negotiations prior to the formation of the trade agreement while also reflecting the idea that the mechanisms through which political considerations influence the executives' preferences appear to be quite different from those at the legislative level---the electoral college and primary/caucus systems for instance.
-
-This setup does not strictly require that the executives are not lobbied.
-The assumption that the weight the executives place on import-competing profits is constant only requires that the extent to which they favor this industry cannot be \textit{directly} influenced by lobbying effort.
-The assumption is innocuous as long as the executives' preferences are not directly altered in a significant way by lobbying over trade.
-
-In the case of the post-war United States, where the Congress has consistently been significantly more protectionist than the President, this seems to reasonably reflect the political reality.
-For trade policy, where there are concentrated benefits but harm is diffuse, there are good reasons for this to be the case.
-Because the President has the largest constituency possible, delegating authority to the executive branch may simply be a mechanism for ``concentrating'' the benefits since consumers seem unable to overcome the free-riding problem.
-In fact, a strong argument can be made that power over trade policy has been delegated to the executive branch precisely \textit{because} it is less susceptible to the influence of special interests (Destler (2005)).
-
-Therefore, in line with both the theoretical and empirical literature, I make one more assumption on $\ga(e,\ta)$:
-
-\begin{assumption}
-  $\ga(e,\ta) \geq \ga_E \geq 1 \ \forall e,\ta$.
-  \label{as:ga_l_e}
-\end{assumption}
-
-That is, even for the least favorable outcome of the lobbying process, the legislature will be at least slightly more protectionist than the executive regardless of the lobby's effort choice.
-This assumption is not essential, but it matches well the empirical findings that politicians with larger constituencies are less sensitive to special interests, and it simplifies the analysis considerably, in particular ensuring that trade agreement tariffs are less than trade war tariffs (See Destler (2005) and footnote~\ref{fn:ga_l_e} above).
-
-Although the political process here most closely matches that of the United States in the post-war era,\footnote{In particular, the binary decision by the legislature about whether to abide by or break the trade agreement is modeled on the ``Fast Track Authority'' that the U.S.
-Congress granted to the Executive branch almost continuously from 1974-1994 and then again as ``Trade Promotion Authority'' from 2002-2007.} the model is broadly applicable because the central dynamic that emerges---that trade agreements are formed in the \textit{shadow} of lobbying pressure instead of in direct response to it---appears to be present in a wide range of countries in which the formation and maintenance of trade policy requires \textit{approval} by a politically susceptible body either at home or in a trading partner.
-
-We can see that the results apply beyond separation-of-powers presidential systems because the presence of veto points is what is essential for the dynamic posited here to occur.
-Although the number of formal veto points tends to be lower in, for instance, parliamentary systems, veto points are typically not absent altogether, at least in democracies (Henisz and Mansfield 2006).
-In reality, constraints on the executive are almost always present in democratic regimes, although they may operate in different ways, e.g.
-over questions of government dissolution and formation in parliamentary regimes.
-It is important to consider that constraints may be formal or informal, and they may come from legislative, regulatory, judicial, sub-national or even other executive bodies.
-
-%I believe the one major fudge between my model and Destler's evidence is acceptable: that the executive is actually the one who breaks trade agreements, but they do it largely to prevent the legislature from doing something far worse.
 

### 2016-06-10 letter-round2.tex
condense and move sec 2.4 text; closes #14; closes #15

@@
 The paper still degenerates far too often into a series of discussions that bring little additional insight.}
-	\item \textit{Section 2.4 either belongs to section 1.1 or should move as section 1.2. Section 2.4 is far too long.}
+	\item \textit{Section 2.4 either belongs to section 1.1 or should move as section 1.2. Section 2.4 is far too long.} \\
+	The text of Section 2.4 has been shorted and redistributed.
+I inserted the few paragraphs that concerned assumptions into Section 2.2 (preferences) since they could not come before the model was introduced.
+The rest (in condensed form) integrated nicely into the introduction and literature review.
+I have bolded them where inserted.
+
 	\item \textit{Your recall stuff far too often. Not all readers suffer from memory losses.}

### 2016-06-11 draft.tex
rewrite pages 1-2

@@
 Lobbying is in the style of Grossman and Helpman (1994) with a new feature: all actors face uncertainty arising from the complexity of the legislative process.
-I demonstrate that the lower the executives set tariffs in a trade agreement, the more effort lobbies put forth to prevent its ratification.
+I demonstrate that the higher the executives set tariffs in a trade agreement, the less effort lobbies put forth to prevent its ratification.
 Thus trade agreements act as a domestic political commitment device: executives set relatively high tariffs to discourage lobbying and increase the chance that the agreement will be ratified.
@@
 \textit{JEL classification:} D72, D80, F13, F53, F55 \\
-\textit{Keywords:} trade policy, lobbying, political economy, protection, international trade agreements, uncertainty}
+\textit{Keywords:} trade agreements, protection, lobbying, political economy, uncertainty}
 \end{quote}
@@
 Much of trade policy is determined in the context of international agreements that must be ratified by some domestic body.
-This paper seeks to shed light on such a policy-making process by incorporating Protection-for-Sale (Grossman and Helpman 1994, hereafter PFS) style lobbying into a model of trade agreements in which the executive branches of the governments set tariff levels in anticipation of political pressure upon a ratifying body, assumed here to be a legislature.
-This political pressure may be interpreted either as lobbying against ratification of the trade agreement or lobbying to break the agreement once in force.
+This paper sheds light on such a policy-making process by incorporating Protection-for-Sale (Grossman and Helpman 1994, hereafter PFS) style lobbying into a model of trade agreements in which the executive branches of the governments set tariff levels in anticipation of political pressure upon a ratifying body, assumed here to be a legislature.
 Important realism is achieved not only through the addition of the ratification constraint but also by the incorporation of political uncertainty into the model.
 
-Assuming that lobbies represent only import-competing interests, lobbies in this environment would like to prevent or derail a trade agreement that would reduce their tariff protection.
-They work toward this end by exerting effort to convince their legislatures to vote against the trade agreement but face some uncertainty about whether their efforts will result in a voting outcome consistent with their interests.
 I show that the probability that the trade agreement is derailed increases in lobbying effort and decreases in the trade agreement tariff.
@@
 
-To see how this works, consider first the case of political certainty, and assume that the executive is more pro-social than the legislature, as has been the case throughout the post-war period in the United States.\footnote{The standard assumption that the executive is less protectionist than the legislature is a special case of the finding that susceptibility to special interests generally declines with constituency size.
-One simple illustration from the realm of trade policy is the following: a legislator whose district has a large concentration of a particular industry does not take into account the impact of tariffs on the welfare of consumers in other districts, while the executive, whose constituency encompasses the whole country, will internalize these diffuse consumption effects.
-See Lohmann and O'Halloran (1994) for a detailed argument and Grossman and Helpman (2005) for a formal model that demonstrates the point for the legislature.\label{fn:ga_l_e}} For any trade-agreement tariff set by the executives, the lobby can calculate the effort level required to prevent the agreement from being ratified: it must shift the preferences of the median legislator just far enough so that he will choose the non-cooperative tariffs over the trade agreement tariffs.
-Because the lobby will pay this price so long as its benefit outweighs the cost, a trade war will occur unless the executive sets the trade-agreement tariff so that lobbying is not worthwhile.\footnote{As will be made clear below, non-ratification is equivalent to breaking a trade agreement once in place; likewise, the non-cooperative tariffs that result from failing to ratify the trade agreement are the same as the tariffs that would arise in trade war, so this model applies to both situations.}
+To see how this works, let's begin by considering the case of political certainty under the assumption that the executive is more pro-social than the legislature, as has been the case throughout the post-war period in the United States.
+Lobbies, assumed for simplicity to represent import-competing interests only, would like to prevent or derail a trade agreement that would reduce their tariff protection.\footnote{In this model, non-ratification is equivalent to breaking a trade agreement once in place; likewise, the non-cooperative tariffs that result from failing to ratify the trade agreement are the same as the tariffs that would arise in trade war, so this model applies to both situations.} They work toward this end by exerting effort to convince their respective legislatures to vote against the trade agreement.
+
+For any trade-agreement tariff set by the executives, the lobby can calculate the effort level required to prevent the agreement from being ratified: it must shift the preferences of the median legislator just far enough so that he will choose the non-cooperative tariffs over the trade agreement tariffs.
+Because the lobby will pay this price so long as its benefit outweighs the cost, a trade war will occur unless the executive sets the trade-agreement tariff so that lobbying is not worthwhile.
 
 Thus, any trade agreement under political certainty involves a tariff high enough to disengage the lobby completely; that is, even welfare-maximizing executives set high tariffs and induce zero contributions in equilibrium because they set policy in the shadow of a ratification decision influenced by lobbying.
-Careful modeling of the political process demonstrates that the relationship between deadweight loss calculations and the tariffs that are provided in a trade agreement may be mediated in important ways by the ratification constraint.
+Careful modeling of the political process demonstrates that the relationship between deadweight loss calculations and the tariffs that are provided in a trade agreement may be mediated in important ways by a ratification constraint.
 
@@
 It matches the empirical findings that politicians with larger constituencies are less sensitive to special interests and seems to represent particularly well the post-war U.S.
-where Congress has consistently been significantly more protectionist than the President (See Destler (2005) and footnote~\ref{fn:ga_l_e} above).}
+where Congress has consistently been significantly more protectionist than the President.}\footnote{cfr.
+Lohmann and O'Halloran (1994) for the general argument, Destler (2005) on U.S.
+trade politics, and Grossman and Helpman (2005) for a formal model.}
 

### 2016-06-13 draft.tex
revise rest of intro, remove parts of lit review

@@
 I show that the probability that the trade agreement is derailed increases in lobbying effort and decreases in the trade agreement tariff.
-Further, the lobby decreases its effort as the tariff it would face under the trade agreement rises: it has less to lose the higher is the trade agreement tariff.
-When forming the trade agreement tariff, the executives therefore trade off the welfare they receive under the agreement with the probability that the agreement actually remains in force.
-Ultimately, I find that observed lobbying effort may often be low because the executive branches of the governments set trade agreement tariffs sufficiently high to reduce political pressure on the legislatures.
-
-To see how this works, let's begin by considering the case of political certainty under the assumption that the executive is more pro-social than the legislature, as has been the case throughout the post-war period in the United States.
-Lobbies, assumed for simplicity to represent import-competing interests only, would like to prevent or derail a trade agreement that would reduce their tariff protection.\footnote{In this model, non-ratification is equivalent to breaking a trade agreement once in place; likewise, the non-cooperative tariffs that result from failing to ratify the trade agreement are the same as the tariffs that would arise in trade war, so this model applies to both situations.} They work toward this end by exerting effort to convince their respective legislatures to vote against the trade agreement.
-
-For any trade-agreement tariff set by the executives, the lobby can calculate the effort level required to prevent the agreement from being ratified: it must shift the preferences of the median legislator just far enough so that he will choose the non-cooperative tariffs over the trade agreement tariffs.
-Because the lobby will pay this price so long as its benefit outweighs the cost, a trade war will occur unless the executive sets the trade-agreement tariff so that lobbying is not worthwhile.
+Further, the lobby decreases its effort as the trade agreement tariff rises: it has less to lose the higher is the trade agreement tariff.
+When choosing the trade agreement tariffs, the executives therefore trade off the welfare they receive under the agreement with the probability that the agreement remains in force.
+Ultimately, I find that observed lobbying effort may often be low because the executives set trade agreement tariffs sufficiently high to reduce political pressure on the legislatures.
+
+To see how this works, let's begin by considering the case of political certainty.
+I assume the executive to be more pro-social than the legislature, as has been the case throughout the post-war period in the United States.
+Lobbies, assumed for simplicity to represent import-competing interests only, would like to prevent or derail a trade agreement that would reduce their tariff protection.\footnote{In this model, non-ratification is equivalent to breaking a trade agreement once in place.
+Likewise, the non-cooperative tariffs that result from failing to ratify the trade agreement are the same as the tariffs that would arise in trade war, so this model applies to both situations.} They work toward this end by exerting effort to convince their respective legislatures to vote against the trade agreement.
+
+For any trade-agreement tariff set by the executives, a lobby can calculate the effort level required to shift the preferences of the median legislator just far enough so that he will choose the non-cooperative tariffs over the trade agreement tariffs.
+Because the lobby will pay this price so long as its benefit outweighs the cost, a trade war will occur unless the executives set the trade-agreement tariff so that lobbying is not worthwhile.
 
 Thus, any trade agreement under political certainty involves a tariff high enough to disengage the lobby completely; that is, even welfare-maximizing executives set high tariffs and induce zero contributions in equilibrium because they set policy in the shadow of a ratification decision influenced by lobbying.
-Careful modeling of the political process demonstrates that the relationship between deadweight loss calculations and the tariffs that are provided in a trade agreement may be mediated in important ways by a ratification constraint.
+Careful modeling of the political process demonstrates that the relationship between deadweight loss calculations and the tariffs that are provided in a trade agreement are mediated in important ways by a ratification constraint.
 
 The fact that zero contributions are predicted in equilibrium is not particularly satisfying.
-The addition of political uncertainty provides the missing realism, smooths the results and delivers additional intuition.
-I assume all actors are uncertain about the weight the median legislator in a non-unitary legislature places on the profits of the lobbying industry.
-\textbf{Similar to probabilistic voting models that incorporate policy motivation such as Roemer (1997),} this represents the idea that the lobbying and voting ``game'' that goes on within legislatures is often complex enough that none of the actors know precisely what the outcome will be---that is, no one can predict exactly which legislator will be the median and therefore what weight will be placed on import-industry profits at the time of the vote.
-
-\textbf{One can conceptualize this uncertainty as being a result of the lobby's strategic choice to engage with only some key members of the legislature who wield significant influence but who cannot deliver policy decisions without the cooperation of others within the legislature, which is in turn uncertain.} For instance, perhaps a small number of lobbied legislators cannot deliver the votes of their non-lobbied colleagues with certainty.
-One can think of this uncertainty as an additional influence on the legislature's preferences about which all parties---executives, legislators and lobbies alike---are uncertain.
-
-In this case of political uncertainty, I show that the legislatures disrupt trade agreements with a higher probability and set higher trade war tariffs when lobbying effort increases.
-Because the lobbies respond to higher trade agreement tariffs by decreasing effort (and therefore the probability the agreement will fail to be ratified or will be broken), the executives must trade off the level of social welfare derived while a trade agreement is in force---since they prefer low tariffs---and the chances that the agreement will remain in force.
-
-The executives therefore set trade-agreement tariffs with an eye to discouraging lobbying activity and the accompanying probability of abrogation.
-It can do this because the trade agreement increases lobbying costs: the legislature now has to be compensated for the loss of the reduced foreign tariff if it breaks the trade agreement.
-Moreover, in addition to the standard terms-of-trade internalization, the executive's agenda-setting power allows it to reduce tariffs until lobbying activity is at the optimal level relative to a non-cooperative setting, reducing the amount of political pressure faced by the legislatures and the resultant expected tariffs.
-
-One contribution of this model, therefore, is the idea that trade agreements can act as a kind of domestic political commitment device.
-Rather than a time inconsistency such as that in Maggi and Rodr\'{i}guez-Clare (2007) where the welfare of the unitary policymaker is improved via commitment to a trade agreement by changing the investment decisions of firms, here we have an inconsistency between the preferences of the executive and legislative branches.\footnote{Indeed, depending on the modeling choices and parameters, the utility of the legislative branch taken as a whole may be lower under the trade agreement scenario.} Although the legislature has the final word both with and without the trade agreement, the executives are able to use the trade agreement to reduce lobbying effort, thus muting the legislature's protectionist bias and bringing the outcome more in line with their relatively liberal preferences.
-
-To the best of my knowledge, this is the first attempt to model lobbying specifically aimed at derailing trade agreements.
+The model becomes more realistic and delivers additional intuition when I allow all actors to be uncertain about the weight the median legislator places on the profits of the lobbying industry.
+\textbf{Similar to probabilistic voting models that incorporate policy motivation as in Roemer (1997),} this represents the idea that the lobbying and voting ``game'' that goes on within legislatures is often complex enough that none of the actors know precisely what the outcome will be---that is, no one can predict exactly which legislator will be the median and therefore what weight will be placed on import-industry profits at the time of the vote.
+
+\textbf{One can conceptualize this uncertainty as being a result of the lobby's strategic choice to engage with only some key members of the legislature.
+For instance, the lobbied legislators may promise to deliver the votes of their non-lobbied colleagues but may not be able to do so with certainty because those colleagues are subject to cross-pressuring from various sources.}
+
+In the face of this political uncertainty, lobbies respond to higher trade agreement tariffs by decreasing effort and in turn the legislatures disrupt trade agreements with a lower probability.
+Since they prefer low tariffs, the executives must trade off the level of social welfare derived while a trade agreement is in force with the chances that the agreement will remain in force.
+
+One contribution of this model is that it shows how trade agreements can act as a domestic political commitment device in a broad set of environments.
+The executives set trade-agreement tariffs with an eye to discouraging lobbying activity and the accompanying probability of abrogation.
+In addition to the standard terms-of-trade internalization, the executive's agenda-setting power allows it to reduce tariffs until lobbying activity is at the optimal level relative to a non-cooperative setting, reducing the amount of political pressure faced by the legislatures and the resultant expected tariffs.
+
+This is not a time inconsistency \`{a} la Maggi and Rodr\'{i}guez-Clare (2007) where the welfare of the unitary policymaker is improved via commitment to a trade agreement by changing firms' investment decisions.
+Instead, it is an inconsistency between the preferences of the executive and legislative branches.
+Although the legislatures have the final word, the executives use the trade agreement to reduce lobbying effort to mute the legislature's protectionist bias and bring the outcome more in line with their relatively liberal preferences.
+
+A further contribution is the demonstration that lobbying effort and the probability of trade disruptions vary in nuanced ways with the amount of political uncertainty present.
+Lobbying incentives are crucial but their impact is not necessarily as straightforward as predicted by models with unitary governments.
+Gawande, Krishna and Olarreaga (2009) provide cross-country empirical evidence that measures such as the number of checks and balances on the power of the legislature have predictive power for the PFS ``welfare-mindedness'' parameter.
+Thus, although empirical tests are outside the scope of this paper, there is suggestive evidence that political uncertainty is indeed an important factor empirically.
+
+To the best of my knowledge, this is also the first attempt to model lobbying specifically aimed at derailing trade agreements.
 Although this kind of politically-driven failure is not rare, prior models have not allowed exploration of the endogenous dynamics behind them.
-Careful consideration of the political process seems to be important for understanding why governments so often fail to cooperate.
-
-Moreover, I show that trade-agreement tariffs, lobbying effort and the probability of trade disruptions vary in nuanced ways with the amount of political uncertainty present: they are influenced strongly by lobbying incentives, but not in the straightforward ways predicted by models with unitary governments.
-Saiegh (2009) argues that political uncertainty is in part related to governmental and political structure.
-Although they do not discuss uncertainty specifically, Gawande, Krishna and Olarreaga (2009) provide cross-country empirical evidence that measures such as the number of checks and balances on the power of the legislature and the gap between the policy positions of the main political parties have predictive power for the PFS ``welfare-mindedness'' parameter.
-Thus, although empirical tests are outside the scope of this paper, there is suggestive evidence that political uncertainty is indeed an important factor empirically.
 
@@
 
-Attempts to more fully model the lobbying process have successfully reduced the parameter estimates for the government's welfare considerations somewhat, but the smallest estimates still indicate that the U.S.
-government, for instance, values social welfare about twenty times as much as contributions.\footnote{cfr.
+Attempts to more fully model the lobbying process have reduced the parameter estimates for the government's welfare considerations somewhat, but they still indicate that the U.S.
+government, for instance, values social welfare at least twenty times as much as contributions.\footnote{cfr.
 Gawande, Krishna and Robbins (2006); Mitra, Thomakos and Ulubasoglu (2006); Bombardini (2008); and Gawande, Krishna and Olarreaga (2012) among others.} Taking a different approach, Gawande and Hoekman (2006) suggest that one can reconcile the empirical results with the model by acknowledging the complexity of the policy-making process and the uncertainty that arises from it.
 
-In the non-unitary government model, these estimates and the earlier-cited stylized facts become consistent with each other.
-The ``case of the missing contributions,'' as Gawande and Hoekman (2006) refer to it, turns out to be well explained by an equilibrium phenomenon arising from the agenda-setting power of the executive branch.
-In particular, in this model, there is no inconsistency for the approximately 15$\%$ of sectors that receive protection but apparently put forth no lobbying effort.
+In the non-unitary government model, these estimates and the stylized facts are not inconsistent.
+The ``missing contributions,'' (Gawande and Hoekman 2006), can be explained as an equilibrium phenomenon arising from the agenda-setting power of the executive branch.
+In particular, in this model, there is no inconsistency for the 15$\%$ or so of sectors that receive protection but apparently put forth no lobbying effort.
 In this framework, even a welfare-maximizing executive and lobbying effort at zero can be reconciled with high tariff levels.
 
-\textbf{Although the political process here most closely matches that of the United States in the post-war era,\footnote{In particular, the binary decision by the legislature about whether to abide by or break the trade agreement is modeled on the ``Fast Track Authority'' that the U.S.
-Congress granted to the Executive branch almost continuously from 1974-1994 and then again as ``Trade Promotion Authority'' from 2002-2007 and 2015 to present.} the model is broadly applicable because the central dynamic that emerges---that trade agreements are formed in the \textit{shadow} of lobbying pressure instead of in direct response to it---appears to be present in a wide range of countries in which the formation and maintenance of trade policy requires approval by a politically susceptible body either at home or in a trading partner.}
-
-\textbf{The results apply beyond separation-of-powers presidential systems because the presence of veto points is what is essential for the dynamic posited here to occur.
-Although the number of formal veto points tends to be lower in, for instance, parliamentary systems, veto points are typically not absent altogether, at least in democracies (Henisz and Mansfield 2006).
+\textbf{Although the political process here most closely matches that of the United States in the post-war era,\footnote{In particular, the legislatures' decision to abide by or break the trade agreement is modeled on the ``Fast Track / Trade Promotion Authority'' that the U.S.
+Congress periodically grants to the Executive branch.} the model is broadly applicable because the central dynamic that emerges---that trade agreements are formed in the \textit{shadow} of lobbying pressure instead of in direct response to it---appears to be present in a wide range of countries.
+The results apply beyond separation-of-powers presidential systems because the presence of veto points is what is essential for this dynamic to occur.
+Although the number of formal veto points tends to be lower in, for instance, parliamentary systems, veto points are typically not absent altogether (Henisz and Mansfield 2006).
 In reality, constraints on the executive are almost always present in democratic regimes, although they may operate in different ways.}
@@
 \subsection{Related Literature}
-The foundations of this work rest on the `political support function' introduced by Hillman (1982) and Grossman and Helpman's (1994) `Protection for Sale.' Two related papers `Trade Wars and Trade Talks' (Grossman and Helpman (1995b)) and `The Politics of Free-Trade Agreements' (Grossman and Helpman (1995a)) first employed the PFS approach in the study of trade agreements.
+The foundations of this work rest on Grossman and Helpman's (1994) `Protection for Sale.' Two related papers `Trade Wars and Trade Talks' (Grossman and Helpman (1995b)) and `The Politics of Free-Trade Agreements' (Grossman and Helpman (1995a)) first employed the PFS approach in the study of trade agreements.
 The first considers ``Trade Wars'' and ``Trade Talks'' separately, whereas my approach allows the ``Trade Talks'' to be influenced by the desire to avoid a ``Trade War'' and thus views the ``Trade War'' as a crucial subgame.
@@
 The government objective function introduced in this paper is intended to be a bridge between the two.
+The objective function I introduce starts from the Baldwin (1987) government objective function and adds flexibility: the weight the legislature places on profits is endogenous (PFS), potentially non-linear (Findlay and Wellisz 1982; Dixit, Grossman and Helpman 1997) and stochastic.
 Buzard (2016) is an example of how this modeling approach can be used to integrate  endogenous political activity and exogenous shocks into the analysis of questions regarding optimal trade agreement design.}
 
-\textbf{The objective function I introduce starts from the \textit{deus ex machina} government objective function popularized by Baldwin (1987).
-It is more flexible in that the weight the legislature places on profits is endogenous (PFS) as well as potentially non-linear (Findlay and Wellisz 1982); Dixit, Grossman and Helpman 1997) and stochastic.}
-
-\textbf{Modeling the objective function so closely on the standard in the trade agreements literature allows for direct comparisons to the large extant body of work that studies exogenous shocks only, revealing cleanly the effects of the addition of endogenous lobbying---and if one wants---this different kind of political uncertainty.} To preview, this model is closest to that of Bagwell and Staiger (2005), with two main changes: in place of their unitary government signing a trade agreement and having different preferences ex-ante and ex-post, this model has two branches of government with differing preferences, where---in the case of the legislature---the political economy weights are determined endogenously.
-
-A very recent literature has begun to study the effects of trade agreements for reducing the uncertainty that firms (Handley 2014 and Handley and Lim\~{a}o 2015) and governments (Lim\~{a}o and Maggi 2015) face.
+\textbf{Modeling the objective function so closely on the standard in the trade agreements literature allows for direct comparisons to the large body of work that studies exogenous shocks only, revealing cleanly the effects of the addition of endogenous lobbying.} To preview, this model is closest to that of Bagwell and Staiger (2005), with two main changes: in place of their unitary government signing a trade agreement and having different preferences ex-ante and ex-post, this model has two branches of government with differing preferences, where the political economy weights of the legislature are determined endogenously.
 
@@
 Dai (2006) typifies a second approach, which, in line with the Schelling conjecture, argues that a country's legislative constraint assists its own executive in maintaining higher barriers.
-Ethier (2002) treats the effect of a different separation of powers on the international trade system: that between trade negotiators and the government officials who administer protection.
-
-The above work is patterned on the model of Milner and Rosendorff (1997), which explores how uncertainty can affect trade policy and the probability of ratification failure when political preferences, and therefore uncertainty, are exogenous;  Iida (1996) presents a similar model.
+Ethier (2002) treats the effect of a separation of powers between trade negotiators and the government officials who administer protection.
+
+The above work is patterned on Milner and Rosendorff (1997), who explore how uncertainty can affect trade policy and ratification failure when political preferences, and therefore uncertainty, are exogenous.
 Related is Le Breton and Salanie (2003), which studies lobbying when the lobby is uncertain about the preferences of a unitary decision maker.
@@
 
+%\begin{comment}
 \section{Appendix}
@@
 
-\item Handley, K., 2014. Exporting under Trade Policy Uncertainty: Theory and Evidence. Journal of international Economics 94, 50-66.
-
-\item Handley, K., Lim\~{a}o, N., 2015. Trade and Investment under Policy Uncertainty: Theory and Firm Evidence. American Economic Journal: Economic Policy 7, 189-222.
-
 \item Henisz, W., Mansfield, E., 2006. Votes and Vetoes: The Political Determinants of Commercial Openness. International Studies Quarterly 50, 189-211.
 
-\item Hillman, A., 1982. Declining Industries and Political-Support Protectionist Motives. The American Economic Review 5, 1180-1187.
-
 \item Horn, H., Maggi, G., Staiger, R. W., 2010. Trade Agreements as Endogenously Incomplete Contracts. American Economic Review 100, 394-419.
 
-\item Iida, K., 1996. Involuntary Defection in Two-Level Games. Public Choice 89, 283-303.
-
 \item Imai, S., Katayama, H., Krishna, K., 2009. Is protection really for sale? A survey and directions for future research. International Review of Economics and Finance 18, 181-191.
@@
 \item Le Breton, M., Zaporozhets, V., 2007. Legislative Lobbying under Political Uncertainty. Available at SSRN: \url{http://ssrn.com/abstract=1024686}.
-
-\item Lim\~{a}o, N., Maggi, G., 2015. Uncertainty and Trade Agreements. American Economic Journal: Microeconomics 7, 1-42.
 
@@
 
-\item Saiegh, S., 2009. Political  Prowess or Lady Luck? Evaluating Chief Executives' Legislative Success Rates. The Journal of Politics 71, 1342-1356.
-
 \item Song, Y., 2008. Protection for Sale: Agenda-Setting and Ratification in the Presence of Lobbying. Korea Institute for International Trade Policy Working Paper Series. Vol. 2008-22, August 2008.
@@
 \end{list}
+%\end{comment}
 

### 2016-06-13 draft.tex
more lit review condensing and deleting; down to five pg total for sec 1

@@
 government, for instance, values social welfare at least twenty times as much as contributions.\footnote{cfr.
-Gawande, Krishna and Robbins (2006); Mitra, Thomakos and Ulubasoglu (2006); Bombardini (2008); and Gawande, Krishna and Olarreaga (2012) among others.} Taking a different approach, Gawande and Hoekman (2006) suggest that one can reconcile the empirical results with the model by acknowledging the complexity of the policy-making process and the uncertainty that arises from it.
+Gawande, Krishna and Robbins (2006); Mitra, Thomakos and Ulubasoglu (2006); Bombardini (2008); and Gawande, Krishna and Olarreaga (2012) among others.
+Gawande and Magee (2012) provide a summary of the extensive literature, while Imai, Katayama and Krishna (2009) argue that many of the classification schemes present serious challenges for the validity of results.} Taking a different approach, Gawande and Hoekman (2006) suggest that one can reconcile the empirical results with the model by acknowledging the complexity of the policy-making process and the uncertainty that arises from it.
 
@@
 Mansfield, Milner and Rosendorff (2000) construct a model in which democracies trade more with each other because the domestic legislature's influence requires the foreign government to make deeper tariff cuts to ensure its cooperation.
-Dai (2006) typifies a second approach, which, in line with the Schelling conjecture, argues that a country's legislative constraint assists its own executive in maintaining higher barriers.
+Dai (2006) argues that a country's legislative constraint assists its own executive in maintaining higher barriers.
 Ethier (2002) treats the effect of a separation of powers between trade negotiators and the government officials who administer protection.
@@
 The above work is patterned on Milner and Rosendorff (1997), who explore how uncertainty can affect trade policy and ratification failure when political preferences, and therefore uncertainty, are exogenous.
-Related is Le Breton and Salanie (2003), which studies lobbying when the lobby is uncertain about the preferences of a unitary decision maker.
-Le Breton and Zaporozhets (2007) go a step further and replace the unitary decision maker with a legislature with multiple actors.
-Song (2008) presents a model in which policy-making power is shared between an executive and a unitary legislature where lobbies are able to endogenously influence the political preferences of the legislature in a context of unilateral policy making with no uncertainty, and Coates and Ludema (2001) study trade policy leadership in a model with endogenous lobbying in the presence of political uncertainty with imperfect monitoring.
-
-The extensive literature that explores the empirical implications of the PFS model is well-summarized by Gawande and Magee (2012).
-After the first generation of work consistently found very high estimates for the government's welfare-mindedness (see footnote~\ref{fn:gh_lobby} above), a second generation has explored several modifications of the model and estimating techniques.
-Most of these have focused on an improved accounting of the lobbying process, such as including lobbying by up- and downstream firms (Gawande, Krishna and Olarreaga (2012)), better identifying which firms in an industry lobby (Bombardini (2008)), taking into account lobbying by foreign interests (Gawande, Krishna and Robbins (2006)) or changing the way in which organized sectors are determined (Mitra, Thomakos and Ulubasoglu (2006)).
-Indeed, Imai, Katayama and Krishna (2009) argue that many of the classification schemes present serious challenges for the validity of results.
-
-I will begin by describing the model in detail.
-Section~\ref{sec:main} then presents the main results.
-In Section~ \ref{sec:example}, an extended example demonstrates these results as well as the impact of changing the institutional environment.
-Section~\ref{sec:dis} explores the connection between the model under consideration here and the `Protection for Sale' theoretical framework and includes in-depth discussions of the roles of uncertainty and the separation-of-powers structure.
-Section~\ref{sec:concl} concludes.
+Le Breton and Salanie (2003) study lobbying when the lobby is uncertain about the preferences of a unitary decision maker.
+Le Breton and Zaporozhets (2007) replace the unitary decision maker with a legislature with multiple actors.
+In Song (2008), policy-making power is shared between an executive and a unitary legislature where lobbies are able to endogenously influence the political preferences of the legislature in a context of unilateral policy making with no uncertainty, and Coates and Ludema (2001) study trade policy leadership in a model with endogenous lobbying in the presence of political uncertainty with imperfect monitoring.
 

### 2016-06-13 draft.tex
revise basic setup section

@@
 I assume the executive to be more pro-social than the legislature, as has been the case throughout the post-war period in the United States.
-Lobbies, assumed for simplicity to represent import-competing interests only, would like to prevent or derail a trade agreement that would reduce their tariff protection.\footnote{In this model, non-ratification is equivalent to breaking a trade agreement once in place.
-Likewise, the non-cooperative tariffs that result from failing to ratify the trade agreement are the same as the tariffs that would arise in trade war, so this model applies to both situations.} They work toward this end by exerting effort to convince their respective legislatures to vote against the trade agreement.
+Lobbies, assumed for simplicity to represent import-competing interests only, would like to prevent or derail a trade agreement that would reduce their tariff protection.
+They work toward this end by exerting effort to convince their respective legislatures to vote against or ``break'' the trade agreement.\footnote{In this model, non-ratification is equivalent to breaking a trade agreement once in place because tariffs in a ``trade war'' are assumed to be set non-cooperatively just as if there were no trade agreement.
+For simplicity, I refer to this as the ``break'' decision throughout.}
 
@@
 In each country, preferences are linear in good $N$, which is denoted the numeraire, while the demand functions for $X$ and $Y$ are assumed strictly decreasing and twice continuously differentiable.
-Further, the demand functions for $X$ and $Y$ are taken to be identical and denoted $D(P_i)$ in home and $D(P_i^*)$ in foreign.
-$P_i$ denotes the home price of good $i \in \left\{N,X,Y\right\}$ and $P_i^*$ denotes the foreign price of good $i$.
-
-Good $N$ is produced with labor alone with constant returns to scale technology so that $Q_N = l_N$.
-I assume the aggregate labor supply $l$ is large enough to ensure that the output of good $N$ is large enough to guarantee balanced trade.
+The demand functions for $X$ and $Y$ are taken to be identical and written $D(P_i)$ in home and $D(P_i^*)$ in foreign.
+$P_i$ ($P_i^*$) denotes the home (foreign) price of good $i \in \left\{N,X,Y\right\}$.
+
+Good $N$ is produced with labor alone so that $Q_N = l_N$.
+I assume the aggregate labor supply is large enough to ensure that the output of good $N$ is enough to guarantee balanced trade.
 The supply functions for good $X$ are $Q_X(P_X)$ and $Q_X^*(P_X^*)$ and are assumed strictly increasing and twice continuously differentiable for all prices that elicit positive supply.
 I also assume $Q_X^*(P_X) > Q_X(P_X)$ for any such $P_X$ so that the home country is a net importer of good $X$.
-The production structure for good $Y$ is taken to be symmetric, with both demand and supply such that the economy is separable in goods $X$ and $Y$.
-It is assumed that the production of goods $X$ and $Y$ requires, in addition to labor, the possession of a sector-specific factor that is available in inelastic supply and is non-tradable so that the income of owners of the specific factors is tied to the price of the good in whose production their factor is used.
+The production structure for good $Y$ is symmetric, with demand and supply such that the economy is separable in goods $X$ and $Y$.
+It is assumed that the production of goods $X$ and $Y$ requires labor and a sector-specific factor that is available in inelastic supply and is non-tradable so that the income of owners of the specific factors is tied to the price of the good in whose production their factor is used.
 
@@
 $$E_Y(P_Y)=Q_Y(P_Y)-D(P_Y) = D(P_Y^*)-Q_Y^*(P_Y) = M_Y^*(P_Y^*)$$
-where $M_X$ are home-county imports and $E_X^*$ are foreign exports of good $X$ and $E_Y$ are home-county exports and $M_Y^*$ are foreign imports of good $Y$.
-Given the production technology and no trade taxes, the price of the numeraire is equal to one in both countries and on the world market.
+where e.g. $M_X$ are home-county imports and $E_X^*$ are foreign exports of good $X$. The price of the numeraire is equal to one in both countries and on the world market.
 
 $P_X^W$ and $P_Y^W$ are decreasing in $\tau$ and $\tau^*$ respectively, while $P_X$ and $P_Y^*$ are increasing in the respective domestic tariff.
-This gives rise to a standard terms-of-trade externality.
-As profits and producer surplus (identical in this model) in a sector are increasing in the price of its good, profits in the import-competing sector are also increasing in the domestic tariff.
-This economic fact, combined with the assumptions on specific factor ownership, is what motivates political activity by import-competing lobbies.
-
-I now turn to describing the strategic players in the game.
-Each country's government is composed of two branches: an executive who can conclude trade agreements and a legislature that has final say on trade policy.
-In order to focus attention on protectionist political forces\textbf{, highlight the model's central mechanisms and minimize the number of actors that must be modeled}, I assume that only the import-competing industry in each country is politically-organized.\footnote{The model's central insights do not depend on this structure and can be extended to the case of multiple lobbies.
-In particular, adding an export lobby in the simplest case of no uncertainty predicts that import lobbies will remain disengaged while the executive will be able to reduce tariffs by recruiting support from exporters.
-The Panama Trade Promotion Act seems to follow this pattern, where there was virtually no lobbying against, significant lobbying in favor, and numerous industries receiving import tariff protection.\label{fn:lobby}} Further, I assume that this industry is represented by a single lobbying organization.
-
-Figure~\ref{fig:ext} illustrates the timing of the game from the perspective of the home country.\footnote{This extensive form is for the case where the foreign legislature does not break the trade agreement.} First, the executives set trade policy cooperatively within the context of an international agreement, that is they choose the trade agreement tariffs $\bta = \left(\tau^a,\tau^{*a}\right)$.
-After the trade agreement is concluded, the lobbies attempt to persuade the legislators in their respective countries to vote ``no'' on the ratification of the agreement; or alternatively, if there is no ratification hurdle, to break the trade agreement.\footnote{As will be seen shortly, these two interpretations are equivalent in this setting because tariffs in a ``trade war'' are assumed to be set non-cooperatively in the same manner as if there were no trade agreement.
-For simplicity, I refer to this as the ``break'' decision throughout.} They do this by choice of lobbying effort $e_b$, which can take on any non-negative value.
-
-Next, nature determines with probabilities $o$ and $o^*$ whether each legislature will have the opportunity to take a vote to break the agreement.
-I assume that these probabilities are less than one-half and mutually exclusive.\footnote{While this assumption is made for purposes of exposition and tractability, one can think of $o$ and $o^*$ as being determined by events beyond the executives' or legislatures' control that determine whether or not the legislatures will be willing to consider the lobby's request; for example, they would be affected by the occurrence of an economic crisis that diverts the legislature's attention from less-pressing matters.} This allows me to focus on the key interaction between the domestic actors and abstract from the strategic interaction between the lobbies in the two countries.\footnote{When the legislatures in both countries have an opportunity to break the agreement simultaneously, an incentive for each lobby to free ride on the effort of the other makes the optimization problem of the executives more complex as they are faced with the problem of how best to exploit the free-riding dynamic.
-\textbf{Alternatively, one can take the break decisions in the two countries to be sequential with the executives behind a veil of ignorance about which legislature will move first.
-In that set up, the end results are the same as in the current model but additional intermediate results are needed.}}
-
-After this, the uncertainty about the the median legislator's identity is resolved.
-This is modeled as the realization of a random variable $\ta_b$, the value of which all players observe simultaneously.
-The ``break stage'' concludes with the legislature making a choice to abide by the agreement or to break the agreement.
-
-In the event that the trade agreement does not enter into or remain in force, there is a final stage of lobbying, resolution of the uncertainty surrounding this decision, and voting to set the trade-war tariffs.\footnote{Alternatively, one can model the lobbying for the ratification decision as determining the potential non-cooperative tariff; this changes behavior slightly but does not change the fundamental dynamics.
-Yet another possibility is to replace the trade-war phase with the ex-ante status quo, which would most naturally be the non-cooperative equilibrium.
-This is precisely the outcome that obtains in the trade-war phase, so the results of the model are unaltered.} Once all political decisions are taken, producers and consumers make their decisions.
+This gives rise to a terms-of-trade externality.
+Profits and producer surplus (identical in this model) in a sector are increasing in the price of its good and also in the domestic tariff.
+This fact, combined with the assumptions on specific factor ownership, motivates political activity by import-competing lobbies.
+
+Each country's government is composed of an executive who concludes trade agreements and a legislature that has final say on trade policy.
+To focus attention on protectionist political forces\textbf{, highlight the model's central mechanisms and minimize the number of actors that must be modeled}, I assume that only the import-competing industry in each country is politically-organized and that it is represented by a single lobbying organization.\footnote{The model extends easily to the case of multiple lobbies.
+To illustrate, adding an export lobby with no uncertainty predicts import lobbies remain disengaged while the executives can reduce tariffs by recruiting support from exporters.
+The Panama Trade Promotion Act follows this pattern, with virtually no lobbying against, significant lobbying in favor, and numerous industries receiving import protection.\label{fn:lobby}}
 
@@
 \end{figure}
+
+Figure~\ref{fig:ext} illustrates the timing of the game from the perspective of the home country.\footnote{This extensive form is for the case where the foreign legislature does not break the trade agreement.} First, the executives set trade policy cooperatively within the context of an international agreement by choosing the trade agreement tariffs $\bta = \left(\tau^a,\tau^{*a}\right)$.
+Then the lobbies attempt to persuade the legislators in their respective countries to break the trade agreement by choice of lobbying effort $e_b$, which can take on any non-negative value.
+
+Next, nature determines with probabilities $o$ and $o^*$ whether each legislature will have the opportunity to take a vote to break the agreement.
+I assume that these probabilities are less than one-half and mutually exclusive.\footnote{While this assumption is made for purposes of exposition and tractability, one can think of $o$ and $o^*$ as being determined by events beyond the executives' or legislatures' control that determine whether or not the legislatures will be willing to consider the lobby's request; for example, they would be affected by the occurrence of an economic crisis that diverts the legislature's attention from less-pressing matters.} This allows me to focus on the key interaction between the domestic actors and abstract from the strategic interaction between the lobbies in the two countries.\footnote{When the legislatures in both countries have an opportunity to break the agreement simultaneously, an incentive for each lobby to free ride on the effort of the other makes the optimization problem of the executives more complex as they are faced with the problem of how best to exploit the free-riding dynamic.
+\textbf{Alternatively, one can take the break decisions in the two countries to be sequential with the executives behind a veil of ignorance about which legislature will move first.
+In that set up, the end results are the same as in the current model but additional intermediate results are needed.}}
+
+After this, the uncertainty about the the median legislator's identity is resolved.
+This is modeled as the realization of a random variable $\ta_b$, the value of which all players observe simultaneously.
+The ``break stage'' concludes with the legislature making a choice to abide by the agreement or to break the agreement.
+
+In the event that the trade agreement does not enter into or remain in force, there is a final stage of lobbying, resolution of the uncertainty surrounding this decision, and voting to set the trade-war tariffs.\footnote{Alternatively, one can model the lobbying for the ratification decision as determining the potential non-cooperative tariff; this changes behavior slightly not the fundamental dynamics.
+Another possibility is to replace the trade-war phase with the ex-ante status quo, which would most naturally be the non-cooperative equilibrium.
+This is precisely the outcome that obtains in the trade-war phase, so the results of the model are unaltered.} Once all political decisions are taken, producers and consumers make their decisions.
 

### 2016-06-13 draft.tex
finish revising model section

@@
 
-Figure~\ref{fig:ext} illustrates the timing of the game from the perspective of the home country.\footnote{This extensive form is for the case where the foreign legislature does not break the trade agreement.} First, the executives set trade policy cooperatively within the context of an international agreement by choosing the trade agreement tariffs $\bta = \left(\tau^a,\tau^{*a}\right)$.
-Then the lobbies attempt to persuade the legislators in their respective countries to break the trade agreement by choice of lobbying effort $e_b$, which can take on any non-negative value.
+Figure~\ref{fig:ext} illustrates the timing of the game from the perspective of the home country.\footnote{This extensive form is for the case where the foreign legislature does not break the trade agreement.} First, the executives set trade policy cooperatively within the context of an international agreement by choosing the trade agreement tariffs $\bta = \left(\tau^a,\tau^{*a}\right)$.\footnote{I use the convention throughout of representing a vector of tariffs for both countries $(\tau,\tau^*)$ as a single bold $\bm{\tau}$.} Then the lobbies attempt to persuade the legislators in their respective countries to break the trade agreement by choice of lobbying effort $e_b$, which can take on any non-negative value.
 
@@
 \label{sec:pref}
-As the economy is fully separable and the economic and political structures are symmetric, I focus throughout the paper on the home country and the $X$-sector.
-The details are analogous for $Y$ and foreign.\footnote{These are the relevant sectors for analysis because they are the sectors in which there are lobbies.}
-
-I specify the welfare function of the home legislature as
+As the economy is fully separable and the economic and political structures are symmetric, I focus throughout on the home country and the $X$-sector.
+The details are analogous for $Y$ and foreign.\footnote{These are the relevant sectors for analysis because they are the sectors in which there are lobbies.} I specify the welfare function of the home legislature as
 \begin{equation}
@@
 \end{equation}
-where $\mathit{CS}$ is consumer surplus, $\mathit{PS}$ is producer surplus, $\ga(e,\ta)$ is the weight placed on producer surplus (profits) in the import-competing industry, $e$ is lobbying effort, $\ta$ is a random variable, and $\mathit{TR}$ is tariff revenue.\footnote{Labor income $l$ could also be included in both median legislator and executive welfare.
-I omit it because its inclusion alters none of the results and only serves to complicate the exposition.} I model the decisions of the legislature as being taken by a median legislator.
-Notice that the weight the median legislator places on the profits of the import-competing industry, $\ga(e,\ve)$, is affected by the level of lobbying effort and a random variable $\ta$.
+where $\mathit{CS}$ is consumer surplus, $\mathit{PS}$ is producer surplus, $\ga(e,\ta)$ is the weight placed on producer surplus in the import-competing industry, and $\mathit{TR}$ is tariff revenue.\footnote{Labor income $l$ could also be included in both median legislator and executive welfare.
+I omit it because its inclusion alters none of the results and only serves to complicate the exposition.} I model the decisions of the legislature as being taken by a median legislator with the weight the median legislator places on import-competing industry profits affected by the level of lobbying effort $e$ and a random variable $\ta$.
 I make the following assumptions on $\ga(e,\ta)$:
@@
 
-Assumption~\ref{as:ga_c} formalizes the intuition that the legislature favors the import-competing industry more the higher is its lobbying effort, but that there are diminishing returns to lobbying activity.\footnote{The diminishing returns here take the form of declining increments to the lobby's influence as effort increases; in Ethier (2012), the returns to lobbying decline as the level of protection increases.} It allows the political uncertainty inherent in $\ta$ to be correlated in interesting ways with the level of $e$, ruling out only (i) that higher effort levels make lower $\ga$-weights more likely and (ii) that effort and uncertainty are correlated in such a way that increasing lobbying effort changes the structure of uncertainty so that higher $\ga$-weights become more likely at an accelerating pace.
-Assumption~\ref{as:ga_ta} \textbf{simply provides for an intuitive labelling so that} larger realization of the uncertainty variable $\ta$ increase the value of the political economy weight.
-
-Given its expectations and the legislature's preferences, the home lobby chooses its lobbying effort ($e_b$ to influence the break/ratification decision, which for simplicity, I will refer to henceforth simply as ``break,'' and $e_n$ to influence the trade war tariff) to maximize the welfare function:
+Assumption~\ref{as:ga_c} formalizes the intuition that the legislature favors the import-competing industry more the higher is its lobbying effort, but that there are diminishing returns to lobbying activity.\footnote{In Ethier (2012), the returns to lobbying decline as the level of protection increases.} It allows political uncertainty to be correlated with the level of $e$, ruling out only (i) that higher effort levels make lower $\ga$-weights more likely and (ii) that increasing lobbying effort changes the structure of uncertainty so that higher $\ga$-weights become more likely at an accelerating pace.
+Assumption~\ref{as:ga_ta} \textbf{simply provides for an intuitive labeling so that} larger realization of the uncertainty variable $\ta$ increase the value of the political economy weight.
+
+Given its expectations and the legislature's preferences, the home lobby chooses its lobbying effort ($e_b$ to influence the break decision and $e_n$ to influence the trade war tariff) to maximize the welfare function:
 \begin{equation}
@@
 where $\pi(\cdot)$ is the current-period profit and $\tau^a \ (\tau^\mathit{n})$ is the home country's tariff on the import good under a trade agreement (war).
-I use the convention throughout of representing a vector of tariffs for both countries $(\tau,\tau^*)$ as a single bold $\bm{\tau}$.
-
-In the first stage of the game, the executives choose the trade agreement tariffs $\bta=\left(\tau^a,\tau^{*a} \right)$ via a negotiating process that I assume to be efficient.
-In this symmetric environment, it therefore maximizes the expected joint payoffs:
+
+In the first stage, the executives choose the trade agreement tariffs.
+I model this choice via the Nash bargaining solution with a disagreement point of executive welfare in the non-cooperative game between the legislatures.
+In this symmetric environment, the executives maximize their expected joint payoffs:
 \begin{multline}
@@
 
-I model the executives' choice via the Nash bargaining solution where the disagreement point is the executives' welfare resulting from the Nash equilibrium in the non-cooperative game (i.e.
-in the absence of a trade agreement) between the legislatures.
-
 The home executive's welfare is specified as follows:
@@
 \]
-Note that this is identical to the welfare function for the legislature aside from the weight $\ga_E$ the executive places on the profits of the import industry, which is not a function of lobbying effort.
-\textbf{This setup does not strictly require that the executives are not lobbied, only that the extent to which the executives favor this industry cannot be \textit{directly} influenced by lobbying effort.
-The assumption is innocuous as long as the executives' preferences are not directly altered in a significant way by lobbying over trade.}
-
-\textbf{I make one more assumption on $\ga(e,\ta)$:}
+This is identical to the legislature's welfare function except the weight the executive places on the profits of the import industry is not a function of lobbying effort.
+\textbf{This setup does not require that the executives are not lobbied, only that the extent to which the executives favor this industry cannot be \textit{directly} influenced by lobbying effort.
+The assumption is innocuous as long as the executives' preferences are not directly altered in a significant way by lobbying over trade.
+I make one more related assumption on $\ga(e,\ta)$:}
 
@@
 
-\textbf{That is, even for the least favorable outcome of the lobbying process, the legislature will be at least slightly more protectionist than the executive regardless of the lobby's effort choice.
-This assumption is not essential, but it simplifies the analysis considerably, ensuring that trade agreement tariffs are less than trade war tariffs.
-It matches the empirical findings that politicians with larger constituencies are less sensitive to special interests and seems to represent particularly well the post-war U.S.
-where Congress has consistently been significantly more protectionist than the President.}\footnote{cfr.
+\textbf{That is, even for the least favorable outcome of the lobbying process, the legislature is more protectionist than the executive regardless of the lobby's effort choice.
+This is not essential but it simplifies the analysis, ensuring that trade agreement tariffs are smaller than trade war tariffs.
+It matches the empirical finding that politicians with larger constituencies are less sensitive to special interests and represents well the post-war U.S.
+where Congress has consistently been more protectionist than the President.}\footnote{cfr.
 Lohmann and O'Halloran (1994) for the general argument, Destler (2005) on U.S.
@@
 \label{sec:info}
-I examine a simple class of equilibria that have the following features.
-
-First, on information.
-Although uncertainty is present, information about it is symmetric.
+I examine a simple class of equilibria that have three key features.
+First, although uncertainty is present, information about it is symmetric.
 However, in line with the literature, I assume the lobby's contribution is not observable to the foreign legislature.
-The implication is that the lobby can directly influence only the home legislature, and so the influence of one country's lobby on the other country's legislature occurs only through the tariffs selected.\footnote{cfr.
-Grossman and Helpman (1995b), page 685.} Given the game's structure and the assumption that lobbying effort is not observable across international borders, players in the same country can take advantage of more information than those who are in different countries.
-That is, the solution concept employed here is perfect public equilibrium (PPE).
-
-Second, whenever there is a possibility of multiple equilibria, I focus on the one that maximizes the welfare of the executives.
-Given the assumption that the executives are social welfare maximizers, this selection criterion puts results in terms of the maximal level of trade policy cooperation that is possible.
+This implies that the influence of one country's lobby on the other country's legislature occurs only through the tariffs selected.\footnote{cfr.
+Grossman and Helpman (1995b), page 685.} Since players in the same country can take advantage of more information than those who are in different countries, the solution concept employed here is perfect public equilibrium (PPE).
+
+Second, whenever there is a possibility of multiple equilibria, I focus on the one that maximizes the executives' welfare.
+Since I assume the executives are social welfare maximizers, this selection criterion puts results in terms of the maximal level of trade policy cooperation that is possible.
 It also allows the question of whether governments use trade agreements as political commitment devices to be answered in a straightforward way.
 
-Finally, I assume that an external authority can ensure enforcement of the agreement in the one-shot game analyzed here.\footnote{Buzard (2015) considers a similar model in a repeated-game context and can therefore remove this restriction.}
+Finally, I assume that an external authority can ensure enforcement of the agreement.\footnote{Buzard (2015) considers a similar model in a repeated-game context and can therefore remove this restriction.}
 

### 2016-06-13 draft.tex
revise trade war tariffs section

@@
 \label{sec:main}
-To understand how the executives optimally structure trade agreements, we must first examine the incentives of the lobbies and how the legislatures make decisions regarding breach of the trade agreement, including how trade-war tariffs are set.
-The symmetric, separable structure of the model permits restriction of attention to the home country.
+To understand how the executives optimally structure trade agreements, we must first examine the lobbies' incentives and the legislatures' decisions regarding breach of the agreement, including how trade-war tariffs are set.
+The symmetric, separable model structure permits restriction of attention to the home country.
 
@@
 In the event that the trade agreement is broken, the legislature sets its tariff $\tau$ unilaterally by maximizing Equation~\ref{eq:ml} given $\tau^*$.
-Because there are no interactions between the home and foreign tariffs, the home and foreign trade-war tariffs are independent and the home country's tariff in a trade war maximizes weighted home-country welfare in the $X$-sector only.
-The foreign legislature's decision problem is analogous, and unilateral optimization leads to what I refer to as the (expected) Nash or trade-war tariffs $\tau^n$ as the solution to the following first order condition:\footnote{Note that the random variable in the median legislator's weight on import profits is written as $\ta_n$ at this stage to distinguish it from $\ta_b$ at the break stage: as two separate votes are taken, there are two distinct realizations of uncertainty.
-It is not necessary to restrict them to come from the same distribution, although in applications it would seem unlikely that the distribution would change.}
+Because there are no interactions between the home and foreign tariffs, the home country's tariff maximizes weighted home-country welfare in the $X$-sector only.
+Unilateral optimization leads to what I refer to as the (expected) Nash or trade-war tariffs $\tau^n$ as the solution to the following first order condition:\footnote{Note that the random variable in the median legislator's weight on import profits is written as $\ta_n$ at this stage to distinguish it from $\ta_b$ at the break stage: as two separate votes are taken, there are two distinct realizations of uncertainty.}
 \begin{equation}
@@
 \end{equation}
-That is, at this stage, the lobby chooses the level of effort that equates its expected marginal increase in profits with its marginal payment.
-
-In accordance with intuition, because profits are increasing in the tariff, trade war tariffs are increasing in the weight attached to the profits of the import-competing industry.
+That is, the lobby equates the expected marginal increase in profits with its marginal payment.
+
+Because profits are increasing in the tariff, trade war tariffs are increasing in the weight attached to the profits of the import-competing industry.
 This can be seen by rearranging Equation~\ref{eq:lobtw} as
@@
 \end{equation}
-since $\frac{\partial \ga}{\partial e_n}$ is positive by Assumption~\ref{as:ga_c} and profits are increasing in prices which are in turn increasing in the tariff.
-This expression can also demonstrate that the second order condition for the legislature's problem is always satisfied.
-By the implicit function theorem, $\frac{\partial \tau^n}{\partial \ga} = -\frac{\frac{\partial \text{Equation}~\ref{eq:legfoc}}{\partial \ga}}{\frac{\partial \text{Equation}~\ref{eq:legfoc}}{\partial \tau^n}} = -\frac{\frac{\partial \pi(\tau^n)}{\partial \tau^n}}{SOC}$ where ``SOC'' stands for `Second Order Condition', the second derivative of the legislature's objective function which should be everywhere negative.
+since $\frac{\partial \ga}{\partial e_n}$ is positive by Assumption~\ref{as:ga_c} and profits are increasing in the tariff.
+This expression demonstrates that the second order condition for the legislature's problem is always satisfied.
+By the implicit function theorem, $\frac{\partial \tau^n}{\partial \ga} = -\frac{\frac{\partial \text{Equation}~\ref{eq:legfoc}}{\partial \ga}}{\frac{\partial \text{Equation}~\ref{eq:legfoc}}{\partial \tau^n}} = -\frac{\frac{\partial \pi(\tau^n)}{\partial \tau^n}}{SOC}$ where ``SOC'' is the second derivative of the legislature's objective function that should be everywhere negative.
 Setting this expression equal to Equation~\ref{eq:lem1}, the second order condition must be satisfied since $\frac{\partial \ga}{\partial e_n}$ is positive by Assumption~\ref{as:ga_c}.

### 2016-06-13 draft.tex
revise break section

@@
 \label{sec:break}
-With trade-war decisions specified, we can proceed by backward induction to analyze the incentives regarding the legislature's decision to uphold or break the trade agreement.
-I assume that the form of the political weighting function is the same regardless of the break decision, but the model could be extended to allow the two to differ.
-What is in principle different are the lobbying efforts toward influencing the break decision, denoted $e_b$ at this stage to distinguish them from those that are made to influence the Nash tariff, $e_n$.
-The realization of uncertainty about the median legislator's identity will similarly be denoted $\ve_b$ so as to distinguish it from $\ve_n$ at the trade-war stage.
-
-The legislature will break the agreement and set the tariff at $\tau^n$ if the median legislator's utility from the Nash tariffs is higher than his utility from the trade agreement tariffs, i.e. if
+We can now proceed by backward induction to analyze the legislature's incentives to uphold or break the trade agreement.
+The legislature will break the agreement and set the tariff at $\tau^n$ if the median legislator's utility from the Nash tariffs is higher than his utility from the trade agreement tariffs, i.e.
+if
 \begin{equation}
@@
 \end{equation}
-\noindent Recall that a bold $\bm{\tau}$ represents a vector of tariffs for both countries $(\tau,\tau^*)$.
-
-The legislature's decisions are stochastic, so the outcome of the vote on whether or not to break the trade agreement, as well as the level at which the tariff will be set in the case of a disruption, is not known to \textit{any} player until the uncertainty over the identity of the median legislator is resolved at the moment the vote takes place on each issue.\footnote{We might think of the trade-war tariff as being set in the legislation before a vote takes place in the case of administered protection for instance; the results are robust to this variant.} I represent the probability that the home legislature breaks the trade agreement and sets the tariff at $\tau^n$ as:
+
+The outcome of the vote on whether or not to break the trade agreement is not known to \textit{any} player until the uncertainty over the identity of the median legislator is resolved at the moment the vote takes place.\footnote{We might think of the trade-war tariff as being set in the legislation before a vote takes place in the case of administered protection for instance; the results are robust to this variant.} I represent the probability that the home legislature breaks the trade agreement and sets the tariff at $\tau^n$ as:
 \begin{multline}
-  b(e_b,\bta,\btn) = \expect_{\ga|e_b} \bm{1} [ W_{ML}(\btn,\ga(e_b,\ve_b)) > W_{ML}\left(\bta,\ga(e_b,\ve_b)\right) ] \\ = \Pr [ W_{ML}(\btn,\ga(e_b,\ve_b)) > W_{ML}\left(\bta,\ga(e_b,\ve_b)\right) | e_b]
+  b(e_b,\bta) = \expect_{\ga|e_b} \bm{1} [ W_{ML}(\btn,\ga(e_b,\ve_b)) > W_{ML}\left(\bta,\ga(e_b,\ve_b)\right) ] \\ = \Pr [ W_{ML}(\btn,\ga(e_b,\ve_b)) > W_{ML}\left(\bta,\ga(e_b,\ve_b)\right) | e_b]
   \label{eq:b}
@@
 Lobbying affects only the weight the legislature places on the profits of the import-competing industry.
-These profits are higher in a trade war than under a trade agreement, so given Assumption~\ref{as:ga_c}, the expectation of $\ga$ is increasing in lobbying effort, implying that the legislature becomes more favorably inclined toward the high trade-war tariff and associated profits as lobbying increases and therefore more likely to break the trade agreement.
-However, there are decreasing returns for the lobby: increasing effort leads to less than proportional increases in the probability that the trade agreement will be broken.
-
-The probability of a trade war, just like the trade-war tariffs, is increasing in $\ga$, so it's the natural assumption that $\ga$ is increasing in lobbying effort that implies that the lobby can increase both its trade war tariff and the probability of a trade war by raising its lobbying effort.
+These profits are higher in a trade war than under a trade agreement.
+Assumption~\ref{as:ga_c} implies that the legislature becomes more favorably inclined---albeit at a decreasing rate---toward the high trade-war tariff and associated profits as lobbying increases and therefore more likely to break the trade agreement.
 
@@
 
-Because the home country is an exporter of good $Y$, the lower world price for $Y$ has a negative effect on producer surplus that is larger than the positive effect on consumer surplus.
+The lower world price for home's export good has a negative effect on producer surplus that is larger than the positive effect on consumer surplus.
 Thus the net effect of an increase in foreign trade-agreement tariffs on both home legislative and home total social welfare is negative, leading to an increased probability that the trade agreement will be broken.
 
-The relationship between the home country's trade-agreement tariff $\tau^a$ and the legislature's probability of breaking the agreement is slightly more complex because $\tau^a$ interacts directly with $\ga(e_b,\ta_b)$.
-The intuition is as follows.
-The median legislator will vote to break the agreement if it receives higher welfare from the trade war tariff than the trade agreement tariff.
-For any given level of effort $e_b$, when the trade agreement tariff is very low, only a small set of realizations of $\ta_b$ that lead to the lowest values of $\ga(e_b,\ta_b)$ will lead to approval of the trade agreement.
-Thus the break probability is high.
+The relationship between the home country's trade-agreement tariff and the legislature's probability of breaking the agreement is more complex.
+For any given level of effort $e_b$, when the trade agreement tariff is very low, only a small set of realizations of $\ta_b$ that lead to the lowest values of $\ga(e_b,\ta_b)$ will lead to approval of the trade agreement, implying a high break probability.
 As the trade agreement tariff rises, larger values of $\ga(e_b,\ta_b)$ are consistent with approving the trade agreement, so the set of $\ta_b$'s that lead to trade agreement approval is larger.
 That is, the break probability decreases as $\tau^a$ increases.
-Even if the trade agreement tariff is so high that it is disliked by most, or even all, potential median legislators, it is preferable to the --- even higher --- trade war tariff, so the probability of breaking either becomes zero or continues to decrease in $\tau^a$.
+Even if the trade agreement tariff is so high that it is disliked by all potential median legislators, it is preferable to the trade war tariff, which is even higher.
+Thus the probability of a break eventually falls to zero.
 
@@
 
-In the symmetric equilibria that are focal in the analysis that follows, any increase in $\tau^a$ will be accompanied by an equal increase in $\tau^{*a}$, so it is useful to state a result about the combined impact of the home and foreign tariff when the two are constrained to be equal:
+In the symmetric equilibria that are focal in the analysis that follows, any increase in $\tau^a$ is accompanied by an equal increase in $\tau^{*a}$.
+Combining the impact of the home and foreign tariff when the two are constrained to be equal:
 \begin{lemma}
@@
 
-\noindent When any change in one of the trade agreement tariffs must be met with an equal change in the other (as in a symmetric agreement), the negative impact on the legislature's propensity to break the trade agreement of raising the home tariff always outweighs the positive impact of raising the foreign tariff because of the legislature's bias toward the import-competing industry.
+\noindent When a change in one of the trade agreement tariffs must be met with an equal change in the other, the negative impact on the legislature's propensity to break the agreement of raising the home tariff outweighs the positive impact of raising the foreign tariff because of the legislature's bias toward the import-competing industry.
 

### 2016-06-14 draft.tex
revise lobby section

@@
 \label{sec:lob_un}
-The lobby chooses its level of effort $e_b$ as a function of $\bta$, given the implications of that choice on the legislature's probability of breaking the agreement.
-The lobby's decision is taken according to
+The lobby chooses its level of effort as a function of $\bta$, given the implications of that choice on the legislature's probability of breaking the agreement.
+The lobby maximizes probability-weighted profits net of effort:
 \[
@@
 \]
-The lobby simply maximizes probability-weighted profits net of effort.
-From its first order condition, we can see that the lobby balances the cost of an extra dollar of lobbying expenditure with the higher profits from a trade war weighted by the increase in the probability of receiving the higher profit level that expenditure induces:
+The first order condition shows that the lobby balances the cost of an extra dollar of expenditure with the higher profits from a trade war weighted by the increase in the probability of receiving the higher profit level induced by that expenditure:
 \begin{equation}
@@
 \end{equation}
-The second order condition is satisfied given Assumption~\ref{as:ga_l_e} and Result~\ref{res:bincC}. However, the solution is not guaranteed to be interior, for which we need
+The second order condition is satisfied given Assumption~\ref{as:ga_l_e} and Result~\ref{res:bincC}. To guarantee an interior solution, we need
   \begin{equation}
@@
   \end{equation}
-For instance, if the executives were to set $\tau^a = \tau^n$, there would be no incentive for the lobby to make a positive contribution.
-We will see that there are some cases in which it may be in the executives' joint interest to set trade agreement tariffs so as to completely disengage the lobby.
-The following results thus only hold when Condition~\ref{ine:lobint} is satisfied, that is, when the marginal impact of the first lobbying dollar on the break probability is sufficiently high to make engaging in the political process worthwhile for the lobby.\footnote{This is primarily a condition on the curvature of the political weighting function but also involves the production and demand structure.}
+If the executives were to set $\tau^a = \tau^n$, there would be no incentive for the lobby to make a positive contribution.
+There are some cases in which it is in the executives' interest to set trade agreement tariffs so as to completely disengage the lobby.
+The following results only hold when the marginal impact of the first lobbying dollar on the break probability is sufficiently high to make engaging in the political process worthwhile for the lobby.
 
@@
 
-\noindent Raising the trade agreement tariff decreases the benefit of inducing a break in the trade agreement by raising profits under the trade agreement.
-This is in many ways the key result of this paper: the executives can reduce lobbying effort by setting higher tariffs in their trade agreement.
+\noindent Raising the trade agreement tariffs decreases the benefit of inducing a break in the trade agreement by raising profits under the trade agreement.\footnote{Note that the symmetric increase in the foreign tariff does not impact profits in the import-competing industry, but it can impact lobbying incentives through the break probability.
+However, there is an analogous effect on the break probability from raising $\tau^a$ and the net effect is negative given Assumption~\ref{as:ga_c}.
+This assumption can be significantly weakened: this effect can be positive as long as it is smaller in absolute value than the negative effect on trade agreement profits.} This key result implies that the executives can reduce lobbying effort by setting higher tariffs in their trade agreement.
 We will see in the next section how this shapes the executives' joint decision.
-
-Note that the symmetric increase in the foreign tariff does not impact profits in the import-competing industry, but it can impact lobbying incentives through the break probability.
-However, there is an analogous effect on the break probability from raising $\tau^a$ and the net effect is negative given Assumption~\ref{as:ga_c} on the probability distribution function of $\ga(e,\ta)$, which ensures that the effect of lobbying effort on the break probability does not increase as $\bta$ increases.
-This assumption can be significantly weakened: this effect can be positive as long as it is smaller in absolute value than the negative effect on trade agreement profits.
 

### 2016-06-15 draft.tex
move executives' NBS defn to sec 2.2

@@
 
-In the event that the trade agreement does not enter into or remain in force, there is a final stage of lobbying, resolution of the uncertainty surrounding this decision, and voting to set the trade-war tariffs.\footnote{Alternatively, one can model the lobbying for the ratification decision as determining the potential non-cooperative tariff; this changes behavior slightly not the fundamental dynamics.
+In the event that the trade agreement does not enter into or remain in force, there is a final stage of lobbying, resolution of the uncertainty surrounding this decision, and voting to set the trade-war tariffs $\btn$.\footnote{Alternatively, one can model the lobbying for the ratification decision as determining the potential non-cooperative tariff; this changes behavior slightly not the fundamental dynamics.
 Another possibility is to replace the trade-war phase with the ex-ante status quo, which would most naturally be the non-cooperative equilibrium.
@@
 \end{equation}
-where $\pi(\cdot)$ is the current-period profit and $\tau^a \ (\tau^\mathit{n})$ is the home country's tariff on the import good under a trade agreement (war).
+where $\pi(\cdot)$ is the current-period profit.
 
 In the first stage, the executives choose the trade agreement tariffs.
-I model this choice via the Nash bargaining solution with a disagreement point of executive welfare in the non-cooperative game between the legislatures.
-In this symmetric environment, the executives maximize their expected joint payoffs:
+With executive welfare in a trade war the disagreement point, the division of surplus according to the Nash bargaining solution is
+\[
+  V_E(\bta) + t = W_E(\btn) + \frac{1}{2} \left( MV(\bta) - W_E(\btn) - W_E^*(\btn) \right)
+\]
+\[
+  V_E^*(\bta) - t = W_E^*(\btn) + \frac{1}{2} \left( MV(\bta) - W_E(\btn) - W_E^*(\btn) \right)
+\]
+where expectation operators are dropped for convenience, $MV(\bta)$ is the maximized joint value and $t$ is an intergovernmental transfer.\footnote{While direct monetary transfers have to date rarely been used in practice, it seems appropriate to interpret linked concessions on non-trade issues as indirect transfers (Klimenko, Ramey and Watson (2008); Maggi and Staiger (2011)).
+Here, transfers do not occur as long as the full set of symmetry assumptions are maintained but are an important consideration in asymmetric environments.} The solution to this system of equations determines the trade agreement tariffs.
+
+I simplify the problem by assuming that political constraints prevent the executives from choosing asymmetric tariffs.
+This implies that the executives simply maximize their expected joint payoffs:\footnote{I have assumed that $W_E(\tau) = W_E^*(\tau) \ \forall \tau$; further in a symmetric equilibrium, $\tau^n =\tau^{*n}$ and $\tau^a =\tau^{*a}$, so
+\[
+  V_E(\bta) + t = W_E(\btn) + \frac{1}{2} \left( MV(\bta) - W_E(\btn) - W_E^*(\btn) \right) = W_E(\btn) + \frac{1}{2} \left( MV(\bta) - 2W_E(\btn) \right) = \frac{1}{2}\expect MV(\bta)
+\]}
 \begin{multline}
@@
 \end{multline}
-
 The home executive's welfare is specified as follows:
@@
 
-The lower world price for home's export good has a negative effect on producer surplus that is larger than the positive effect on consumer surplus.
-Thus the net effect of an increase in foreign trade-agreement tariffs on both home legislative and home total social welfare is negative, leading to an increased probability that the trade agreement will be broken.
+The lower world price for home's export good has a larger negative effect on producer surplus than the positive effect on consumer surplus.
+The net effect of an increase in foreign trade-agreement tariffs on home legislative welfare is negative, leading to an increased probability that the trade agreement will be broken.
 
@@
 \label{sec:ta}
-The executives choose trade agreement tariffs $\bta=\left(\tau^a,\tau^{*a} \right)$ to maximize Equation \ref{eq:jv}.
-I call the maximized joint value $MV(\bta)$ and write the division of surplus according to the Nash bargaining solution, which, due to symmetry, can be simplified from its general multiplicative form as
-\[
-  V_E(\bta) + t = W_E(\btn) + \frac{1}{2} \left( MV(\bta) - W_E(\btn) - W_E^*(\btn) \right)
-\]
-\[
-  V_E^*(\bta) - t = W_E^*(\btn) + \frac{1}{2} \left( MV(\bta) - W_E(\btn) - W_E^*(\btn) \right)
-\]
-where expectation operators have been dropped for convenience and $t$ is an intergovernmental transfer.\footnote{While direct monetary transfers have to date rarely been used in practice, it seems appropriate to interpret linked concessions on non-trade issues as indirect transfers (Klimenko, Ramey and Watson (2008); Maggi and Staiger (2011)).
-Here, transfers do not occur as long as the full set of symmetry assumptions are maintained but are an important consideration in asymmetric environments.} The solution to this system of equations determines the tariffs $\bta$ that will be set in the trade agreement.
-
-I simplify the problem by assuming that political constraints prevent the executives from choosing asymmetric tariffs.
-This restricts attention to symmetric solutions in the fully symmetric environment under consideration, which means that the trade agreement tariffs for which we're looking are simply those that maximize the joint welfare of the executives.\footnote{I have assumed that $W_E(\tau) = W_E^*(\tau) \ \forall \tau$; further in a symmetric equilibrium, $\tau^n =\tau^{*n}$ and $\tau^a =\tau^{*a}$, so
-\[
-  V_E(\bta) + t = W_E(\btn) + \frac{1}{2} \left( MV(\bta) - W_E(\btn) - W_E^*(\btn) \right) = W_E(\btn) + \frac{1}{2} \left( MV(\bta) - 2W_E(\btn) \right) = \frac{1}{2}\expect MV(\bta)
-\]}
-In order to solve for them, I represent the probability that the trade agreement will be broken as $B(\bta)=b(e(\bta),\bta)$ where $e(\bta)$ is the best response function implicit in Equation \ref{eq:lobbyfoc}, the lobby's first order condition.
-
-The problem then becomes to maximize the following modified version of Equation \ref{eq:jv}:
+I represent the probability that the trade agreement will be broken as $B(\bta)=b(e(\bta),\bta)$ where $e(\bta)$ is the best response function implicit in Equation \ref{eq:lobbyfoc}, the lobby's first order condition.
+
+To solve for the trade agreement tariffs, one must maximize the following modified version of Equation \ref{eq:jv}:
 \begin{equation}
@@
 \end{equation}
-where I have written the sum of the home and foreign executives' utilities as $\bm{W_E}(\cdot)$.
-Here $o$ and $o^*$ are the probabilities that the home and foreign legislatures, respectively, will have the opportunity to break the trade agreement; these are assumed to be mutually exclusive events that are realized after the conclusion of the trade agreement.
+where $\bm{W_E}(\cdot)$ is the sum of the home and foreign executives' utilities.
 

### 2016-06-15 draft.tex
finish revising trade agreements part of results section

@@
 
-Of central concern is how the break probabilities behave as a function of the trade agreement tariffs given the lobby's optimal response.
-Recalling the definition of the total break probability, $B(\bta)=b(e_b(\bta),\bta)$, we must combine the direct effect of tariffs on the legislature's break decision with the indirect effect through the impact on the lobby's choice of $e_b$.
-
-The direct effect is negative by Lemma~\ref{res:bcomb}.
-The fact that the indirect effect of raising tariffs leads to a reduction in lobbying (Result~\ref{res:lobby}) and therefore a further reduction in the propensity of the legislature to break the agreement (Result~\ref{res:bincC}) leads to the following key result: the overall probability that the trade agreement will be broken is decreasing in the trade agreement tariffs.
+Of central concern is how the break probability varies as both a direct and indirect function of $\bta$ given the lobby's optimal response.
+Result~\ref{res:bcomB} takes into effect both the direct and indirect effects:
 
@@
 
-\noindent That is, when the executives raise the trade agreement tariffs the legislature becomes less likely to abrogate the agreement, for three reasons.
-First, the legislature prefers a higher domestic tariff \textbf{(Lemma~\ref{res:leg_a})}; second, the higher tariff discourages lobbying \textbf{(Result~\ref{res:lobby})}; and finally, the lower lobbying effort reduces the legislature's preferred tariff further \textbf{(Lemma~\ref{res:bcomb})}.
+\noindent When the executives raise $\bta$ the legislature becomes less likely to abrogate the agreement, for three reasons.
+First, the legislature prefers a higher domestic tariff \textbf{(Lemma~\ref{res:leg_a})}; second, the higher tariff discourages lobbying, reducing $B(\bta)$ indirectly \textbf{(Result~\ref{res:lobby})}; and finally, the lower lobbying effort directly reduces the legislature's preferred tariff further \textbf{(Lemma~\ref{res:bcomb})}.
 Thus, beyond promising a lower tariff from the trading partner, we can think of the trade agreement as a sort of political commitment device that can be used to reduce political pressure and therefore get the legislature to maintain a lower tariff than it otherwise would.
@@
 \end{lemma}
-I define the tariff the executives prefer in the absence of legislative constraints as $\bm{\tau^E}$ and show in the \hyperlink{int_soln}{Appendix} that it is never optimal in a symmetric equilibrium for the executives to choose $\bm{\tau^a} < \bm{\tau^E}$ if there is a positive probability that the legislatures will have an opportunity to break the trade agreement.
-This is the result of an envelope-like result: there are only second-order losses from raising tariffs slightly from the most-preferred level yet there are first-order gains in reducing the probability that the agreement will be broken.
-
-The executives will also always choose $\bta < \btn$ unless the legislature breaks even agreements with tariffs very close to the Nash level with certainty, in which case the problem is not interesting so I will ignore this case.
+I show in the \hyperlink{int_soln}{Appendix} that it is never optimal in a symmetric equilibrium for the executives to choose $\bm{\tau^a}$ below the executives' most preferred level if the legislatures will have an opportunity to break the trade agreement.
+This is because there are only second-order losses from raising tariffs slightly from the most-preferred level yet there are first-order gains in reducing the probability that the agreement will be broken.
+
+The executives will also always choose $\bta < \btn$ unless the legislature breaks even agreements with tariffs very close to the Nash level with certainty.
+In this case the problem is not interesting so I ignore this case.
 Thus there is an interior solution in all cases of interest, but this solution may be of two different forms.
 First, it could be at the point that maximizes the concave portion of the executives' welfare function (Equation~\ref{eq:jv2}).
-Here, the optimal tariffs are chosen to balance concerns for trade-agreement welfare against those for discouraging lobbying activity aimed at breaking the agreement.
-
-But there is an alternative.
-For some specifications, high enough tariffs can disengage the lobby sufficiently so that the probability that the trade agreement will be broken is reduced to zero.
-Interestingly, if this occurs at a low enough tariff level, executive welfare can be maximized at this point.
+
+For some specifications, however, high enough tariffs can disengage the lobby sufficiently so that the probability that the trade agreement will be broken is zero.
+If this occurs at a low enough tariff level, executive welfare can be maximized at this point.
 \begin{result}
@@
 \end{result}
-%Can I say when one would happen versus the other?
-This seems to accord well with observations of trade policy politics, in that some lobbies appear to exert significant effort toward disrupting trade agreements (often in the form of some kind of individual administrative remedy such as anti-dumping duties) whereas others apparently do not contest the levels of protection granted them.
+This accords well with observations of trade policy politics, in that some lobbies exert significant effort toward disrupting trade agreements whereas others apparently do not engage in the political process.
 The current model points to differences across industries in production and demand structure, as well as political weighting ($\ga$), to help explain these variations.

### 2016-06-16 draft.tex
revise most of example section

@@
 \label{sec:example}
-It is instructive to examine a simple parameterization of the model economy.
-The fundamentals here are chosen to match those of Bagwell and Staiger (2005).
-Home country demand, supply and profits are given by $D(P_i) = 1 - P_i$, $Q_X(P_X) = \frac{P_X}{2}$, $Q_Y(P_Y) = P_Y$, $\Pi_X(P_X) = \frac{(P_X)^2}{4}$, and $\Pi_Y(P_Y) = \frac{(P_Y)^2}{2}$ where $P_i$ is the price of good $i$ in the home country market.
-%Note that production function consistent with these supply functions are $F_X = \sqrt{l}$ and $F_Y = \sqrt{2l}$ where we assume $w=1$.
+Examining a simple parameterization of the model economy is instructive.
+The fundamentals are chosen to match Bagwell and Staiger (2005) with $D(P_i) = 1 - P_i$, $Q_X(P_X) = \frac{P_X}{2}$, $Q_Y(P_Y) = P_Y$, $\Pi_X(P_X) = \frac{(P_X)^2}{4}$, and $\Pi_Y(P_Y) = \frac{(P_Y)^2}{2}$.
+Home-country imports of $X$ and exports of $Y$ are $M_X(P_X)= 1 - \frac{3}{2}P_X$ and $E_Y(P_Y)= 2P_Y -1$.
+Market clearing implies that world and home prices of $X$ are $P_X^W = \frac{4-3\tau}{7}$ and $P_X = \frac{4+4\tau}{7}$.
 Foreign is taken to be symmetric.
-
-This implies Home-country imports of $X$ and exports of $Y$ of $M_X(P_X)= 1 - \frac{3}{2}P_X$ and $E_Y(P_Y)= 2P_Y -1$, with foreign imports of $Y$ and exports of $X$ given by $M_Y^*(P_Y^*)= 1 - \frac{3}{2}P_Y^*$ and $E_X(P_X^*)= 2P_X^* -1$.
-With the only trade policy instruments being tariffs on import competing goods, world prices are $P_X = P_X^W + \tau$, $P_X^* = P_X^W$, $P_Y^* = P_Y^W + \tau^*$, and $P_Y = P_Y^W$.
-Market clearing implies that world and home prices of $X$ are $P_X^W = \frac{4-3\tau}{7}$ and $P_X = \frac{4+4\tau}{7}$, symmetric for $Y$.
-\\
 
@@
 
-When setting the trade-war tariff, the legislature maximizes $W_{\mathit{ML}}(\tau, \tau^*) = W_{\mathit{ML}}^X(\tau) + W_{\mathit{ML}}^Y(\tau^*)$ by choice of $\tau$ given $\tau^*$.
-As there are no interactions between $\tau$ and $\tau^*$, the legislature maximizes $W_{\mathit{ML}}^X(\tau)$ only (in curly braces above) and sets the trade war tariff
+The Nash tariff that results from unilateral maximization is\footnote{The second order condition is satisfied for all realizations of $\ga < 17/2$.
+Because $\ga = 7/4$ is enough to achieve the prohibitive tariff of $1/6$, I will assume this condition is satisfied.}
 \[
@@
 \]
-I refer to this as the Nash tariff because it is the result of Nash equilibrium in the non-cooperative game between the legislatures.
-$\tau^n$ is increasing in $e$ and the second order condition is satisfied for all realizations of $\ga < 17/2$.
-Because $\ga = 7/4$ is enough to achieve the prohibitive tariff of $1/6$, I will assume this condition is satisfied.
-
-In the event of a trade war and facing this tariff-setting behavior by the legislature, the lobby maximizes $\pi\left(\tau^n\left(\ga\left(e_n,\ve_n\right)\right)\right) - e_n$.
+
 In order to predict the Nash tariff, the political weighting function and form of political uncertainty must be specified.
@@
 
-Next we move to the legislature's decision on whether or not to break the trade agreement.
 Figure~\ref{fig:br} depicts the probability that the legislature will vote to break the trade agreement as a function of the tariff levels set in the trade agreement (with the restriction that $\tau^a = \tau^{*a}$) and the lobby's effort.
@@
 \subsection{Trade Agreement}
-Because the probability of legislative break never reaches zero for this parameterization, the executives' welfare function is concave everywhere.
-Assuming that each legislature has the opportunity to vote to break the agreement with probability $\frac{1}{2}$ and that the executives are social-welfare maximizers (i.e.
-$\ga_E = \ga_E^* = 1$), they will set trade agreements tariffs of $\tau^a = \tau^{*a} = 0.078$, with lobbying expenditures at the trade maintence phase of $0.0007$ and a total break probability of $0.505$.
+The probability of legislative break never reaches zero for this parameterization so the executives' welfare function is concave everywhere.
+Assuming that each legislature has the opportunity to vote to break the agreement with probability $\frac{1}{2}$ and that the executives are social-welfare maximizers, they will set trade agreements tariffs of $\tau^a = \tau^{*a} = 0.078$, with lobbying expenditures at the trade maintenance phase of $0.0007$ and a total break probability of $0.505$.
 Total lobbying expenditure is $1.54\%$ of the expected profits of $0.98591$.
@@
 We can compare this against several different benchmarks.
-The most stark is the trade-war outcome itself, which can also be interpreted as the outcome that would prevail in the absence of executive involvement in trade policy \textit{and} the absence of any effort to make a trade agreement.
-The tariff in the executive-formed trade agreement is about 60$\%$ of the Nash tariff of $0.129$, and the expected level given the probability that the agreement will be broken is 80$\%$ of the Nash level.
+The most stark is the trade-war outcome, which is the outcome that would prevail in the absence of any trade agreement.
+The tariff in the executive-formed trade agreement is about 60$\%$ of the Nash tariff of $0.129$, and the expected tariff given the probability that the agreement will be broken is 80$\%$ of the Nash level.
 Lobbying expenditures in the agreement-maintenance phase are about 40$\%$ of the those in the Nash game, while expected expenditures in the executive-led trade-agreement scenario are 90$\%$ of what they would be if the legislatures set tariffs unilaterally.
-
 Although the welfare-maximizing governments here are not able to set and maintain tariffs at zero as they would like, they are able to achieve significant reductions in tariff levels through the use of the trade agreement.
@@
 Another interesting benchmark is the scenario in which no executives are involved in trade policy, but in which the politically-susceptible legislatures can make use of a trade agreement.
-As in Bagwell and Staiger (2005), I find when maximizing their joint welfare in a legislature-led trade agreement ($W_{\mathit{ML}}(\tau, \tau^*) = W_{\mathit{ML}}^X(\tau) + W_{\mathit{ML}}^Y(\tau^*)$), the cooperative tariff levels will be set at
+As in Bagwell and Staiger (2005), I find when maximizing their joint welfare in a legislature-led trade agreement, the cooperative tariff levels will be set at
 \[
@@
 The legislatures are able to use the agreement to internalize the terms of trade externality, making political influence more expensive for the lobbies.
-Indeed, I find that lobbying expenditures rise to $0.0027$---60$\%$ higher than in the Nash case and $2.64\%$ of the gross profits of $0.10204$---while the agreement tariffs are set at $0.118$, only slightly lower than the Nash tariffs of $0.129$ and still higher than even the expected tariffs from the trade agreement scenario in which the executives choose the tariffs.
+Indeed, lobbying expenditures rise to $0.0027$---60$\%$ higher than in the Nash case and $2.64\%$ of the gross profits of $0.10204$---while the agreement tariffs are set at $0.118$, only slightly lower than the Nash tariffs of $0.129$ and still higher than even the expected tariffs from the trade agreement scenario in which the executives choose the tariffs.
 Comparing net (expected profits) across the three scenarios, they are highest in the Nash game ($0.1024$), lowest under the trade agreement ($0.9789$) and intermediate when the legislatures make policy directly ($0.0993$).

### 2016-06-16 draft.tex
remove autocracy / democracy discussion from example section

@@
 
-If one interprets the addition of a second policy-making actor as consistent with a move from autocracy to democracy, this result helps to explain the extensive empirical evidence that democracies trade more with each other than autocracies (Laver and Shepsle 1991, Aidt and Gassebner 2010).
-In particular, Frye and Mansfield (2003) argue that the reduction in trade barriers becomes more likely as the number of veto points increases as countries move from autarky to democracy as exemplified by the transition of Communist regimes in the post-Cold-War era.
-Here, a more open trading stance emerged as groups with a liberal stance were integrated into the political process following the fall of the Berlin Wall.
-
-Crucially, Henisz and Mansfield (2006) demonstrate that the impact of adding veto players depends on the status quo policy.
-In the case of these Communist regimes, Cold War policy was near autarkic.
-The additional actors had an interest in liberalization.
-Henisz and Mansfield's careful empirical work shows that, among democracies, systems with more veto players as in a separation-of-powers structure will be more robust to retreats from liberal policies in the face of protectionist pressures than are those with fewer veto players.
-Conversely, it stands to reason that protectionist separation-of-power regimes would also be less susceptible to liberalization than unitary democractic regimes, \textit{ceteris paribus}.
-
 \section{Discussion}

### 2016-06-16 draft.tex
revise separation of powers discussion

@@
 \label{sec:sep_powers}
-Consider the model of Section~\ref{sec:main} for the case of $\ve_b=0$---that is, no uncertainty at the break stage (as will be discussed in Section~\ref{sec:uncertainty}, in the case of mean-zero uncertainty, strategic behavior at the trade-war stage is not altered by uncertainty).
-This allows the isolation of the results that derive from the assumption that power over trade policy is shared between the executive and legislative branches.
-Again, we focus on the home lobby.
-
-Now $\ga(e_b)$ is deterministic, so the lobby knows the precise contribution it must make for any given trade agreement tariffs $\bta$ to induce the legislature to break the agreement (see Inequality~\ref{eq:lwcg}).
+We can isolate the results that derive from the assumption that power over trade policy is shared between the executive and legislative branches if we assume that there is no uncertainty at the break stage.
+
+Now $\ga(e_b)$ is deterministic, so the lobby knows the precise contribution it must make for any given trade agreement tariffs $\bta$ to induce the legislature to break the agreement.
 Here, the lobby's contribution \textit{increases} in $\bta$ since the higher are the trade agreement tariffs, the larger is the political weight that is required to induce the legislature to find them unsatisfactorily low.
@@
 
-I return to the example of Section~\ref{sec:example} to illustrate the effects of changing the environment to one without political uncertainty.
-Recall that in the main example with $\ve$ distributed uniformly on $[-0.25,0.25]$, the expected Nash tariff is $0.129$ and the executive-led trade agreement tariff is $0.078$, with lobbying expenditures of $0.0007$ and a total break probability of $0.505$.
-The expected tariff is then $0.103$.
-
-In this example, if the executives set the trade agreement facing no uncertainty, the tariff level is $0.106$, the lobby exerts zero effort, and the agreement remain in force with probability 1.
+In the example of Section~\ref{sec:example}, if the executives set the trade agreement facing no uncertainty, the tariff level is $0.106$, the lobby exerts zero effort, and the agreement remain in force with probability 1.
 If instead the executives were to set trade-agreement tariffs of $0.105$, the lobby would contribute $0.0025$ and the legislature would break the agreement with probability 1.

### 2016-06-16 draft.tex
revise PFS section

@@
 \begin{center}
-\includegraphics[height=3in, width=4in]{brprob2.jpg}
+\includegraphics[height=2.5in, width=3in]{brprob2.jpg}
 \end{center}
@@
 \subsection{Trade Agreement}
-The probability of legislative break never reaches zero for this parameterization so the executives' welfare function is concave everywhere.
-Assuming that each legislature has the opportunity to vote to break the agreement with probability $\frac{1}{2}$ and that the executives are social-welfare maximizers, they will set trade agreements tariffs of $\tau^a = \tau^{*a} = 0.078$, with lobbying expenditures at the trade maintenance phase of $0.0007$ and a total break probability of $0.505$.
+The break probability never reaches zero for this parameterization so the executives' welfare function is concave everywhere.
+Assuming that each legislature has the opportunity to break the agreement with probability $\frac{1}{2}$ and that the executives are social-welfare maximizers, they set trade agreements tariffs of $\tau^a = \tau^{*a} = 0.078$, with lobbying expenditures at the trade maintenance phase of $0.0007$ and a total break probability of $0.505$.
 Total lobbying expenditure is $1.54\%$ of the expected profits of $0.98591$.
@@
 \label{sec:gh}
-The legislative welfare function employed in this paper permits flexibility to take account of a wide range of political processes within a non-unitary legislature; alternatively, it can be a model of a unitary policy-maker when special interests face non-constant returns to lobbying activity.
-As noted in Section 2, the non-linear relationship between lobbying effort and tariffs follows Dixit, Grossman and Helpman (1997) and is a departure from the `Protection for Sale' welfare function, which assumes that a unitary government maximizes the sum of contributions and some fraction, $a$, of social welfare.
-
-To see the relationship between the two forms, write the PFS welfare function (replacing $C$ with $e$) as
-\[
-  e + aW = e + a \left[ \mathit{CS}_X(\tau) + \mathit{PS}_X(\tau) + \mathit{CS}_Y(\tau^*) + \mathit{PS}_Y(\tau^*) + \mathit{TR}(\tau) \right]
-\]
-The relationship between $e$ and the weight the legislature places on profits is endogenous; because the PFS government welfare function has such an elegant and completely specified form, it is possible to solve for equilibrium behavior to retrieve this relationship directly.
-
-It is easy to describe the equilibrium behavior and thus how lobbying effort will be related to the weight on profits in equilibrium in the PFS model when there is one lobby who makes a take-it-or-leave-it offer.
-Take the case of no uncertainty and assume that the legislature chooses $\tau=0$ in the absence of lobbying and that $\ga(0)=1$.
-For any $\tau$ that the lobby desires, it must pay according to the  indifference condition $e(\tau) = a \left[ W(0) - W(\tau) \right]$.
-That is, the lobby must pay for the $a$-weighted welfare loss caused by the tariff it requests.
-
-In the model of this paper, this indifference condition, rearranged to solve for $\ga(e)$, is
-		    \[
-		      \ga(e) = \frac{\mathit{CS}_X(0) + \pi_X(0) + \mathit{CS}_Y(0) + \pi_Y(0) + \mathit{TR}(0) - \mathit{CS}_X(\tau) - \mathit{CS}_Y(\tau^*) - \pi_Y(\tau^*) - \mathit{TR}(\tau)}{\pi_X(\tau)}
-		    \] 
-In order to match the form of the PFS framework, we must therefore have
-		    \begin{equation}
-		      \ga(e) = 1 + \frac{e}{a\pi_X(\tau)}
+I have introduced a legislative welfare function that admits a wide range of political processes within a non-unitary legislature as well as unitary policy-making when special interests face non-constant returns to lobbying activity.
+The non-linear relationship between lobbying effort and tariffs follows Dixit, Grossman and Helpman (1997) and is a departure from the `Protection for Sale' welfare function, which assumes that a unitary government maximizes the sum of contributions and some fraction, $a$, of social welfare.
+
+I show in the \hyperlink{psf}{Appendix} that choosing the following political weighting function aligns the political objective function used in the current model with the PFS framework:
+		    \begin{equation}
+		      \ga(e) = 1 + \frac{e}{a\pi_X(\tau)} .
 		      \label{eq:gh}
 		    \end{equation}
-
-Choosing the political weighting function identified in Equation~\ref{eq:gh} aligns the political objective function used in the current model with the PFS framework.
-Although using the $\ga$-weighted government welfare function is a reduced-form approach in the sense that legislative dynamics are not fully modeled, a drawback of the PFS welfare function is that it embodies the restriction that returns to lobbying effort must be constant.
-
-As noted in Dixit, Grossman and Helpman (1997), the linear form also implies transferable utility between the lobby and unitary government; this leads to equilibrium indeterminacies and an inability to speak to distributional questions, prompting them to generalize the PFS form.
-From their results, we have that the same equilibrium analysis holds in the more general environment.
-
-For the trade policy literature, generalizing the linear form in a sense would obviate many of the questions that have been raised by empirical investigations of the PFS model about the relationship of the magnitude of welfare losses relative to the weight placed on social welfare by governments as the $a$ parameter for the constant relative marginal utility provided by social welfare would no longer be present in the model.
-
-The linear PFS form provides for an almost unmatched simplicity and elegance that is able to capture the key cross-industry predictions on import tariffs.
+Although using the $\ga$-weighted government welfare function is a reduced-form approach in the sense that legislative dynamics are not fully modeled, a benefit is that it does not require returns to lobbying effort to be constant.
+The more flexible form of the weighting function also removes the equilibrium indeterminacies and inability to speak to distributional questions that are implied by the linear form and its transferable utility (Dixit, Grossman and Helpman 1997).
+
+The linear PFS form provides for an almost unmatched simplicity and elegance that is able to capture key cross-industry predictions on import tariffs.
 A perhaps unintended byproduct in the subsequent literature seems to have arisen in the form of a focus on matching fine details of governments' welfare-mindedness using a model that has a very simple unitary government and is thus better suited for other purposes.
@@
 
+\vskip.4in
+\noindent \textbf{\hypertarget{pfs}{Connection to Grossman and Helpman's `Protection for Sale'}} \\
+To see the relationship between the two forms, write the PFS welfare function (replacing $C$ with $e$) as
+\[
+  e + aW = e + a \left[ \mathit{CS}_X(\tau) + \mathit{PS}_X(\tau) + \mathit{CS}_Y(\tau^*) + \mathit{PS}_Y(\tau^*) + \mathit{TR}(\tau) \right]
+\]
+The relationship between $e$ and the weight the legislature places on profits is endogenous; because the PFS government welfare function has such an elegant and completely specified form, it is possible to solve for equilibrium behavior to retrieve this relationship directly.
+
+It is easy to describe the equilibrium behavior and thus how lobbying effort will be related to the weight on profits in equilibrium in the PFS model when there is one lobby who makes a take-it-or-leave-it offer.
+Take the case of no uncertainty and assume that the legislature chooses $\tau=0$ in the absence of lobbying and that $\ga(0)=1$.
+For any $\tau$ that the lobby desires, it must pay according to the  indifference condition $e(\tau) = a \left[ W(0) - W(\tau) \right]$.
+That is, the lobby must pay for the $a$-weighted welfare loss caused by the tariff it requests.
+
+In the model of this paper, this indifference condition, rearranged to solve for $\ga(e)$, is
+		    \[
+		      \ga(e) = \frac{\mathit{CS}_X(0) + \pi_X(0) + \mathit{CS}_Y(0) + \pi_Y(0) + \mathit{TR}(0) - \mathit{CS}_X(\tau) - \mathit{CS}_Y(\tau^*) - \pi_Y(\tau^*) - \mathit{TR}(\tau)}{\pi_X(\tau)}
+		    \]
+
 %\section{Acknowledgements}

### 2016-06-16 draft.tex
reverse uncertainty and pfs sections

@@
 
-\subsection{Relation to Grossman and Helpman's `Protection for Sale'}
-\label{sec:gh}
-I have introduced a legislative welfare function that admits a wide range of political processes within a non-unitary legislature as well as unitary policy-making when special interests face non-constant returns to lobbying activity.
-The non-linear relationship between lobbying effort and tariffs follows Dixit, Grossman and Helpman (1997) and is a departure from the `Protection for Sale' welfare function, which assumes that a unitary government maximizes the sum of contributions and some fraction, $a$, of social welfare.
-
-I show in the \hyperlink{psf}{Appendix} that choosing the following political weighting function aligns the political objective function used in the current model with the PFS framework:
-		    \begin{equation}
-		      \ga(e) = 1 + \frac{e}{a\pi_X(\tau)} .
-		      \label{eq:gh}
-		    \end{equation}
-Although using the $\ga$-weighted government welfare function is a reduced-form approach in the sense that legislative dynamics are not fully modeled, a benefit is that it does not require returns to lobbying effort to be constant.
-The more flexible form of the weighting function also removes the equilibrium indeterminacies and inability to speak to distributional questions that are implied by the linear form and its transferable utility (Dixit, Grossman and Helpman 1997).
-
-The linear PFS form provides for an almost unmatched simplicity and elegance that is able to capture key cross-industry predictions on import tariffs.
-A perhaps unintended byproduct in the subsequent literature seems to have arisen in the form of a focus on matching fine details of governments' welfare-mindedness using a model that has a very simple unitary government and is thus better suited for other purposes.
-
-It is clear that trade policy is often shaped in a complicated process involving multiple actors (see, for instance, Destler (2005) for the U.S.).
-The model presented herein is a first attempt to more richly model the policy-making process and thus shed light on open questions surrounding how government preferences and the details of political institutions impact trade policy outcomes.
-
 \subsection{The Role of Political Uncertainty}
 \label{sec:uncertainty}
-In this model, if there were no political uncertainty, there would be no lobbying in equilibrium.
-Thus, if nothing else, adding the realism of political uncertainty allows for positive lobbying on the equilibrium path.
-If we examine more closely the assumption that there is significant uncertainty surrounding the legislative lobbying process, additional insights come to light.
-It is most clear in the case of mean-zero uncertainty, so I will assume this throughout this section unless otherwise noted.
+Adding political uncertainty allows for positive lobbying on the equilibrium path, but this is not the only additional insight.
+These are most clear in the case of mean-zero uncertainty, so I will assume this throughout this section unless otherwise noted.
 
@@
 \begin{center}
-\includegraphics[height=3in, width=4in]{lobby_br.jpg}
+\includegraphics[height=2.5in, width=3in]{lobby_br.jpg}
 \end{center}
@@
 The provided tariff levels are strongly influenced by lobbying incentives, but not in the straightforward ways predicted by unitary models.
+
+\subsection{Relation to Grossman and Helpman's `Protection for Sale'}
+\label{sec:gh}
+I have introduced a legislative welfare function that admits a wide range of political processes within a non-unitary legislature as well as unitary policy-making when special interests face non-constant returns to lobbying activity.
+The non-linear relationship between lobbying effort and tariffs follows Dixit, Grossman and Helpman (1997) and is a departure from the `Protection for Sale' welfare function, which assumes that a unitary government maximizes the sum of contributions and some fraction, $a$, of social welfare.
+
+I show in the \hyperlink{pfs}{Appendix} that choosing the following political weighting function aligns the political objective function used in the current model with the PFS framework:
+		    \begin{equation}
+		      \ga(e) = 1 + \frac{e}{a\pi_X(\tau)} .
+		      \label{eq:gh}
+		    \end{equation}
+Although using the $\ga$-weighted government welfare function is a reduced-form approach in the sense that legislative dynamics are not fully modeled, a benefit is that it does not require returns to lobbying effort to be constant.
+The more flexible form of the weighting function also removes the equilibrium indeterminacies and inability to speak to distributional questions that are implied by the linear form and its transferable utility (Dixit, Grossman and Helpman 1997).
+
+The linear PFS form provides for an almost unmatched simplicity and elegance that is able to capture key cross-industry predictions on import tariffs.
+A perhaps unintended byproduct in the subsequent literature seems to have arisen in the form of a focus on matching fine details of governments' welfare-mindedness using a model that has a very simple unitary government and is thus better suited for other purposes.
+
+It is clear that trade policy is often shaped in a complicated process involving multiple actors (see, for instance, Destler (2005) for the U.S.).
+The model presented herein is a first attempt to more richly model the policy-making process and thus shed light on open questions surrounding how government preferences and the details of political institutions impact trade policy outcomes.
 

### 2016-06-16 draft.tex
revise uncertainty section

@@
 Adding political uncertainty allows for positive lobbying on the equilibrium path, but this is not the only additional insight.
-These are most clear in the case of mean-zero uncertainty, so I will assume this throughout this section unless otherwise noted.
-
-In the event of a trade war, the lobbies' behavior, and therefore the expected trade-war tariffs, are not altered by mean-zero uncertainty (see Equation~\ref{eq:lobtw}).
-
-At the earlier stage of lobbying, uncertainty alters optimizing behavior by both the lobby and the executives.
-In contrast to the previous section with no uncertainty, consider adding a very small amount of uncertainty to the previous example: take $\ve$ distributed uniformly on $[-0.01,0.01]$.
-Here, the executives find it optimal to set the tariffs to completely disengage the lobby and ensure the agreement remains in force.
-However, the tariff level will be different because instead of making the decision to lobby according to the certainty condition
-\[\pi(\tau^n) - e_n - \pi(\tau^a) > e_b,
-\]
-the lobby makes this decision according to Condition~\ref{ine:lobint}, which can be re-written as
-\[
-\pi(\tau^n) - e_n - \pi(\tau^a) > \left[\frac{\partial b(0,\bta,\btn(e_n,e_n^*))}{\partial e_b}\right]^{-1}.
-\]
-
-Although the relationship between the tariff under certainty and a small amount of uncertainty will depend on the structure of the economy and the legislative process, in this case, the tariff can be reduced slightly to $0.105$.
-As we have seen above, with significant uncertainty, the tariff in this example is reduced to $0.078$.
-Interestingly, the optimal tariff does not decrease monotonically between these two values as uncertainty increases.
+For simplicity, I assume uncertainty is mean-zero.
+In particular, lobbying behavior in a trade war is not altered by mean-zero uncertainty and so neither are expected trade-war tariffs.
+
+At earlier stages, uncertainty alters optimizing behavior by both the lobby and the executives.
+Consider adding a very small amount of uncertainty to the example of the previous section: take $\ve$ distributed uniformly on $[-0.01,0.01]$.
+Here, the executives find it optimal to set the tariffs to completely disengage the lobby, ensuring the agreement remains in force.
+However, the tariff level will be different than when $\ta_b=0$ because instead of making the lobbying decision according to the certainty condition $\pi(\tau^n) - e_n - \pi(\tau^a) > e_b$, the lobby makes this decision according to Condition~\ref{ine:lobint}.
+Relative to the example with $\ta_b=0$, the tariff can be reduced slightly to $0.105$.
+Interestingly, the optimal tariff does not decrease monotonically.
 
@@
 
-It is particularly instructive to examine the lobby's reaction function at this very low level of uncertainty as well as the intermediate level when $\ve$ is distributed uniformly on $[-0.14,0.14]$ because it is at the latter that the executives set the highest tariff level of $0.107$.
-For the former, when facing very little uncertainty, the range of tariff levels to which the lobby responds with an intermediate contribution that leaves open the possibility of a break is very small: only those between 0.103 and 0.104.
-The lobby disengages completely at $\bta=0.105$ (see Figure~\ref{fig:lobby_br}).
-
+Figure~\ref{fig:lobby_br} shows the lobby's reaction function at this very low level of uncertainty as well as the intermediate level when $\ve$ is distributed uniformly on $[-0.14,0.14]$ where the executives set the highest tariff level of $0.107$.
+For the former, the range of tariff levels to which the lobby responds with an intermediate contribution that leaves open the possibility of a break is very small: only those between 0.103 and 0.104.
+The lobby disengages completely at $\bta=0.105$.
 In contrast, when uncertainty is increased to the interval $[-0.14,0.14]$, the lobby begins to leave open the possibility of a break at $\bta = 0.048$.
 It disengages at $0.107$ and this is the tariff level that the executives find optimal.
-Interestingly, at intermediate uncertainly levels, both portions of the best response curve are steeper and the lobby disengages at slightly lower tariff levels---as low as $0.099$ when $\ve \sim U[-0.06,0.06]$.
-
-The preceding discussion illustrates Part (a) of Result~\ref{res:execsoln}: here the executives maximize their welfare by raising tariffs to the point where the import-competing industry ceases to lobby and thus the agreement remains in force for sure.
-Note, however, that although lobbying expenditures are zero, the potential for lobbying behavior is essential in shaping the trade agreement.
+At intermediate uncertainly levels, both portions of the best response curve are steeper and the lobby disengages at slightly lower tariff levels---as low as $0.099$ when $\ve \sim U[-0.06,0.06]$.
+
+This illustrates Part (a) of Result~\ref{res:execsoln}: here the executives maximize their welfare by raising tariffs to the point where the import-competing industry ceases to lobby and the agreement remains in force for sure.
+Although lobbying expenditures are zero, the potential for lobbying behavior is essential in shaping the trade agreement.
 
@@
 
-When uncertainty rises above this threshold of $\ve \sim U[-0.14,0.14]$, the executives' choices are made according to Part (b) of Result~\ref{res:execsoln}: that is, they trade off reductions in welfare under the agreement with reductions in the probability that the agreement will be abrogated by the legislature.
+When uncertainty rises above this threshold of $\ve \sim U[-0.14,0.14]$, the executives' choices are made according to Part (b) of Result~\ref{res:execsoln}: they trade off reductions in welfare under the agreement with reductions in the probability that the agreement will be abrogated by the legislature.
 %In terms of the lobby's best response function, the executives now choose the optimal point on the downward-sloping portion of the curve instead of the point where it reaches zero.
@@
 
-As uncertainty increases further, the general pattern is that tariffs, and therefore lobbying effort, are reduced, while the break probability increases.
-However, this pattern is not smooth or absolute, in contrast to the welfare level achieved by the executives, which increases monotonically in the amount of uncertainty present.
-%It is not clear how general a result this is, but it is particularly interesting in contrast to the non-monotone pattern over the lower-range of uncertainty where executive welfare is minimized at the edges of the range that lead to disengagement of the lobby and maximized in the middle of that range.
-
-Several salient points emerge from this example.
-First, different amounts of uncertainty lead to very different outcomes---both in terms of the tariffs that are set in the trade agreement and the probability of disruption.
-Second, these outcomes vary in nuanced ways with the underlying behavior of the lobbies and the executives.
-The provided tariff levels are strongly influenced by lobbying incentives, but not in the straightforward ways predicted by unitary models.
+We see from this example that variations in uncertainty lead to very different outcomes in terms of lobbying, trade agreement tariffs and the probability of disruption.
+The provided tariff levels are strongly influenced by lobbying incentives, but in more nuanced ways than those predicted by unitary models.
 

### 2016-06-16 draft.tex
revise conclusion

@@
 \label{sec:concl}
-I have shown that the legislature both breaks trade agreements with a higher probability and sets higher trade war tariffs when lobbying activity increases, while the probability with which it breaks agreements (or fails to ratify them) decreases in the domestic trade agreement tariff.
+I have shown that the legislature both breaks trade agreements with a higher probability and sets higher trade war tariffs when lobbying activity increases, while the probability with which it breaks agreements decreases in the domestic trade agreement tariff.
 Because the lobby decreases its effort in response to higher trade agreement tariffs, the executives face a trade-off between the welfare derived while a trade agreement is in force and the probability with which the agreement actually enters into force.
@@
 I have also shown that in a policy-making structure in which power is shared, a less politically-motivated executive can utilize an international trade agreement to reduce the political pressure on the ratifying body and therefore increase the probability that the agreement will remain in force.
-Thus, in a model with a richer description of government structure, a political-commitment role for trade agreements can arise.
+Thus, in a model with a richer description of government structure, a broad political-commitment role for trade agreements can arise.
 

### 2016-06-16 draft.tex
more cuts from intro and lit review

@@
 To see how this works, let's begin by considering the case of political certainty.
-I assume the executive to be more pro-social than the legislature, as has been the case throughout the post-war period in the United States.
+I assume the executive is more pro-social than the legislature, as has been the case throughout the post-war period in the United States.
 Lobbies, assumed for simplicity to represent import-competing interests only, would like to prevent or derail a trade agreement that would reduce their tariff protection.
@@
 Thus, any trade agreement under political certainty involves a tariff high enough to disengage the lobby completely; that is, even welfare-maximizing executives set high tariffs and induce zero contributions in equilibrium because they set policy in the shadow of a ratification decision influenced by lobbying.
-Careful modeling of the political process demonstrates that the relationship between deadweight loss calculations and the tariffs that are provided in a trade agreement are mediated in important ways by a ratification constraint.
 
 The fact that zero contributions are predicted in equilibrium is not particularly satisfying.
-The model becomes more realistic and delivers additional intuition when I allow all actors to be uncertain about the weight the median legislator places on the profits of the lobbying industry.
-\textbf{Similar to probabilistic voting models that incorporate policy motivation as in Roemer (1997),} this represents the idea that the lobbying and voting ``game'' that goes on within legislatures is often complex enough that none of the actors know precisely what the outcome will be---that is, no one can predict exactly which legislator will be the median and therefore what weight will be placed on import-industry profits at the time of the vote.
-
+The model becomes more realistic, has smoother results and delivers additional intuition when I allow all actors to be uncertain about the weight the median legislator places on the profits of the lobbying industry.
+\textbf{Similar to probabilistic voting models that incorporate policy motivation as in Roemer (1997),} this represents the idea that the lobbying and voting ``game'' that goes on within legislatures is often so complex that none of the actors know precisely what the outcome will be---that is, no one can predict exactly which legislator will be the median and therefore what weight will be placed on import-industry profits at the time of the vote.
 \textbf{One can conceptualize this uncertainty as being a result of the lobby's strategic choice to engage with only some key members of the legislature.
 For instance, the lobbied legislators may promise to deliver the votes of their non-lobbied colleagues but may not be able to do so with certainty because those colleagues are subject to cross-pressuring from various sources.}
-
-In the face of this political uncertainty, lobbies respond to higher trade agreement tariffs by decreasing effort and in turn the legislatures disrupt trade agreements with a lower probability.
-Since they prefer low tariffs, the executives must trade off the level of social welfare derived while a trade agreement is in force with the chances that the agreement will remain in force.
 
@@
 
-This is not a time inconsistency \`{a} la Maggi and Rodr\'{i}guez-Clare (2007) where the welfare of the unitary policymaker is improved via commitment to a trade agreement by changing firms' investment decisions.
-Instead, it is an inconsistency between the preferences of the executive and legislative branches.
-Although the legislatures have the final word, the executives use the trade agreement to reduce lobbying effort to mute the legislature's protectionist bias and bring the outcome more in line with their relatively liberal preferences.
+This is not a time inconsistency \`{a} la Maggi and Rodr\'{i}guez-Clare (2007) where the welfare of the unitary policymaker is improved by changing firms' investment decisions via commitment to a trade agreement.
+Instead, this is an inconsistency between the preferences of the executive and legislative branches.
+Although the legislatures have the final word, the executives use the trade agreement to reduce lobbying effort, muting the legislature's protectionist bias and bringing the outcome more in line with their relatively liberal preferences.
 
@@
 Lobbying incentives are crucial but their impact is not necessarily as straightforward as predicted by models with unitary governments.
-Gawande, Krishna and Olarreaga (2009) provide cross-country empirical evidence that measures such as the number of checks and balances on the power of the legislature have predictive power for the PFS ``welfare-mindedness'' parameter.
-Thus, although empirical tests are outside the scope of this paper, there is suggestive evidence that political uncertainty is indeed an important factor empirically.
+Gawande, Krishna and Olarreaga (2009) provide cross-country evidence that measures such as the number of checks and balances on the power of the legislature have predictive power for the PFS ``welfare-mindedness'' parameter.
+Thus, although empirical tests are outside the scope of this paper, there is suggestive evidence that political uncertainty is indeed empirically important.
 
@@
 Gawande, Krishna and Robbins (2006); Mitra, Thomakos and Ulubasoglu (2006); Bombardini (2008); and Gawande, Krishna and Olarreaga (2012) among others.
-Gawande and Magee (2012) provide a summary of the extensive literature, while Imai, Katayama and Krishna (2009) argue that many of the classification schemes present serious challenges for the validity of results.} Taking a different approach, Gawande and Hoekman (2006) suggest that one can reconcile the empirical results with the model by acknowledging the complexity of the policy-making process and the uncertainty that arises from it.
-
-In the non-unitary government model, these estimates and the stylized facts are not inconsistent.
-The ``missing contributions,'' (Gawande and Hoekman 2006), can be explained as an equilibrium phenomenon arising from the agenda-setting power of the executive branch.
+Gawande and Magee (2012) provide a summary of the extensive literature, while Imai, Katayama and Krishna (2009) argue that many of the classification schemes present serious challenges for the validity of results.} As suggested by Gawande and Hoekman (2006), in the non-unitary government model, these estimates and the stylized facts are not inconsistent.
+The ``missing contributions,'' can be explained as an equilibrium phenomenon arising from the agenda-setting power of the executive branch.
 In particular, in this model, there is no inconsistency for the 15$\%$ or so of sectors that receive protection but apparently put forth no lobbying effort.
@@
 \subsection{Related Literature}
-The foundations of this work rest on Grossman and Helpman's (1994) `Protection for Sale.' Two related papers `Trade Wars and Trade Talks' (Grossman and Helpman (1995b)) and `The Politics of Free-Trade Agreements' (Grossman and Helpman (1995a)) first employed the PFS approach in the study of trade agreements.
-The first considers ``Trade Wars'' and ``Trade Talks'' separately, whereas my approach allows the ``Trade Talks'' to be influenced by the desire to avoid a ``Trade War'' and thus views the ``Trade War'' as a crucial subgame.
-Their model of preferential trade agreements (Grossman and Helpman (1995a)), although treating the government as a unitary actor, is closer to this approach.
+This work rests on the foundation of Grossman and Helpman's (1994) `Protection for Sale.' Two related papers `Trade Wars and Trade Talks' (Grossman and Helpman (1995b)) and `The Politics of Free-Trade Agreements' (Grossman and Helpman (1995a)) first employed the PFS approach in the study of trade agreements.
+The first considers ``Trade Wars'' and ``Trade Talks'' separately, whereas my approach views the ``Trade War'' as a crucial subgame.
+The model of preferential trade agreements (Grossman and Helpman (1995a)), although treating the government as a unitary actor, is closer to this approach.
 
@@
 These studies (e.g.
-Feenstra and Lewis (1991); Milner and Rosendorff (2001); Bagwell and Staiger (2005); Beshkar (2010); Horn, Maggi and Staiger (2010); Beshkar and Bond (2012); Amador and Bagwell (2013)) derive various implications for the design of international trade agreements using a Baldwin (1987)-style government welfare function with exogenous shocks to the political-pressure parameter.
-
-\textbf{Since trade policies are overwhelmingly determined within the context of trade agreements, it is useful to have a framework to bring together the endogenous political pressure of PFS-style modeling with the trade agreements approach.
+Feenstra and Lewis (1991); Milner and Rosendorff (2001); Bagwell and Staiger (2005); Beshkar (2010); Horn, Maggi and Staiger (2010); Beshkar and Bond (2012)) derive various implications for the design of international trade agreements using a Baldwin (1987)-style government welfare function with exogenous shocks to the political-pressure parameter.
+
+\textbf{Since trade policies are overwhelmingly determined in the context of trade agreements, it is useful to have a framework to bring together the endogenous political pressure of PFS-style modeling with the trade agreements approach.
 The government objective function introduced in this paper is intended to be a bridge between the two.
-The objective function I introduce starts from the Baldwin (1987) government objective function and adds flexibility: the weight the legislature places on profits is endogenous (PFS), potentially non-linear (Findlay and Wellisz 1982; Dixit, Grossman and Helpman 1997) and stochastic.
+This objective function starts from the Baldwin (1987) government objective function and adds flexibility: the weight the legislature places on profits is endogenous (PFS), potentially non-linear (Findlay and Wellisz 1982; Dixit, Grossman and Helpman 1997) and stochastic.
 Buzard (2016) is an example of how this modeling approach can be used to integrate  endogenous political activity and exogenous shocks into the analysis of questions regarding optimal trade agreement design.}
@@
 
-\item Amador, M., and Bagwell, K., 2013. The Theory of Optimal Delegation with an Application to Tariff Caps. Econometrica 81, 1541-1599.
+%\item Amador, M., and Bagwell, K., 2013. The Theory of Optimal Delegation with an Application to Tariff Caps. Econometrica 81, 1541-1599.
 

### 2016-06-16 draft.tex
misc cuts

@@
 
-Several papers have studied the impact of executive/legislative interactions on international agreements in the case of exogenously-determined preferences.
-Mansfield, Milner and Rosendorff (2000) construct a model in which democracies trade more with each other because the domestic legislature's influence requires the foreign government to make deeper tariff cuts to ensure its cooperation.
-Dai (2006) argues that a country's legislative constraint assists its own executive in maintaining higher barriers.
+Mansfield, Milner and Rosendorff (2000) study the impact of executive/legislative interactions on international agreements in the case of exogenously-determined preferences.
+They show that democracies trade more with each other because the domestic legislature only cooperates if the foreign government makes deep tariff cuts.
 Ethier (2002) treats the effect of a separation of powers between trade negotiators and the government officials who administer protection.
 
-The above work is patterned on Milner and Rosendorff (1997), who explore how uncertainty can affect trade policy and ratification failure when political preferences, and therefore uncertainty, are exogenous.
-Le Breton and Salanie (2003) study lobbying when the lobby is uncertain about the preferences of a unitary decision maker.
+Milner and Rosendorff (1997) explore how uncertainty can affect trade policy and ratification failure when political preferences, and therefore uncertainty, are exogenous.
+In Le Breton and Salanie (2003), the lobby is uncertain about the preferences of a unitary decision maker.
 Le Breton and Zaporozhets (2007) replace the unitary decision maker with a legislature with multiple actors.
-In Song (2008), policy-making power is shared between an executive and a unitary legislature where lobbies are able to endogenously influence the political preferences of the legislature in a context of unilateral policy making with no uncertainty, and Coates and Ludema (2001) study trade policy leadership in a model with endogenous lobbying in the presence of political uncertainty with imperfect monitoring.
+In Song (2008), policy-making power is shared between an executive and a unitary legislature.
+The lobbies influence the preferences of the legislature in a context of unilateral policy making with no uncertainty, and Coates and Ludema (2001) study trade policy leadership with endogenous lobbying in the presence of political uncertainty with imperfect monitoring.
 
@@
 The supply functions for good $X$ are $Q_X(P_X)$ and $Q_X^*(P_X^*)$ and are assumed strictly increasing and twice continuously differentiable for all prices that elicit positive supply.
-I also assume $Q_X^*(P_X) > Q_X(P_X)$ for any such $P_X$ so that the home country is a net importer of good $X$.
+For any such $P_X$, I assume $Q_X^*(P_X) > Q_X(P_X)$ so that the home country is a net importer of good $X$.
 The production structure for good $Y$ is symmetric, with demand and supply such that the economy is separable in goods $X$ and $Y$.
-It is assumed that the production of goods $X$ and $Y$ requires labor and a sector-specific factor that is available in inelastic supply and is non-tradable so that the income of owners of the specific factors is tied to the price of the good in whose production their factor is used.
+The production of goods $X$ and $Y$ requires labor and a sector-specific factor that is available in inelastic supply and is non-tradable so that the income of owners of the specific factors is tied to the price of the good in whose production their factor is used.
 
@@
 
-Each country's government is composed of an executive who concludes trade agreements and a legislature that has final say on trade policy.
+In each country, there is an executive who concludes trade agreements and a legislature that has final say on trade policy.
 To focus attention on protectionist political forces\textbf{, highlight the model's central mechanisms and minimize the number of actors that must be modeled}, I assume that only the import-competing industry in each country is politically-organized and that it is represented by a single lobbying organization.\footnote{The model extends easily to the case of multiple lobbies.
@@
 
-Figure~\ref{fig:ext} illustrates the timing of the game from the perspective of the home country.\footnote{This extensive form is for the case where the foreign legislature does not break the trade agreement.} First, the executives set trade policy cooperatively within the context of an international agreement by choosing the trade agreement tariffs $\bta = \left(\tau^a,\tau^{*a}\right)$.\footnote{I use the convention throughout of representing a vector of tariffs for both countries $(\tau,\tau^*)$ as a single bold $\bm{\tau}$.} Then the lobbies attempt to persuade the legislators in their respective countries to break the trade agreement by choice of lobbying effort $e_b$, which can take on any non-negative value.
-
+Figure~\ref{fig:ext} illustrates the timing of the game from the perspective of the home country.\footnote{This extensive form is for the case where the foreign legislature does not break the trade agreement.} First, the executives set trade policy cooperatively within the context of an international agreement by choosing the trade agreement tariffs $\bta = \left(\tau^a,\tau^{*a}\right)$.\footnote{I use the convention throughout of representing a vector of tariffs for both countries $(\tau,\tau^*)$ as a single bold $\bm{\tau}$.} Then the lobbies attempt to persuade the legislators in their respective countries to break the trade agreement by choice of lobbying effort $e_b$.
 Next, nature determines with probabilities $o$ and $o^*$ whether each legislature will have the opportunity to take a vote to break the agreement.
-I assume that these probabilities are less than one-half and mutually exclusive.\footnote{While this assumption is made for purposes of exposition and tractability, one can think of $o$ and $o^*$ as being determined by events beyond the executives' or legislatures' control that determine whether or not the legislatures will be willing to consider the lobby's request; for example, they would be affected by the occurrence of an economic crisis that diverts the legislature's attention from less-pressing matters.} This allows me to focus on the key interaction between the domestic actors and abstract from the strategic interaction between the lobbies in the two countries.\footnote{When the legislatures in both countries have an opportunity to break the agreement simultaneously, an incentive for each lobby to free ride on the effort of the other makes the optimization problem of the executives more complex as they are faced with the problem of how best to exploit the free-riding dynamic.
-\textbf{Alternatively, one can take the break decisions in the two countries to be sequential with the executives behind a veil of ignorance about which legislature will move first.
-In that set up, the end results are the same as in the current model but additional intermediate results are needed.}}
-
-After this, the uncertainty about the the median legislator's identity is resolved.
-This is modeled as the realization of a random variable $\ta_b$, the value of which all players observe simultaneously.
+I assume that these probabilities are less than one-half and mutually exclusive.\footnote{While this assumption is made for purposes of exposition and tractability, one can think of $o$ and $o^*$ as being determined by events beyond the executives' or legislatures' control that determine whether or not the legislatures will be willing to consider the lobby's request; for example, they would be affected by the occurrence of an economic crisis that diverts the legislature's attention from less-pressing matters.} This allows me to focus on the key interaction between the domestic actors and abstract from the strategic interaction between the lobbies.\footnote{\textbf{If the break decisions in the two countries are sequential with the executives behind a veil of ignorance about which legislature moves first, the end results are the same but additional intermediate results are needed.} The optimization problem of the executives becomes more complex as they decide how best to exploit the second lobby's incentive to free ride on the effort of the first-mover lobby.}
+
+After this, uncertainty about the the median legislator's identity is resolved.
+All players simultaneously observe the realization of the random variable $\ta_b$ that represents this uncertainty.
 The ``break stage'' concludes with the legislature making a choice to abide by the agreement or to break the agreement.
-
-In the event that the trade agreement does not enter into or remain in force, there is a final stage of lobbying, resolution of the uncertainty surrounding this decision, and voting to set the trade-war tariffs $\btn$.\footnote{Alternatively, one can model the lobbying for the ratification decision as determining the potential non-cooperative tariff; this changes behavior slightly not the fundamental dynamics.
-Another possibility is to replace the trade-war phase with the ex-ante status quo, which would most naturally be the non-cooperative equilibrium.
-This is precisely the outcome that obtains in the trade-war phase, so the results of the model are unaltered.} Once all political decisions are taken, producers and consumers make their decisions.
+In the event that the legislature breaks the agreement, there is a final stage of lobbying, resolution of the uncertainty surrounding this decision, and voting to set the trade-war tariffs $\btn$.\footnote{Alternatively, one can model the non-cooperative tariff as being determined by the break decision lobbying or the ex-ante status quo.
+This does not change the fundamental dynamics of the model.} Finally, producers and consumers make their decisions.
 
@@
 \label{sec:pref}
-As the economy is fully separable and the economic and political structures are symmetric, I focus throughout on the home country and the $X$-sector.
-The details are analogous for $Y$ and foreign.\footnote{These are the relevant sectors for analysis because they are the sectors in which there are lobbies.} I specify the welfare function of the home legislature as
+With the structure of politics and the economy symmetric and the latter fully separable, I focus on the home country's and the $X$-sector.
+The details are analogous for $Y$ and foreign.
+The home legislature's welfare function is
 \begin{equation}
@@
 
-Assumption~\ref{as:ga_c} formalizes the intuition that the legislature favors the import-competing industry more the higher is its lobbying effort, but that there are diminishing returns to lobbying activity.\footnote{In Ethier (2012), the returns to lobbying decline as the level of protection increases.} It allows political uncertainty to be correlated with the level of $e$, ruling out only (i) that higher effort levels make lower $\ga$-weights more likely and (ii) that increasing lobbying effort changes the structure of uncertainty so that higher $\ga$-weights become more likely at an accelerating pace.
-Assumption~\ref{as:ga_ta} \textbf{simply provides for an intuitive labeling so that} larger realization of the uncertainty variable $\ta$ increase the value of the political economy weight.
+Assumption~\ref{as:ga_c} formalizes the intuition that the legislature favors the import-competing industry more the higher is its lobbying effort, but that there are diminishing returns to lobbying activity.
+It rules out higher effort making lower weights more likely and that the structure of uncertainty changes with increasing effort so that higher weights become more likely at an accelerating pace.
+Assumption~\ref{as:ga_ta} \textbf{simply provides for an intuitive labeling so that} larger realizations of $\ta$ increase the value of the political economy weight.
 
@@
 \begin{equation}
-  \expect \left[U_L \right] = \Pr\left[ \text{TradeWar}(\bta,e_b,e_n) \right] \left[ \pi(\tau^{\mathit{n}}) - e_n \right] + \Pr\left[ \text{TradeAgreement}(\bta,e_b,e_n) \right] \pi(\tau^a) - e_b
+  \expect \left[U_L \right] = \Pr\left[ \text{TradeWar} \right] \left[ \pi(\tau^{\mathit{n}}) - e_n \right] + \Pr\left[ \text{TradeAgreement} \right] \pi(\tau^a) - e_b
   \label{eq:lv}
@@
 
-I simplify the problem by assuming that political constraints prevent the executives from choosing asymmetric tariffs.
-This implies that the executives simply maximize their expected joint payoffs:\footnote{I have assumed that $W_E(\tau) = W_E^*(\tau) \ \forall \tau$; further in a symmetric equilibrium, $\tau^n =\tau^{*n}$ and $\tau^a =\tau^{*a}$, so
-\[
-  V_E(\bta) + t = W_E(\btn) + \frac{1}{2} \left( MV(\bta) - W_E(\btn) - W_E^*(\btn) \right) = W_E(\btn) + \frac{1}{2} \left( MV(\bta) - 2W_E(\btn) \right) = \frac{1}{2}\expect MV(\bta)
-\]}
+I simplify the problem by assuming that political constraints prevent the executives from choosing asymmetric tariffs. This implies that the executives simply maximize their expected joint payoffs:
 \begin{multline}
-  \expect \left[W_E(\bta) \right] + \expect \left[ W_E^{*}(\bta) \right] = \Pr\left[ \text{TradeWar}(\bta) \right] \left[W_E(\bm{\tau}^{\mathit{n}}) + W_E^*(\bm{\tau}^{\mathit{n}}) \right] + \\ \Pr\left[ \text{TradeAgreement}(\bta) \right] \left[W_E(\bta) + W_E^*(\bta) \right]
+  \expect \left[W_E(\bta) \right] + \expect \left[ W_E^{*}(\bta) \right] = \Pr\left[ \text{TradeWar} \right] \left[W_E(\bm{\tau}^{\mathit{n}}) + W_E^*(\bm{\tau}^{\mathit{n}}) \right] + \\ \Pr\left[ \text{TradeAgreement} \right] \left[W_E(\bta) + W_E^*(\bta) \right]
   \label{eq:jv}
 \end{multline}
-The home executive's welfare is specified as follows:
-\[
-  W_E = \mathit{CS}_X(\tau) + \ga_E \cdot \mathit{PS}_X(\tau) + \mathit{CS}_Y(\tau^*) + \mathit{PS}_Y(\tau^*) + \mathit{TR}(\tau)
+The home executive's welfare is specified as
+\[
+  W_E(\bm{\tau}) = \mathit{CS}_X(\tau) + \ga_E \cdot \mathit{PS}_X(\tau) + \mathit{CS}_Y(\tau^*) + \mathit{PS}_Y(\tau^*) + \mathit{TR}(\tau)
 \]
 This is identical to the legislature's welfare function except the weight the executive places on the profits of the import industry is not a function of lobbying effort.
-\textbf{This setup does not require that the executives are not lobbied, only that the extent to which the executives favor this industry cannot be \textit{directly} influenced by lobbying effort.
-The assumption is innocuous as long as the executives' preferences are not directly altered in a significant way by lobbying over trade.
+\textbf{This setup does not require that the executives are not lobbied; the assumption is innocuous as long as the executives' preferences are not directly altered in a significant way by lobbying over trade.
 I make one more related assumption on $\ga(e,\ta)$:}
@@
 I examine a simple class of equilibria that have three key features.
-First, although uncertainty is present, information about it is symmetric.
+First, information about political uncertainty is symmetric.
 However, in line with the literature, I assume the lobby's contribution is not observable to the foreign legislature.
-This implies that the influence of one country's lobby on the other country's legislature occurs only through the tariffs selected.\footnote{cfr.
-Grossman and Helpman (1995b), page 685.} Since players in the same country can take advantage of more information than those who are in different countries, the solution concept employed here is perfect public equilibrium (PPE).
+Thus the influence of one country's lobby on the other country's legislature occurs only through the tariffs selected.\footnote{cfr.
+Grossman and Helpman (1995b), page 685.} Since players in the same country can take advantage of more information than those who are in different countries, the solution concept is perfect public equilibrium (PPE).
 
@@
 \label{sec:main}
-To understand how the executives optimally structure trade agreements, we must first examine the lobbies' incentives and the legislatures' decisions regarding breach of the agreement, including how trade-war tariffs are set.
-The symmetric, separable model structure permits restriction of attention to the home country.
+To understand how the executives optimally structure trade agreements, I first examine the lobbies' incentives and the legislatures' decisions regarding breach of the agreement, including how trade-war tariffs are set.
 
@@
 
-\item Dai, X., 2006. Dyadic Myth and Monadic Advantage: Conceptualizing the Effect of Democratic Constraints on Trade. Journal of Theoretical Politics 18, 267-297.
+%\item Dai, X., 2006. Dyadic Myth and Monadic Advantage: Conceptualizing the Effect of Democratic Constraints on Trade. Journal of Theoretical Politics 18, 267-297.
 
@@
 
-\item Ethier, W., 2012. The Political-Support Approach to Protection. Global Journal of Economics 1, 1-14.
+%\item Ethier, W., 2012. The Political-Support Approach to Protection. Global Journal of Economics 1, 1-14.
 

### 2016-06-16 draft.tex
more revisions in results section

@@
 The ``break stage'' concludes with the legislature making a choice to abide by the agreement or to break the agreement.
-In the event that the legislature breaks the agreement, there is a final stage of lobbying, resolution of the uncertainty surrounding this decision, and voting to set the trade-war tariffs $\btn$.\footnote{Alternatively, one can model the non-cooperative tariff as being determined by the break decision lobbying or the ex-ante status quo.
+In the event that the legislature breaks the agreement, there is a final stage of lobbying, resolution of the uncertainty surrounding this decision, and voting to set the trade-war tariffs $\btn$.\footnote{Alternatively, one can model the non-cooperative tariff as being determined by the break decision lobbying, through a separate process in the case of administered protection, or the ex-ante status quo.
 This does not change the fundamental dynamics of the model.} Finally, producers and consumers make their decisions.
@@
 
-The outcome of the vote on whether or not to break the trade agreement is not known to \textit{any} player until the uncertainty over the identity of the median legislator is resolved at the moment the vote takes place.\footnote{We might think of the trade-war tariff as being set in the legislation before a vote takes place in the case of administered protection for instance; the results are robust to this variant.} I represent the probability that the home legislature breaks the trade agreement and sets the tariff at $\tau^n$ as:
+The outcome of the vote on whether or not to break the trade agreement is not known to \textit{any} player until the uncertainty over the identity of the median legislator is resolved at the moment the vote takes place.
+I represent the probability that the home legislature breaks the trade agreement and sets the tariff at $\tau^n$ as:\footnote{I suppress the Nash tariffs in the expression of the break probability since they do not vary from the point of view of earlier stages.}
 \begin{multline}
@@
 \end{multline}
-
-Note that the legislature's best response functions---both for whether to break the agreement, and at what level to set the tariff if it does---will be affected by the shape of $\ga(e,\ve)$.
-I suppress the Nash tariffs in the expression of the break probability since they do not vary from the point of view of earlier stages.
 
@@
 The relationship between the home country's trade-agreement tariff and the legislature's probability of breaking the agreement is more complex.
-For any given level of effort $e_b$, when the trade agreement tariff is very low, only a small set of realizations of $\ta_b$ that lead to the lowest values of $\ga(e_b,\ta_b)$ will lead to approval of the trade agreement, implying a high break probability.
+For any given level of effort $e_b$, when the trade agreement tariff is very low, only a small set of realizations of $\ta_b$ associated with the lowest values of $\ga(e_b,\ta_b)$ will lead to approval of the trade agreement, implying a high break probability.
 As the trade agreement tariff rises, larger values of $\ga(e_b,\ta_b)$ are consistent with approving the trade agreement, so the set of $\ta_b$'s that lead to trade agreement approval is larger.
 That is, the break probability decreases as $\tau^a$ increases.
-Even if the trade agreement tariff is so high that it is disliked by all potential median legislators, it is preferable to the trade war tariff, which is even higher.
-Thus the probability of a break eventually falls to zero.
 
@@
 
-In the symmetric equilibria that are focal in the analysis that follows, any increase in $\tau^a$ is accompanied by an equal increase in $\tau^{*a}$.
+Since I focus on symmetric equilibria, any increase in $\tau^a$ is accompanied by an equal increase in $\tau^{*a}$.
 Combining the impact of the home and foreign tariff when the two are constrained to be equal:
@@
 
-\noindent When a change in one of the trade agreement tariffs must be met with an equal change in the other, the negative impact on the legislature's propensity to break the agreement of raising the home tariff outweighs the positive impact of raising the foreign tariff because of the legislature's bias toward the import-competing industry.
+\noindent The legislature's bias toward the import-competing industry overweights the negative impact of the home tariff on the break probability and ensures this result.
 
@@
 \]
-The first order condition shows that the lobby balances the cost of an extra dollar of expenditure with the higher profits from a trade war weighted by the increase in the probability of receiving the higher profit level induced by that expenditure:
+The first order condition shows that the lobby balances the cost of an extra dollar of expenditure with the higher profits from a trade war weighted by the increase in the probability of provoking the trade war:
 \begin{equation}
@@
 If the executives were to set $\tau^a = \tau^n$, there would be no incentive for the lobby to make a positive contribution.
-There are some cases in which it is in the executives' interest to set trade agreement tariffs so as to completely disengage the lobby.
+There are some other cases in which it is in the executives' interest to set trade agreement tariffs so as to disengage the lobby.
 The following results only hold when the marginal impact of the first lobbying dollar on the break probability is sufficiently high to make engaging in the political process worthwhile for the lobby.
 
-We can now proceed to the central result concerning the lobby's behavior. 
 \begin{result}
@@
 
-\noindent Raising the trade agreement tariffs decreases the benefit of inducing a break in the trade agreement by raising profits under the trade agreement.\footnote{Note that the symmetric increase in the foreign tariff does not impact profits in the import-competing industry, but it can impact lobbying incentives through the break probability.
-However, there is an analogous effect on the break probability from raising $\tau^a$ and the net effect is negative given Assumption~\ref{as:ga_c}.
-This assumption can be significantly weakened: this effect can be positive as long as it is smaller in absolute value than the negative effect on trade agreement profits.} This key result implies that the executives can reduce lobbying effort by setting higher tariffs in their trade agreement.
+\noindent Raising the trade agreement tariffs decreases the benefit of inducing a break in the trade agreement by raising profits under the trade agreement.
+This key result implies that the executives can reduce lobbying effort by setting higher tariffs in their trade agreement.
 We will see in the next section how this shapes the executives' joint decision.
@@
 Of central concern is how the break probability varies as both a direct and indirect function of $\bta$ given the lobby's optimal response.
-Result~\ref{res:bcomB} takes into effect both the direct and indirect effects:
+Result~\ref{res:bcomB} takes into account both the direct and indirect effects:
 
@@
 \end{lemma}
-I show in the \hyperlink{int_soln}{Appendix} that it is never optimal in a symmetric equilibrium for the executives to choose $\bm{\tau^a}$ below the executives' most preferred level if the legislatures will have an opportunity to break the trade agreement.
+I show in the \hyperlink{int_soln}{Appendix} that it is never optimal in a symmetric equilibrium for the executives to choose $\bm{\tau^a}$ at or below the executives' most preferred level if the legislatures will have an opportunity to break the trade agreement.
 This is because there are only second-order losses from raising tariffs slightly from the most-preferred level yet there are first-order gains in reducing the probability that the agreement will be broken.
 
-The executives will also always choose $\bta < \btn$ unless the legislature breaks even agreements with tariffs very close to the Nash level with certainty.
+The executives will always choose $\bta < \btn$ unless the legislature breaks even agreements with tariffs very close to the Nash level with certainty.
 In this case the problem is not interesting so I ignore this case.
 Thus there is an interior solution in all cases of interest, but this solution may be of two different forms.
-First, it could be at the point that maximizes the concave portion of the executives' welfare function (Equation~\ref{eq:jv2}).
-
-For some specifications, however, high enough tariffs can disengage the lobby sufficiently so that the probability that the trade agreement will be broken is zero.
+The solution could be at the point that maximizes the concave portion of the executives' welfare function.
+For some specifications, however, high enough tariffs can disengage the lobby so that the probability that the trade agreement will be broken is zero.
 If this occurs at a low enough tariff level, executive welfare can be maximized at this point.

### 2016-06-16 draft.tex
revise secs 4-6 again

@@
 \noindent All proofs are in the Appendix.
-
 Lobbying affects only the weight the legislature places on the profits of the import-competing industry.
-These profits are higher in a trade war than under a trade agreement.
-Assumption~\ref{as:ga_c} implies that the legislature becomes more favorably inclined---albeit at a decreasing rate---toward the high trade-war tariff and associated profits as lobbying increases and therefore more likely to break the trade agreement.
+These profits are higher in a trade war than a trade agreement.
+Assumption~\ref{as:ga_c} implies that the legislature becomes more favorably inclined---albeit at a decreasing rate---toward the high trade-war tariff and associated profits as lobbying increases and thus more likely to break the trade agreement.
 
@@
 
-The relationship between the home country's trade-agreement tariff and the legislature's probability of breaking the agreement is more complex.
-For any given level of effort $e_b$, when the trade agreement tariff is very low, only a small set of realizations of $\ta_b$ associated with the lowest values of $\ga(e_b,\ta_b)$ will lead to approval of the trade agreement, implying a high break probability.
+The relationship between $\tau^a$ and break probability is more complex.
+For any effort level $e_b$, when the trade agreement tariff is very low, only a small set of realizations of $\ta_b$ associated with the lowest values of $\ga(e_b,\ta_b)$ will lead to approval of the trade agreement, implying a high break probability.
 As the trade agreement tariff rises, larger values of $\ga(e_b,\ta_b)$ are consistent with approving the trade agreement, so the set of $\ta_b$'s that lead to trade agreement approval is larger.
@@
 \end{equation}
-The second order condition is satisfied given Assumption~\ref{as:ga_l_e} and Result~\ref{res:bincC}. To guarantee an interior solution, we need
+Assumption~\ref{as:ga_l_e} and Result~\ref{res:bincC} ensure the second order condition. To guarantee an interior solution, we need
   \begin{equation}
@@
 
-\noindent Raising the trade agreement tariffs decreases the benefit of inducing a break in the trade agreement by raising profits under the trade agreement.
+\noindent Raising the trade agreement tariffs decreases the benefit of a break in the trade agreement by raising trade agreement profits.
 This key result implies that the executives can reduce lobbying effort by setting higher tariffs in their trade agreement.
@@
 \label{sec:example}
-Examining a simple parameterization of the model economy is instructive.
-The fundamentals are chosen to match Bagwell and Staiger (2005) with $D(P_i) = 1 - P_i$, $Q_X(P_X) = \frac{P_X}{2}$, $Q_Y(P_Y) = P_Y$, $\Pi_X(P_X) = \frac{(P_X)^2}{4}$, and $\Pi_Y(P_Y) = \frac{(P_Y)^2}{2}$.
+In this simple parameterization of the model, the fundamentals are chosen to match Bagwell and Staiger (2005): $D(P_i) = 1 - P_i$, $Q_X(P_X) = \frac{P_X}{2}$, $Q_Y(P_Y) = P_Y$, $\Pi_X(P_X) = \frac{(P_X)^2}{4}$, and $\Pi_Y(P_Y) = \frac{(P_Y)^2}{2}$.
 Home-country imports of $X$ and exports of $Y$ are $M_X(P_X)= 1 - \frac{3}{2}P_X$ and $E_Y(P_Y)= 2P_Y -1$.
@@
 
-In order to predict the Nash tariff, the political weighting function and form of political uncertainty must be specified.
-Take for instance $\ga(e,\ve) = 1.25 + e^{.2} + \ta$ with $\ve$ distributed uniformly on $[-0.25,0.25]$.
-Facing this specification of the political process, the lobby maximizes its objective function at $e_n = 0.00166$, which produces an expected Nash tariff of $0.129$.
-Here net profits are $0.10239$ in terms of the numeraire and lobbying effort is $1.60\%$ of gross profits.
+To predict the Nash tariff, the political weighting function and form of political uncertainty must be specified.
+Take $\ga(e,\ve) = 1.25 + e^{.2} + \ta$ with $\ve$ distributed uniformly on $[-0.25,0.25]$.
+Facing this specification of the political process, the lobby maximizes its objective function at $e_n = 0.00166$, producing an expected Nash tariff of $0.129$.
+Net profits are $0.10239$ in terms of the numeraire and lobbying effort is $1.60\%$ of gross profits.
 
@@
 
-We can compare this against several different benchmarks.
+We can compare this against different benchmarks.
 The most stark is the trade-war outcome, which is the outcome that would prevail in the absence of any trade agreement.
@@
 
-Another interesting benchmark is the scenario in which no executives are involved in trade policy, but in which the politically-susceptible legislatures can make use of a trade agreement.
+Another benchmark is the scenario in which the legislatures make use of a trade agreement without the involvement of the executives.
 As in Bagwell and Staiger (2005), I find when maximizing their joint welfare in a legislature-led trade agreement, the cooperative tariff levels will be set at
@@
 The legislatures are able to use the agreement to internalize the terms of trade externality, making political influence more expensive for the lobbies.
-Indeed, lobbying expenditures rise to $0.0027$---60$\%$ higher than in the Nash case and $2.64\%$ of the gross profits of $0.10204$---while the agreement tariffs are set at $0.118$, only slightly lower than the Nash tariffs of $0.129$ and still higher than even the expected tariffs from the trade agreement scenario in which the executives choose the tariffs.
-Comparing net (expected profits) across the three scenarios, they are highest in the Nash game ($0.1024$), lowest under the trade agreement ($0.9789$) and intermediate when the legislatures make policy directly ($0.0993$).
+Lobbying expenditures rise to $0.0027$---60$\%$ higher than in the Nash case and $2.64\%$ of the gross profits of $0.10204$---while the agreement tariffs are set at $0.118$, only slightly lower than the Nash tariffs of $0.129$ and still higher than in the executive-led trade agreement.
+Comparing net expected profits across the three scenarios, they are highest in the Nash game, lowest under the trade agreement and intermediate when the legislatures make policy directly.
 
@@
 
-Now $\ga(e_b)$ is deterministic, so the lobby knows the precise contribution it must make for any given trade agreement tariffs $\bta$ to induce the legislature to break the agreement.
+When $\ga(e_b)$ is deterministic, the lobby knows the precise contribution it must make for any given trade agreement tariffs $\bta$ to induce the legislature to break the agreement.
 Here, the lobby's contribution \textit{increases} in $\bta$ since the higher are the trade agreement tariffs, the larger is the political weight that is required to induce the legislature to find them unsatisfactorily low.
@@
 
-Moreover, this case highlights the manipulation of the tariff level to discourage lobbying and therefore the disconnect that can arise between the tariffs that are chosen and the preferences of the governmental actors.
-In the example, the executive is a social welfare maximizer and so prefers zero tariffs, while the legislators with $\ga(e) = 1.25 + e^{.2}$ with $e_b = 0$ prefer $\tau^a = 0.05$.
-However, the trade agreement tariffs are set at $0.106$ precisely to ensure that $e_b =0$ so that we have a relatively free-trading legislature that upholds the agreement.
+This case highlights the manipulation of tariffs to discourage lobbying and the disconnect that can arise between the tariffs that are chosen and the preferences of governmental actors.
+The executives here prefers zero tariffs while the legislators prefer $\tau^a = 0.05$ when $e_b = 0$.
+However, the trade agreement tariffs are set at $0.106$ precisely to ensure that $e_b =0$ so that we have a legislature that upholds the agreement.
 
@@
 For simplicity, I assume uncertainty is mean-zero.
-In particular, lobbying behavior in a trade war is not altered by mean-zero uncertainty and so neither are expected trade-war tariffs.
-
+Lobbying behavior in a trade war is not altered by mean-zero uncertainty and so neither are expected trade-war tariffs.
 At earlier stages, uncertainty alters optimizing behavior by both the lobby and the executives.
@@
 Relative to the example with $\ta_b=0$, the tariff can be reduced slightly to $0.105$.
-Interestingly, the optimal tariff does not decrease monotonically.
 
@@
 Figure~\ref{fig:lobby_br} shows the lobby's reaction function at this very low level of uncertainty as well as the intermediate level when $\ve$ is distributed uniformly on $[-0.14,0.14]$ where the executives set the highest tariff level of $0.107$.
-For the former, the range of tariff levels to which the lobby responds with an intermediate contribution that leaves open the possibility of a break is very small: only those between 0.103 and 0.104.
+For the former, the range of tariff levels to which the lobby responds with a contribution that leaves open the possibility of a break is very small: only those between 0.103 and 0.104.
 The lobby disengages completely at $\bta=0.105$.
-In contrast, when uncertainty is increased to the interval $[-0.14,0.14]$, the lobby begins to leave open the possibility of a break at $\bta = 0.048$.
-It disengages at $0.107$ and this is the tariff level that the executives find optimal.
-At intermediate uncertainly levels, both portions of the best response curve are steeper and the lobby disengages at slightly lower tariff levels---as low as $0.099$ when $\ve \sim U[-0.06,0.06]$.
+When uncertainty is increased to the interval $[-0.14,0.14]$, the lobby begins to leave open the possibility of a break at $\bta = 0.048$.
+It disengages at the optimal tariff level of $0.107$.
+At intermediate uncertainly levels, both portions of the best response curve are steeper and the lobby disengages at lower tariff levels---as low as $0.099$ when $\ve \sim U[-0.06,0.06]$.
+The optimal tariff does not decrease monotonically.
 
@@
 
-I show in the \hyperlink{pfs}{Appendix} that choosing the following political weighting function aligns the political objective function used in the current model with the PFS framework:
-		    \begin{equation}
-		      \ga(e) = 1 + \frac{e}{a\pi_X(\tau)} .
-		      \label{eq:gh}
-		    \end{equation}
+I show in the \hyperlink{pfs}{Appendix} that choosing $\ga(e) = 1 + \frac{e}{a\pi_X(\tau)}$ aligns the political objective function used in the current model with the PFS framework.
 Although using the $\ga$-weighted government welfare function is a reduced-form approach in the sense that legislative dynamics are not fully modeled, a benefit is that it does not require returns to lobbying effort to be constant.
@@
 
-The linear PFS form provides for an almost unmatched simplicity and elegance that is able to capture key cross-industry predictions on import tariffs.
+The elegant PFS form is able to capture key cross-industry predictions on import tariffs.
 A perhaps unintended byproduct in the subsequent literature seems to have arisen in the form of a focus on matching fine details of governments' welfare-mindedness using a model that has a very simple unitary government and is thus better suited for other purposes.
@@
 The executives' incentive to raise tariffs in order to reduce lobbying effort provides insight into the empirical puzzle in the Protection for Sale literature that levels of protection and associated deadweight losses are too high relative to lobbying expenditure given the high estimates for governments' weighting of social welfare.
-I have shown that this serves to mediate the relationship between preferences for contributions relative to social welfare and the tariffs that are provided.
 The intuition is clear: the observed lobbying expenditure levels may in fact be low \textit{because} tariffs have been raised sufficiently high to prevent political pressure and the increased risk of a costly trade disruption it engenders.

### 2016-06-17 draft.tex
revise round 2 letter and draft for submission

@@
 In each country, there is an executive who concludes trade agreements and a legislature that has final say on trade policy.
-To focus attention on protectionist political forces\textbf{, highlight the model's central mechanisms and minimize the number of actors that must be modeled}, I assume that only the import-competing industry in each country is politically-organized and that it is represented by a single lobbying organization.\footnote{The model extends easily to the case of multiple lobbies.
+\textbf{To focus attention on protectionist political forces, highlight the model's central mechanisms and minimize the number of actors that must be modeled}, I assume that only the import-competing industry in each country is politically-organized and that it is represented by a single lobbying organization.\footnote{The model extends easily to the case of multiple lobbies.
 To illustrate, adding an export lobby with no uncertainty predicts import lobbies remain disengaged while the executives can reduce tariffs by recruiting support from exporters.
@@
 Thus the influence of one country's lobby on the other country's legislature occurs only through the tariffs selected.\footnote{cfr.
-Grossman and Helpman (1995b), page 685.} Since players in the same country can take advantage of more information than those who are in different countries, the solution concept is perfect public equilibrium (PPE).
+Grossman and Helpman (1995b), page 685.} \textbf{Since players in the same country can take advantage of more information than those who are in different countries, the solution concept is perfect public equilibrium (PPE).}
 

### 2016-06-17 letter-round2.tex
revise round 2 letter and draft for submission

--- 2016-06-10 letter-round2.tex
@@
 \begin{center}
-\large Response to Editor and Referee \\
-\normalsize Month ???, 2016
+\large Response to Editor \\
+\normalsize June 2016
 \end{center}
@@
 \vskip.2in
-Thank you. \\
-
-Logistics. Points were not numbered, so pasting them here\\
+Thank you for the second opportunity to bring this paper above your bar.
+I believe the paper is now much, much tighter and better organized.
+Below I detail how I responded to each of your points.
+As they were not numbered, I have listed each of your points in order (italicized) with the corresponding response immediately below.
+Because the changes in the text are so dramatic, bolding every one became nearly meaningless.
+I therefore bolded only those passages that have changed in response to a specific suggestion.
+\\
 
@@
 The idea is sufficiently straightforward that it does not need 28 pages of exposition.
-Please reduce the length of the paper by half.}
+Please reduce the length of the paper by half.} \\
+I believe I have achieved what you have asked for through a combination of jettisoning superfluous material, moving essential material so that it was not repeated, and generally tightening the exposition.
+I've been able to cut the length of the body of the paper from 28 pages to $18\frac{1}{2}$.
 
@@
 \renewcommand\labelitemi{-}
-	\item \textit{Avoid changes of notation as you go along. This is totally unhelpful to the reader. Please pick one and stick to it throughout the paper.}
+	\item \textit{Avoid changes of notation as you go along. This is totally unhelpful to the reader. Please pick one and stick to it throughout the paper.} \\
+	I have worked hard to standardize notation.
+I would welcome any suggestions about remaining inconsistencies.
+The one potential duplication of which I am aware is using ``PS'' (producer surplus) and $\pi$ in the government welfare functions and lobby's objective function respectively although they represent the same quantity.
+This is a purposeful stylistic choice but it can be reversed to use $\pi$ throughout.
 	\item \textit{Avoid digressions, lengthy discussions, and overreach: you stretch your model far too much.
 Stick to the main story.
-The paper still degenerates far too often into a series of discussions that bring little additional insight.}
-	\item \textit{Section 2.4 either belongs to section 1.1 or should move as section 1.2. Section 2.4 is far too long.}
-	\item \textit{Your recall stuff far too often. Not all readers suffer from memory losses.}
+The paper still degenerates far too often into a series of discussions that bring little additional insight.} \\
+	I have ruthlessly excised all of the above, leaving only those that were added as a result of the refereeing process at this journal or that have arisen consistently in the past.
+	\item \textit{Section 2.4 either belongs to section 1.1 or should move as section 1.2. Section 2.4 is far too long.} \\
+	The text of Section 2.4 has been shorted and redistributed.
+I inserted the few paragraphs that concern assumptions into Section 2.2 (preferences) since they could not come before the model was introduced.
+All the rest that I deemed essential have been integrated in condensed form into the introduction and literature review.
+I have bolded them where inserted.
+
+	\item \textit{Your recall stuff far too often. Not all readers suffer from memory losses.} \\
+	All recalls have been removed as far as I can tell.
 \end{itemize}
@@
 	\item \textit{Assuming that lobbies represent only import-competing interests is a strong assumption and it requires theoretical and/or empirical justification.} \\
-	Done, on page [8]
+	Done, on page 6. Footnote 6 explains that extending to multiple lobbies is possible.
 	\item \textit{Your model is static, so claiming that it rationalizes the four-year delay and eventual ratification of the US-SK FTA is a stretch.
@@
   \item \textit{Your assumption that the identity of the median voter is not known a priori is standard in probability voting models, so you may want to stress the analogy between your setting and such models.} \\
-	Thank you for this suggestion. I have incorporated a short reference on page [ ].
+	Thank you for this suggestion. I have incorporated a reference on page 2.
 
@@
 		I agree that the idea you propose concerning fast-track authority / TPA is an important one and it is an idea that a colleague and I are exploring in a separate project as I believe it is beyond the scope of the current paper.
-The extension of the model to one with multiple sectors and lobbies is one that I believe is worthwhile for many purposes.
+The extension of the model to one with multiple sectors and lobbies is one that is worthwhile for many purposes.
 However, the essential idea of the current paper is, as you have judged, `sufficiently straightforward' that in my view the complications involved in the multi-sector model are not warranted.
@@
 All the results go through (with one small additional assumption on magnitudes: it must be ensured that the lobby's direct response to an increase in trade agreement tariffs is stronger than a new effect wherein the lobby might increase effort to compensate for the trading partner's legislature reducing its break probability in response to that same tariff increase).
-However, the model is significantly complicated as two actors have to be modeled and results added since they can no longer be accounted for by symmetry.
+However, the model is significantly complicated as two additional actors have to be modeled and results added since the foreign lobby and legislature can no longer be accounted for by symmetry.
 This added considerable length without compensating insight, so I opted to add a footnote explaining the near-equivalence of the models in line with your desire to simplify the presentation and not distract from the central point.
-The new text reads as follows can can be found in \textbf{Footnote ???
-on Page???}:
+The new text reads as follows can can be found in Footnote 10 on Page 7:
 	\begin{quote}
-		Alternatively, one can take the break decisions in the two countries to be sequential with the executives behind a veil of ignorance about which legislature will move first.
-In that set up, the end results are the same as in the current model but additional intermediate results are needed.
+		If the break decisions in the two countries are sequential with the executives behind a veil of ignorance about which legislature moves first, the end results are the same but additional intermediate results are needed.
 	\end{quote}
@@
   \item \textit{``the uncertainty about the median legislator's identity is resolved.'' What is this uncertainty really about?} \\
-	This is explained in Section [currently 2.4, but will move], so I did not repeat it or reference it here in line with the earlier stylistic advice.
+	This is now explained on page 2, so I did not repeat it or reference it here in line with the earlier stylistic advice.
 
   \item \textit{The equilibrium selection criterion requires justification and motivation.} \\
-	Done. See page [10]
+	Done. See page 9.
 
   \item \textit{Assumption 2 does not formalize any intuition: it is just a labelling.} \\
-	Quite right. The text has been modified appropriately.
+	Quite right. The text has been modified appropriately. This is now near the top of page 8.
+
   \item \textit{I feel uneasy about Assumption 3. I would rather have it as an outcome and, if it is not the only possibly one, a discussion of the conditions under which this arises.} \\
@@
 	The issue is quite a nuanced, in particular among democracies.
-I have discussed the empirical evidence on democracies---separate from the claim that I made about moving from autocracy to democracy---from Henisz and Mansfield (2006) in a new paragraph.
-I have also added a reference to Frye and Mansfield (2003) to support the original claim.
-I don't consider any of this essential and would be happy to remove either or both paragraphs.
+In line with the earlier suggestion to remove digressions, I removed the paragraph at the end of section 4 regarding moves from autocracy to democracy.
+I have, however, added a short discussion regarding democracies citing from Henisz and Mansfield (2006) at the end of the introduction to support the argument that the model is broadly applicable.
 \end{itemize}
 
+\vskip.3in
 Best wishes,\\

### 2016-06-23 draft.tex
remove bolding for previous round submission

--- 2016-06-16 draft.tex
@@
 Thus trade agreements act as a domestic political commitment device: executives set relatively high tariffs to discourage lobbying and increase the chance that the agreement will be ratified.
-The model sheds light on the empirical puzzle surrounding governments' welfare weights in the Grossman and Helpman (1994) model and \textbf{provides a new explanation for failures to ratify trade agreements}.
+The model sheds light on the empirical puzzle surrounding governments' welfare weights in the Grossman and Helpman (1994) model and provides a new explanation for failures to ratify trade agreements.
 
@@
 The model becomes more realistic, has smoother results and delivers additional intuition when I allow all actors to be uncertain about the weight the median legislator places on the profits of the lobbying industry.
-\textbf{Similar to probabilistic voting models that incorporate policy motivation as in Roemer (1997),} this represents the idea that the lobbying and voting ``game'' that goes on within legislatures is often so complex that none of the actors know precisely what the outcome will be---that is, no one can predict exactly which legislator will be the median and therefore what weight will be placed on import-industry profits at the time of the vote.
-\textbf{One can conceptualize this uncertainty as being a result of the lobby's strategic choice to engage with only some key members of the legislature.
-For instance, the lobbied legislators may promise to deliver the votes of their non-lobbied colleagues but may not be able to do so with certainty because those colleagues are subject to cross-pressuring from various sources.}
+Similar to probabilistic voting models that incorporate policy motivation as in Roemer (1997), this represents the idea that the lobbying and voting ``game'' that goes on within legislatures is often so complex that none of the actors know precisely what the outcome will be---that is, no one can predict exactly which legislator will be the median and therefore what weight will be placed on import-industry profits at the time of the vote.
+One can conceptualize this uncertainty as being a result of the lobby's strategic choice to engage with only some key members of the legislature.
+For instance, the lobbied legislators may promise to deliver the votes of their non-lobbied colleagues but may not be able to do so with certainty because those colleagues are subject to cross-pressuring from various sources.
 
@@
 
-\textbf{Although the political process here most closely matches that of the United States in the post-war era,\footnote{In particular, the legislatures' decision to abide by or break the trade agreement is modeled on the ``Fast Track / Trade Promotion Authority'' that the U.S.
+Although the political process here most closely matches that of the United States in the post-war era,\footnote{In particular, the legislatures' decision to abide by or break the trade agreement is modeled on the ``Fast Track / Trade Promotion Authority'' that the U.S.
 Congress periodically grants to the Executive branch.} the model is broadly applicable because the central dynamic that emerges---that trade agreements are formed in the \textit{shadow} of lobbying pressure instead of in direct response to it---appears to be present in a wide range of countries.
@@
 Although the number of formal veto points tends to be lower in, for instance, parliamentary systems, veto points are typically not absent altogether (Henisz and Mansfield 2006).
-In reality, constraints on the executive are almost always present in democratic regimes, although they may operate in different ways.}
+In reality, constraints on the executive are almost always present in democratic regimes, although they may operate in different ways.
 
@@
 
-\textbf{Since trade policies are overwhelmingly determined in the context of trade agreements, it is useful to have a framework to bring together the endogenous political pressure of PFS-style modeling with the trade agreements approach.
+Since trade policies are overwhelmingly determined in the context of trade agreements, it is useful to have a framework to bring together the endogenous political pressure of PFS-style modeling with the trade agreements approach.
 The government objective function introduced in this paper is intended to be a bridge between the two.
 This objective function starts from the Baldwin (1987) government objective function and adds flexibility: the weight the legislature places on profits is endogenous (PFS), potentially non-linear (Findlay and Wellisz 1982; Dixit, Grossman and Helpman 1997) and stochastic.
-Buzard (2016) is an example of how this modeling approach can be used to integrate  endogenous political activity and exogenous shocks into the analysis of questions regarding optimal trade agreement design.}
-
-\textbf{Modeling the objective function so closely on the standard in the trade agreements literature allows for direct comparisons to the large body of work that studies exogenous shocks only, revealing cleanly the effects of the addition of endogenous lobbying.} To preview, this model is closest to that of Bagwell and Staiger (2005), with two main changes: in place of their unitary government signing a trade agreement and having different preferences ex-ante and ex-post, this model has two branches of government with differing preferences, where the political economy weights of the legislature are determined endogenously.
+Buzard (2016) is an example of how this modeling approach can be used to integrate  endogenous political activity and exogenous shocks into the analysis of questions regarding optimal trade agreement design.
+
+Modeling the objective function so closely on the standard in the trade agreements literature allows for direct comparisons to the large body of work that studies exogenous shocks only, revealing cleanly the effects of the addition of endogenous lobbying.
+To preview, this model is closest to that of Bagwell and Staiger (2005), with two main changes: in place of their unitary government signing a trade agreement and having different preferences ex-ante and ex-post, this model has two branches of government with differing preferences, where the political economy weights of the legislature are determined endogenously.
 
@@
 In each country, there is an executive who concludes trade agreements and a legislature that has final say on trade policy.
-To focus attention on protectionist political forces\textbf{, highlight the model's central mechanisms and minimize the number of actors that must be modeled}, I assume that only the import-competing industry in each country is politically-organized and that it is represented by a single lobbying organization.\footnote{The model extends easily to the case of multiple lobbies.
+To focus attention on protectionist political forces, highlight the model's central mechanisms and minimize the number of actors that must be modeled, I assume that only the import-competing industry in each country is politically-organized and that it is represented by a single lobbying organization.\footnote{The model extends easily to the case of multiple lobbies.
 To illustrate, adding an export lobby with no uncertainty predicts import lobbies remain disengaged while the executives can reduce tariffs by recruiting support from exporters.
@@
 Next, nature determines with probabilities $o$ and $o^*$ whether each legislature will have the opportunity to take a vote to break the agreement.
-I assume that these probabilities are less than one-half and mutually exclusive.\footnote{While this assumption is made for purposes of exposition and tractability, one can think of $o$ and $o^*$ as being determined by events beyond the executives' or legislatures' control that determine whether or not the legislatures will be willing to consider the lobby's request; for example, they would be affected by the occurrence of an economic crisis that diverts the legislature's attention from less-pressing matters.} This allows me to focus on the key interaction between the domestic actors and abstract from the strategic interaction between the lobbies.\footnote{\textbf{If the break decisions in the two countries are sequential with the executives behind a veil of ignorance about which legislature moves first, the end results are the same but additional intermediate results are needed.} The optimization problem of the executives becomes more complex as they decide how best to exploit the second lobby's incentive to free ride on the effort of the first-mover lobby.}
+I assume that these probabilities are less than one-half and mutually exclusive.\footnote{While this assumption is made for purposes of exposition and tractability, one can think of $o$ and $o^*$ as being determined by events beyond the executives' or legislatures' control that determine whether or not the legislatures will be willing to consider the lobby's request; for example, they would be affected by the occurrence of an economic crisis that diverts the legislature's attention from less-pressing matters.} This allows me to focus on the key interaction between the domestic actors and abstract from the strategic interaction between the lobbies.\footnote{If the break decisions in the two countries are sequential with the executives behind a veil of ignorance about which legislature moves first, the end results are the same but additional intermediate results are needed.
+The optimization problem of the executives becomes more complex as they decide how best to exploit the second lobby's incentive to free ride on the effort of the first-mover lobby.}
 
@@
 It rules out higher effort making lower weights more likely and that the structure of uncertainty changes with increasing effort so that higher weights become more likely at an accelerating pace.
-Assumption~\ref{as:ga_ta} \textbf{simply provides for an intuitive labeling so that} larger realizations of $\ta$ increase the value of the political economy weight.
+Assumption~\ref{as:ga_ta} simply provides for an intuitive labeling so that larger realizations of $\ta$ increase the value of the political economy weight.
 
@@
 This is identical to the legislature's welfare function except the weight the executive places on the profits of the import industry is not a function of lobbying effort.
-\textbf{This setup does not require that the executives are not lobbied; the assumption is innocuous as long as the executives' preferences are not directly altered in a significant way by lobbying over trade.
-I make one more related assumption on $\ga(e,\ta)$:}
+This setup does not require that the executives are not lobbied; the assumption is innocuous as long as the executives' preferences are not directly altered in a significant way by lobbying over trade.
+I make one more related assumption on $\ga(e,\ta)$:
 
@@
 
-\textbf{That is, even for the least favorable outcome of the lobbying process, the legislature is more protectionist than the executive regardless of the lobby's effort choice.
+That is, even for the least favorable outcome of the lobbying process, the legislature is more protectionist than the executive regardless of the lobby's effort choice.
 This is not essential but it simplifies the analysis, ensuring that trade agreement tariffs are smaller than trade war tariffs.
 It matches the empirical finding that politicians with larger constituencies are less sensitive to special interests and represents well the post-war U.S.
-where Congress has consistently been more protectionist than the President.}\footnote{cfr.
+where Congress has consistently been more protectionist than the President.\footnote{cfr.
 Lohmann and O'Halloran (1994) for the general argument, Destler (2005) on U.S.
@@
 \noindent When the executives raise $\bta$ the legislature becomes less likely to abrogate the agreement, for three reasons.
-First, the legislature prefers a higher domestic tariff \textbf{(Lemma~\ref{res:leg_a})}; second, the higher tariff discourages lobbying, reducing $B(\bta)$ indirectly \textbf{(Result~\ref{res:lobby})}; and finally, the lower lobbying effort directly reduces the legislature's preferred tariff further \textbf{(Lemma~\ref{res:bcomb})}.
+First, the legislature prefers a higher domestic tariff (Lemma~\ref{res:leg_a}); second, the higher tariff discourages lobbying, reducing $B(\bta)$ indirectly (Result~\ref{res:lobby}); and finally, the lower lobbying effort directly reduces the legislature's preferred tariff further (Lemma~\ref{res:bcomb}).
 Thus, beyond promising a lower tariff from the trading partner, we can think of the trade agreement as a sort of political commitment device that can be used to reduce political pressure and therefore get the legislature to maintain a lower tariff than it otherwise would.
@@
 
-\item \textbf{Roemer, J., 1997. Political-economic Equilibrium when Parties Represent Constituents: The Unidimensional Case. Social Choice and Welfare 14, 479-502.}
+\item Roemer, J., 1997. Political-economic Equilibrium when Parties Represent Constituents: The Unidimensional Case. Social Choice and Welfare 14, 479-502.
 

### 2016-06-23 draft.tex
remove G&G appendix and ref to isomorphism in sec:gh
change 'producer surplus' to 'profits' throughout text and appendix

@@
 This gives rise to a terms-of-trade externality.
-Profits and producer surplus (identical in this model) in a sector are increasing in the price of its good and also in the domestic tariff.
+Profits in a sector are increasing in the price of its good and also in the domestic tariff.
 This fact, combined with the assumptions on specific factor ownership, motivates political activity by import-competing lobbies.
@@
 \begin{equation}
-  W_{\mathit{ML}} = \mathit{CS}_X(\tau) + \ga(e,\ve) \cdot \mathit{PS}_X(\tau) + \mathit{CS}_Y(\tau^*) + \mathit{PS}_Y(\tau^*) + \mathit{TR}(\tau)
+  W_{\mathit{ML}} = \mathit{CS}_X(\tau) + \ga(e,\ve) \cdot \pi_X(\tau) + \mathit{CS}_Y(\tau^*) + \pi_Y(\tau^*) + \mathit{TR}(\tau)
   \label{eq:ml}
 \end{equation}
-where $\mathit{CS}$ is consumer surplus, $\mathit{PS}$ is producer surplus, $\ga(e,\ta)$ is the weight placed on producer surplus in the import-competing industry, and $\mathit{TR}$ is tariff revenue.\footnote{Labor income $l$ could also be included in both median legislator and executive welfare.
+where $\mathit{CS}$ is consumer surplus, $\pi$ are \textbf{profits}, $\ga(e,\ta)$ is the weight placed on \textbf{profits} in the import-competing industry, and $\mathit{TR}$ is tariff revenue.\footnote{Labor income $l$ could also be included in both median legislator and executive welfare.
 I omit it because its inclusion alters none of the results and only serves to complicate the exposition.} I model the decisions of the legislature as being taken by a median legislator with the weight the median legislator places on import-competing industry profits affected by the level of lobbying effort $e$ and a random variable $\ta$.
@@
 \end{equation}
-where $\pi(\cdot)$ is the current-period profit.
+where $\pi(\cdot)$ is the current-period profit \textbf{and the subscript $X$ has been dropped for notational convenience}.
 
@@
 
-The lower world price for home's export good has a larger negative effect on producer surplus than the positive effect on consumer surplus.
+The lower world price for home's export good has a larger negative effect on \textbf{profits} than the positive effect on consumer surplus.
 The net effect of an increase in foreign trade-agreement tariffs on home legislative welfare is negative, leading to an increased probability that the trade agreement will be broken.
@@
 
-I show in the \hyperlink{pfs}{Appendix} that choosing $\ga(e) = 1 + \frac{e}{a\pi_X(\tau)}$ aligns the political objective function used in the current model with the PFS framework.
 Although using the $\ga$-weighted government welfare function is a reduced-form approach in the sense that legislative dynamics are not fully modeled, a benefit is that it does not require returns to lobbying effort to be constant.
@@
 
-\noindent The price of good $Y$ is decreasing in $\tau^{*a}$, so consumer surplus is increasing and producer surplus is decreasing in $\tau^{*a}$.
-Because $Y$ is being exported, the decrease in producer surplus is larger than the increase in consumer surplus, making the numerator negative.
-Since producer surplus in the denominator is increasing in $\tau$, the expression in Equation~\ref{appex:astar} is negative for all $\tau^{*a}$.
+\noindent The price of good $Y$ is decreasing in $\tau^{*a}$, so consumer surplus is increasing and \textbf{profits} are decreasing in $\tau^{*a}$.
+Because $Y$ is being exported, the decrease in \textbf{profits} is larger than the increase in consumer surplus, making the numerator negative.
+Since \textbf{profits} in the denominator are increasing in $\tau$, the expression in Equation~\ref{appex:astar} is negative for all $\tau^{*a}$.
 $\hfill\blacksquare$
@@
 
-Because total social welfare is maximized at $\tau^a = \tau^{*a} = 0$, $W(\btn)-W(\bta)$ is always negative, whereas producer surplus is increasing in $\tau^a$, so the second summand is positive everywhere.
+Because total social welfare is maximized at $\tau^a = \tau^{*a} = 0$, $W(\btn)-W(\bta)$ is always negative, whereas \textbf{profits} are increasing in $\tau^a$, so the second summand is positive everywhere.
 With a positive denominator, we thus have that the derivative is positive on $[0,\tau^O]$.
@@
 Thus the first summand in the numerator is zero at $\bta = \bm{0}$ and increasingly negative as $\bta$ increases.
-The second summand is positive everywhere because social welfare, $W$, is lowest at $\btn$ and producer surplus is increasing everywhere.
+The second summand is positive everywhere because social welfare, $W$, is lowest at $\btn$ and \textbf{profits} are increasing everywhere.
 Thus the numerator is positive at $\bm{0}$ and at least for very small $\bta$
@@
 
+\begin{comment}
 \vskip.4in
@@
 		    \] 
+\end{comment}
 

### 2016-06-23 draft.tex
change PS to \pi notation throughout

@@
 \[
-  W_E(\bm{\tau}) = \mathit{CS}_X(\tau) + \ga_E \cdot \mathit{PS}_X(\tau) + \mathit{CS}_Y(\tau^*) + \mathit{PS}_Y(\tau^*) + \mathit{TR}(\tau)
+  W_E(\bm{\tau}) = \mathit{CS}_X(\tau) + \ga_E \cdot \pi_X(\tau) + \mathit{CS}_Y(\tau^*) + \pi_Y(\tau^*) + \mathit{TR}(\tau)
 \]
@@
 \begin{equation}
-		\frac{\partial \mathit{CS}_X(\tau)}{\partial \tau^n} + \ga(e_n,\ve_n) \cdot \frac{\partial \mathit{PS}_X(\tau)}{\partial \tau^n} +  \frac{\partial \mathit{TR}(\tau)}{\partial \tau^n} = 0 .
+		\frac{\partial \mathit{CS}_X(\tau)}{\partial \tau^n} + \ga(e_n,\ve_n) \cdot \frac{\partial \pi_X(\tau)}{\partial \tau^n} +  \frac{\partial \mathit{TR}(\tau)}{\partial \tau^n} = 0 .
 		\label{eq:legfoc}
@@
 \begin{multline}  
-  b(e_b,\bta,\btn) = \Pr [ \mathit{CS}_X(\tau^n) + \ga(e_b,\ve_b) \cdot \mathit{PS}_X(\tau^n) + \mathit{CS}_Y(\tau^{*n}) + \mathit{PS}_Y(\tau^{*n}) + \mathit{TR}(\tau^n) >  \\ \mathit{CS}_X(\tau^a) + \ga(e_b,\ve_b) \cdot \mathit{PS}_X(\tau^a) + \mathit{CS}_Y(\tau^{*a}) + \mathit{PS}_Y(\tau^{*a}) + \mathit{TR}(\tau^a) ]
+  b(e_b,\bta,\btn) = \Pr [ \mathit{CS}_X(\tau^n) + \ga(e_b,\ve_b) \cdot \pi_X(\tau^n) + \mathit{CS}_Y(\tau^{*n}) + \pi_Y(\tau^{*n}) + \mathit{TR}(\tau^n) >  \\ \mathit{CS}_X(\tau^a) + \ga(e_b,\ve_b) \cdot \pi_X(\tau^a) + \mathit{CS}_Y(\tau^{*a}) + \pi_Y(\tau^{*a}) + \mathit{TR}(\tau^a) ]
 \end{multline}
@@
 \begin{multline}  
-  \textstyle \Pr \Big[ \frac{\mathit{CS}_X(\tau^n) + \mathit{PS}_X(\tau^n) + \mathit{CS}_Y(\tau^{*n}) + \mathit{PS}_Y(\tau^{*n}) + \mathit{TR}(\tau^n)  -\mathit{CS}_X(\tau^a) - \mathit{PS}_X(\tau^a) - \mathit{CS}_Y(\tau^{*a}) - \mathit{PS}_Y(\tau^{*a}) - \mathit{TR}(\tau^a)}{\mathit{PS}_X(\tau^a) - \mathit{PS}_X(\tau^n)} + 1 \\ \textstyle < \ga(e_b,\ve_b) \Big]
+  \textstyle \Pr \Big[ \frac{\mathit{CS}_X(\tau^n) + \pi_X(\tau^n) + \mathit{CS}_Y(\tau^{*n}) + \pi_Y(\tau^{*n}) + \mathit{TR}(\tau^n)  -\mathit{CS}_X(\tau^a) - \pi_X(\tau^a) - \mathit{CS}_Y(\tau^{*a}) - \pi_Y(\tau^{*a}) - \mathit{TR}(\tau^a)}{\pi_X(\tau^a) - \pi_X(\tau^n)} + 1 \\ \textstyle < \ga(e_b,\ve_b) \Big]
   \label{eq:b_ex}
@@
 \begin{equation}
-  \frac{-\left(\mathit{PS}_X(\tau^a) - \mathit{PS}_X(\tau^n)\right)\left(\frac{\partial \mathit{CS}_Y(\tau^{*a})}{\partial \tau^{*a}} + \frac{\partial \mathit{PS}_Y(\tau^{*a})}{\partial \tau^{*a}}\right)}{\left(\mathit{PS}_X(\tau^a) - \mathit{PS}_X(\tau^n)\right)^2} = \frac{\frac{\partial \mathit{CS}_Y(\tau^{*a})}{\partial \tau^{*a}} + \frac{\partial \mathit{PS}_Y(\tau^{*a})}{\partial \tau^{*a}}}{\mathit{PS}_X(\tau^n) - \mathit{PS}_X(\tau^a)}.
+  \frac{-\left(\pi_X(\tau^a) - \pi_X(\tau^n)\right)\left(\frac{\partial \mathit{CS}_Y(\tau^{*a})}{\partial \tau^{*a}} + \frac{\partial \pi_Y(\tau^{*a})}{\partial \tau^{*a}}\right)}{\left(\pi_X(\tau^a) - \pi_X(\tau^n)\right)^2} = \frac{\frac{\partial \mathit{CS}_Y(\tau^{*a})}{\partial \tau^{*a}} + \frac{\partial \pi_Y(\tau^{*a})}{\partial \tau^{*a}}}{\pi_X(\tau^n) - \pi_X(\tau^a)}.
   \label{appex:astar}
@@
 \begin{equation}
-  \frac{\left(\mathit{PS}_X(\tau^n) - \mathit{PS}_X(\tau^a)\right)\left(\frac{\partial \mathit{CS}_X(\tau^{a})}{\partial \tau^{a}} + \frac{\partial \mathit{PS}_X(\tau^a)}{\partial \tau^{a}} + \frac{\partial \mathit{TR}(\tau^{a})}{\partial \tau^{a}}\right) - \left[ W(\btn)-W(\bta)\right] \frac{\partial \mathit{PS}_X(\tau^a)}{\partial \tau^a}}{\left(\mathit{PS}_X(\tau^a) - \mathit{PS}_X(\tau^n)\right)^2}.
+  \frac{\left(\pi_X(\tau^n) - \pi_X(\tau^a)\right)\left(\frac{\partial \mathit{CS}_X(\tau^{a})}{\partial \tau^{a}} + \frac{\partial \pi_X(\tau^a)}{\partial \tau^{a}} + \frac{\partial \mathit{TR}(\tau^{a})}{\partial \tau^{a}}\right) - \left[ W(\btn)-W(\bta)\right] \frac{\partial \pi_X(\tau^a)}{\partial \tau^a}}{\left(\pi_X(\tau^a) - \pi_X(\tau^n)\right)^2}.
   \label{appex:a}
@@
 
-\noindent $\left(\mathit{PS}_X(\tau^n) - \mathit{PS}_X(\tau^a)\right)$ is always positive by Assumption~\ref{as:ga_l_e}.
-Because the optimal unilateral tariff for large welfare-maximizing governments is positive (call it $\tau^O$), $\left(\frac{\partial \mathit{CS}_X(\tau^{a})}{\partial \tau^{a}} + \frac{\partial \mathit{PS}_X(\tau^a)}{\partial \tau^{a}} + \frac{\partial \mathit{TR}(\tau^{a})}{\partial \tau^{a}}\right)$ is increasing up to $\tau^O$ and decreasing above it.
+\noindent $\left(\pi_X(\tau^n) - \pi_X(\tau^a)\right)$ is always positive by Assumption~\ref{as:ga_l_e}.
+Because the optimal unilateral tariff for large welfare-maximizing governments is positive (call it $\tau^O$), $\left(\frac{\partial \mathit{CS}_X(\tau^{a})}{\partial \tau^{a}} + \frac{\partial \pi_X(\tau^a)}{\partial \tau^{a}} + \frac{\partial \mathit{TR}(\tau^{a})}{\partial \tau^{a}}\right)$ is increasing up to $\tau^O$ and decreasing above it.
 Thus the first summand is increasing up until $\tau^O$ and decreasing thereafter.
@@
 It is also positive over the remaining $(\tau^O,\tau^n)$.
-To see this, add $\left(\tilde{\Gamma} - 1 \right) \frac{\partial \mathit{PS}_X(\tau^a)}{\partial \tau^a} \left(\mathit{PS}_X(\tau^n) - \mathit{PS}_X(\tau^a)\right)$ to the first summand and subtract it from the second.
+To see this, add $\left(\tilde{\Gamma} - 1 \right) \frac{\partial \pi_X(\tau^a)}{\partial \tau^a} \left(\pi_X(\tau^n) - \pi_X(\tau^a)\right)$ to the first summand and subtract it from the second.
 For any particular value of $\tilde{\tau}^a$, one can choose the $\tilde{\Gamma}$ weight that would make $\tilde{\tau}^a$ the preferred unilateral tariff; this makes the derivative in the first summand zero.
@@
 \[
-  \frac{\left(\mathit{PS}_X(\tau^n) - \mathit{PS}_X(\tau^a)\right)\left(\frac{\partial \mathit{W}_X(\bta)}{\partial \tau^{a}} + \frac{\partial \mathit{W}_X(\bta)}{\partial \tau^{*a}}\right) - \left[ W(\btn)-W(\bta)\right] \frac{\partial \mathit{PS}_X(\tau^a)}{\partial \tau^a}}{\left(\mathit{PS}_X(\tau^n) - \mathit{PS}_X(\tau^a)\right)^2}.
-\]
-where $\frac{\partial \mathit{W}_X(\bta)}{\partial \tau^{a}} + \frac{\partial \mathit{W}_X(\bta)}{\partial \tau^{*a}} = \frac{\partial \mathit{CS}_X(\tau^{a})}{\partial \tau^{a}} + \frac{\partial \mathit{PS}_X(\tau^a)}{\partial \tau^{a}} + \frac{\partial \mathit{TR}(\tau^{a})}{\partial \tau^{a}} + \frac{\partial \mathit{CS}_Y(\tau^{*a})}{\partial \tau^{*a}} + \frac{\partial \mathit{PS}_Y(\tau^{*a})}{\partial \tau^{*a}}$ is the total derivative of social welfare.
+  \frac{\left(\pi_X(\tau^n) - \pi_X(\tau^a)\right)\left(\frac{\partial \mathit{W}_X(\bta)}{\partial \tau^{a}} + \frac{\partial \mathit{W}_X(\bta)}{\partial \tau^{*a}}\right) - \left[ W(\btn)-W(\bta)\right] \frac{\partial \pi_X(\tau^a)}{\partial \tau^a}}{\left(\pi_X(\tau^n) - \pi_X(\tau^a)\right)^2}.
+\]
+where $\frac{\partial \mathit{W}_X(\bta)}{\partial \tau^{a}} + \frac{\partial \mathit{W}_X(\bta)}{\partial \tau^{*a}} = \frac{\partial \mathit{CS}_X(\tau^{a})}{\partial \tau^{a}} + \frac{\partial \pi_X(\tau^a)}{\partial \tau^{a}} + \frac{\partial \mathit{TR}(\tau^{a})}{\partial \tau^{a}} + \frac{\partial \mathit{CS}_Y(\tau^{*a})}{\partial \tau^{*a}} + \frac{\partial \pi_Y(\tau^{*a})}{\partial \tau^{*a}}$ is the total derivative of social welfare.
 Since social welfare is maximized at $\bta = (0,0)$,\footnote{Note, this is identical to the result that joint social welfare is maximized at zero tariffs because of symmetry.}  this is negative $\forall \bm{\tau} \in (\bm{0},\btn]$; note that it is 0 at $\bm{0}$ and vanishingly small for very small tariffs.
@@
 It is also positive for all other values of $\bta$ strictly below $\btn$. Just as in the proof of Lemma~\ref{res:leg_a}, one can add
-$\left(\tilde{\Gamma} - 1 \right) \frac{\partial \mathit{PS}_X(\tau^a)}{\partial \tau^a} \left(\mathit{PS}_X(\tau^n) - \mathit{PS}_X(\tau^a)\right)$ to the first summand and subtract it from the second.
+$\left(\tilde{\Gamma} - 1 \right) \frac{\partial \pi_X(\tau^a)}{\partial \tau^a} \left(\pi_X(\tau^n) - \pi_X(\tau^a)\right)$ to the first summand and subtract it from the second.
 For any particular value of $\tilde{\bta}$, one can choose the $\tilde{\Gamma}$ weight that would make $\tilde{\bta}$ the politically optimal tariff; this makes the derivative in the first summand zero.
@@
 \[
-  e + aW = e + a \left[ \mathit{CS}_X(\tau) + \mathit{PS}_X(\tau) + \mathit{CS}_Y(\tau^*) + \mathit{PS}_Y(\tau^*) + \mathit{TR}(\tau) \right]
+  e + aW = e + a \left[ \mathit{CS}_X(\tau) + \pi_X(\tau) + \mathit{CS}_Y(\tau^*) + \pi_Y(\tau^*) + \mathit{TR}(\tau) \right]
 \]

### 2016-07-04 draft.tex
remove bolding added for last revision

@@
 \end{equation}
-where $\mathit{CS}$ is consumer surplus, $\pi$ are \textbf{profits}, $\ga(e,\ta)$ is the weight placed on \textbf{profits} in the import-competing industry, and $\mathit{TR}$ is tariff revenue.\footnote{Labor income $l$ could also be included in both median legislator and executive welfare.
+where $\mathit{CS}$ is consumer surplus, $\pi$ are profits, $\ga(e,\ta)$ is the weight placed on profits in the import-competing industry, and $\mathit{TR}$ is tariff revenue.\footnote{Labor income $l$ could also be included in both median legislator and executive welfare.
 I omit it because its inclusion alters none of the results and only serves to complicate the exposition.} I model the decisions of the legislature as being taken by a median legislator with the weight the median legislator places on import-competing industry profits affected by the level of lobbying effort $e$ and a random variable $\ta$.
@@
 \end{equation}
-where $\pi(\cdot)$ is the current-period profit \textbf{and the subscript $X$ has been dropped for notational convenience}.
+where $\pi(\cdot)$ is the current-period profit and the subscript $X$ has been dropped for notational convenience.
 
@@
 
-The lower world price for home's export good has a larger negative effect on \textbf{profits} than the positive effect on consumer surplus.
+The lower world price for home's export good has a larger negative effect on profits than the positive effect on consumer surplus.
 The net effect of an increase in foreign trade-agreement tariffs on home legislative welfare is negative, leading to an increased probability that the trade agreement will be broken.
@@
 
-\noindent The price of good $Y$ is decreasing in $\tau^{*a}$, so consumer surplus is increasing and \textbf{profits} are decreasing in $\tau^{*a}$.
-Because $Y$ is being exported, the decrease in \textbf{profits} is larger than the increase in consumer surplus, making the numerator negative.
-Since \textbf{profits} in the denominator are increasing in $\tau$, the expression in Equation~\ref{appex:astar} is negative for all $\tau^{*a}$.
+\noindent The price of good $Y$ is decreasing in $\tau^{*a}$, so consumer surplus is increasing and profits are decreasing in $\tau^{*a}$.
+Because $Y$ is being exported, the decrease in profits is larger than the increase in consumer surplus, making the numerator negative.
+Since profits in the denominator are increasing in $\tau$, the expression in Equation~\ref{appex:astar} is negative for all $\tau^{*a}$.
 $\hfill\blacksquare$
@@
 
-Because total social welfare is maximized at $\tau^a = \tau^{*a} = 0$, $W(\btn)-W(\bta)$ is always negative, whereas \textbf{profits} are increasing in $\tau^a$, so the second summand is positive everywhere.
+Because total social welfare is maximized at $\tau^a = \tau^{*a} = 0$, $W(\btn)-W(\bta)$ is always negative, whereas profits are increasing in $\tau^a$, so the second summand is positive everywhere.
 With a positive denominator, we thus have that the derivative is positive on $[0,\tau^O]$.
@@
 Thus the first summand in the numerator is zero at $\bta = \bm{0}$ and increasingly negative as $\bta$ increases.
-The second summand is positive everywhere because social welfare, $W$, is lowest at $\btn$ and \textbf{profits} are increasing everywhere.
+The second summand is positive everywhere because social welfare, $W$, is lowest at $\btn$ and profits are increasing everywhere.
 Thus the numerator is positive at $\bm{0}$ and at least for very small $\bta$
